### 2024-03-07 analysis2_manuscript.tex
stubbing out

\documentclass[hidelinks]{article}
\usepackage[letterpaper,margin=1.0in]{geometry}
\usepackage[utf8]{inputenc}
\pagenumbering{arabic}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage[singlelinecheck=false]{caption} % singlelinecheck makes single line caption left aligned instead of centered
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage[round]{natbib}
\usepackage{fancyhdr}
\usepackage{longtable}
\usepackage{booktabs}
% hyperlinks
\usepackage{hyperref}

\usepackage{xspace}
\usepackage{mathrsfs}
\usepackage{graphicx}

\pagestyle{fancy}
\fancyhead[R]{\textbf{Expanding \stdpopsim}}

% for highlighting text
\usepackage{xcolor}
\usepackage{soul}

% bibliography
\usepackage[round]{natbib}   % omit 'round' option if you prefer square brackets
\bibliographystyle{plainnat}

\newcommand{\Stdpopsim}{\texttt{Stdpopsim}\xspace}
\newcommand{\stdpopsim}{\texttt{stdpopsim}\xspace}

%commands to format figure and table references in the supplement
\newcommand{\beginsupplement}{%
        \fancyhead[L]{Supplemental Material}
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
\newcommand{\stopsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{\arabic{figure}}%
     }

\makeatletter
\newcommand{\labelname}[1]{\def\@currentlabelname{#1}}
\makeatother

% Avoid pandoc bug when there are lists in the body.
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\title{Adding selection to \stdpopsim: oh how the tables have turned}

% \author[1,+]{M. Elise Lauterbur}
% \author[2,*]{Maria Izabel A. Cavassim}
% \author[3,*]{Ariella L. Gladstein}
% \author[4,*]{Graham Gower}
% \author[5*]{Nathaniel S. Pope}
% \author[6,*]{Georgia Tsambos}
% \author[5,7]{Jeff Adrion}
% \author[5]{Saurabh Belsare}
% \author[8]{Arjun Biddanda}
% \author[5]{Victoria Caudill}
% \author[9]{Jean Cury}
% \author[10]{Ignacio Echevarria}
% \author[11]{Benjamin C. Haller}
% \author[12,13]{Ahmed R. Hasan}
% \author[14,15]{Xin Huang}
% \author[16]{Leonardo Nicola Martin Iasi}
% \author[17]{Ekaterina Noskova}
% \author[18]{Jana Obšteter}
% \author[19]{Vitor Antonio Corrêa Pavinato}
% \author[20,21]{Alice Pearson}
% \author[22,23]{David Peede}
% \author[24]{Manolo F. Perez}
% \author[5]{Murillo F. Rodrigues}
% \author[5]{Chris C. R. Smith}
% \author[25]{Jeffrey P. Spence}
% \author[5]{Anastasia Teterina}
% \author[5]{Silas Tittes}
% \author[26]{Per Unneberg}
% \author[27]{Juan Manuel Vazquez}
% \author[28]{Ryan K. Waples}
% \author[29]{Anthony Wilder Wohns}
% \author[30]{Yan Wong}
% \author[31]{Franz Baumdicker}
% \author[32]{Reed A. Cartwright}
% \author[33]{Gregor Gorjanc}
% \author[34]{Ryan N. Gutenkunst}
% \author[30]{Jerome Kelleher}
% \author[5]{Andrew D. Kern}
% \author[35]{Aaron P. Ragsdale}
% \author[5,36]{Peter L. Ralph}
% \author[37]{Daniel R. Schrider}
% \author[38,+]{Ilan Gronau}

% \affil[*]{\small{These authors contributed equally to the paper.}}
% \affil[+]{\small{Corresponding authors: lauterbur@gmail.com ; ilan.gronau@runi.ac.il.}}
% \affil[1]{\small{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ 85719, USA}}
% \affil[2]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
% \affil[3]{\small{Embark Veterinary, Inc., Boston MA 02111, USA}}
% \affil[4]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
% \affil[5]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
% \affil[6]{\small{School of Mathematics and Statistics, University of Melbourne, Australia}}
% \affil[7]{\small{AncestryDNA, San Francisco CA 94107, USA}}
% \affil[8]{\small{54Gene, Inc., Washington DC 20005, USA}}
% \affil[9]{\small{Université Paris-Saclay, CNRS, INRIA, Laboratoire Interdisciplinaire des Sciences du Numérique, UMR 9015 Orsay, France}}
% \affil[10]{\small{School of Life Sciences, University of Glasgow, Glasgow, UK}}
% \affil[11]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
% \affil[12]{\small{Department of Cell and Systems Biology, University of Toronto, Toronto ON, Canada}}
% \affil[13]{\small{Department of Biology, University of Toronto Mississauga, Mississauga ON, Canada}}
% \affil[14]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
% \affil[15]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
% \affil[16]{\small{Department of Evolutionary Genetics, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany}}
% \affil[17]{\small{Computer Technologies Laboratory, ITMO University, St Petersburg, Russia}}
% \affil[18]{\small{Agricultural Institute of Slovenia, Department of Animal Science, Ljubljana, Slovenia}}
% \affil[19]{\small{Entomology Department, The Ohio State University, Wooster OH, USA}}
% \affil[20]{\small{Department of Genetics, University of Cambridge, Cambridge, UK}}
% \affil[21]{\small{Department of Zoology, University of Cambridge, Cambridge, UK}}
% \affil[22]{\small{Department of Ecology, Evolution, and Organismal Biology, Brown University, Providence RI, USA}}
% \affil[23]{\small{Center for Computational Molecular Biology, Brown University, Providence RI, USA}}
% \affil[24]{\small{Department of Genetics and Evolution, Federal University of Sao Carlos, Sao Carlos 13565905, Brazil}}
% \affil[25]{\small{Department of Genetics, Stanford University School of Medicine, Stanford CA 94305, USA}}
% \affil[26]{\small{Department of Cell and Molecular Biology, National Bioinformatics Infrastructure Sweden, Science for Life Laboratory, Uppsala University, Husargatan 3, SE-752 37 Uppsala, Sweden}}
% \affil[27]{\small{Department of Integrative Biology, University of California, Berkeley, Berkeley CA, USA}}
% \affil[28]{\small{Department of Biostatistics, University of Washington, Seattle WA, USA}}
% \affil[29]{\small{Broad Institute of MIT and Harvard, Cambridge MA 02142, USA}}
% \affil[30]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
% \affil[31]{\small{Cluster of Excellence - Controlling Microbes to Fight Infections, Eberhard Karls Universität Tübingen, Tübingen, Baden-Württemberg, Germany}}
% \affil[32]{\small{School of Life Sciences and The Biodesign Institute, Arizona State University, Tempe AZ, USA}}
% \affil[33]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
% \affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson AZ 85721, USA}}
% \affil[35]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
% \affil[36]{\small{Department of Mathematics, University of Oregon, Eugene OR 97402, USA}}
% \affil[37]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
% \affil[38]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}

\date{\small{\today{}}}

\begin{document}

\maketitle

\section*{Abstract}

\section*{Introduction}
    \label{introduction}

\section*{Additions to stdpopsim}
    \label{additions}

\section*{Conclusion}
    \label{conclusion}

\section*{Data availability}\label{data_availability}

\section*{Acknowledgments}\label{acknowledgements}

\section*{Funding}
    \label{funding}

%\bibliography{}
\end{document}

### 2024-03-07 analysis2_manuscript.tex
draft of first two paragraphs

@@
 % \author[3,*]{Ariella L. Gladstein}
-% \author[4,*]{Graham Gower}
-% \author[5*]{Nathaniel S. Pope}
+ \author[4,*]{Graham Gower}
+ \author[5*]{Nathaniel S. Pope}
 % \author[6,*]{Georgia Tsambos}
@@
 
-% \affil[*]{\small{These authors contributed equally to the paper.}}
+ \affil[*]{\small{These authors contributed equally to the paper.}}
 % \affil[+]{\small{Corresponding authors: lauterbur@gmail.com ; ilan.gronau@runi.ac.il.}}
@@
 % \affil[3]{\small{Embark Veterinary, Inc., Boston MA 02111, USA}}
-% \affil[4]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
-% \affil[5]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
+ \affil[4]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
+ \affil[5]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
 % \affil[6]{\small{School of Mathematics and Statistics, University of Melbourne, Australia}}
@@
     \label{introduction}
+As the complexity of population genetic datasets has increased,
+simulation has become a critical tool for interpretation, 
+analysis, and exploration of realistic evolutionary models.
+Simulation in population genetics has a long history 
+including both backward in time coalescent simulations
+\citep{kingman1982genealogy,hudson1983testing, hudson1990gene}
+and forward in time simulations of complex demography and selection
+\citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.
+The development of simulation tools has been driven by the need to
+understand the effects of complex evolutionary processes on genetic
+variation, to provide a null model for hypothesis testing, to
+to explore the power and limitations of statistical methods, and 
+increasingly to provide a basis for machine learning and other
+simulation-based inference methods.
+
+A lingering challenge with simulation in population genetics has been
+reproducibility and the ability to share and compare results among 
+researchers. This challenge has been addressed in part by the development
+of community resources for sharing and distributing simulation software
+via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
+provides a standardized interface for accessing a wide range of
+population genetic models, and has begun to be be widely adopted by the community.
+While that is so, the current version of \stdpopsim does not include
+models of selection, which is a major limitation for many applications
+of population genetic simulation. In particular, modeling selection
+through simulation is critical for understanding the processes such
+as adaptation, the effects of selective sweeps, and the impact of
+background selection on genetic diversity. Ideally, one would like
+to have a single, unified framework for simulating both neutral and
+non-neutral evolutionary processes, and to be able to compare the
+results of these simulations to empirical data in a manner that is
+accessible to a wide range of researchers and highly reproducible. 
+Further the framework should include the complex realities of 
+genomes, including heterogenous recombination rates, 
+variation in the density of functional elements, and relevant
+population size histories.
 
@@
     \label{additions}
+
+    \begin{figure}[ht]
+        \centering
+        \includegraphics[width=\textwidth]{figures/HomSap/estimated_Ne_t_final}
+        \caption{
+        \label{fig:demography}
+        Performance of methods to infer human OOA model
+        with and without background selection on exons.
+        }
+    \end{figure}
+
+    \begin{figure}[ht]
+        \centering
+        \includegraphics[width=\textwidth]{figures/HomSap/dfe.inference.benchmark}
+        \caption{
+        \label{fig:dfe_humans}
+        Performance of methods to infer distribution of fitness effects (DFE).
+        }
+    \end{figure}
+
+    \begin{figure}[ht]
+        \centering
+        \includegraphics[width=\textwidth]{figures/PhoSin/dfe.inference.benchmark}
+        \caption{
+        \label{fig:dfe_vaquita}
+        Performance of methods to infer distribution of fitness effects (DFE).
+        }
+    \end{figure}
 
@@
 
-%\bibliography{}
+\bibliography{references.bib}
 \end{document}

### 2024-05-23 analysis2_manuscript.tex

--- 2024-03-07 analysis2_manuscript.tex
@@
 
+    \begin{figure}[ht]
+        \centering
+        \includegraphics[width=\textwidth]{figures/HomSap/estimated_Ne_t_final}
+        \caption{
+        \label{fig:demography}
+        Performance of methods to infer human OOA model
+        with and without background selection on exons.
+        }
+    \end{figure}
+
+    \begin{figure}[ht]
+        \centering
+        \includegraphics[width=\textwidth]{figures/HomSap/dfe.inference.benchmark}
+        \caption{
+        \label{fig:dfe_humans}
+        Performance of methods to infer distribution of fitness effects (DFE).
+        }
+    \end{figure}
+
+    \begin{figure}[ht]
+        \centering
+        \includegraphics[width=\textwidth]{figures/PhoSin/dfe.inference.benchmark}
+        \caption{
+        \label{fig:dfe_vaquita}
+        Performance of methods to infer distribution of fitness effects (DFE).
+        }
+    \end{figure}
+
 \section*{Conclusion}

### 2024-05-23 analysis2_manuscript.tex

--- 2024-03-07 analysis2_manuscript.tex
@@
     \label{introduction}
+    % natural selection
+    Natural selection is a fundamental force in evolution, shaping the
+    genetic diversity of populations and driving the adaptation of
+    species to their environments. The effects of natural selection
+    on genetic variation are complex, and can be difficult to disentangle
+    from other evolutionary processes such as mutation, recombination,
+    and genetic drift. The impact of selection on genetic variation
+    is often studied using population genetic models, which can be
+    used to generate predictions about the patterns of genetic variation
+
+    % simulation
     As the complexity of population genetic datasets has increased,
@@
 
+    % stdpopsim
     A lingering challenge with simulation in population genetics has been
@@
     population size histories.
+
+\section*{outline}
+    \label{outline}
+    \begin{enumerate}
+        \item Single-population demographic inference methods
+        \begin{enumerate}
+            \item mscm2
+            \item stairwaiplot
+            \item GONE
+            \item SMC++
+        \end{enumerate}
+        \item Multi-population demographic inference methods
+        \begin{enumerate}
+            \item dadi
+            \item fastsimcoal
+            \item momi2
+        \end{enumerate}
+        \item DFE inference methods
+        \begin{enumerate}
+            \item dadi
+            \item polyDFE
+            \item DFE-alpha
+            \item GRAPES
+        \end{enumerate} 
+        \item sweeps
+        \begin{enumerate}
+            \item compare effect of BGS on sweep detection
+            \item compare power in different pops
+            \item compare methods
+        \end{enumerate}
+    \end{enumerate}
 

### 2024-05-23 analysis2_manuscript.tex
intro clean up

@@
     from other evolutionary processes such as mutation, recombination,
-    and genetic drift. The impact of selection on genetic variation
-    is often studied using population genetic models, which can be
-    used to generate predictions about the patterns of genetic variation
+    and genetic drift \cite[e.g.,][]{gillespie1991causes}.
+    For instance changes in population size can lead to fluctuations
+    in genetic diversity across a recombining chromosome 
+    that can mimic the effects of selection \citep{simonsen1995properties, barton1998effect},
+    and lead to spurious inferences about the strength and targets of genetic adaptation
+    \cite{simonsen1995properties,akey2004population,nielsen2005genomic}.
+    In turn selection can confound our ability to infer demographic 
+    history from allele frequencies \citep{ewing2016consequences} and
+    estimates of inverse coalescent rate \citep{schrider2016effects}.
+    Thus it is imperative to joingly account for the effects of selection
+    and demography when inferring evolutionary history from genetic data
+    however this is a challenging task from a modeling perspective.
 
     % simulation
-    As the complexity of population genetic datasets has increased,
-    simulation has become a critical tool for interpretation, 
+    To meet this need the field of population genetics
+    has turned to simulation.
+    Simulation has become a critical tool for interpretation, 
     analysis, and exploration of realistic evolutionary models.
@@
     understand the effects of complex evolutionary processes on genetic
-    variation, to provide a null model for hypothesis testing, to
-    to explore the power and limitations of statistical methods, and 
-    increasingly to provide a basis for machine learning and other
-    simulation-based inference methods.
+    variation \citep[e.g.,][]{galloway2020few}, to provide a null model for hypothesis testing
+    \citep[e.g.,][]{hudson1992statistical}, to
+    to explore the power and limitations of statistical methods \cite[e.g.,]{przeworski2002signature},
+    and increasingly to provide a basis for machine learning and other
+    simulation-based inference methods \citep[e.g.,]{kern2018diplos}.
+    While this is so, joint simulation of complex demography and selection
+    is challenging, and requires a deep understanding of the underlying
+    evolutionary processes, as well as a number of parameter choices including
+    the strength of selection, the distribution of fitness effects, and the
+    recombination rate, as well as a parameterized model of demography.
+    This is dounting task for many researchers, and can be a barrier to
+    the adoption of simulation-based methods in population genetics.
+    Further, sharing and comparing simulation results among studies can 
+    itself be challenging, as different simulation tools may use different
+    models and parameterizations, and may not be directly comparable.
+    This is a major limitation for the field, as it can make it difficult
+    to assess the robustness of results, and to compare the results of
+    different studies.
 

### 2024-05-29 analysis2_manuscript.tex
working on results and fig 1

@@
     reproducibility and the ability to share and compare results among 
-    researchers. This challenge has been addressed in part by the development
+    researchers \citep[][e.g.,]{ragsdale2020lessons}.
+    This challenge has been addressed in part by the development
     of community resources for sharing and distributing simulation software
@@
     provides a standardized interface for accessing a wide range of
-    population genetic models, and has begun to be be widely adopted by the community.
+    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
     While that is so, the current version of \stdpopsim does not include
-    models of selection, which is a major limitation for many applications
-    of population genetic simulation. In particular, modeling selection
+    models of selection, which is a major limitation for more empirical
+    applications of population genetic simulation. In particular, modeling selection
     through simulation is critical for understanding the processes such
@@
     results of these simulations to empirical data in a manner that is
-    accessible to a wide range of researchers and highly reproducible. 
+    both accessible to a wide range of researchers and highly reproducible. 
     Further the framework should include the complex realities of 
@@
     population size histories.
+
+    % goals
+    In this manuscript we provide an overview of a major new addition
+    to \stdpopsim-- the inclusion of models of selection.
+    We begin by describing the models of selection that we have implemented,
+    as well as the parameter choices that are available to the user.
+    This includes models of background selection, selective sweeps, and
+    models of the distribution of fitness effects.
+    Further we describe how these models can be combined with genomic
+    annotations that are specific to a given species available
+    through the \stdpopsim API, and combined with a parameterized model of
+    demography and recombination to provide a complete simulation of
+    genetic variation in a population.
+    We then provide a series of examples of how these models can be used
+    to benchmark and compare the performance of different methods for
+    inferring demographic history, selection, and the distribution of
+    fitness effects from genetic data.
+
+\section*{Implementing Selection in \stdpopsim}
+    \label{selection}
+
+    % implementing selection
+    \stdpopsim previously provided a wide range of models of neutral
+    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
+    This was accomplished through the use of a standardized interface
+    and data structure that allowed for curated addition of new
+    species and models, and relied on backend simulation engines,
+    including msprime \citep{Baumdicker2022} and SLiM \citep{haller2019slim},
+    for efficient simulation of genetic variation.
+    To extend this to include models of selection, 
+    we introduce an important new class representing a specfic distribution of fitness effects
+    to the \stdpopsim API, the \texttt{DFE} class. 
+    This class provides an interface for specifying
+    a distribution of the fitness effects of new mutations, and
+    is added to a \texttt{Contig} object to provide a description
+    of the fitness effects of mutations along a chromosome.
+    A \texttt{DFE} object can apply to a specific interval set of a chromosome,
+    and can be combined with other \texttt{DFE} objects to provide, 
+    a rich model of how selection may differ along a chromosome. 
+    Once specified a model of selection is then run in the same way
+    but currently is limited to using SLiM simulation engine, which
+    is probably the most flexible simulation engine for modeling selection currently available.
+
+    As with other \stdpopsim models, we aim to hew as closely as possible
+    to realistic genomes, and provide public annotations of specices-specific functional
+    elements that can be used along with a \texttt{DFE} object to
+    readily provide realistic model of selection for a specific genome and set of
+    populations. 
+    % A figure here would be great to show how this works
+
+    % current numbers of DFEs / species / etc
+    the \stdpopsim
 

### 2024-05-29 analysis2_manuscript.tex

@@
 
+    \begin{figure}[t!]
+        \centering
+        \includegraphics[width=0.2\textwidth]{figures/schematics/catalog.pdf}
+        \caption{
+        \label{fig:schematic}
+        A schematic of the \stdpopsim catalog for simulating genetic variation
+        in a population. The user specifies a species, a portion of the genome to simulate,
+        and optionally a genetic map, a model of demography, and a model of selection that 
+        itself is composed of a distribution of fitness effects (DFE) and a set of functional
+        annotations.
+        }
+    \end{figure}
+
     As with other \stdpopsim models, we aim to hew as closely as possible
@@
     readily provide realistic model of selection for a specific genome and set of
-    populations. 
-    % A figure here would be great to show how this works
+    populations. A schematic of the \stdpopsim catalog is shown in Figure \ref{fig:schematic}.
+    From a users perspective, one can specify a species, a portion of the genome to simulate,
+    a genetic map if available, a model of demography, and a model of selection that is
+    composed of a set of annotations and a \texttt{DFE} object.
+
+    While the vast majority of published DFE estimates are based on a limited number of species,
+    in particular humans and \textit{Drosophila}, these are mix and match with the other species
+    in the catalog. For instance, one could simulate a butterfly species with a Gamma DFE originally
+    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
 
@@
 
-\section*{Additions to stdpopsim}
+\section*{Example applications}
     \label{additions}

### 2024-11-11 analysis2_manuscript.tex
first two paras

--- 2024-05-23 analysis2_manuscript.tex
@@
 \usepackage{amsmath}
-\usepackage[round]{natbib}
 \usepackage{fancyhdr}
@@
 \usepackage{hyperref}
-
+\usepackage{wrapfig}
 \usepackage{xspace}
@@
 \usepackage{graphicx}
+\usepackage{lipsum}
+\usepackage{makecell}
 
 \pagestyle{fancy}
-\fancyhead[R]{\textbf{Expanding \stdpopsim}}
+\fancyhead[R]{\textbf{\stdpopsim with selection}}
 
@@
 % bibliography
-\usepackage[round]{natbib}   % omit 'round' option if you prefer square brackets
-\bibliographystyle{plainnat}
+\usepackage[
+    backend=biber,
+    natbib=true,
+    style=authoryear,
+    ]{biblatex}
+\addbibresource{references.bib}
+
+%\usepackage[round]{natbib}   % omit 'round' option if you prefer square brackets
+%\bibliographystyle{plainnat}
 
@@
 
+% add commands for msmc2, stairway plot, gone, and smc++
+\newcommand{\msmc}{\texttt{msmc2}\xspace}
+\newcommand{\stairway}{\texttt{stairwayplot}\xspace}
+\newcommand{\gone}{\texttt{GONE}\xspace}
+\newcommand{\smcpp}{\texttt{SMC++}\xspace}
+
+% add commands for the DFE inference methods
+\newcommand{\dfe}{\texttt{dfe-alpha}\xspace}
+\newcommand{\polydfe}{\texttt{polyDFE}\xspace}
+\newcommand{\dadi}{$\partial$\texttt{a}$\partial$\texttt{i}\xspace}
+\newcommand{\grapes}{\texttt{GRAPES}\xspace}
+
+% add commands for the sweep detection methods
+\newcommand{\sweepfinder}{\texttt{sweepfinder2}\xspace}
+\newcommand{\diploshic}{\texttt{diploshic}\xspace}
+
 % Avoid pandoc bug when there are lists in the body.
@@
 % \author[1,+]{M. Elise Lauterbur}
-% \author[2,*]{Maria Izabel A. Cavassim}
 % \author[3,*]{Ariella L. Gladstein}
@@
  \author[5*]{Nathaniel S. Pope}
-% \author[6,*]{Georgia Tsambos}
+ \author[5*]{Murillo F. Rodrigues}
+\author[6]{Georgia Tsambos}
+\author[34]{Linh N. Tran}
+\author[2]{Maria Izabel A. Cavassim}
 % \author[5,7]{Jeff Adrion}
@@
 % \author[12,13]{Ahmed R. Hasan}
-% \author[14,15]{Xin Huang}
+\author[14,15]{Xin Huang}
 % \author[16]{Leonardo Nicola Martin Iasi}
@@
 % \author[24]{Manolo F. Perez}
-% \author[5]{Murillo F. Rodrigues}
-% \author[5]{Chris C. R. Smith}
+\author[5]{Chris C. R. Smith}
 % \author[25]{Jeffrey P. Spence}
 % \author[5]{Anastasia Teterina}
-% \author[5]{Silas Tittes}
+\author[5]{Silas Tittes}
+\author[5]{Scott T. Small}
 % \author[26]{Per Unneberg}
@@
 % \author[33]{Gregor Gorjanc}
-% \author[34]{Ryan N. Gutenkunst}
-% \author[30]{Jerome Kelleher}
-% \author[5]{Andrew D. Kern}
-% \author[35]{Aaron P. Ragsdale}
-% \author[5,36]{Peter L. Ralph}
-% \author[37]{Daniel R. Schrider}
-% \author[38,+]{Ilan Gronau}
+\author[4]{Kirk E. Lohmueller}
+\author[34]{Ryan N. Gutenkunst}
+\author[30]{Jerome Kelleher}
+\author[35]{Aaron P. Ragsdale}
+
+\author[37]{Daniel R. Schrider}
+\author[38]{Ilan Gronau}
+\author[5,36]{Peter L. Ralph}
+\author[5]{Andrew D. Kern}
 
@@
 % \affil[29]{\small{Broad Institute of MIT and Harvard, Cambridge MA 02142, USA}}
-% \affil[30]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
+\affil[30]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
 % \affil[31]{\small{Cluster of Excellence - Controlling Microbes to Fight Infections, Eberhard Karls Universität Tübingen, Tübingen, Baden-Württemberg, Germany}}
@@
 % \affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson AZ 85721, USA}}
-% \affil[35]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
-% \affil[36]{\small{Department of Mathematics, University of Oregon, Eugene OR 97402, USA}}
-% \affil[37]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
-% \affil[38]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
+\affil[35]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
+\affil[36]{\small{Department of Mathematics, University of Oregon, Eugene OR 97402, USA}}
+\affil[37]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
+\affil[38]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
 
@@
     that can mimic the effects of selection \citep{simonsen1995properties, barton1998effect},
+    % DRS -- I'm not sure Barton reference is right for this?
     and lead to spurious inferences about the strength and targets of genetic adaptation
@@
     In turn selection can confound our ability to infer demographic 
-    history from allele frequencies \citep{ewing2016consequences} and
-    estimates of inverse coalescent rate \citep{schrider2016effects}.
+    history from allele frequencies \citep{ewing2016consequences,schrider2016effects} and
+    estimates of inverse coalescent rate \citep{schrider2016effects, johri2021impact, cousins2024accurate}.
     Thus it is imperative to joingly account for the effects of selection
-    and demography when inferring evolutionary history from genetic data
-    however this is a challenging task from a modeling perspective.
+    and demography when inferring evolutionary history from genetic data \citep{sheehan2016deep,johri2020toward}.
+    However this is a challenging task from a modeling perspective \citep{johri2022prospect}.
 
     % simulation
-    To meet this need the field of population genetics
-    has turned to simulation.
-    Simulation has become a critical tool for interpretation, 
-    analysis, and exploration of realistic evolutionary models.
+    % DRS: lots more refs will be needed in this paragraph. added a few and am happy to help add more later if needed
+    To meet the growing need for the interpretation,
+    analysis, and exploration of realistic and complex evolutionary models
+    the field of population genetics has increasingly turned to simulation.
     Simulation in population genetics has a long history 
-    including both backward in time coalescent simulations
+    including both backward-in-time coalescent simulations
     \citep{kingman1982genealogy,hudson1983testing, hudson1990gene}
-    and forward in time simulations of complex demography and selection
+    and forward-in-time simulations of complex demography and selection
     \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.
@@
     variation \citep[e.g.,][]{galloway2020few}, to provide a null model for hypothesis testing
-    \citep[e.g.,][]{hudson1992statistical}, to
-    to explore the power and limitations of statistical methods \cite[e.g.,]{przeworski2002signature},
+    \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting}, to
+    to explore the power and limitations of statistical methods \cite[e.g.,][]{przeworski2002signature},
     and increasingly to provide a basis for machine learning and other
-    simulation-based inference methods \citep[e.g.,]{kern2018diplos}.
+    simulation-based inference methods \citep[e.g.,][]{pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}.
     While this is so, joint simulation of complex demography and selection
@@
     recombination rate, as well as a parameterized model of demography.
-    This is dounting task for many researchers, and can be a barrier to
+    This is a daunting task for many researchers, and can be a barrier to
     the adoption of simulation-based methods in population genetics.
-    Further, sharing and comparing simulation results among studies can 
-    itself be challenging, as different simulation tools may use different
-    models and parameterizations, and may not be directly comparable.
+    Furthermore, sharing and comparing simulation results among studies can 
+    itself be challenging, as different simulation tools may use different structures
+    and parameter scalins for their models, and may not be directly comparable.
     This is a major limitation for the field, as it can make it difficult
@@
     % stdpopsim
+    % DRS: lots more refs will be needed in this paragraph. added a few and am happy to help add more later if needed
     A lingering challenge with simulation in population genetics has been
     reproducibility and the ability to share and compare results among 
-    researchers. This challenge has been addressed in part by the development
+    researchers \citep[e.g.,][]{ragsdale2020lessons}.
+    This challenge has been addressed in part by the development
     of community resources for sharing and distributing simulation software
@@
     provides a standardized interface for accessing a wide range of
-    population genetic models, and has begun to be be widely adopted by the community.
-    While that is so, the current version of \stdpopsim does not include
-    models of selection, which is a major limitation for many applications
-    of population genetic simulation. In particular, modeling selection
-    through simulation is critical for understanding the processes such
-    as adaptation, the effects of selective sweeps, and the impact of
-    background selection on genetic diversity. Ideally, one would like
+    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
+    While that is so, the original version of \stdpopsim did not include
+    models of selection, which is a major limitation for more empirical
+    applications of population genetic simulation. In particular, modeling selection
+    through simulation is critical for understanding processes such
+    as adaptation, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
+    background selection on genetic diversity
+    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
+    Ideally, one would like
     to have a single, unified framework for simulating both neutral and
@@
     results of these simulations to empirical data in a manner that is
-    accessible to a wide range of researchers and highly reproducible. 
+    both accessible to a wide range of researchers and highly reproducible. 
     Further the framework should include the complex realities of 
     genomes, including heterogenous recombination rates, 
-    variation in the density of functional elements, and relevant
-    population size histories.
+    variation in the size and density of functional elements, and
+    non-equilibrium demographic histories. Ideally, one should be able
+    to model all of these features after estimates/annotations obtained
+    from a species of interest \citep[e.g.][]{schrider2020background}.
 
-\section*{outline}
-    \label{outline}
-    \begin{enumerate}
-        \item Single-population demographic inference methods
-        \begin{enumerate}
-            \item mscm2
-            \item stairwaiplot
-            \item GONE
-            \item SMC++
-        \end{enumerate}
-        \item Multi-population demographic inference methods
-        \begin{enumerate}
-            \item dadi
-            \item fastsimcoal
-            \item momi2
-        \end{enumerate}
-        \item DFE inference methods
-        \begin{enumerate}
-            \item dadi
-            \item polyDFE
-            \item DFE-alpha
-            \item GRAPES
-        \end{enumerate} 
-        \item sweeps
-        \begin{enumerate}
-            \item compare effect of BGS on sweep detection
-            \item compare power in different pops
-            \item compare methods
-        \end{enumerate}
-    \end{enumerate}
+    % goals
+    In this manuscript we provide an overview of a major new addition
+    to \stdpopsim---the inclusion of models of selection.
+    We begin by describing the models of selection that we have implemented,
+    as well as the parameter choices that are available to the user.
+    This includes models of background selection, selective sweeps, and
+    models of the distribution of fitness effects (DFE).
+    Further we describe how these models can be combined with genomic
+    annotations that are specific to a given species available
+    through the \stdpopsim API, and combined with a parameterized model of
+    demography and recombination to provide a realistic simulation of
+    genetic variation in a population.
+    We then provide a series of examples of how these models can be used
+    to benchmark and compare the performance of different methods for
+    inferring demographic history, the targets of recent positive selection,
+    and the distribution of fitness effects from genetic data.
 
-\section*{Additions to stdpopsim}
-    \label{additions}
+\section*{Implementing Selection in \stdpopsim}
+    \label{selection}
 
-    \begin{figure}[ht]
+    % implementing selection
+    \stdpopsim previously provided a wide range of models of neutral
+    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
+    This was accomplished through the use of a standardized interface
+    and data structure that allows for curated addition of new
+    species and models, and can use either
+    msprime \citep{Baumdicker2022} or SLiM \citep{haller2019slim}
+    as the backend engine for efficient population genetic simulation.
+    To extend \stdpopsim's functionality to include models of selection, 
+    we introduce an important new class representing a specfic distribution of fitness effects
+    to the \stdpopsim API: the \texttt{DFE} class. 
+    This class provides an interface for specifying
+    a distribution of the fitness effects of new mutations, and
+    is added to a \texttt{Contig} object to provide a description
+    of the fitness effects of mutations along a chromosome.
+    A \texttt{DFE} object can apply to a specific interval set of a chromosome,
+    and can be combined with other \texttt{DFE} objects to provide, 
+    a rich model of how selection may differ along a chromosome. 
+    Once specified a model of selection is then run in the same way
+    but currently is limited to using SLiM simulation engine, which
+    is perhaps the most flexible simulation engine for modeling selection currently available.
+
+    As with other \stdpopsim models, when incorporating selection we aim to hew as closely as possible
+    to realistic genomes, and provide public annotations of specices-specific functional
+    elements that can be used along with a \texttt{DFE} object to
+    readily provide realistic model of selection for a specific genome and set of
+    populations. A schematic of the \stdpopsim catalog is shown in Figure \ref{fig:schematic}.
+    From a users perspective, one can specify a species, a portion of the genome to simulate,
+    a genetic map if available, a model of demography, and a model of selection that is
+    composed of a set of annotations and a \texttt{DFE} object.
+
+    % DRS: This figure could really use some Python API and comand line examples like in Figs 1B,C
+    % form the original stdpopsim paper
+    \begin{wrapfigure}[]{l}{0.4\textwidth}
+        \vspace{-0.0cm}
+        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
+        \caption{\label{fig:schematic}
+        A schematic of the \stdpopsim catalog for simulating genetic variation
+        in a population. The user specifies a species, a portion of the genome to simulate,
+        and optionally a genetic map, a model of demography, and a model of selection that 
+        itself is composed of a distribution of fitness effects (DFE) and a set of functional
+        annotations.}
+    \end{wrapfigure}
+
+    The vast majority of published DFE estimates are based on a limited number of species,
+    so \stdpopsim currently has implemented published DFEs from
+    four species: \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
+    humans, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
+    While that is so, these DFEs can be mixed and matched with the other species
+    in the catalog. For instance, one could simulate a butterfly species with a Gamma DFE originally
+    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
+    Furthermore the user can specify a custom DFE, and provide their own annotations
+    of functional elements to simulate selection in a species for which we do not yet have 
+    a published DFE included in the catalog. This flexibility allows for a wide range of
+    models of selection to be simulated, and for the results of these simulations to be
+    compared to empirical data or used as a null model for hypothesis testing.
+
+    % sweep interface
+    In addition to the \texttt{DFE} class, we have also implemented a class (\texttt{stdpopsim.ext.selective\_sweep})
+    that enables selective sweeps to be simulated in \stdpopsim.
+    This class augments a demographic model with an ``extended event''
+    which conditions on the introduction of a selected mutation at a given time and position
+    and with a given selection coefficient. Further the user can specify the minimum frequency
+    of the selected allele at the time of sampling. As these extended events are implemented
+    on top of the existing \stdpopsim API, they can be combined with other models of selection
+    and demography to provide a rich model of the combined effects of multiple disparate evolutionary processes
+    on genetic variation in a population.
+
+    % current numbers of DFEs / species / etc
+    At this release the \stdpopsim catalog comprises 24 species (i.e. genome representations)
+    for which we have implemented XX demographic models, have XX genetic maps, and XX DFEs.
+    Additions to the catalog are ongoing -- we point the reader to our previously published 
+    report detailing this effort \citep{lauterbur2023expanding} and we welcome contributions from the
+    community---we encourage those interested in contributing to visit https://popsim-consortium.github.io/stdpopsim-docs
+    or contact one of the PopSim Consortium members for guidance.
+
+\section*{Example Applications with Selection}
+    \label{applications}
+    In this section we provide a series of examples of how the new models of selection
+    in \stdpopsim can be used to benchmark and compare the performance of different
+    methods for population genetic inference.
+    We focus on three main areas: demographic inference in the 
+    single and multi-population settings, inference of the distribution of fitness effects,
+    and the detection of selective sweeps, particularly in the context of background selection.
+
+    \section*{Inference of $N_e(t)$ in the context of selection}
+    One of the most common applications of population genetic inference is to estimate
+    the effective population size over time, $N_e(t)$, from genetic data. This can be done
+    using a variety of methods, including the pairwise sequentially Markovian coalescent
+    (PSMC) \citep{li2011inference}, through use of the site frequency spectrum (SFS),
+    as well as through identity by decent information. 
+    While that is so, it has been well described that selection, because of its effects on
+    local and more global patterns of genetic variation, can drive estimates of $N_e(t)$
+    further from true census population sizes simply because selection increases the 
+    rate of coalescence in the genome.
+
+    Using \stdpopsim we can simulate genetic data under a model of selection and then 
+    examine the performance of different methods for inferring $N_e(t)$ from these simulated 
+    data. In Figure \ref{fig:1pop-human-demography} we show the results of such a benchmark
+    where we have simulated genomes under a model of human out of Africa (OOA) demography
+    with and without background selection on exons. For this simulation we have used
+    a genetic map from the HapMap Project \citep{international2007second} along with a
+    DFE for deleterious alleles (estimated by \cite{huber2017determining})
+    that acts on exonic positions defined from the HAVANA group release 104 for the human genome. 
+    These simulations were then used to estimate $N_e(t)$ using four methods: 1) \msmc \citep{Schiffels2020}, 
+    2)\stairway method \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
+    In this and what follows we have performed 3 replicate whole genome simulations, so uncertainty in estimates might
+    be roughly assessed, however we are mindful of wasted compute as thorough benchmarking
+    isn't the main purpose of this manuscript.
+
+    In Figure \ref{fig:1pop-human-demography} we show six panels, each of which shows the estimated $N_e(t)$
+    from each independent sample in a five-population model of human out-of-Africa demography
+    from Ragsdale \textit{et al.} (2019)
+    with samples taken from three modern geographic groups (CEU, CHB, and YRI from left to right).
+    The top row shows the results of simulations without selection, while the bottom row shows the results
+    of simulations with a Gamma distributed DFE for the deleterious effects of mutations occurring exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
+    in various colors for the four methods. Dotted lines on the bottom row are to visually reinforce that inference is done
+    from a model with selection. We see that these methods generally perform reasonably well in the absence of selection, 
+    although GONE seems to overestimate $N_e(t)$ in the more recent past for the CEU and CHB populations.
+    Presumably this could be due to patterns of linkage disequilibrium that result from the combination
+    of bottleneck and growth in these populations.
+    In the presence of selection, different methods behave differently. 
+    For instance stairwaiplot which uses the SFS to infer $N_e(t)$ seems to slightly underestimate $N_e(t)$
+    but is reasonably close to the true $N_e(t)$ in the absence of selection.
+    Likewise \smcpp seems to be more robust to the presence of selection,
+    but there is some added noise to the returned estimates. 
+    \msmc seems to be the most affected by the presence of selection, 
+    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$.
+    % DRS: I don't agree with the interpretation in the previous two sentences (starting with "Likewise").
+    % MSMC is already barfing a little bit even without selection, and it is not clear to me from these
+    % plots that selection makes things any worse. Also, SMC++ is doing quite well without selection so maybe
+    % the minor drop in accuracy is meaningful in relative terms.
+
+    \begin{figure}[t]
         \centering
-        \includegraphics[width=\textwidth]{figures/HomSap/estimated_Ne_t_final}
+        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
         \caption{
-        \label{fig:demography}
-        Performance of methods to infer human OOA model
-        with and without background selection on exons.
+        \label{fig:1pop-human-demography}
+        Performance of methods to infer $N_e(t)$ from a human out-of-Africa model \citep{ragsdale2019models}
+        with and without background selection on exons. The top row shows estimates of $N_e(t)$ from simulations
+        without selection, while the bottom row shows estimates of $N_e(t)$ from simulations with a Gamma distributed   
+        DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
+        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.  
         }
@@
 
-    \begin{figure}[ht]
+    To highlight the ease of comparisons between different species and model enabled by \stdpopsim,
+    we have also performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
+    In Figure \ref{fig:1pop-vaquita-demography} we show the results of $N_e(t)$ inference
+    where we have simulated genomes under a model of vaquita demography from \textcite{robinson2022critically}.
+    For these simulations we have used the genome structure and exon annotations from the vaquita genome assembly
+    and a DFE from \textcite{robinson2022critically} that acts on exonic positions, however no genetic map 
+    is currently available, so recombination rates are assumed to be constant across the genome.
+    The demographic model is a simple two-epoch model where population size has declined in the recent past. 
+    In this case we see that all methods perform well both in the absence and presence of selection,
+    with a close match between the true $N_e(t)$ and the estimated $N_e(t)$ in all cases.
+
+\begin{figure}[t]
     \centering
-        \includegraphics[width=\textwidth]{figures/HomSap/dfe.inference.benchmark}
+    \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final}
     \caption{
-        \label{fig:dfe_humans}
-        Performance of methods to infer distribution of fitness effects (DFE).
+    \label{fig:1pop-vaquita-demography}
+    Performance of methods to infer $N_e(t)$ from simulations of the vaquita porpoise genome under a single population
+    model of declining population size \citep{robinson2022critically} with and without background selection on exons. 
+    The top panel shows estimates of $N_e(t)$ from simulations
+    without selection, while the bottom panel shows estimates of $N_e(t)$ from simulations with a Gamma distributed   
+    DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
+    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.  
     }
@@
 
-    \begin{figure}[ht]
+\section*{Estimation of the Distribution of Fitness effects}
+    \label{dfe}
+    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE)
+    from genetic data. Here we show how the new models of selection in \stdpopsim can be used to benchmark and compare
+    the performance of different methods for inferring the DFE from genetic data. Here we have performed simulations using the 
+    human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
+    To remind the reader, the model of selection in each case is a Gamma distributed DFE for deleterious mutations.
+
+    \begin{figure}[htbp]
         \centering
-        \includegraphics[width=\textwidth]{figures/PhoSin/dfe.inference.benchmark}
+        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_OutOfAfricaArchaicAdmixture_5R19_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
+        \caption{A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
+        Simulations were performed using a human out-of-Africa model \citep{ragsdale2019models} with a Gamma distributed DFE
+        acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
+        parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
+        The human model has three extant populations (CEU, CHB, and YRI), and parameter estimates from each
+        sample are shown in different colors.
+        Estimates from three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
+        and 3) \dadi \citep{gutenkunst2009inferring}.}
+        \label{fig:homsap-dfe.ooa}
+    \end{figure}
+
+    In Figures \ref{fig:homsap-dfe.constant} and \ref{fig:homsap-dfe.ooa} we show a comparison of estimates
+    from three software packages for inferring the DFE from genetic data-- 
+    \dadi \citep{gutenkunst2009inferring}, \polydfe \citep{tataru2020polydfe}, 
+    and \grapes \citep{galtier2016adaptive}. 
+    Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
+    while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
+    In a constant size population, all methods perform reasonably well, although \dadi seems to
+    perform slightly worse than the other methods examined, at least for the small number of replicates we simulated here.
+    Notably only \polydfe overlapped the true mean selection coefficient, whereas 
+    the other methods slightly underestimated the strength of selection. 
+    For the shape parameter in constant size simulations \grapes and \polydfe are roughly
+    equivalent in their performance whereas \dadi seems to yield overestimates.
+
+    In the out-of-Africa model (fig \ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
+    In particular variation in accuracy of the parameters estimated in each
+    sample population depends on the methods used, and each method seems to 
+    have its own biases with respect to the way in which population history 
+    affects parameter estimates. For instance for \dadi we see that estimates of 
+    the shape parameter are most accurate in the YRI sample, while for \polydfe
+    estimates are consistently least accurate in the YRI sample. 
+    With respect to the strength of selection, as in constant size populations,
+    all methods are performing reasonably well, although \dadi seems to
+    trail in terms of accuracy.
+
+    %add PhoSin DFE figure here
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
+\includegraphics{figures/PhoSin/Vaquita2Epoch_1R22/hs_hist.pdf}
         \caption{
-        \label{fig:dfe_vaquita}
-        Performance of methods to infer distribution of fitness effects (DFE).
+        \label{fig:vaquita-dfe}
+        A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
+        Simulations were performed using a model of vaquita porpoise demography \citep{robinson2022critically} with a Gamma distributed DFE
+        acting on exons parameterized by a mean selection coefficient, a shape parameter, and a relationship between selection and domiance. Estimates of the 
+        parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
+        This vaquita model has a single extant population, labelled pop 0, and parameter estimates from each
+        of the three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
+        and 3) \dadi \citep{gutenkunst2009inferring,kim2017inference}.
+        C) Shown is the simulated distribution of $2 h s$, where $h$ is the dominance coefficient, along with the simulated distribution of $s$ and the median inferred DFEs.
+The distribution of $2 h s$ is multimodal because the simulated relationship between $h$ and $s$ is not continuous.
+        % RNG: For C, need to extract inferred values from pipeline.
         }
@@
 
-\section*{Conclusion}
-    \label{conclusion}
+    In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
+    from the same three software packages for inferring the DFE from genetic data, but for simulations of the
+    vaquita porpoise genome. In the constant size model (fig \ref{fig:vaquita-dfe.constant}) we see that all methods
+    perform uniformly worse that in the human genome simulations, which consistent underestimation of the mean selection coefficient    
+    and overestimation of the shape parameter. 
+A critical distinction is that the previous human simulations employed a DFE that assumed all selected mutations had additive effects (no dominance, $h = 1/2$), whereas the vaquita simulations use a DFE in which more deleterious mutations are more recessive ($h < 1/2$).
+For all three software packages, the inferred mean and shape of the distribution of selection coefficients is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient and the selection coefficient (Fig.~\ref{fig:vaquita-dfe}C).
+This is likely because deleterious alleles are typically at low frequency and thus heterozygous, where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
+Strongly deleterious alleles are often observed to be recessive in many species \citep{}, so these results suggest caution in interpreting DFE inferences.
+%RNG: Could use Kirk's help with the best references here
+
+\section*{Detection of Selective Sweeps}
+    \label{sweeps}
+    We next highlight how the new models of selection in \stdpopsim can be used to benchmark and compare
+    the performance of different methods for detecting selective sweeps from genetic data. 
+    Here we have performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring}
+    with purifying selection acting on exons under the Gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second},    
+    but overlay on top of this a model of selective sweeps. In particular we have conditioned simulations on the introduction of a beneficial mutation
+    at a given location (to be varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
+    For each sweep location we performed 200 replicate simulations, sampling from them only if the selected allele reached a frequency of
+    0.95 or greater. Rather than simulate the entire chromosome arm, replicate simulations were performed on 5cM regions of the chromosome
+    at each of 100 different locations. As recombination rate varies across the genome, the size of the region in base pairs simulated varies,
+    however the recombination distance, which is critical for the linked effects of a sweep, is held constant.
+
+    %sweeep power figure
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.8 \textwidth]{figures/sweeps/chr1_power.pdf}
+        \caption{
+        Power to detect selective sweeps at 100 locations along human chromosome 1 under a three population out-of-Africa model \citep{gutenkunst2009inferring}  
+        with and without purifying selection acting on exons under the Gamma-distributed DFE described above. 
+        The HapMap II genetic map \citep{international2007second} was used in each simulation. 
+        Single beneficial mutations were introduced at each location with a selection coefficient of $s = 0.03$,
+        and only sampled if they reached a terminal frequency of 0.95 or greater. 
+        Three methods for detecting sweeps were used: 1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row), 
+        2) \diploshic \citep{kern2018diplos} (middle row), and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity (bottom row).
+        True positive rate is shown for these methods for samples from the CEU and YRI samples (left and right respectively). 
+        Under each method we show as a heat map exon density along the chromosome, along with the local recombination rate in cM/Mb.
+        Power is shown as the proportion of replicates in which a sweep was detected at each location. 
+        Each panel shows the results of simulations with purifying selection acting on exons (red; labelled BGS)
+        and without purifying selection acting on exons (blue; labelled Neutral).
+        }
+        \label{fig:chr1_power}
+    \end{figure}
+
+    ** this section needs revision based on new figures that Murillo will add **
+    In Figure \ref{fig:chr1_power} we show the power to detect selective sweeps at 100 locations along human chromosome 1 from
+    the simulations described above. We show the results of three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
+    2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity.
+    A few things are immediately apparent from this figure. First, the power to detect sweeps is generally higher in the YRI sample
+    than in the CEU sample, which is consistent with the fact that the YRI sample has a larger effective population size and thus
+    more variation to detect a sweep. Further the population size of YRI has been relatively constant over the last 100,000 years,
+    thus the out-of-Africa bottleneck has does not mute the signal of a sweep as much as in the CEU sample. This is consistent with
+    what we know about the effects of demography on the power to detect sweeps \citep[e.g.,][]{simonsen1995properties}.
+    Second, the power to detect sweeps is slightly lower in the presence of purifying selection acting on exons but the effect is not
+    as strong as one might expect. This is likely because the strength of selection acting on exons is relatively weak, and the
+    sweep detection methods we have used are relatively robust to the effects of background selection. 
+    Finally and most importantly, the power to detect sweeps varies considerably among regions of the genome. 
+    Indeed variation in power along the chromosome is much greater than the variation in power between the CEU and YRI samples or between
+    models with or without purifying selection. This is likely due to the effects of recombination rate heterogeniety as well
+    as the density of functional elements along the chromosome.
+
+    %sweep power vs recombination rate
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.8 \textwidth]{figures/sweeps/relationship_power_cM.pdf}
+        \caption{
+        Power to detect selective sweeps at 100 locations along human chromosome 1 under a three population out-of-Africa model \citep{gutenkunst2009inferring}  
+        with and without purifying selection as a function of local recombination rate.
+        Simulations are as described in Figure \ref{fig:chr1_power}.
+        Each panel shows the results of simulations with purifying selection acting on exons (red; labelled BGS)
+        and without purifying selection acting on exons (blue; labelled Neutral).
+        True positive rate is shown for these methods for samples from the CEU and YRI samples (left and right respectively).
+        Points represent the power at each location and trend lines along with 95\% confidence intervals are shown fit to those points.
+        }
+        \label{fig:power-recomb}
+    \end{figure}
+
+    To further explore the effects of recombination rate and exon density on the power to detect sweeps, we show in Figure \ref{fig:power-recomb}
+    the relationship between power and local recombination rate for our sweep simulations. We see that power to detect sweeps is generally
+    a decreasing function of local recombination rate, which is consistent with sweeps having a larger genomic footprint in regions of low
+    recombination. We also see that the presence of purifying selection acting on exons has a relatively small effect on the power to detect sweeps,    
+    but one that is present across the entire range of recombinations rates observed in human chromosome 1. 
+    We also note that the different methods for detecting sweeps that we have explored here respond differently to the effects of recombination rate.
+    Unsurprisingly, the empirical cutoff of $\pi$ to detect sweeps is the less robust to the effects of recombination rate than the other two methods,
+    that are specifically designed to find selective sweeps. Whereas \sweepfinder2 has a clear decline in power with increasing recombination rate,
+    \diploshic 's behavior is more variable, with a slight increase in power at intermediate recombination rates in the CEU sample and a decline
+    in power at high recombination rates in the YRI sample.
+
+    In Figure \ref{fig:power-exon} we show the relationship between power and local exon density for our sweep simulations.
+    Here there is little relationship between power and exon density, which is somewhat surprising given that the power to detect sweeps
+    is generally lower in the presence of purifying selection acting on exons. However while this is the case we see that 
+    generally all methods have more power to detect sweeps in regions of lower exon density, which is consistent with the idea that
+    background selection acting on nearby exons can reduce the signature of a sweep. However this is far from a simple pattern:
+    for instance sweepfinder2 seems to have minimum power in regions of intermediate exon density, while \diploshic is roughly constant
+    or perhaps has a slight increase in power at intermediate exon densities. Noise is the estimates of power with respect to exon density
+    are prohibitively high for the empirical cutoff of $\pi$ to detect sweeps, so it's not clear what conclusions can be drawn.
+
+\section*{Discussion}
+    \label{Discussion}
+    In this paper we have presented an important new additional to the \stdpopsim{} library
+    that allows for simulating genetic variation in the presence of selection.
+    We have demonstrated the utility of this new API by showing how it can be used to benchmark
+    the performance of different methods for demographic inference, the inference of the distribution
+    of fitness effects of mutations, and the detection of selective sweeps.
+    In general, this is an important step forward for the field of population genetics,
+    as it allows for a wide range of models of selection to be easily simulated and compared to
+    empirical data, with reproducibility, computational efficiency, and rigor.
+
+    We have shown how the new API can be used to benchmark the performance of different methods
+    although we acknowledge that the benchmarks shown here merely scratch the surface of what
+    can be done with the new API. We envision moving forward that many researchers will use
+    the new API to simulate data under a wide range of models of selection, and to develop
+    new methods for the analysis of genetic data in the presence of selection. 
+    Further the critical step of benchmarking methods for the analysis of genetic data
+    can now be done on common ground, using the same simulations and parameterizations of
+    models of selection in a way that was not possible before.
+
+\section*{Methods}
+    \label{methods}
+
+    \subsection*{Simulations}
+    In this study we performed simulations of the human and vaquita porpoise genomes
+    using the \stdpopsim{} framework. As stated above, simulations included a model 
+    of selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
+    Three replications of full genome simulations were performed for each species, with 
+    a sample size of 100 diploid individuals. Demographic models are as stated above, each
+    using the exon annotations from the human and vaquita porpoise genomes respectively.
+    For simulations with selection, a SLiM scaling factor of 1 (corresponding to no scaling)
+    was used, with a burn-in of 10 ticks.
+
+    \subsection*{Demographic inference}
+    With simulations complete, demographic inference was performed using a collection of methods:
+    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+    Each of these methods was run using the same set of simulations in replicate.
+    In each case, the non-recombining portion
+    of the genome was masked to remove any bias from inference due to recombination rate
+    variation. Comparisons between inferences made with and without selection were 
+    of particular interest. For \msmc, we used a random sample of $N=6$ genomes with 20
+    iterations of the EM algorithm. For \gone, we set max\_snps = 500000,
+    the number of generations to 2000, and the number of bins = 400, all representing default settings.
+
+    \subsection*{Distribution of fitness effects inference}
+    We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
+    \dadi \citep{gutenkunst2009inferring,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
+    Each software package was run using the same set of simulations in replicate.   
+    For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
+    A key distinction between our analyses and those done by the empirical biologist
+    is that we have complete knowledge of which sites are under selection in the simulation,
+    thus our inferences is a best case scenario. In practice, empirical inference of the DFE is a much
+    harder problem as not all mutations assumed to be free of selection
+    may in reality be under selection. As above, all methods we run using default settings.
+
+    \subsection*{Selective sweeps detection}
+    We ran simulations of 5Mbp regions of chromosome 1 of the human genome with an without
+    purifying selection acting on exonic regions, and under a the three population
+    out-of-Africa model of human demography \citep{gutenkunst2009inferring}, 
+    along with the HapMap II genetic map \citep{international2007second}.
+    Single beneficial mutations were introduced at each location with a selection coefficient of $s = 0.03$,
+    and only sampled if they reached a terminal frequency of 0.95 or greater. 
+    We then benchmarked the performance of three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
+    2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity.
+    \diploshic was trained on data simulated using \texttt{discoal} \citep{kern2016discoal}.
+    For training simulations, we simulated 20 haploid genomic regions of 1.21Mbp
+    with a constant population size of 10,000 individuals, a mutation rate of 1e-8, and a recombination rate of 1e-8.
+    For \texttt{discoal} simulations that included selective sweeps,
+    selection coefficients were drawn from a uniform distribution between 0.001 and 0.05.
+    The time of the sweep was chosen from a uniform distribution between 0 and 0.02 Ne generations in the past.
+    For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between 1e-4 and 0.1.
+    All \texttt{discoal} training simulations were done under a model of equilibrium population size, 
+    so we expected it to perform suboptimally given the variation in population size in the out-of-Africa model.
 
@@
 
-\bibliography{references.bib}
+\printbibliography
+
+%%%%%%%% supplementary material
+\clearpage
+\beginsupplement
+
+\section*{Supplementary Material}
+
+% constant size DFE figure
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figures/HomSap/Constant/HomSap_Constant_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
+    \caption{
+    Performance of methods to infer the distribution of fitness effects (DFE) from genetic data.
+    Simulations were performed using a human constant size model with a Gamma distributed DFE
+    acting on exons parameterized by a mean selection coefficient and a shape parameter. On each row we show the true parameter 
+    as a red line, and the estimates as boxplots from replicate simulations. The human model has three extant populations (CEU, CHB, and YRI).  
+    Each row of the figure shows estimates from a different method: 1) dadi \citep{gutenkunst2009inferring}, 2) polyDFE \citep{tataru2020polydfe},  
+    3) DFE-alpha \citep{eyre2009estimating}, and 4) GRAPES \citep{galtier2016adaptive}.  
+    }
+    \label{fig:homsap-dfe.constant}    
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/PhoSin/Constant/PhoSin_Constant_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
+    \caption{
+    \label{fig:vaquita-dfe.constant}
+    A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
+    Simulations were performed using a model of vaquita porpoise genome under constant population size with a Gamma distributed DFE
+    acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
+    parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
+    This vaquita model has a single extant population, labelled pop 0, and parameter estimates from each
+    of the three different methods are shown: 1) GRAPES \cite{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe},
+    and 3) dadi \citep{gutenkunst2009inferring}.}
+\end{figure}
+
+% constant all site size DFE figure
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figures/HomSap/Constant/HomSap_Constant_Gamma_K17_all_sites_DFE_plot}
+    \caption{
+    \label{fig:homsap-dfe.constant.all_sites}
+    Performance of methods to infer the distribution of fitness effects (DFE) from genetic data
+    where all sites along the simulated chromosome can experience selected and neutral mutation
+    as opposed to exonic regions only.
+    }
+\end{figure}
+
+\begin{table}[ht]
+\centering
+\small
+\caption{\bf{Performance of DFE methods across the simulation scenarios}. 
+Bold rows show the lowest mean absolute error (MAE) for the two DFE parameters
+for each species.}
+\begin{tabular}{lllrrrrrr}
+\toprule
+species ID & demography & annotation & \makecell{MAE \\ $E|s|$ \\ dadi} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
+\midrule
+HomSap & Constant & all sites & 0.0027 & 0.0049 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
+HomSap & Constant & exons & 0.0012 & \bf{0.00081} & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
+HomSap & OOA 5R19 & all sites & 0.012 & 0.0079 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
+HomSap & OOA 5R19 & exons & 0.014 & \bf{0.0049} & 0.0072 & 0.031 & 0.051 & \bf{0.024} \\
+PhoSin & Constant & all sites & 0.024 & \bf{0.024} & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
+PhoSin & Constant & exons & 0.024 & \bf{0.024} & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
+PhoSin & Vaquita2Epoch 1R22 & all sites & 0.024 & \bf{0.023} & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
+PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & \bf{0.023} & 0.024 & \bf{0.18} & 0.23 & 0.21 \\
+\bottomrule
+\end{tabular}
+\label{tab:dfe_table}
+\end{table}
+
+%sweep power vs exon density
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8 \textwidth]{figures/sweeps/relationship_power_exon.pdf}
+    \caption{
+    Power to detect selective sweeps at 100 locations along human chromosome 1 under a three population out-of-Africa model \citep{gutenkunst2009inferring}  
+    with and without purifying selection as a function of local exon density.
+    Simulations are as described in Figure \ref{fig:chr1_power}.
+    Each panel shows the results of simulations with purifying selection acting on exons (red; labelled BGS)
+    and without purifying selection acting on exons (blue; labelled Neutral).
+    True positive rate is shown for these methods for samples from the CEU and YRI samples (left and right respectively).
+    Points represent the power at each location and trend lines along with 95\% confidence intervals are shown fit to those points.
+    }
+    \label{fig:power-exon}
+\end{figure}
+
+\stopsupplement
 \end{document}

### 2024-11-15 analysis2_manuscript.tex
revised sweep section

@@
 
-    ** this section needs revision based on new figures that Murillo will add **
     In Figure \ref{fig:chr1_power} we show the power to detect selective sweeps at 100 locations along human chromosome 1 from
@@
 
-    %sweep power vs recombination rate
+    Using this set of simulations we can also compare the power of different methods to detect sweeps.
+    In general, the power of the methods varies with respect to the population history of the sample.
+    For instance \sweepfinder2 has marginally higher power to detect sweeps in the CEU sample than \diploshic,
+    likely as a results of the fact that we here trained \diploshic on simulations under a constant size population,
+    while the population size history of CEU is characterized by a strong recent bottleneck.
+    Conversely, \diploshic has higher power to detect sweeps in the YRI sample than \sweepfinder2,
+    likely because that population size history more closely matches the population size history used to train \diploshic.
+    We can further explore the performance of the methods by plotting the receiver operating characteristic (ROC) curves for the methods.
+    In Figure \ref{fig:roc-curves} we show the ROC curves for the three methods for detecting sweeps.
+    False positive rates were established using neutral simulations that contained no sweeps
+    but were conditioned on the same population history as the CEU and YRI samples.
+    These suggest that the $\pi$ cutoff is a suprisingly effective method for detecting sweeps,
+    in these genomic regions at least, but that it lags behind \diploshic and \sweepfinder2 in terms of specificity
+    generally. 
+    %While this is not exhaustive benchmarking, we aim to give a sense to the reader
+    %of how \stdpopsim can be used to benchmark the performance of different methods easily and efficiently.
+
     \begin{figure}
@@
     To further explore the effects of recombination rate and exon density on the power to detect sweeps, we show in Figure \ref{fig:power-recomb}
-    the relationship between power and local recombination rate for our sweep simulations. We see that power to detect sweeps is generally
-    a decreasing function of local recombination rate, which is consistent with sweeps having a larger genomic footprint in regions of low
-    recombination. We also see that the presence of purifying selection acting on exons has a relatively small effect on the power to detect sweeps,    
+    the relationship between power and local recombination rate for our sweep simulations. We see that power to detect sweeps is
+    a decreasing function of local recombination rate for \sweepfinder2 and $pi$ outliers,
+    which is consistent with sweeps having a larger genomic footprint in regions of low
+    recombination.
+    \diploshic on the other hand shows a slight increase in power with recombination rates.
+    We also see that the presence of purifying selection acting on exons has a relatively small effect on the power to detect sweeps,    
     but one that is present across the entire range of recombinations rates observed in human chromosome 1. 
-    We also note that the different methods for detecting sweeps that we have explored here respond differently to the effects of recombination rate.
     Unsurprisingly, the empirical cutoff of $\pi$ to detect sweeps is the less robust to the effects of recombination rate than the other two methods,
-    that are specifically designed to find selective sweeps. Whereas \sweepfinder2 has a clear decline in power with increasing recombination rate,
-    \diploshic 's behavior is more variable, with a slight increase in power at intermediate recombination rates in the CEU sample and a decline
-    in power at high recombination rates in the YRI sample.
+    that are specifically designed to find selective sweeps.
 
@@
     for instance sweepfinder2 seems to have minimum power in regions of intermediate exon density, while \diploshic is roughly constant
-    or perhaps has a slight increase in power at intermediate exon densities. Noise is the estimates of power with respect to exon density
+    or perhaps has a slight increase in power at intermediate exon densities. Noise in the estimates of power with respect to exon density
     are prohibitively high for the empirical cutoff of $\pi$ to detect sweeps, so it's not clear what conclusions can be drawn.
@@
 
+% roc curves for sweeps
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8 \textwidth]{figures/sweeps/roc_neutral_null.pdf}
+    \caption{
+    \label{fig:roc-curves}
+    Receiver operating characteristic (ROC) curves for the three methods for detecting sweeps.
+    False positive rates were established using neutral simulations that contained no sweeps
+    but were conditioned on the same population history as the CEU and YRI samples.
+    }
+\end{figure}
+
 %sweep power vs exon density

### 2024-11-18 analysis2_manuscript.tex
added abstract draft

@@
  \author[5*]{Murillo F. Rodrigues}
+ \author[5*]{Silas Tittes}
 \author[6]{Georgia Tsambos}
@@
 % \author[5]{Anastasia Teterina}
-\author[5]{Silas Tittes}
+
 \author[5]{Scott T. Small}
@@
 \section*{Abstract}
+Natural selection is a fundamental evolutionary force that shapes patterns of genetic variation across species. 
+However, simulating realistic models that incorporate both selection and complex demographic histories
+has remained challenging, limiting our ability to benchmark statistical methods and explore theoretical predictions.
+Here we present a major extension to the \stdpopsim{} framework that enables simulation of various modes
+of natural selection, including background selection, selective sweeps, and arbitrary distributions of fitness effects (DFE).
+This extension maintains \stdpopsim's core principles of reproducibility and standardization while adding support
+for species-specific genomic annotations and published DFE estimates. 
+We demonstrate the utility of this framework by benchmarking methods for demographic inference,
+DFE estimation, and selective sweep detection across different species and scenarios. 
+Our results highlight how selection can bias demographic inference, 
+reveal the sensitivity of DFE inference methods to model assumptions, 
+and show how genomic features like recombination rate and functional density influence power to detect selective sweeps.
+This extension to \stdpopsim{} provides a powerful new resource for the population genetics community
+to explore the interplay between selection and other evolutionary forces in a reproducible framework.
 

### 2024-11-18 analysis2_manuscript.tex
edits to other sections

@@
 \section*{Abstract}
+    \label{abstract}
     Natural selection is a fundamental evolutionary force that shapes patterns of genetic variation across species. 
@@
     in genetic diversity across a recombining chromosome 
-    that can mimic the effects of selection \citep{simonsen1995properties, barton1998effect},
-    % DRS -- I'm not sure Barton reference is right for this?
+    that can mimic the effects of selection \citep{simonsen1995properties},
     and lead to spurious inferences about the strength and targets of genetic adaptation
@@
     % simulation
-    % DRS: lots more refs will be needed in this paragraph. added a few and am happy to help add more later if needed
     To meet the growing need for the interpretation,
@@
     % stdpopsim
-    % DRS: lots more refs will be needed in this paragraph. added a few and am happy to help add more later if needed
     A lingering challenge with simulation in population genetics has been
@@
     At this release the \stdpopsim catalog comprises 24 species (i.e. genome representations)
-    for which we have implemented XX demographic models, have XX genetic maps, and XX DFEs.
+    for which we have implemented 28 demographic models, have 37 genetic maps, and 7 DFEs. %someone should double check the numbers
     Additions to the catalog are ongoing -- we point the reader to our previously published 
@@
     using a variety of methods, including the pairwise sequentially Markovian coalescent
-    (PSMC) \citep{li2011inference}, through use of the site frequency spectrum (SFS),
-    as well as through identity by decent information. 
+    (PSMC) \citep{li2011inference}, through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
+    as well as through identity by decent information \citep{santiago2020recent}. 
     While that is so, it has been well described that selection, because of its effects on
@@
     further from true census population sizes simply because selection increases the 
-    rate of coalescence in the genome.
+    rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}.
 
@@
     In this and what follows we have performed 3 replicate whole genome simulations, so uncertainty in estimates might
-    be roughly assessed, however we are mindful of wasted compute as thorough benchmarking
+    be roughly assessed, however we are mindful of wasted computational resources as thorough benchmarking
     isn't the main purpose of this manuscript.
@@
     The top row shows the results of simulations without selection, while the bottom row shows the results
-    of simulations with a Gamma distributed DFE for the deleterious effects of mutations occurring exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
+    of simulations with a Gamma distributed DFE for the deleterious effects of mutations occurring exons.
+    In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
     in various colors for the four methods. Dotted lines on the bottom row are to visually reinforce that inference is done
@@
     \msmc seems to be the most affected by the presence of selection, 
-    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$.
+    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$,
+    although estimates from \msmc have quite a bit of added noise even in the absence of selection.
+
     % DRS: I don't agree with the interpretation in the previous two sentences (starting with "Likewise").
@@
     % the minor drop in accuracy is meaningful in relative terms.
+    % ADK: I edited the text-- what do you think Dan?
 
@@
     and \grapes \citep{galtier2016adaptive}.
+    In each case we note that the DFE is estimated from segregating sites only,
+    without the use of substitution data, as a current limitation of \stdpopsim
+    currently is that we are not able to simulate an outgroup sequence to estimate
+    divergence. 
     Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
@@
     vaquita porpoise genome. In the constant size model (fig \ref{fig:vaquita-dfe.constant}) we see that all methods
-    perform uniformly worse that in the human genome simulations, which consistent underestimation of the mean selection coefficient    
+    perform uniformly worse that in the human genome simulations, with consistent underestimation of the mean selection coefficient    
     and overestimation of the shape parameter. 
-A critical distinction is that the previous human simulations employed a DFE that assumed all selected mutations had additive effects (no dominance, $h = 1/2$), whereas the vaquita simulations use a DFE in which more deleterious mutations are more recessive ($h < 1/2$).
-For all three software packages, the inferred mean and shape of the distribution of selection coefficients is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient and the selection coefficient (Fig.~\ref{fig:vaquita-dfe}C).
-This is likely because deleterious alleles are typically at low frequency and thus heterozygous, where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
-Strongly deleterious alleles are often observed to be recessive in many species \citep{}, so these results suggest caution in interpreting DFE inferences.
+    A critical distinction is that the previous human simulations employed a DFE that assumed all selected mutations had additive effects
+    (no dominance, $h = 1/2$), whereas the vaquita simulations use a DFE in which more deleterious mutations are more recessive ($h < 1/2$).
+    For all three software packages, the inferred mean and shape of the distribution of selection coefficients 
+    is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient
+    and the selection coefficient (Fig.~\ref{fig:vaquita-dfe}C).
+    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
+    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
+    Strongly deleterious alleles are often observed to be recessive in many species \citep{mukai1972mutation},
+    so these results suggest caution in interpreting DFE inferences.
     %RNG: Could use Kirk's help with the best references here
@@
     \label{funding}
-
+    Multiple funding sources supported this work.
+    MFR, ST, NP, and ADK were supported in part by NIH award R35GM148253.
+    %other funding sources?
 \printbibliography

### 2024-11-25 analysis2_manuscript.tex

@@
 
+    % DRS: This figure could really use some Python API and comand line examples like in Figs 1B,C
+    % form the original stdpopsim paper
+    \begin{wrapfigure}[37]{l}{0.4\textwidth}
+        \vspace{-0.0cm}
+        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
+        \caption{\label{fig:schematic}
+        A schematic of the \stdpopsim catalog and the different components that
+        can be used in simulations. The user specifies a species, a portion of
+        the genome to simulate, and optionally a genetic map, a model of
+        demography, and a model of selection that itself is composed of a
+        distribution of fitness effects (DFE) and a set of functional
+        annotations.}
+    \end{wrapfigure}
+    \hfill
+
     % current numbers of DFEs / species / etc
@@
         \caption{
-        Power to detect selective sweeps at 100 locations along human chromosome 1 under a three population out-of-Africa model \citep{gutenkunst2009inferring}  
-        with and without purifying selection acting on exons under the Gamma-distributed DFE described above. 
-        The HapMap II genetic map \citep{international2007second} was used in each simulation. 
-        Single beneficial mutations were introduced at each location with a selection coefficient of $s = 0.03$,
-        and only sampled if they reached a terminal frequency of 0.95 or greater. 
-        Three methods for detecting sweeps were used: 1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row), 
-        2) \diploshic \citep{kern2018diplos} (middle row), and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity (bottom row).
-        True positive rate is shown for these methods for samples from the CEU and YRI samples (left and right respectively). 
-        Under each method we show as a heat map exon density along the chromosome, along with the local recombination rate in cM/Mb.
-        Power is shown as the proportion of replicates in which a sweep was detected at each location. 
-        Each panel shows the results of simulations with purifying selection acting on exons (red; labelled BGS)
-        and without purifying selection acting on exons (blue; labelled Neutral).
+        Power to detect selective sweeps at 100 locations along human
+        chromosome 1 under a three population out-of-Africa model
+        \citep{gutenkunst2009inferring} when using null models with and without
+        purifying selection acting on exons under the Gamma-distributed DFE
+        described above.  The HapMap II genetic map
+        \citep{international2007second} was used in each simulation.  Single
+        beneficial mutations were introduced at each location with a selection
+        coefficient of $s = 0.03$, and only sampled if they reached a terminal
+        frequency of 0.95 or greater.  Three methods for detecting sweeps were
+        used: 1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row), 2)
+        \diploshic \citep{kern2018diplos} (middle row), and 3) an empirical
+        cutoff of $\pi$ to find regions of reduced diversity (bottom row).
+        True positive rate is shown for these methods for samples from the CEU
+        and YRI samples (left and right respectively).  Under each method we
+        show as a heat map exon density along the chromosome, along with the
+        local recombination rate in cM/Mb.  Power is shown as the proportion of
+        replicates in which a sweep was detected at each location.  To define
+        thresholds for calling a sweep, two different null models were used: a
+        neutral model (blue lines), and a model with purifying selection acting
+        on exons (red lines).
         }
@@
         \caption{
-        Power to detect selective sweeps at 100 locations along human chromosome 1 under a three population out-of-Africa model \citep{gutenkunst2009inferring}  
-        with and without purifying selection as a function of local recombination rate.
-        Simulations are as described in Figure \ref{fig:chr1_power}.
-        Each panel shows the results of simulations with purifying selection acting on exons (red; labelled BGS)
-        and without purifying selection acting on exons (blue; labelled Neutral).
-        True positive rate is shown for these methods for samples from the CEU and YRI samples (left and right respectively).
-        Points represent the power at each location and trend lines along with 95\% confidence intervals are shown fit to those points.
+        Power to detect selective sweeps at 100 locations along human
+        chromosome 1 under a three population out-of-Africa model
+        \citep{gutenkunst2009inferring}  when using null models with and
+        without purifying selection as a function of local recombination rate.
+        Simulations are as described in Figure \ref{fig:chr1_power}.  True
+        positive rate is shown for these methods for samples from the CEU and
+        YRI samples (left and right respectively).  To define thresholds for
+        calling a sweep, two different null models were used: a neutral model
+        (blue lines), and a model with purifying selection acting on exons (red
+        lines).  Points represent the power at each location and trend lines
+        along with 95\% confidence intervals are shown fit to those points.
         }

### 2024-11-27 analysis2_manuscript.tex
second round of edits from DRS (#33)

--- 2024-11-18 analysis2_manuscript.tex
@@
     However, simulating realistic models that incorporate both selection and complex demographic histories
-    has remained challenging, limiting our ability to benchmark statistical methods and explore theoretical predictions.
+    is challenging, limiting our ability to benchmark statistical methods and explore theoretical predictions.
+    \stdpopsim is a tool for facilitating the simulation of a contiually expanding catalog of population genetic models.
     Here we present a major extension to the \stdpopsim{} framework that enables simulation of various modes
@@
     reveal the sensitivity of DFE inference methods to model assumptions, 
-    and show how genomic features like recombination rate and functional density influence power to detect selective sweeps.
+    and show how genomic features like recombination rate and the density of functional sequence influence power to detect selective sweeps.
     This extension to \stdpopsim{} provides a powerful new resource for the population genetics community
@@
     from other evolutionary processes such as mutation, recombination,
-    and genetic drift \cite[e.g.,][]{gillespie1991causes}.
+    and genetic drift \citep[e.g.,][]{gillespie1991causes}.
     For instance changes in population size can lead to fluctuations
@@
     and lead to spurious inferences about the strength and targets of genetic adaptation
-    \cite{simonsen1995properties,akey2004population,nielsen2005genomic}.
+    \citep{simonsen1995properties,akey2004population,nielsen2005genomic}.
     In turn selection can confound our ability to infer demographic 
@@
     variation \citep[e.g.,][]{galloway2020few}, to provide a null model for hypothesis testing
-    \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting}, to
-    to explore the power and limitations of statistical methods \cite[e.g.,][]{przeworski2002signature},
+    \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
+    to explore the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature},
     and increasingly to provide a basis for machine learning and other
-    simulation-based inference methods \citep[e.g.,][]{pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}.
+    simulation-based inference methods \citep[e.g.,][]{beaumont2002approximate,pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}.
     While this is so, joint simulation of complex demography and selection
     is challenging, and requires a deep understanding of the underlying
-    evolutionary processes, as well as a number of parameter choices including
-    the strength of selection, the distribution of fitness effects, and the
-    recombination rate, as well as a parameterized model of demography.
+    evolutionary processes, and necessitates a number of parameter choices including
+    the strength of selection and the
+    recombination rate, as well as a parameterized demographic model.
     This is a daunting task for many researchers, and can be a barrier to
     the adoption of simulation-based methods in population genetics.
-    Furthermore, sharing and comparing simulation results among studies can 
-    itself be challenging, as different simulation tools may use different structures
-    and parameter scalins for their models, and may not be directly comparable.
-    This is a major limitation for the field, as it can make it difficult
-    to assess the robustness of results, and to compare the results of
-    different studies.
+    Furthermore, different simulation and inference tools may use different structures
+    and parameter scalings for their models.
 
     % stdpopsim
-    A lingering challenge with simulation in population genetics has been
+    In light of the complexities associated with estimating and simulating
+    population genetic models, it is no surprise that
+    a lingering challenge with simulation in population genetics has been
     reproducibility and the ability to share and compare results among 
@@
     of the fitness effects of mutations along a chromosome.
-    A \texttt{DFE} object can apply to a specific interval set of a chromosome,
-    and can be combined with other \texttt{DFE} objects to provide, 
+    A \texttt{DFE} object can apply to specified intervals within a chromosome,
+    and can be combined with other \texttt{DFE} objects to provide 
     a rich model of how selection may differ along a chromosome. 
-    Once specified a model of selection is then run in the same way
+    Once specified, a model of selection can then be run using \stdpopsim's standard interfaces
     but currently is limited to using SLiM simulation engine, which
@@
     As with other \stdpopsim models, when incorporating selection we aim to hew as closely as possible
-    to realistic genomes, and provide public annotations of specices-specific functional
+    to realistic genomes, and provide public annotations of specices' functional genomic
     elements that can be used along with a \texttt{DFE} object to
@@
     populations. A schematic of the \stdpopsim catalog is shown in Figure \ref{fig:schematic}.
-    From a users perspective, one can specify a species, a portion of the genome to simulate,
+    From a user's perspective, one can specify a species, a portion of the genome to simulate,
     a genetic map if available, a model of demography, and a model of selection that is
-    composed of a set of annotations and a \texttt{DFE} object.
+    composed of a set of annotations and a \texttt{DFE} object, all in an automated and user-friendly fashion.
 
@@
     four species: \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
-    humans, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
-    While that is so, these DFEs can be mixed and matched with the other species
-    in the catalog. For instance, one could simulate a butterfly species with a Gamma DFE originally
+    humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
+    While that is so, these DFEs can be applied to other species
+    in the catalog. For instance, one could simulate a butterfly species with a gamma DFE originally
     estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
@@
     a published DFE included in the catalog. This flexibility allows for a wide range of
-    models of selection to be simulated, and for the results of these simulations to be
-    compared to empirical data or used as a null model for hypothesis testing.
+    models of selection to be simulated. %, and for the results of these simulations to be
+    %compared to empirical data or used as a null model for hypothesis testing.
+    % DRS: commenting that stuff out because there are obviously other use cases that could be
+    % listed here (e.g. ML/ABC, and if we don't know enough about a species to do that then
+    % we probably have no business doing hypothesis testing either, which could be just as
+    % misleading or maybe even more so), so the simplest solution is to cut this as it is
+    % not needed here.
 
@@
     % current numbers of DFEs / species / etc
+    % DRS: are we giving this a version number?
     At this release the \stdpopsim catalog comprises 24 species (i.e. genome representations)
     for which we have implemented 28 demographic models, have 37 genetic maps, and 7 DFEs. %someone should double check the numbers
-    Additions to the catalog are ongoing -- we point the reader to our previously published 
-    report detailing this effort \citep{lauterbur2023expanding} and we welcome contributions from the
-    community---we encourage those interested in contributing to visit https://popsim-consortium.github.io/stdpopsim-docs
+    Additions to the catalog are ongoing---we point the reader to our previously published 
+    report detailing this effort \citep{lauterbur2023expanding}. We welcome contributions from the
+    community and encourage those interested in contributing to visit https://popsim-consortium.github.io/stdpopsim-docs
     or contact one of the PopSim Consortium members for guidance.
@@
     In this section we provide a series of examples of how the new models of selection
-    in \stdpopsim can be used to benchmark and compare the performance of different
+    in \stdpopsim can be used to benchmark different
     methods for population genetic inference.
     We focus on three main areas: demographic inference in the 
-    single and multi-population settings, inference of the distribution of fitness effects,
+    single and multi-population settings, inference of the DFE,
     and the detection of selective sweeps, particularly in the context of background selection.
@@
     While that is so, it has been well described that selection, because of its effects on
-    local and more global patterns of genetic variation, can drive estimates of $N_e(t)$
-    further from true census population sizes simply because selection increases the 
+    patterns of genetic variation, can bias estimates of $N_e(t)$
+    away from true census population sizes simply because selection increases the 
     rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}.
 
-    Using \stdpopsim we can simulate genetic data under a model of selection and then 
+    Using \stdpopsim we can easily simulate genetic data under a model of selection and then 
     examine the performance of different methods for inferring $N_e(t)$ from these simulated 
@@
     where we have simulated genomes under a model of human out of Africa (OOA) demography
-    with and without background selection on exons. For this simulation we have used
+    with and without purifying selection on exons. For this simulation we have used
     a genetic map from the HapMap Project \citep{international2007second} along with a
@@
     that acts on exonic positions defined from the HAVANA group release 104 for the human genome.
+    % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
+    % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
     These simulations were then used to estimate $N_e(t)$ using four methods: 1) \msmc \citep{Schiffels2020}, 
@@
     In Figure \ref{fig:1pop-human-demography} we show six panels, each of which shows the estimated $N_e(t)$
-    from each independent sample in a five-population model of human out-of-Africa demography
+    from data simulated under the five-population model of human out-of-Africa demography
     from Ragsdale \textit{et al.} (2019)
     with samples taken from three modern geographic groups (CEU, CHB, and YRI from left to right).
-    The top row shows the results of simulations without selection, while the bottom row shows the results
-    of simulations with a Gamma distributed DFE for the deleterious effects of mutations occurring exons.
+    The top row shows the results of $N_e(t)$ estimates made from simulations without selection, while the bottom row shows the results
+    from simulations with deleterious mutatinos occurring in exons.
     In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
@@
     % ADK: I edited the text-- what do you think Dan?
+    % DRS: I am still finding this description pretty inaccurate. Happy to discuss over zoom if you think
+    % that might be faster.
 
@@
         Performance of methods to infer $N_e(t)$ from a human out-of-Africa model \citep{ragsdale2019models}
-        with and without background selection on exons. The top row shows estimates of $N_e(t)$ from simulations
-        without selection, while the bottom row shows estimates of $N_e(t)$ from simulations with a Gamma distributed   
+        with and without purifying selection on exons. The top row shows estimates of $N_e(t)$ from simulations
+        without selection, while the bottom row shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
         DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
@@
     \end{figure}
+    %DRS-fig: reminder to update the legend to just say ``exons'' instead of ensembl_havana_whatever...
 
-    To highlight the ease of comparisons between different species and model enabled by \stdpopsim,
+    To highlight the ease of comparisons between different species and models enabled by \stdpopsim,
     we have also performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
@@
     For these simulations we have used the genome structure and exon annotations from the vaquita genome assembly
-    and a DFE from \textcite{robinson2022critically} that acts on exonic positions, however no genetic map 
+    and a DFE from \textcite{robinson2022critically} that acts on exons, however no genetic map 
     is currently available, so recombination rates are assumed to be constant across the genome.
@@
     The top panel shows estimates of $N_e(t)$ from simulations
-    without selection, while the bottom panel shows estimates of $N_e(t)$ from simulations with a Gamma distributed   
+    without selection, while the bottom panel shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
     DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
@@
     from genetic data. Here we show how the new models of selection in \stdpopsim can be used to benchmark and compare
-    the performance of different methods for inferring the DFE from genetic data. Here we have performed simulations using the 
+    the performance of different methods for inferring the DFE. We performed simulations using the 
     human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
-    To remind the reader, the model of selection in each case is a Gamma distributed DFE for deleterious mutations.
 
@@
         \caption{A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
-        Simulations were performed using a human out-of-Africa model \citep{ragsdale2019models} with a Gamma distributed DFE
+        Simulations were performed using a human out-of-Africa model \citep{ragsdale2019models} with a gamma-distributed DFE
         acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
@@
     In Figures \ref{fig:homsap-dfe.constant} and \ref{fig:homsap-dfe.ooa} we show a comparison of estimates
-    from three software packages for inferring the DFE from genetic data-- 
+    from three software packages for inferring the DFE from genetic data:
     \dadi \citep{gutenkunst2009inferring}, \polydfe \citep{tataru2020polydfe}, 
@@
     the shape parameter are most accurate in the YRI sample, while for \polydfe
-    estimates are consistently least accurate in the YRI sample. 
+    estimates are least accurate in the YRI sample. 
     With respect to the strength of selection, as in constant size populations,
@@
         A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
-        Simulations were performed using a model of vaquita porpoise demography \citep{robinson2022critically} with a Gamma distributed DFE
+        Simulations were performed using a model of vaquita porpoise demography \citep{robinson2022critically} with a gamma-distributed DFE
         acting on exons parameterized by a mean selection coefficient, a shape parameter, and a relationship between selection and domiance. Estimates of the 
         parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
-        This vaquita model has a single extant population, labelled pop 0, and parameter estimates from each
+        This vaquita model has a single extant population, and parameter estimates from each
         of the three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
@@
     In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
-    from the same three software packages for inferring the DFE from genetic data, but for simulations of the
+    from the same three DFE-inference packages on simulations of the
     vaquita porpoise genome. In the constant size model (fig \ref{fig:vaquita-dfe.constant}) we see that all methods
@@
     Here we have performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring}
-    with purifying selection acting on exons under the Gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second},    
-    but overlay on top of this a model of selective sweeps. In particular we have conditioned simulations on the introduction of a beneficial mutation
+    with purifying selection acting on exons under the gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second},    
+    but also include sweeps. In particular we introduced a beneficial mutation
     at a given location (to be varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
@@
         Power to detect selective sweeps at 100 locations along human chromosome 1 under a three population out-of-Africa model \citep{gutenkunst2009inferring}  
-        with and without purifying selection acting on exons under the Gamma-distributed DFE described above. 
+        with and without purifying selection acting on exons under the gamma-distributed DFE described above. 
         The HapMap II genetic map \citep{international2007second} was used in each simulation. 
@@
         2) \diploshic \citep{kern2018diplos} (middle row), and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity (bottom row).
-        True positive rate is shown for these methods for samples from the CEU and YRI samples (left and right respectively). 
-        Under each method we show as a heat map exon density along the chromosome, along with the local recombination rate in cM/Mb.
+        True positive rate is shown for these methods for the CEU and YRI samples (left and right respectively). 
+        Under each method we show heatmaps of exon density and local recombination rate along the chromosome.
         Power is shown as the proportion of replicates in which a sweep was detected at each location. 
@@
     A few things are immediately apparent from this figure. First, the power to detect sweeps is generally higher in the YRI sample
-    than in the CEU sample, which is consistent with the fact that the YRI sample has a larger effective population size and thus
+    than in the CEU sample, which is consistent with the fact that the YRI sample has a larger effective population size and thus both stronger selection and
     more variation to detect a sweep. Further the population size of YRI has been relatively constant over the last 100,000 years,
@@
     what we know about the effects of demography on the power to detect sweeps \citep[e.g.,][]{simonsen1995properties}.
-    Second, the power to detect sweeps is slightly lower in the presence of purifying selection acting on exons but the effect is not
+    Second, the power to detect sweeps is slightly lower in the presence of purifying selection acting on exons (consistent with \cite{schrider2020background}) but the effect is not
     as strong as one might expect. This is likely because the strength of selection acting on exons is relatively weak, and the
@@
     generally.
+    % DRS: the ROC curve doesn't tell us which methods have better/worse specificity, just how their sensitivities compare at a given specificity
+    % DRS: but we could say something about how pi might be "less robust to heterogeneity in mapping/genotyping quality across the genome than the other methods"
+    % DRS: or something like that. I know we didn't test that, but it must be the case.
     %While this is not exhaustive benchmarking, we aim to give a sense to the reader
@@
     Unsurprisingly, the empirical cutoff of $\pi$ to detect sweeps is the less robust to the effects of recombination rate than the other two methods,
-    that are specifically designed to find selective sweeps.
+    which are specifically designed to find selective sweeps.
 
@@
     are prohibitively high for the empirical cutoff of $\pi$ to detect sweeps, so it's not clear what conclusions can be drawn.
+    % DRS: the exon-density results look pretty freakin flat to me. I think we could just curtail the description in this paragraph and
+    % DRS: say that there are no obvious patterns that stand out from the noise?
 
@@
     We have shown how the new API can be used to benchmark the performance of different methods
-    although we acknowledge that the benchmarks shown here merely scratch the surface of what
+    although the benchmarks shown here merely scratch the surface of what
     can be done with the new API. We envision moving forward that many researchers will use
@@
     new methods for the analysis of genetic data in the presence of selection. 
-    Further the critical step of benchmarking methods for the analysis of genetic data
-    can now be done on common ground, using the same simulations and parameterizations of
-    models of selection in a way that was not possible before.
+    The additions to \stdpopsim described here further expand this tool's utility for both 
+    characterizing the patterns of diversity expected under a wide range of population genetic
+    models, and for benchmarking methods for the analysis of genetic data
+    on common ground
+    %, using the same simulations and parameterizations of models of selection
+    % DRS: this sentence got pretty bloated with my edits so I commented part of it out
+    in a way that was not possible before.
 
@@
     of selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
-    Three replications of full genome simulations were performed for each species, with 
+    Three replicates of whole-genome simulations were performed for each species, with 
     a sample size of 100 diploid individuals. Demographic models are as stated above, each
@@
     \subsection*{Demographic inference}
-    With simulations complete, demographic inference was performed using a collection of methods:
+    We assessed the accuracy of demographic inferences from
     \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
@@
     In each case, the non-recombining portion
-    of the genome was masked to remove any bias from inference due to recombination rate
+    of the genome was masked to remove any bias due to recombination rate % DRS: might need more detail here
     variation. Comparisons between inferences made with and without selection were 
     of particular interest. For \msmc, we used a random sample of $N=6$ genomes with 20
-    iterations of the EM algorithm. For \gone, we set max\_snps = 500000,
-    the number of generations to 2000, and the number of bins = 400, all representing default settings.
+    iterations of the EM algorithm. For \gone, we set max\_snps to 500000,
+    the number of generations to 2000, and the number of bins to 400, all representing default settings.
 
@@
     \dadi \citep{gutenkunst2009inferring,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
-    Each software package was run using the same set of simulations in replicate.   
+    Each software package was run using the same set of simulations.
     For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
-    A key distinction between our analyses and those done by the empirical biologist
+    A key distinction between our analyses and those done in practice
     is that we have complete knowledge of which sites are under selection in the simulation,
-    thus our inferences is a best case scenario. In practice, empirical inference of the DFE is a much
-    harder problem as not all mutations assumed to be free of selection
-    may in reality be under selection. As above, all methods we run using default settings.
+    thus our inferences is a best case scenario. Empirical inference of the DFE is a much
+    harder problem as some mutations assumed to be free of selection
+    may in reality be under selection. As above, all methods were run using default settings.
 
     \subsection*{Selective sweeps detection}
-    We ran simulations of 5Mbp regions of chromosome 1 of the human genome with an without
-    purifying selection acting on exonic regions, and under a the three population
+    We ran simulations of 5Mbp regions of chromosome 1 of the human genome with and without
+    purifying selection acting on exonic regions, and under a three population
     out-of-Africa model of human demography \citep{gutenkunst2009inferring}, 
@@
     Performance of methods to infer the distribution of fitness effects (DFE) from genetic data.
-    Simulations were performed using a human constant size model with a Gamma distributed DFE
+    Simulations were performed using a human constant size model with a gamma-distributed DFE
     acting on exons parameterized by a mean selection coefficient and a shape parameter. On each row we show the true parameter 
-    as a red line, and the estimates as boxplots from replicate simulations. The human model has three extant populations (CEU, CHB, and YRI).  
+    as a black line, and the estimates as boxplots from replicate simulations. The human model has three extant populations (CEU, CHB, and YRI).
+    %DRS-fig: only one population is shown here, but 3 referenced in the legend. Wrong legend or wrong fig?
+    %DRS-fig: If it is the right fig, get rid of the pop0 legend inset
     Each row of the figure shows estimates from a different method: 1) dadi \citep{gutenkunst2009inferring}, 2) polyDFE \citep{tataru2020polydfe},  
@@
     A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
-    Simulations were performed using a model of vaquita porpoise genome under constant population size with a Gamma distributed DFE
+    Simulations were performed using a model of vaquita porpoise genome under constant population size with a gamma-distributed DFE
     acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
     parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
-    This vaquita model has a single extant population, labelled pop 0, and parameter estimates from each
-    of the three different methods are shown: 1) GRAPES \cite{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe},
+    This vaquita model has a single extant population, and parameter estimates from each
+    of the three different methods are shown: 1) GRAPES \citep{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe},
     and 3) dadi \citep{gutenkunst2009inferring}.}
+    %DRS-fig: remove the pop 0 legend again
 \end{figure}
+% DRS-fig: ditch the unnecessary legends from panels A and B. Also, are we deciding which version of panel C to include here? I guess I prefer
+% to bottom one but with the legend of the top one (``Original s DFE'' sounds a little clunky).
 
@@
     }
+    %DRS-fig: human? vaquita? need more info in this figure legend, and also get rid of the pop 0 thing again
 \end{figure}
@@
     }
+    %DRS-fig: can we include AUCs here?
 \end{figure}
@@
     \caption{
-    Power to detect selective sweeps at 100 locations along human chromosome 1 under a three population out-of-Africa model \citep{gutenkunst2009inferring}  
+    Power to detect selective sweeps at 100 locations along human chromosome 1 under a three-population out-of-Africa model \citep{gutenkunst2009inferring}  
     with and without purifying selection as a function of local exon density.
@@
     Each panel shows the results of simulations with purifying selection acting on exons (red; labelled BGS)
-    and without purifying selection acting on exons (blue; labelled Neutral).
-    True positive rate is shown for these methods for samples from the CEU and YRI samples (left and right respectively).
+    and without purifying selection (blue; labelled Neutral).
+    True positive rate is shown for these methods for the CEU and YRI samples (left and right respectively).
     Points represent the power at each location and trend lines along with 95\% confidence intervals are shown fit to those points.

### 2024-11-27 analysis2_manuscript.tex
changes

--- 2024-11-25 analysis2_manuscript.tex
@@
     and genetic drift \cite[e.g.,][]{gillespie1991causes}.
-    For instance changes in population size can lead to fluctuations
+    For instance, changes in population size can lead to fluctuations
     in genetic diversity across a recombining chromosome 
@@
     \cite{simonsen1995properties,akey2004population,nielsen2005genomic}.
-    In turn selection can confound our ability to infer demographic 
+    In turn, selection can confound our ability to infer demographic 
     history from allele frequencies \citep{ewing2016consequences,schrider2016effects} and
     estimates of inverse coalescent rate \citep{schrider2016effects, johri2021impact, cousins2024accurate}.
-    Thus it is imperative to joingly account for the effects of selection
+    Thus, it is imperative to jointly account for the effects of selection
     and demography when inferring evolutionary history from genetic data \citep{sheehan2016deep,johri2020toward}.
-    However this is a challenging task from a modeling perspective \citep{johri2022prospect}.
+    However, this is a challenging task from a modeling perspective \citep{johri2022prospect}. % MFR this is a bit unclear
 
@@
     itself be challenging, as different simulation tools may use different structures
-    and parameter scalins for their models, and may not be directly comparable.
+    and parameter scalings for their models, and may not be directly comparable.
     This is a major limitation for the field, as it can make it difficult
@@
     to model all of these features after estimates/annotations obtained
-    from a species of interest \citep[e.g.][]{schrider2020background}.
+    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.
 
     % goals
-    In this manuscript we provide an overview of a major new addition
+    In this manuscript, we provide an overview of a major new addition
     to \stdpopsim---the inclusion of models of selection.
@@
 
-    % DRS: This figure could really use some Python API and comand line examples like in Figs 1B,C
-    % form the original stdpopsim paper
-    \begin{wrapfigure}[]{l}{0.4\textwidth}
-        \vspace{-0.0cm}
-        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
-        \caption{\label{fig:schematic}
-        A schematic of the \stdpopsim catalog for simulating genetic variation
-        in a population. The user specifies a species, a portion of the genome to simulate,
-        and optionally a genetic map, a model of demography, and a model of selection that 
-        itself is composed of a distribution of fitness effects (DFE) and a set of functional
-        annotations.}
-    \end{wrapfigure}
-
     The vast majority of published DFE estimates are based on a limited number of species,
@@
     estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
-    Furthermore the user can specify a custom DFE, and provide their own annotations
+    Furthermore, the user can specify a custom DFE, and provide their own annotations
     of functional elements to simulate selection in a species for which we do not yet have 
@@
     \label{applications}
-    In this section we provide a series of examples of how the new models of selection
+    In this section, we present a series of examples of how the new models of selection
     in \stdpopsim can be used to benchmark and compare the performance of different
@@
     single and multi-population settings, inference of the distribution of fitness effects,
-    and the detection of selective sweeps, particularly in the context of background selection.
+    and the detection of selective sweeps, particularly in the context of realistic recombination maps and background selection.
 
@@
     (PSMC) \citep{li2011inference}, through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
-    as well as through identity by decent information \citep{santiago2020recent}. 
+    as well as through identity by descent information \citep{santiago2020recent}. 
     While that is so, it has been well described that selection, because of its effects on
@@
 
-    Using \stdpopsim we can simulate genetic data under a model of selection and then 
+    Using \stdpopsim, we can simulate genetic data under a model of selection and then 
     examine the performance of different methods for inferring $N_e(t)$ from these simulated 
-    data. In Figure \ref{fig:1pop-human-demography} we show the results of such a benchmark
+    data. In Figure \ref{fig:1pop-human-demography}, we show the results of such a benchmark
     where we have simulated genomes under a model of human out of Africa (OOA) demography
-    with and without background selection on exons. For this simulation we have used
+    with and without background selection on exons. For this simulation, we used
     a genetic map from the HapMap Project \citep{international2007second} along with a
     DFE for deleterious alleles (estimated by \cite{huber2017determining})
-    that acts on exonic positions defined from the HAVANA group release 104 for the human genome. 
+    that acts on exonic positions, as defined from the HAVANA group release 104 for the human genome. 
     These simulations were then used to estimate $N_e(t)$ using four methods: 1) \msmc \citep{Schiffels2020}, 
     2)\stairway method \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
-    In this and what follows we have performed 3 replicate whole genome simulations, so uncertainty in estimates might
+    In this and what follows we performed 3 replicate whole genome simulations, so uncertainty in estimates might
     be roughly assessed, however we are mindful of wasted computational resources as thorough benchmarking
-    isn't the main purpose of this manuscript.
+    is not the main purpose of this manuscript.
 
-    In Figure \ref{fig:1pop-human-demography} we show six panels, each of which shows the estimated $N_e(t)$
+    In Figure \ref{fig:1pop-human-demography}, we show six panels, each of which shows the estimated $N_e(t)$
     from each independent sample in a five-population model of human out-of-Africa demography
@@
     of simulations with a Gamma distributed DFE for the deleterious effects of mutations occurring exons.
-    In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
+    In each panel, we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
     in various colors for the four methods. Dotted lines on the bottom row are to visually reinforce that inference is done
@@
     To highlight the ease of comparisons between different species and model enabled by \stdpopsim,
-    we have also performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
-    In Figure \ref{fig:1pop-vaquita-demography} we show the results of $N_e(t)$ inference
+    we also performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
+    In Figure \ref{fig:1pop-vaquita-demography}, we show the results of $N_e(t)$ inference
     where we have simulated genomes under a model of vaquita demography from \textcite{robinson2022critically}.
-    For these simulations we have used the genome structure and exon annotations from the vaquita genome assembly
+    For these simulations, we used the genome structure and exon annotations from the vaquita genome assembly
     and a DFE from \textcite{robinson2022critically} that acts on exonic positions, however no genetic map 
@@
     The demographic model is a simple two-epoch model where population size has declined in the recent past. 
-    In this case we see that all methods perform well both in the absence and presence of selection,
+    In this case, we see that all methods perform well both in the absence and presence of selection,
     with a close match between the true $N_e(t)$ and the estimated $N_e(t)$ in all cases.
+    % MFR: Should we say somewhere that people use to expect from BGS?
+That is, that it can be seen as a reduction in local rates of coalescences and thus that we would expect these methods to be somewhat downwardly biased in the presence of BGS?
 
@@
     Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE)
-    from genetic data. Here we show how the new models of selection in \stdpopsim can be used to benchmark and compare
-    the performance of different methods for inferring the DFE from genetic data. Here we have performed simulations using the 
-    human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
+    from genetic data. Here, we show how the new models of selection in \stdpopsim can be used to benchmark and compare
+    the performance of different methods for inferring the DFE from genetic data. We performed simulations using the 
+    human and vaquita porpoise models described above with complex demographic histories and models of selection acting on exons.
     To remind the reader, the model of selection in each case is a Gamma distributed DFE for deleterious mutations.
@@
 
-    In Figures \ref{fig:homsap-dfe.constant} and \ref{fig:homsap-dfe.ooa} we show a comparison of estimates
+    % MFR: This is mentioning a constant scenario that is not currently shown here. Are we going to add it?
+    In Figures \ref{fig:homsap-dfe.constant} and \ref{fig:homsap-dfe.ooa}, we show a comparison of estimates
     from three software packages for inferring the DFE from genetic data-- 
@@
     without the use of substitution data, as a current limitation of \stdpopsim
-    currently is that we are not able to simulate an outgroup sequence to estimate
+    is that we are not able to simulate an outgroup sequence to estimate
     divergence. 
@@
     Notably only \polydfe overlapped the true mean selection coefficient, whereas 
-    the other methods slightly underestimated the strength of selection. 
+    the other methods do not. 
     For the shape parameter in constant size simulations \grapes and \polydfe are roughly
@@
     In the out-of-Africa model (fig \ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
-    In particular variation in accuracy of the parameters estimated in each
+    In particular, variation in accuracy of the parameters estimated in each
     sample population depends on the methods used, and each method seems to 
     have its own biases with respect to the way in which population history 
-    affects parameter estimates. For instance for \dadi we see that estimates of 
+    affects parameter estimates. For instance, for \dadi we see that estimates of 
     the shape parameter are most accurate in the YRI sample, while for \polydfe
     estimates are consistently least accurate in the YRI sample. 
-    With respect to the strength of selection, as in constant size populations,
-    all methods are performing reasonably well, although \dadi seems to
-    trail in terms of accuracy.
+    With respect to the strength of selection, \polydfe seems slightly better than the other two methods.
 
@@
     the performance of different methods for detecting selective sweeps from genetic data. 
-    Here we have performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring}
+    Here, we performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring}
     with purifying selection acting on exons under the Gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second},    
@@
     at a given location (to be varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
-    For each sweep location we performed 200 replicate simulations, sampling from them only if the selected allele reached a frequency of
+    For each sweep location, we performed 2000 replicate simulations, sampling from them only if the selected allele reached a frequency of
     0.95 or greater. Rather than simulate the entire chromosome arm, replicate simulations were performed on 5cM regions of the chromosome
@@
 
-    In Figure \ref{fig:chr1_power} we show the power to detect selective sweeps at 100 locations along human chromosome 1 from
+    In Figure \ref{fig:chr1_power}, we show the power to detect selective sweeps at 100 locations along human chromosome 1 from
     the simulations described above. We show the results of three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
@@
     than in the CEU sample, which is consistent with the fact that the YRI sample has a larger effective population size and thus
-    more variation to detect a sweep. Further the population size of YRI has been relatively constant over the last 100,000 years,
+    more genetic variation, making sweeps more apparent. Further, the population size of YRI has been relatively constant over the last 100,000 years,
     thus the out-of-Africa bottleneck has does not mute the signal of a sweep as much as in the CEU sample. This is consistent with

### 2024-11-27 analysis2_manuscript.tex

--- 2024-11-27 analysis2_manuscript.tex
@@
     and genetic drift \citep[e.g.,][]{gillespie1991causes}.
-    For instance changes in population size can lead to fluctuations
+    For instance, changes in population size can lead to fluctuations
     in genetic diversity across a recombining chromosome 
@@
     \citep{simonsen1995properties,akey2004population,nielsen2005genomic}.
-    In turn selection can confound our ability to infer demographic 
+    In turn, selection can confound our ability to infer demographic 
     history from allele frequencies \citep{ewing2016consequences,schrider2016effects} and
     estimates of inverse coalescent rate \citep{schrider2016effects, johri2021impact, cousins2024accurate}.
-    Thus it is imperative to joingly account for the effects of selection
+    Thus, it is imperative to jointly account for the effects of selection
     and demography when inferring evolutionary history from genetic data \citep{sheehan2016deep,johri2020toward}.
-    However this is a challenging task from a modeling perspective \citep{johri2022prospect}.
+    However, this is a challenging task from a modeling perspective \citep{johri2022prospect}. % MFR this is a bit unclear
 
@@
     \label{applications}
-    In this section we provide a series of examples of how the new models of selection
+    In this section, we present a series of examples of how the new models of selection
     in \stdpopsim can be used to benchmark different
@@
     single and multi-population settings, inference of the DFE,
-    and the detection of selective sweeps, particularly in the context of background selection.
+    and the detection of selective sweeps, particularly in the context of realistic recombination maps and background selection.
 
@@
 
-    Using \stdpopsim we can easily simulate genetic data under a model of selection and then 
+    Using \stdpopsim, we can easily simulate genetic data under a model of selection and then 
     examine the performance of different methods for inferring $N_e(t)$ from these simulated 
-    data. In Figure \ref{fig:1pop-human-demography} we show the results of such a benchmark
+    data. In Figure \ref{fig:1pop-human-demography}, we show the results of such a benchmark
     where we have simulated genomes under a model of human out of Africa (OOA) demography
@@
     2)\stairway method \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
-    In this and what follows we have performed 3 replicate whole genome simulations, so uncertainty in estimates might
+    In this and what follows we performed 3 replicate whole genome simulations, so uncertainty in estimates might
     be roughly assessed, however we are mindful of wasted computational resources as thorough benchmarking
-    isn't the main purpose of this manuscript.
+    is not the main purpose of this manuscript.
 
-    In Figure \ref{fig:1pop-human-demography} we show six panels, each of which shows the estimated $N_e(t)$
+    In Figure \ref{fig:1pop-human-demography}, we show six panels, each of which shows the estimated $N_e(t)$
     from data simulated under the five-population model of human out-of-Africa demography
@@
     To highlight the ease of comparisons between different species and models enabled by \stdpopsim,
-    we have also performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
-    In Figure \ref{fig:1pop-vaquita-demography} we show the results of $N_e(t)$ inference
-    where we have simulated genomes under a model of vaquita demography from \textcite{robinson2022critically}.
-    For these simulations we have used the genome structure and exon annotations from the vaquita genome assembly
+    we also performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
+    In Figure \ref{fig:1pop-vaquita-demography}, we show the results of $N_e(t)$ inference
+    where we simulated genomes under a model of vaquita demography from \textcite{robinson2022critically}.
+    For these simulations, we used the genome structure and exon annotations from the vaquita genome assembly
     and a DFE from \textcite{robinson2022critically} that acts on exons, however no genetic map 
@@
     Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE)
-    from genetic data. Here we show how the new models of selection in \stdpopsim can be used to benchmark and compare
+    from genetic data. Here, we show how the new models of selection in \stdpopsim can be used to benchmark and compare
     the performance of different methods for inferring the DFE. We performed simulations using the 
@@
     In the out-of-Africa model (fig \ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
-    In particular variation in accuracy of the parameters estimated in each
+    In particular, variation in accuracy of the parameters estimated in each
     sample population depends on the methods used, and each method seems to 
     have its own biases with respect to the way in which population history 
-    affects parameter estimates. For instance for \dadi we see that estimates of 
+    affects parameter estimates. For instance, for \dadi we see that estimates of 
     the shape parameter are most accurate in the YRI sample, while for \polydfe
@@
     the performance of different methods for detecting selective sweeps from genetic data. 
-    Here we have performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring}
+    Here, we performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring}
     with purifying selection acting on exons under the gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second},    
@@
         \caption{
-        Power to detect selective sweeps at 100 locations along human chromosome 1 under a three population out-of-Africa model \citep{gutenkunst2009inferring}  
-        with and without purifying selection acting on exons under the gamma-distributed DFE described above. 
-        The HapMap II genetic map \citep{international2007second} was used in each simulation. 
-        Single beneficial mutations were introduced at each location with a selection coefficient of $s = 0.03$,
-        and only sampled if they reached a terminal frequency of 0.95 or greater. 
-        Three methods for detecting sweeps were used: 1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row), 
-        2) \diploshic \citep{kern2018diplos} (middle row), and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity (bottom row).
-        True positive rate is shown for these methods for the CEU and YRI samples (left and right respectively). 
-        Under each method we show heatmaps of exon density and local recombination rate along the chromosome.
-        Power is shown as the proportion of replicates in which a sweep was detected at each location. 
-        Each panel shows the results of simulations with purifying selection acting on exons (red; labelled BGS)
-        and without purifying selection acting on exons (blue; labelled Neutral).
+        Power to detect selective sweeps at 100 locations along human
+        chromosome 1 under a three population out-of-Africa model
+        \citep{gutenkunst2009inferring} when using null models with and without
+        purifying selection acting on exons under the gamma-distributed DFE
+        described above.  The HapMap II genetic map
+        \citep{international2007second} was used in each simulation.  Single
+        beneficial mutations were introduced at each location with a selection
+        coefficient of $s = 0.03$, and only sampled if they reached a terminal
+        frequency of 0.95 or greater.  Three methods for detecting sweeps were
+        used: 1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row), 2)
+        \diploshic \citep{kern2018diplos} (middle row), and 3) an empirical
+        cutoff of $\pi$ to find regions of reduced diversity (bottom row).
+        True positive rate is shown for these methods for the CEU and YRI
+        samples (left and right respectively).  Under each method we show
+        heatmaps of exon density and local recombination rate along the
+        chromosome.  Power is shown as the proportion of replicates in which a
+        sweep was detected at each location.  To define
+        thresholds for calling a sweep, two different null models were used: a
+        neutral model (blue lines), and a model with purifying selection acting
+        on exons (red lines).
         }

### 2024-11-27 analysis2_manuscript.tex

@@
     \label{sweeps}
-    We next highlight how the new models of selection in \stdpopsim can be used to benchmark and compare
-    the performance of different methods for detecting selective sweeps from genetic data. 
-    Here, we performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring}
-    with purifying selection acting on exons under the gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second},    
-    but also include sweeps. In particular we introduced a beneficial mutation
-    at a given location (to be varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
-    For each sweep location, we performed 2000 replicate simulations, sampling from them only if the selected allele reached a frequency of
-    0.95 or greater. Rather than simulate the entire chromosome arm, replicate simulations were performed on 5cM regions of the chromosome
-    at each of 100 different locations. As recombination rate varies across the genome, the size of the region in base pairs simulated varies,
-    however the recombination distance, which is critical for the linked effects of a sweep, is held constant.
+    We next highlight how the new models of selection in \stdpopsim can be used to benchmark and compare the performance of different methods for detecting selective sweeps from genetic data.
+    Here, we performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring} with purifying selection acting on exons under the gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second}, but also include sweeps.
+    In particular, we introduced a beneficial mutation at a given location (to be varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
+    For each sweep location, we performed 2000 replicate simulations, sampling from them only if the selected allele reached a frequency of 0.95 or greater.
+    Rather than simulate the entire chromosome arm, replicate simulations were performed on 5cM regions of the chromosome at each of 100 different locations.
+    As recombination rate varies across the genome, the size of the region in base pairs simulated varies, however the recombination distance, which is critical for the linked effects of a sweep, is held constant.
 
@@
     In Figure \ref{fig:chr1_power}, we show the power to detect selective sweeps at 100 locations along human chromosome 1 from
-    the simulations described above. We show the results of three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
+    the simulations described above. We tested three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
     2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity.
@@
     than in the CEU sample, which is consistent with the fact that the YRI sample has a larger effective population size and thus both stronger selection and
-    more variation to detect a sweep. Further the population size of YRI has been relatively constant over the last 100,000 years,
+    more variation to detect a sweep. Further, the population size of YRI has been relatively constant over the last 100,000 years,
     thus the out-of-Africa bottleneck has does not mute the signal of a sweep as much as in the CEU sample. This is consistent with
     what we know about the effects of demography on the power to detect sweeps \citep[e.g.,][]{simonsen1995properties}.
-    Second, the power to detect sweeps is slightly lower in the presence of purifying selection acting on exons (consistent with \cite{schrider2020background}) but the effect is not
+    Second, the power to detect sweeps is slightly lower when we use a null model with purifying selection acting on exons (consistent with \cite{schrider2020background}) but the effect is not
     as strong as one might expect. This is likely because the strength of selection acting on exons is relatively weak, and the
@@
 
-    Using this set of simulations we can also compare the power of different methods to detect sweeps.
+    Using this set of simulations, we can also compare the power of different methods to detect sweeps.
     In general, the power of the methods varies with respect to the population history of the sample.
@@
     likely because that population size history more closely matches the population size history used to train \diploshic.
+    \sweepfinder2 power remains more constant along the chromosome, while the $\pi$ cutoff method varies wildly.
+
     We can further explore the performance of the methods by plotting the receiver operating characteristic (ROC) curves for the methods.
-    In Figure \ref{fig:roc-curves} we show the ROC curves for the three methods for detecting sweeps.
+    In Figure \ref{fig:roc-curves}, we show the ROC curves for the three methods for detecting sweeps.
     False positive rates were established using neutral simulations that contained no sweeps
@@
     %of how \stdpopsim can be used to benchmark the performance of different methods easily and efficiently.
+    These ROC curves, however, hide the previously seen phenomenon of high variation in power to detect sweeps along a realistic looking chromosome.
+    That is, although in aggregate the $\pi$ cutoff may seem effective, it's power is highly heterogeneous across the chromosome.
 
@@
 
-    To further explore the effects of recombination rate and exon density on the power to detect sweeps, we show in Figure \ref{fig:power-recomb}
-    the relationship between power and local recombination rate for our sweep simulations. We see that power to detect sweeps is
+    To further dissect the reasons for high variance (along the chromosome) in power to detect sweeps,
+    we show in Figure \ref{fig:power-recomb} the relationship between power and local recombination rate for our sweep simulations.
+    We see that power to detect sweeps is
     a decreasing function of local recombination rate for \sweepfinder2 and $pi$ outliers,
@@
     recombination.
-    \diploshic on the other hand shows a slight increase in power with recombination rates.
+    \diploshic on the other hand shows a slight increase in power with recombination rates,
+    probably because it is designed to distinguish regions where a sweep happend from regions linked to a sweep.
     We also see that the presence of purifying selection acting on exons has a relatively small effect on the power to detect sweeps,    
@@
 
-    In Figure \ref{fig:power-exon} we show the relationship between power and local exon density for our sweep simulations.
-    Here there is little relationship between power and exon density, which is somewhat surprising given that the power to detect sweeps
-    is generally lower in the presence of purifying selection acting on exons. However while this is the case we see that 
-    generally all methods have more power to detect sweeps in regions of lower exon density, which is consistent with the idea that
-    background selection acting on nearby exons can reduce the signature of a sweep. However this is far from a simple pattern:
-    for instance sweepfinder2 seems to have minimum power in regions of intermediate exon density, while \diploshic is roughly constant
-    or perhaps has a slight increase in power at intermediate exon densities. Noise in the estimates of power with respect to exon density
-    are prohibitively high for the empirical cutoff of $\pi$ to detect sweeps, so it's not clear what conclusions can be drawn.
+    In Figure \ref{fig:power-exon}, we show the relationship between power and local exon density for our sweep simulations.
+    There is little relationship between power (across all methods) and exon density,
+    which is somewhat surprising given that background selection should somewhat interfere with the sweep signature.
+    It is unclear why that is so, but the effects of realistic background selection may not be as similar to sweeps as previously thought \citep{schrider2020background}.
     % DRS: the exon-density results look pretty freakin flat to me. I think we could just curtail the description in this paragraph and
     % DRS: say that there are no obvious patterns that stand out from the noise?
+    % MFR: I changed this a bit: I agree with Dan that there is no clear pattern.
 

### 2024-12-09 analysis2_manuscript.tex
methods; + minor discussion  stuff

@@
     \label{Discussion}
-    In this paper we have presented an important new additional to the \stdpopsim{} library
+    In this paper, we have presented an important new addition to the \stdpopsim{} library
     that allows for simulating genetic variation in the presence of selection.
@@
 
-    We have shown how the new API can be used to benchmark the performance of different methods
-    although the benchmarks shown here merely scratch the surface of what
-    can be done with the new API. We envision moving forward that many researchers will use
+    Although we have shown how the new API can be used to benchmark the performance of different methods,
+    the benchmarks shown here merely scratch the surface of what can be done with the new API.
+    We envision that, moving forward, many researchers will use
     the new API to simulate data under a wide range of models of selection, and to develop
@@
     using the \stdpopsim{} framework. As stated above, simulations included a model 
-    of selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
+    of (negtive) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
     Three replicates of whole-genome simulations were performed for each species, with 
@@
     using the exon annotations from the human and vaquita porpoise genomes respectively.
-    For simulations with selection, a SLiM scaling factor of 1 (corresponding to no scaling)
-    was used, with a burn-in of 10 ticks.
+    For simulations with selection, we sampled 10 diploid individuals from each of two populations (YRI and CEU). 
+    To decrease runtimes, we used a SLiM scaling factor of 2 with a burn-in of 2N generations.
 
@@
     \subsection*{Selective sweeps detection}
-    We ran simulations of 5Mbp regions of chromosome 1 of the human genome with and without
-    purifying selection acting on exonic regions, and under a three population
-    out-of-Africa model of human demography \citep{gutenkunst2009inferring}, 
-    along with the HapMap II genetic map \citep{international2007second}.
-    Single beneficial mutations were introduced at each location with a selection coefficient of $s = 0.03$,
-    and only sampled if they reached a terminal frequency of 0.95 or greater. 
+    We ran simulations of human chromosome 1 using the three population out-of-Africa model
+    (identified as OutOfAfrica\_3G09 in the stdpopsim library; \cite{gutenkunst_inferring_2009})
+    and the genetic map estimated in \citet{frazer_second_2007} (HapMapII\_GRCh38).
+    For computational efficiency, we simulated a 5Mb region flanking 100 evenly distributed points along the chromosome 1.
+    We generated 1000 replicates of each of three classes of simulations:
+    (i) neutral (with the specified demography and genetic map),
+    (ii) background selection, in which we applied the DFE estimated in \citet{kim_inference_2017} (Gamma\_K17) to exon annotations (taken from Ensembl; ensembl\_havana\_104\_exons),
+    and (iii) sweep with background selection, in which we simulated the hard sweep as described with the addition of constraint in exonic regions (same as for \emph{ii}).
+    In the sweep simulations, we placed a single beneficial mutation in the center of the simulated region with a selection coefficient of $s = 0.03$, and only sampled if the beneficial mutation reached terminal frequency of 0.95 or greater.
+
     We then benchmarked the performance of three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
-    2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity.
+    2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff for nucleotide diversity ($\pi$) to find regions of reduced diversity.
     \diploshic was trained on data simulated using \texttt{discoal} \citep{kern2016discoal}.
+    Nucleotide diversity was computed directly from the underlying tree sequences using \tskit \citep{ralph_efficiently_2020} over 10 equally sized and non-overlapping subwindows.
+    The composite likelihood ratio (output by \sweepfinder) was computed at 21 uniformly distributed points along the simulated region.
+    To train the \diploshic classifier, 
     For training simulations, we simulated 20 haploid genomic regions of 1.21Mbp
-    with a constant population size of 10,000 individuals, a mutation rate of 1e-8, and a recombination rate of 1e-8.
+    with a constant population size of 10,000 individuals, a mutation rate of 2e-8, and a recombination rate of 2e-8.
     For \texttt{discoal} simulations that included selective sweeps,
-    selection coefficients were drawn from a uniform distribution between 0.001 and 0.05.
-    The time of the sweep was chosen from a uniform distribution between 0 and 0.02 Ne generations in the past.
+    selection coefficients were drawn from a uniform distribution between 0.001 and 0.10.
+    The time of the sweep was chosen from a uniform distribution between 0 and 0.01 Ne generations in the past.
     For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between 1e-4 and 0.1.
@@
     so we expected it to perform suboptimally given the variation in population size in the out-of-Africa model.
+
+    To call sweeps, we empirically determined the distribution of each statistic (nucleotide diversity, \sweepfinder's CLR, and \diploshic's probability of sweep) under a null model and defined the top $5\%$ quantile as a cutoff to call a sweep (though regions with extremely low recombination rate were excluded from the computation).
+    We considered two null models: neutral and background selection (as explained before).
+    Power (or true positive rate) to detect a sweep is the proportion of simulations with a sweep that were correctly classified as a sweep for each method and at each of the 100 locations along chromosome 1.
 

### 2024-12-09 analysis2_manuscript.tex
minor fix

@@
     In particular, we introduced a beneficial mutation at a given location (to be varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
-    For each sweep location, we performed 2000 replicate simulations, sampling from them only if the selected allele reached a frequency of 0.95 or greater.
+    For each sweep location, we performed 1000 replicate simulations, sampling from them only if the selected allele reached a frequency of 0.95 or greater.
     Rather than simulate the entire chromosome arm, replicate simulations were performed on 5cM regions of the chromosome at each of 100 different locations.

### 2024-12-10 analysis2_manuscript.tex
some revisions (#29)
* fig legends; fix schematic

* changes

* finished results and intro

* methods; + minor discussion  stuff

* update figs (ticks were fed up)

* minor fix

* typo
citation cleanup

@@
 \makeatother
+
+\newcommand{\tskit}{\texttt{tskit}\xspace}
 
@@
     using the \stdpopsim{} framework. As stated above, simulations included a model 
-    of (negtive) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
+    of (negative) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
     Three replicates of whole-genome simulations were performed for each species, with 
@@
     We ran simulations of human chromosome 1 using the three population out-of-Africa model
-    (identified as OutOfAfrica\_3G09 in the stdpopsim library; \cite{gutenkunst_inferring_2009})
-    and the genetic map estimated in \citet{frazer_second_2007} (HapMapII\_GRCh38).
+    (identified as OutOfAfrica\_3G09 in the stdpopsim library; \cite{gutenkunst2009inferring})
+    and the genetic map estimated in \citet{international2007second} (HapMapII\_GRCh38).
     For computational efficiency, we simulated a 5Mb region flanking 100 evenly distributed points along the chromosome 1.
@@
     (i) neutral (with the specified demography and genetic map),
-    (ii) background selection, in which we applied the DFE estimated in \citet{kim_inference_2017} (Gamma\_K17) to exon annotations (taken from Ensembl; ensembl\_havana\_104\_exons),
+    (ii) background selection, in which we applied the DFE estimated in \citet{kim2017inference} (Gamma\_K17) to exon annotations (taken from Ensembl; ensembl\_havana\_104\_exons),
     and (iii) sweep with background selection, in which we simulated the hard sweep as described with the addition of constraint in exonic regions (same as for \emph{ii}).
@@
     \diploshic was trained on data simulated using \texttt{discoal} \citep{kern2016discoal}.
-    Nucleotide diversity was computed directly from the underlying tree sequences using \tskit \citep{ralph_efficiently_2020} over 10 equally sized and non-overlapping subwindows.
+    Nucleotide diversity was computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently} over 10 equally sized and non-overlapping subwindows.
     The composite likelihood ratio (output by \sweepfinder) was computed at 21 uniformly distributed points along the simulated region.

### 2024-12-10 analysis2_manuscript.tex
edits with DRS

@@
     The top row shows the results of $N_e(t)$ estimates made from simulations without selection, while the bottom row shows the results
-    from simulations with deleterious mutatinos occurring in exons.
+    from simulations with deleterious mutations occurring in exons.
     In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
-    in various colors for the four methods. Dotted lines on the bottom row are to visually reinforce that inference is done
-    from a model with selection. We see that these methods generally perform reasonably well in the absence of selection, 
-    although GONE seems to overestimate $N_e(t)$ in the more recent past for the CEU and CHB populations.
+    in various colors for the four methods. Dotted lines on the top row are to visually reinforce that inference is done
+    from a model without selection. We see that these methods generally perform reasonably well in the absence of selection, 
+    although GONE seems to overestimate $N_e(t)$ for the CHB sample and to a lesser extent in the CEU sample. #ADK: is GONE time scaled incorrectly? i.e. gens not years?
     Presumably this could be due to patterns of linkage disequilibrium that result from the combination
@@
     In the presence of selection, different methods behave differently. 
-    For instance stairwaiplot which uses the SFS to infer $N_e(t)$ seems to slightly underestimate $N_e(t)$
-    but is reasonably close to the true $N_e(t)$ in the absence of selection.
-    Likewise \smcpp seems to be more robust to the presence of selection,
-    but there is some added noise to the returned estimates. 
+    For instance stairwayplot, which uses the SFS to infer $N_e(t)$, 
+    is reasonably close to the true $N_e(t)$ both with and without selection,
+    although it seems to infer significant population size changes that are not part of the true history.
+    Likewise \smcpp seems to be robust to the presence of selection,
+    but the timing of population size changes is slightly inaccurate. 
     \msmc seems to be the most affected by the presence of selection, 
-    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$,
+    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$ particularly in the very recent past,
     although estimates from \msmc have quite a bit of added noise even in the absence of selection.
 
-    % DRS: I don't agree with the interpretation in the previous two sentences (starting with "Likewise").
-    % MSMC is already barfing a little bit even without selection, and it is not clear to me from these
-    % plots that selection makes things any worse. Also, SMC++ is doing quite well without selection so maybe
-    % the minor drop in accuracy is meaningful in relative terms.
-    % ADK: I edited the text-- what do you think Dan?
-    % DRS: I am still finding this description pretty inaccurate. Happy to discuss over zoom if you think
-    % that might be faster.
+    % TODO: legend here needs to be changed. The dotted line indicates no selection, the solid selection
 
@@
     These suggest that the $\pi$ cutoff is a suprisingly effective method for detecting sweeps,
-    in these genomic regions at least, but that it lags behind \diploshic and \sweepfinder2 in terms of specificity
-    generally.
-    % DRS: the ROC curve doesn't tell us which methods have better/worse specificity, just how their sensitivities compare at a given specificity
-    % DRS: but we could say something about how pi might be "less robust to heterogeneity in mapping/genotyping quality across the genome than the other methods"
-    % DRS: or something like that. I know we didn't test that, but it must be the case.
+    in these genomic regions at least, but that it lags behind \diploshic and \sweepfinder2 in terms of accuracy in the YRI sample.
     %While this is not exhaustive benchmarking, we aim to give a sense to the reader
     %of how \stdpopsim can be used to benchmark the performance of different methods easily and efficiently.
-    These ROC curves, however, hide the previously seen phenomenon of high variation in power to detect sweeps along a realistic looking chromosome.
+    These ROC curves, however, hide the phenomenon of high variation in power (see Figure \ref{fig:chr1_power}) to detect sweeps along a realistic looking chromosome.
     That is, although in aggregate the $\pi$ cutoff may seem effective, it's power is highly heterogeneous across the chromosome.
@@
     It is unclear why that is so, but the effects of realistic background selection may not be as similar to sweeps as previously thought \citep{schrider2020background}.
-    % DRS: the exon-density results look pretty freakin flat to me. I think we could just curtail the description in this paragraph and
-    % DRS: say that there are no obvious patterns that stand out from the noise?
-    % MFR: I changed this a bit: I agree with Dan that there is no clear pattern.
 

### 2024-12-11 analysis2_manuscript.tex
edits throughout

@@
 
+    \begin{wrapfigure}{l}{0.5\textwidth}
+        \vspace{-0.0cm}
+        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
+        \caption{\label{fig:schematic}
+        \textbf{(A)} A schematic of the \stdpopsim catalog and the different components that
+        can be used in simulations. 
+        \textbf{(B)} The python API and \textbf{(C)} the command line interface.
+        The user specifies a species, a portion of
+        the genome to simulate, and optionally a genetic map, a model of
+        demography, and a model of selection that itself is composed of a
+        distribution of fitness effects (DFE) and a set of functional
+        annotations.}
+    \end{wrapfigure}
+    \hfill
+
     As with other \stdpopsim models, when incorporating selection we aim to hew as closely as possible
@@
 
-    % DRS: This figure could really use some Python API and comand line examples like in Figs 1B,C
-    % form the original stdpopsim paper
-    \begin{wrapfigure}[37]{l}{0.4\textwidth}
-        \vspace{-0.0cm}
-        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
-        \caption{\label{fig:schematic}
-        A schematic of the \stdpopsim catalog and the different components that
-        can be used in simulations. The user specifies a species, a portion of
-        the genome to simulate, and optionally a genetic map, a model of
-        demography, and a model of selection that itself is composed of a
-        distribution of fitness effects (DFE) and a set of functional
-        annotations.}
-    \end{wrapfigure}
-    \hfill
-
     % current numbers of DFEs / species / etc
@@
     where we have simulated genomes under a model of human out of Africa (OOA) demography
-    with and without purifying selection on exons. For this simulation we have used
+    with and without purifying selection acting, either on exons or on every site in the genome.
+    For this simulation we have used
     a genetic map from the HapMap Project \citep{international2007second} along with a
@@
     with samples taken from three modern geographic groups (CEU, CHB, and YRI from left to right).
-    The top row shows the results of $N_e(t)$ estimates made from simulations without selection, while the bottom row shows the results
+    The left column shows the results of $N_e(t)$ estimates made from simulations without selection, 
+    while the right column shows the results
     from simulations with deleterious mutations occurring in exons.
@@
     from a model without selection. We see that these methods generally perform reasonably well in the absence of selection, 
-    although GONE seems to overestimate $N_e(t)$ for the CHB sample and to a lesser extent in the CEU sample. #ADK: is GONE time scaled incorrectly? i.e. gens not years?
+    although GONE seems to overestimate $N_e(t)$ for the CHB sample and to a lesser extent in the CEU sample.
     Presumably this could be due to patterns of linkage disequilibrium that result from the combination
@@
         Performance of methods to infer $N_e(t)$ from a human out-of-Africa model \citep{ragsdale2019models}
-        with and without purifying selection on exons. The top row shows estimates of $N_e(t)$ from simulations
-        without selection, while the bottom row shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
+        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ from simulations
+        without selection, while the right column shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
         DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
@@
     In general, the power of the methods varies with respect to the population history of the sample.
-    For instance \sweepfinder2 has marginally higher power to detect sweeps in the CEU sample than \diploshic,
+    For instance \sweepfinder has marginally higher power to detect sweeps in the CEU sample than \diploshic,
     likely as a results of the fact that we here trained \diploshic on simulations under a constant size population,
     while the population size history of CEU is characterized by a strong recent bottleneck.
-    Conversely, \diploshic has higher power to detect sweeps in the YRI sample than \sweepfinder2,
+    Conversely, \diploshic has higher power to detect sweeps in the YRI sample than \sweepfinder,
     likely because that population size history more closely matches the population size history used to train \diploshic.
-    \sweepfinder2 power remains more constant along the chromosome, while the $\pi$ cutoff method varies wildly.
+    \sweepfinder power remains more constant along the chromosome, while the $\pi$ cutoff method varies wildly.
 
@@
     These suggest that the $\pi$ cutoff is a suprisingly effective method for detecting sweeps,
-    in these genomic regions at least, but that it lags behind \diploshic and \sweepfinder2 in terms of accuracy in the YRI sample.
+    in these genomic regions at least, but that it lags behind \diploshic and \sweepfinder in terms of accuracy in the YRI sample.
     %While this is not exhaustive benchmarking, we aim to give a sense to the reader
@@
     We see that power to detect sweeps is
-    a decreasing function of local recombination rate for \sweepfinder2 and $pi$ outliers,
+    a decreasing function of local recombination rate for \sweepfinder and $pi$ outliers,
     which is consistent with sweeps having a larger genomic footprint in regions of low

### 2024-12-11 analysis2_manuscript.tex

@@
     where we have simulated genomes under a model of human out of Africa (OOA) demography
-    with and without purifying selection acting, either on exons or on every site in the genome.
+    with and without purifying selection acting on exons.
     For this simulation we have used

### 2024-12-11 analysis2_manuscript.tex
Nt figure updates and MAE table updates for grapes vlaues (#39)

@@
     model of declining population size \citep{robinson2022critically} with and without background selection on exons. 
-    The top panel shows estimates of $N_e(t)$ from simulations
-    without selection, while the bottom panel shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
+    The left panel shows estimates of $N_e(t)$ from simulations
+    without selection, while the right panel shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
     DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
@@
     Performance of methods to infer the distribution of fitness effects (DFE) from genetic data
-    where all sites along the simulated chromosome can experience selected and neutral mutation
+    under constant population size  where all sites along the simulated chromosome can experience selected and neutral mutation
     as opposed to exonic regions only.
@@
 \midrule
-HomSap & Constant & all sites & 0.0027 & 0.0049 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
-HomSap & Constant & exons & 0.0012 & \bf{0.00081} & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
-HomSap & OOA 5R19 & all sites & 0.012 & 0.0079 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
-HomSap & OOA 5R19 & exons & 0.014 & \bf{0.0049} & 0.0072 & 0.031 & 0.051 & \bf{0.024} \\
-PhoSin & Constant & all sites & 0.024 & \bf{0.024} & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
-PhoSin & Constant & exons & 0.024 & \bf{0.024} & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
-PhoSin & Vaquita2Epoch 1R22 & all sites & 0.024 & \bf{0.023} & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
-PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & \bf{0.023} & 0.024 & \bf{0.18} & 0.23 & 0.21 \\
+HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
+HomSap & Constant & exons & \bf{0.0012} & 0.0101 & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
+HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
+HomSap & OOA 5R19 & exons & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
+PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
+PhoSin & Constant & exons & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
+PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
+PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
 \bottomrule

### 2024-12-18 analysis2_manuscript.tex
Edits from Ryan, mostly on DFE section.

--- 2024-12-11 analysis2_manuscript.tex
@@
 \newcommand{\polydfe}{\texttt{polyDFE}\xspace}
-\newcommand{\dadi}{$\partial$\texttt{a}$\partial$\texttt{i}\xspace}
+\newcommand{\dadicli}{\texttt{dadi-cli}\xspace}
 \newcommand{\grapes}{\texttt{GRAPES}\xspace}
@@
 
+\newcommand{\rngcomment}[1]{\textcolor{red}{RNG: #1}}
+
 \title{Adding selection to \stdpopsim: oh how the tables have turned}
@@
  \author[5*]{Silas Tittes}
+ \author[34*]{Linh N. Tran}
 \author[6]{Georgia Tsambos}
-\author[34]{Linh N. Tran}
+
 \author[2]{Maria Izabel A. Cavassim}
@@
 % \affil[13]{\small{Department of Biology, University of Toronto Mississauga, Mississauga ON, Canada}}
-% \affil[14]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
-% \affil[15]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
+\affil[14]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
+\affil[15]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
 % \affil[16]{\small{Department of Evolutionary Genetics, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany}}
@@
 % \affil[33]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
-% \affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson AZ 85721, USA}}
+\affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson AZ 85721, USA}}
 \affil[35]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
@@
     While this is so, joint simulation of complex demography and selection
-    is challenging, and requires a deep understanding of the underlying
+    is challenging,  requires a deep understanding of the underlying
     evolutionary processes, and necessitates a number of parameter choices including
@@
     non-equilibrium demographic histories. Ideally, one should be able
-    to model all of these features after estimates/annotations obtained
+    to model all of these features using estimates/annotations obtained
     from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.
@@
         distribution of fitness effects (DFE) and a set of functional
-        annotations.}
+        annotations. \rngcomment{I find the images too small to read the text comfortably.}}
     \end{wrapfigure}
@@
     While that is so, these DFEs can be applied to other species
-    in the catalog. For instance, one could simulate a butterfly species with a gamma DFE originally
+    in the catalog. For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
     estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
-    Furthermore, the user can specify a custom DFE, and provide their own annotations
+    Furthermore, the user can specify a custom DFE and provide their own annotations
     of functional elements to simulate selection in a species for which we do not yet have 
@@
     These simulations were then used to estimate $N_e(t)$ using four methods: 1) \msmc \citep{Schiffels2020}, 
-    2)\stairway method \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
+    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
     In this and what follows we performed 3 replicate whole genome simulations, so uncertainty in estimates might
-    be roughly assessed, however we are mindful of wasted computational resources as thorough benchmarking
+    be roughly assessed.
+    However we are mindful of wasted computational resources as thorough benchmarking
     is not the main purpose of this manuscript.
@@
     from Ragsdale \textit{et al.} (2019)
-    with samples taken from three modern geographic groups (CEU, CHB, and YRI from left to right).
+    with samples taken from three modern geographic groups (CEU, CHB, and YRI from top to bottom).
     The left column shows the results of $N_e(t)$ estimates made from simulations without selection, 
     while the right column shows the results
-    from simulations with deleterious mutations occurring in exons.
+    from simulations with deleterious mutations occurring in exons \rngcomment{(with inference based on simulated synonymous data from exons? or from whole genome?)}.
     In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
@@
         \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.  
+        \rngcomment{Need to update figure legend to remove ``all\_sites'' which seems not to be in plot.}
+        \rngcomment{Why do some methods have confidence intervals (like stairwayplot) and others don't (like msmc)?}
         }
@@
     \label{dfe}
-    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE)
+    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE) of new mutations
     from genetic data. Here, we show how the new models of selection in \stdpopsim can be used to benchmark and compare
@@
     human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
+
+    A challenge when comparing quantitative DFE estimates among tools and with simulations is differing conventions of factors of two.
+    The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
+    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$. %https://academic.oup.com/genetics/article/207/3/1103/5930676
+    Moreover, all three inference tools directly infer the distribution of the selection coefficient scaled by an inferred ancestral effective population size $N_e$, but they assume different scaling.
+    \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$.
+%https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
+    For comparison, here we have normalized all inference results to the convention of SLiM.
+
+    % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?
 
@@
         Estimates from three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
-        and 3) \dadi \citep{gutenkunst2009inferring}.}
+        and 3) \dadicli \citep{Huang2023}.
+            \rngcomment{Update figure legend. dadi=> dadi-cli.}
+}
         \label{fig:homsap-dfe.ooa}
@@
     from three software packages for inferring the DFE from genetic data:
-    \dadi \citep{gutenkunst2009inferring}, \polydfe \citep{tataru2020polydfe}, 
+    \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
     and \grapes \citep{galtier2016adaptive}.
@@
     while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
-    In a constant size population, all methods perform reasonably well, although \dadi seems to
+    In a constant size population, all methods perform reasonably well, although \dadicli seems to
     perform slightly worse than the other methods examined, at least for the small number of replicates we simulated here.
@@
     For the shape parameter in constant size simulations \grapes and \polydfe are roughly
-    equivalent in their performance whereas \dadi seems to yield overestimates.
+    equivalent in their performance whereas \dadicli seems to yield overestimates. 
+    \rngcomment{Has Fig.~\ref{fig:homsap-dfe.constant} been updated for the new GRAPES scaling?}
 
-    In the out-of-Africa model (fig \ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
+    In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
     In particular, variation in accuracy of the parameters estimated in each
@@
     have its own biases with respect to the way in which population history 
-    affects parameter estimates. For instance, for \dadi we see that estimates of 
+    affects parameter estimates. For instance, for \dadicli we see that estimates of 
     the shape parameter are most accurate in the YRI sample, while for \polydfe
@@
     With respect to the strength of selection, as in constant size populations,
-    all methods are performing reasonably well, although \dadi seems to
-    trail in terms of accuracy.
+    all methods are performing reasonably well, although \polydfe seems to
+    lead in terms of accuracy.
 
@@
         \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
-\includegraphics{figures/PhoSin/Vaquita2Epoch_1R22/hs_hist.pdf}
         \caption{
@@
         Simulations were performed using a model of vaquita porpoise demography \citep{robinson2022critically} with a gamma-distributed DFE
-        acting on exons parameterized by a mean selection coefficient, a shape parameter, and a relationship between selection and domiance. Estimates of the 
+        acting on exons parameterized by a mean selection coefficient, a shape parameter, and a relationship between selection and dominance. Estimates of the 
         parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
@@
         of the three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
-        and 3) \dadi \citep{gutenkunst2009inferring,kim2017inference}.
+        and 3) \dadicli \citep{Huang2023,kim2017inference}.
         C) Shown is the simulated distribution of $2 h s$, where $h$ is the dominance coefficient, along with the simulated distribution of $s$ and the median inferred DFEs.
 The distribution of $2 h s$ is multimodal because the simulated relationship between $h$ and $s$ is not continuous.
-        % RNG: For C, need to extract inferred values from pipeline.
+            \rngcomment{Update figure legend. dadi=> dadi-cli.}
         }
@@
     from the same three DFE-inference packages on simulations of the
-    vaquita porpoise genome. In the constant size model (fig \ref{fig:vaquita-dfe.constant}) we see that all methods
+    vaquita porpoise genome. In both situations, we see that all methods
     perform uniformly worse that in the human genome simulations, with consistent underestimation of the mean selection coefficient    
@@
     so these results suggest caution in interpreting DFE inferences.
-    %RNG: Could use Kirk's help with the best references here
 
@@
     Here, we performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring} with purifying selection acting on exons under the gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second}, but also include sweeps.
-    In particular, we introduced a beneficial mutation at a given location (to be varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
+    In particular, we introduced a beneficial mutation at a given location (that was varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
     For each sweep location, we performed 1000 replicate simulations, sampling from them only if the selected allele reached a frequency of 0.95 or greater.
     Rather than simulate the entire chromosome arm, replicate simulations were performed on 5cM regions of the chromosome at each of 100 different locations.
-    As recombination rate varies across the genome, the size of the region in base pairs simulated varies, however the recombination distance, which is critical for the linked effects of a sweep, is held constant.
+    As recombination rate varies across the genome, the size of the region in base pairs simulated varied, however the recombination distance, which is critical for the linked effects of a sweep, was held constant.
 
@@
         cutoff of $\pi$ to find regions of reduced diversity (bottom row).
-        True positive rate is shown for these methods for the CEU and YRI
+        Power (true positive rate) is shown for these methods for the CEU and YRI
         samples (left and right respectively).  Under each method we show
@@
     thus the out-of-Africa bottleneck has does not mute the signal of a sweep as much as in the CEU sample. This is consistent with
-    what we know about the effects of demography on the power to detect sweeps \citep[e.g.,][]{simonsen1995properties}.
+    what is known about the effects of demography on the power to detect sweeps \citep[e.g.,][]{simonsen1995properties}.
     Second, the power to detect sweeps is slightly lower when we use a null model with purifying selection acting on exons (consistent with \cite{schrider2020background}) but the effect is not
-    as strong as one might expect. This is likely because the strength of selection acting on exons is relatively weak, and the
+    as strong as one might expect. 
+    \rngcomment{I'm confused by the terminology ``null model'' throughout this section.
+Are we always testing on simulated data that includes background selection, and ``null model'' refers to how the methods are tuned?
+Or are we comparing power on data simulated with and without background selection?} This is likely because the strength of selection acting on exons is relatively weak, and the
     sweep detection methods we have used are relatively robust to the effects of background selection. 
@@
 
-    We can further explore the performance of the methods by plotting the receiver operating characteristic (ROC) curves for the methods.
+    We further explored the performance of the methods by plotting their receiver operating characteristic (ROC) curves.
     In Figure \ref{fig:roc-curves}, we show the ROC curves for the three methods for detecting sweeps.
     False positive rates were established using neutral simulations that contained no sweeps
-    but were conditioned on the same population history as the CEU and YRI samples.
-    These suggest that the $\pi$ cutoff is a suprisingly effective method for detecting sweeps,
+    but were conditioned on the same population history as the CEU and YRI samples. \rngcomment{Did these simulations contain background selection?}
+    These suggest that the $\pi$ cutoff is a surprisingly effective method for detecting sweeps,
     in these genomic regions at least, but that it lags behind \diploshic and \sweepfinder in terms of accuracy in the YRI sample.
@@
     These ROC curves, however, hide the phenomenon of high variation in power (see Figure \ref{fig:chr1_power}) to detect sweeps along a realistic looking chromosome.
-    That is, although in aggregate the $\pi$ cutoff may seem effective, it's power is highly heterogeneous across the chromosome.
+    That is, although in aggregate the $\pi$ cutoff may seem effective, its power is highly heterogeneous across the chromosome.
 
@@
     We see that power to detect sweeps is
-    a decreasing function of local recombination rate for \sweepfinder and $pi$ outliers,
+    a decreasing function of local recombination rate for \sweepfinder and $\pi$ outliers,
     which is consistent with sweeps having a larger genomic footprint in regions of low
@@
     \diploshic on the other hand shows a slight increase in power with recombination rates,
-    probably because it is designed to distinguish regions where a sweep happend from regions linked to a sweep.
+    probably because it is designed to distinguish regions where a sweep happened from regions linked to a sweep.
     We also see that the presence of purifying selection acting on exons has a relatively small effect on the power to detect sweeps,    
@@
     We have demonstrated the utility of this new API by showing how it can be used to benchmark
-    the performance of different methods for demographic inference, the inference of the distribution
-    of fitness effects of mutations, and the detection of selective sweeps.
+    the performance of different methods for demographic histtory inference, the inference of the distribution
+    of fitness effects of new mutations, and the detection of selective sweeps.
     In general, this is an important step forward for the field of population genetics,
@@
     In this study we performed simulations of the human and vaquita porpoise genomes
-    using the \stdpopsim{} framework. As stated above, simulations included a model 
+    using the \stdpopsim{} framework. As stated above, simulations did or did not include a model 
     of (negative) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
@@
     We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
-    \dadi \citep{gutenkunst2009inferring,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
+    \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
     Each software package was run using the same set of simulations.
@@
     MFR, ST, NP, and ADK were supported in part by NIH award R35GM148253.
+    RNG and LNT were supported in part by NIH award R35GM149235.
+
     %other funding sources?
@@
     %DRS-fig: If it is the right fig, get rid of the pop0 legend inset
-    Each row of the figure shows estimates from a different method: 1) dadi \citep{gutenkunst2009inferring}, 2) polyDFE \citep{tataru2020polydfe},  
-    3) DFE-alpha \citep{eyre2009estimating}, and 4) GRAPES \citep{galtier2016adaptive}.  
+    \rngcomment{Update figure legend. dadi=> dadi-cli.}
+    Each row of the figure shows estimates from a different method: 1) GRAPES \citep{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe}, and 3) \dadicli \citep{Huang2023}.
     }
@@
     of the three different methods are shown: 1) GRAPES \citep{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe},
-    and 3) dadi \citep{gutenkunst2009inferring}.}
+    and 3) \dadicli \citep{Huang2023}.}
+    \rngcomment{Update figure legend. dadi=> dadi-cli.}
     %DRS-fig: remove the pop 0 legend again

### 2024-12-18 analysis2_manuscript.tex
Remove comment fixed in last merge.
Update analysis2_manuscript.tex

@@
         \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.  
-        \rngcomment{Need to update figure legend to remove ``all\_sites'' which seems not to be in plot.}
         \rngcomment{Why do some methods have confidence intervals (like stairwayplot) and others don't (like msmc)?}
@@
     We have demonstrated the utility of this new API by showing how it can be used to benchmark
-    the performance of different methods for demographic histtory inference, the inference of the distribution
+    the performance of different methods for demographic history inference, the inference of the distribution
     of fitness effects of new mutations, and the detection of selective sweeps.

### 2024-12-19 analysis2_manuscript.tex

@@
         \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.  
-        \rngcomment{Why do some methods have confidence intervals (like stairwayplot) and others don't (like msmc)?}
         }
@@
         and 3) \dadicli \citep{Huang2023}.
-            \rngcomment{Update figure legend. dadi=> dadi-cli.}
 }
@@
     equivalent in their performance whereas \dadicli seems to yield overestimates.
-    \rngcomment{Has Fig.~\ref{fig:homsap-dfe.constant} been updated for the new GRAPES scaling?}
 
@@
 The distribution of $2 h s$ is multimodal because the simulated relationship between $h$ and $s$ is not continuous.
-            \rngcomment{Update figure legend. dadi=> dadi-cli.}
         }
@@
     %DRS-fig: If it is the right fig, get rid of the pop0 legend inset
-    \rngcomment{Update figure legend. dadi=> dadi-cli.}
     Each row of the figure shows estimates from a different method: 1) GRAPES \citep{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe}, and 3) \dadicli \citep{Huang2023}.
@@
     and 3) \dadicli \citep{Huang2023}.}
-    \rngcomment{Update figure legend. dadi=> dadi-cli.}
     %DRS-fig: remove the pop 0 legend again
@@
 \toprule
-species ID & demography & annotation & \makecell{MAE \\ $E|s|$ \\ dadi} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
+species ID & demography & annotation & \makecell{MAE \\ $E|s|$ \\ dadi-cli} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi-cli} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
 \midrule

### 2025-01-08 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
 
+\newcommand{\kelcomment}[1]{\textcolor{blue}{KEL: #1}}
+
 \title{Adding selection to \stdpopsim: oh how the tables have turned}
+    \kelcomment{Normally, I like quirky, witty titles, but I'm not sure I get this one...what is "how the tables have turned" in reference to? }
 
@@
     However, simulating realistic models that incorporate both selection and complex demographic histories
-    is challenging, limiting our ability to benchmark statistical methods and explore theoretical predictions.
+    is challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and explore theoretical predictions.
     \stdpopsim is a tool for facilitating the simulation of a contiually expanding catalog of population genetic models.
@@
     and forward-in-time simulations of complex demography and selection
-    \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.
+    \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.  \kelcomment{maybe also cite https://doi.org/10.1093/bioinformatics/btn522}
     The development of simulation tools has been driven by the need to
     understand the effects of complex evolutionary processes on genetic
-    variation \citep[e.g.,][]{galloway2020few}, to provide a null model for hypothesis testing
+    variation \citep[e.g.,][]{galloway2020few},\kelcomment{also cite: PMID: 32487519 } to provide a null model for hypothesis testing
     \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
@@
     simulation-based inference methods \citep[e.g.,][]{beaumont2002approximate,pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}.
+\kelcomment{also cite: PMID: 36617238 }
     While this is so, joint simulation of complex demography and selection
@@
     and can be combined with other \texttt{DFE} objects to provide 
-    a rich model of how selection may differ along a chromosome. 
+    a rich model of how selection may vary along a chromosome. 
     Once specified, a model of selection can then be run using \stdpopsim's standard interfaces
@@
     humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
-    While that is so, these DFEs can be applied to other species
+    While the DFEs are from a limited number of sepecies, these DFEs can be applied to other species
     in the catalog. For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
@@
     using a variety of methods, including the pairwise sequentially Markovian coalescent
-    (PSMC) \citep{li2011inference}, through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
+    (PSMC) \citep{li2011inference}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
     as well as through identity by descent information \citep{santiago2020recent}. 
-    While that is so, it has been well described that selection, because of its effects on
-    patterns of genetic variation, can bias estimates of $N_e(t)$
+    While these methods perform well under neutral evolution, because of its effects on
+    patterns of genetic variation, natural selection on linked sites can bias estimates of $N_e(t)$
     away from true census population sizes simply because selection increases the 
@@
     DFE for deleterious alleles (estimated by \cite{huber2017determining})
-    that acts on exonic positions defined from the HAVANA group release 104 for the human genome.
+    that acts on exonic \kelcomment{do you mean exonic, or nonsynonymous?} positions defined from the HAVANA group release 104 for the human genome.
     % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
@@
     be roughly assessed.
-    However we are mindful of wasted computational resources as thorough benchmarking
-    is not the main purpose of this manuscript.
+    However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simualtion replicates.
 
@@
 
-    A challenge when comparing quantitative DFE estimates among tools and with simulations is differing conventions of factors of two.
+    A challenge when comparing quantitative DFE estimates among tools and with simulations is differing conventions of factors of two in terms of how the selection coefficient ($s$) is delimited.
     The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
@@
     lead in terms of accuracy.
+\kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things.
 

### 2025-01-08 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
 \kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things.
+That said, I think it would be good to plot the empirical (in the case simulated) SFS along with the SFSs from the MLE models.
+I would also do this for the demogrpahic models as well as the DFE models}
 
@@
     from the same three DFE-inference packages on simulations of the
-    vaquita porpoise genome. In both situations, we see that all methods
+    vaquita porpoise genome.
+Importantly, these simulations use the vaquita demogrpahic model shown in Fig.
+X and a DFE that includes recessive deleterious mutations, as suggested in Robinsion et al.
+(2022).
+All methods
     perform uniformly worse that in the human genome simulations, with consistent underestimation of the mean selection coefficient    
     and overestimation of the shape parameter.
-    A critical distinction is that the previous human simulations employed a DFE that assumed all selected mutations had additive effects
-    (no dominance, $h = 1/2$), whereas the vaquita simulations use a DFE in which more deleterious mutations are more recessive ($h < 1/2$).
-    For all three software packages, the inferred mean and shape of the distribution of selection coefficients 
+This is not surprising as a mis-specified model is being fit to the data.
+All DFE inference methods used here assume an addive model (no dominance, $h = 1/2$) while the simulated data from the vaquita  was generated from a DFE in which more deleterious mutations are more recessive ($h < 1/2$)
+
+    Reassuringly, for all three software packages, the inferred mean and shape of the distribution of selection coefficients 
     is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient
@@
     where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
-    Strongly deleterious alleles are often observed to be recessive in many species \citep{mukai1972mutation},
+   \kelcomment{I think we should move this next sentence to the Discussion and expand on it a bit} Strongly deleterious alleles are often observed to be recessive in many species \citep{mukai1972mutation},
     so these results suggest caution in interpreting DFE inferences.
@@
     To decrease runtimes, we used a SLiM scaling factor of 2 with a burn-in of 2N generations.
+\kelcomment{I'm confused about the 100 diploid individuals samples vs.
+the 10 diploid individuals sampled.
+Did some analyses use 100 and others use 10?
+That's fine, but maybe we can specify that more clearly...}
 
@@
     may in reality be under selection. As above, all methods were run using default settings.
+    \kelcomment{I think we need a lot more detail here...how big was the coding sequence length?
+How was demogrpahic inference done for the different models via the different programs?
+How was the unscaling of E[s] done?, etc}
 

### 2025-01-10 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
     be roughly assessed.
-    However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simualtion replicates.
+    However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.
 

### 2025-01-10 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
 
-    A challenge when comparing quantitative DFE estimates among tools and with simulations is differing conventions of factors of two in terms of how the selection coefficient ($s$) is delimited.
+    A challenge when comparing quantitative DFE estimates among tools and with simulations is differing conventions of factors of two in terms of how the selection coefficient ($s$) is defined.
     The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.

### 2025-01-10 analysis2_manuscript.tex
Update analysis2_manuscript.tex
small change to make this build (#43)

@@
 
-\title{Adding selection to \stdpopsim: oh how the tables have turned}
-    \kelcomment{Normally, I like quirky, witty titles, but I'm not sure I get this one...what is "how the tables have turned" in reference to? }
+\title{Accesible, realistic genome simulation with selection using \stdpopsim}
 
@@
 \author[5]{Scott T. Small}
+\author[5]{Clara T. Rehmann}
 % \author[26]{Per Unneberg}
@@
 Importantly, these simulations use the vaquita demogrpahic model shown in Fig.
-X and a DFE that includes recessive deleterious mutations, as suggested in Robinsion et al.
-(2022).
+\ref{fig:1pop-vaquita-demography} and a DFE that includes recessive deleterious mutations, as suggested in \cite{robinson2022critically}.
 All methods

### 2025-01-11 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
     where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
-   \kelcomment{I think we should move this next sentence to the Discussion and expand on it a bit} Strongly deleterious alleles are often observed to be recessive in many species \citep{mukai1972mutation},
-    so these results suggest caution in interpreting DFE inferences.
+   \kelcomment{I think we should move this next sentence to the Discussion and expand on it a bit....update on 1/10--I moved this to the Discussion and fleshed it out}
 
@@
     empirical data, with reproducibility, computational efficiency, and rigor.
+
+    One important finding we made using the new stdpopsim selection API is that inference of the DFE is sensitive to the dominance of deleterious mutations.
+All three of the inference methods of the DFE applied here assume that deleterious mutations act in an additive fashion (h=0.5) by default.
+However, the simulated DFE for the vaquita genome included highly recessive mutations.
+Specifically, these simulations used a DFE where mutations with $s>0.01$ are quite recessive, with $h<=0.01$ and where mutations with $0.001<s<0.01$ have $h=0.1$.
+We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Fig 4).
+This performance is not surprising as decreasing h for a given distribution of s will lead to a decrease in hs, or an apparent decrease in the amount of selection.
+Fitting a mis-specified DFE that assumes $h=0.5$ will then lead to underestimates of $E[s]$.
+Indeed, the distribution of $hs$ is much more accurately estimated (Fig 4b).
+Overall, these results are consistent with prior work showing that it is challenging to separately infer the DFE of $s$ along with $h$ (cite PMID: 24830675; PMID: 39302992; PMID: 34951958 ) and that a large proportion (>5%) of recessive deleterious mutations can confound inferences of $s$ (PMID: 37074880).
+Given the plethora of evidence for recessive deleterious mutations (cite PMID: 4630587, PMID: 39114967 , PMID: 30013096 , PMID: 21098719 ), these results suggest the need for care in DFE inference.
+Kyriazis & Lohmueller (PMID: 39302992 ) inferred the DFE of $s$ for nonsynonymous mutations in humans when considering different values of $h$.
+Intriguingly, additive models show a satisfactory fit to the SFS of nonsynonymous variants in humans (Kyriazis et al) and more complex models of dominance do not substantially improve the fit.
+Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome, show a similar fit to the SFS and may be more biologically realistic than fully additive models, given the evidence for the presence of recessive deleterious mutations.
+More broadly, new computational approaches are needed to precisely quantify the DFE and the covariance of $h$ and $s$.
 

### 2025-02-03 analysis2_manuscript.tex
Ilan's edits and comments (not including sweeps and the dicsussion)

@@
 \newcommand{\kelcomment}[1]{\textcolor{blue}{KEL: #1}}
+
+\newcommand{\igcomment}[1]{\textcolor{orange}{IG: #1}}
 
@@
     as the backend engine for efficient population genetic simulation.
+    \igcomment{edited here to add details on the spec of the DFE and Annotation classes
+    moved below the information on how they are used in simulation. Please check!}
     To extend \stdpopsim's functionality to include models of selection, 
-    we introduce an important new class representing a specfic distribution of fitness effects
-    to the \stdpopsim API: the \texttt{DFE} class. 
-    This class provides an interface for specifying
-    a distribution of the fitness effects of new mutations, and
-    is added to a \texttt{Contig} object to provide a description
-    of the fitness effects of mutations along a chromosome.
-    A \texttt{DFE} object can apply to specified intervals within a chromosome,
-    and can be combined with other \texttt{DFE} objects to provide 
-    a rich model of how selection may vary along a chromosome. 
-    Once specified, a model of selection can then be run using \stdpopsim's standard interfaces
-    but currently is limited to using SLiM simulation engine, which
-    is perhaps the most flexible simulation engine for modeling selection currently available.
+    we introduce two new classes
+    %an important new class representing a specfic distribution of fitness effects
+    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
+    The \texttt{DFE} class provides an interface for specifying
+    a distribution of the fitness effects of new mutations.
+    Each \texttt{DFE} object specifies a list of mutation types and
+    their proportions, where 
+    each mutation type is associated with a dominance coefficient (in $[0,1]$),
+    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
+    and the parameters of this distribution.
+    Each \texttt{Annotation} object is typically associated with a collection of functional
+    genomic elements specified using a GFF3 file. %
+    %    
+    For example, say that we wish to implement a model of selection consistent with the
+    DFE inferred by \citep{kim2017inference} for exons in the human genome.
+    The associated \texttt{DFE} object is defined with two mutation types,
+    neutral and deleterious, corresponding to synonymous and non-synonimous sites.
+    Their proportions are set to $0.3$ and $0.7$ respectively, following their expected
+    prevelanece in exons.
+    The neutral mutation type is associted with a fixed selection coefficient (0),
+    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
+    a gamma-distributed selection coefficient
+    with mean and shape as inferred by \citep{kim2017inference}.
+    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients
+    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below).
 
-    \begin{wrapfigure}{l}{0.5\textwidth}
-        \vspace{-0.0cm}
+    \begin{figure}[t!]
         \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
@@
         distribution of fitness effects (DFE) and a set of functional
-        annotations. \rngcomment{I find the images too small to read the text comfortably.}}
-    \end{wrapfigure}
-    \hfill
+        annotations. \rngcomment{I find the images too small to read the text comfortably.}
+        \igcomment{I changed this to full width. We might want to reduce font size in the panel titles.}}
+    \end{figure}
 
+    \igcomment{modified this paragraph to focus on the catalog and moved simulation details to next paragraph}
     As with other \stdpopsim models, when incorporating selection we aim to hew as closely as possible
-    to realistic genomes, and provide public annotations of specices' functional genomic
-    elements that can be used along with a \texttt{DFE} object to
-    readily provide realistic model of selection for a specific genome and set of
-    populations. A schematic of the \stdpopsim catalog is shown in Figure \ref{fig:schematic}.
-    From a user's perspective, one can specify a species, a portion of the genome to simulate,
-    a genetic map if available, a model of demography, and a model of selection that is
-    composed of a set of annotations and a \texttt{DFE} object, all in an automated and user-friendly fashion.
-
-    The vast majority of published DFE estimates are based on a limited number of species,
-    so \stdpopsim currently has implemented published DFEs from
-    four species: \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
+    to realistic genomes.
+    Thus, the \stdpopsim catalog provides \texttt{Annotation} object based on public annotations
+    of specices' functional genomic elements 
+    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
+    Since the vast majority of published DFE estimates are based on a limited number of species,
+    \stdpopsim currently has implemented DFEs for four species:
+    \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
     humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
@@
     % not needed here.
+
+    From a user's perspective, a model of selection is specified by pairing a \texttt{DFE} object
+    with a collection of genomic segments defined using an \texttt{Annotation} object.
+    Multiple such pairings can be defined in one simulation
+    to provide a rich model of how selection may vary along a chromosome.
+    For example, coding sequences may be associated with one DFE and non-coding sequences in exons with another DFE.
+    Simulations with selection are then implemented by specifying a species, 
+    a portion of the genome to simulate,
+    a genetic map if available, a model of demography, and pairings of DFEs with genomic segments
+    using either the python API or the command line interface (Figure \ref{fig:schematic}B-C).
+    %IG: commented this out because I think it's covered by the text I wrote above
+    %A \texttt{DFE} object can apply to specified intervals within a chromosome,
+    %and can be combined with other \texttt{DFE} objects 
+    Simulations with selection can currently be generated only 
+    using the SLiM simulation engine, which is
+    %is perhaps
+    the most flexible simulation engine for modeling selection available to-date
+    (msprime currently cannot be used to simulate models of selection).
+    %IG: this is a nice statement, but I couldn't figure where to plug it after my changes so I'm leaving it commented
+    %all in an automated and user-friendly fashion.
 
@@
     for which we have implemented 28 demographic models, have 37 genetic maps, and 7 DFEs. %someone should double check the numbers
+    \igcomment{do we want to also mention the number of annotation objects?}
     Additions to the catalog are ongoing---we point the reader to our previously published 
@@
     using a variety of methods, including the pairwise sequentially Markovian coalescent
-    (PSMC) \citep{li2011inference}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
+    (PSMC) \citep{li2011inference}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
+    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
+    through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
     as well as through identity by descent information \citep{santiago2020recent}. 
@@
 
-    Using \stdpopsim, we can easily simulate genetic data under a model of selection and then 
-    examine the performance of different methods for inferring $N_e(t)$ from these simulated 
-    data. In Figure \ref{fig:1pop-human-demography}, we show the results of such a benchmark
-    where we have simulated genomes under a model of human out of Africa (OOA) demography
-    with and without purifying selection acting on exons.
-    For this simulation we have used
-    a genetic map from the HapMap Project \citep{international2007second} along with a
-    DFE for deleterious alleles (estimated by \cite{huber2017determining})
-    that acts on exonic \kelcomment{do you mean exonic, or nonsynonymous?} positions defined from the HAVANA group release 104 for the human genome.
+    \igcomment{Questions about methods here:
+    (1) Which DFE model was used here? The previous version of the text cited Huber, but the production config says Kim;
+    (2) Was DFE applied to all exonic sites or nonsynonymous? (I think that all exonic sites, but if non-syn, we need to clarify);
+    (3) We mention 3 reps, but this is not used in the results (see also comment in next para)
+    }
+    Using \stdpopsim, we can easily examine the effect of selection on the effective
+    population sizes inferred by different methods.
+    For this purpose, we simulated human genomes with and without natural selection.
+    In both scenarios, we ran three replicate simulations
+    under the out of Africa (OOA) demographic model
+    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second}.
+    Simulations with selection were implemented by modeling purifying selection on exons,
+    defined from the HAVANA group release 104 for the human genome,
+    using a DFE inferred by \cite{kim2017inference}.
+    Each of the six simulated datasets contained 300 genomes, 100 for each geographic group: CEU, CHB, and YRI.
+    All simulations were implemented using a relatively short python script (as illustrated in Figure 1B).
     % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
     % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
-    These simulations were then used to estimate $N_e(t)$ using four methods: 1) \msmc \citep{Schiffels2020}, 
+    % IG: added a note about this above.
+    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
     2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
-    In this and what follows we performed 3 replicate whole genome simulations, so uncertainty in estimates might
-    be roughly assessed.
-    However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.
+    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
+    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.
 
-    In Figure \ref{fig:1pop-human-demography}, we show six panels, each of which shows the estimated $N_e(t)$
-    from data simulated under the five-population model of human out-of-Africa demography
-    from Ragsdale \textit{et al.} (2019)
-    with samples taken from three modern geographic groups (CEU, CHB, and YRI from top to bottom).
-    The left column shows the results of $N_e(t)$ estimates made from simulations without selection, 
-    while the right column shows the results
-    from simulations with deleterious mutations occurring in exons \rngcomment{(with inference based on simulated synonymous data from exons? or from whole genome?)}.
-    In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
-    in various colors for the four methods. Dotted lines on the top row are to visually reinforce that inference is done
-    from a model without selection. We see that these methods generally perform reasonably well in the absence of selection, 
-    although GONE seems to overestimate $N_e(t)$ for the CHB sample and to a lesser extent in the CEU sample.
-    Presumably this could be due to patterns of linkage disequilibrium that result from the combination
-    of bottleneck and growth in these populations.
+    \igcomment{Questions about inference and summary:
+    (1) In simulation with selection, do we infer Ne from all sites (exons and non-exons)?
+        This is what I understand from the code, but if not, we should clarify.
+    (2) The effect of selection on inference is difficult to glean from Fig 2.
+        The main thing I see there is that the influence is very small.
+        Maybe we should have a separate plot showing each method in a diff panel with inferences from all 6 datasets?
+        This might allow us to see differences between the two scenarios (with/without selection).  
+    }
+    %IG: I commented this out because it repeats some of the info above and we now start by directly
+    %discussing the results rather than describing the figure
+    %In Figure \ref{fig:1pop-human-demography}, we show six panels, each of which shows the estimated $N_e(t)$
+    %from data simulated under the five-population model of human out-of-Africa demography
+    %from Ragsdale \textit{et al.} (2019)
+    %with samples taken from three modern geographic groups (CEU, CHB, and YRI from top to bottom).
+    %The left column shows the results of $N_e(t)$ estimates made from simulations without selection, 
+    %while the right column shows the results
+    %from simulations with deleterious mutations occurring in exons.
+     % IG: commented this out, because these details are better left for the fig caption
+    %In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
+    %in various colors for the four methods. Dotted lines on the top row are to visually reinforce that inference is done
+    %from a model without selection.
+    We see that all methods generally perform reasonably well in the absence of selection (Figure \ref{fig:1pop-human-demography}; left).
+    The only exception to this is that GONE seems to overestimate $N_e(t)$ from the CHB genomes,
+    %IG: I commented this out because this seems like a very minor blip
+    %and to a lesser extent in the CEU sample.
+    presumably due to patterns of linkage disequilibrium that result from the combination
+    of bottleneck and growth in the populations ancestral to these samples.
+    %
+    \igcomment{I don't really see evidence for the observations listed below in Fig 2. See my suggestion above to enhance the comparison.}
     In the presence of selection, different methods behave differently. 
@@
         \label{fig:1pop-human-demography}
-        Performance of methods to infer $N_e(t)$ from a human out-of-Africa model \citep{ragsdale2019models}
-        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ from simulations
-        without selection, while the right column shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
+        Inference of $N_e(t)$ from a human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
+        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ obtained from genomes simulated
+        without selection, while the right column shows estimates of $N_e(t)$ from genomes simulated with a gamma-distributed   
         DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
         \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+        \igcomment{This is a nice figure but it's difficult to see if there is any subtle influence of selection from it.
+        We should think of one multi-panel main figure to summarize different aspects of the $N_e(t)$ analysis, and push the rest to the supplement.
+        Specifically for this figure, I'm not sure we need to use different line types for left and right since the plots are separated.}  
         }
@@
     %DRS-fig: reminder to update the legend to just say ``exons'' instead of ensembl_havana_whatever...
+    %IG: this seems to have been taken care of
 
-    To highlight the ease of comparisons between different species and models enabled by \stdpopsim,
-    we also performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
-    In Figure \ref{fig:1pop-vaquita-demography}, we show the results of $N_e(t)$ inference
-    where we simulated genomes under a model of vaquita demography from \textcite{robinson2022critically}.
-    For these simulations, we used the genome structure and exon annotations from the vaquita genome assembly
-    and a DFE from \textcite{robinson2022critically} that acts on exons, however no genetic map 
-    is currently available, so recombination rates are assumed to be constant across the genome.
-    The demographic model is a simple two-epoch model where population size has declined in the recent past. 
-    In this case, we see that all methods perform well both in the absence and presence of selection,
-    with a close match between the true $N_e(t)$ and the estimated $N_e(t)$ in all cases.
+    We note that while selection in these simulations directly acts on a small fraction of sites in the analyzed genomes,
+    it has a widespread influence on genetic diversity through background selection \citep[e.g.][]{charlesworth1993effect}.
+    Our results here suggest that the combined effect of direct selection and background selection does not considerably
+    influence infrerence of $N_e(t)$ for human populations under reasonable assumptions about their evolutionary history.
+    %
+    To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
+    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
+    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
+    and since no genetic map is currently available, we assumed a constant recombination rate across the genome.
+    In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
+    annotated from the vaquita genome assembly \igcomment{Maybe cite Morin et al. 2010 for assembly and annotations?}.
+    \igcomment{I added this about the h-s relationship. Please check!}
+    In this DFE model, we implemented a relationship between the dominance coefficient ($h$) and the selection coefficient ($s$)
+    as follows: $s<-0.1 \Rightarrow h=0$; $s\in [-0.1,-0.01) \Rightarrow h=0.01$; $s\in [-0.01,-0.001) \Rightarrow h=0.1$;
+    and $s\geq -0.001 \Rightarrow h=0.4$.
+    This approach is similar to what was used in the simulations cunducted by \textcite{robinson2022critically} and is easily
+    implemented in \stdpopsim's \texttt{DFE} class.
+    Using these simulations, we see that all methods infer $N_e(t)$ reasonably well with or without selection
+    (Figure \ref{fig:1pop-vaquita-demography}), although they do not accurately identify the boundary time
+    between the two epochs.
     % MFR: Should we say somewhere that people use to expect from BGS?
 That is, that it can be seen as a reduction in local rates of coalescences and thus that we would expect these methods to be somewhat downwardly biased in the presence of BGS?
+    % IG: we can address this in the discussion.
 
@@
     \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+    \igcomment{see my comment on fig 2. We should decide what part of this stays in the main figure and what ends up in the supp.}  
     }
@@
     Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE) of new mutations
-    from genetic data. Here, we show how the new models of selection in \stdpopsim can be used to benchmark and compare
-    the performance of different methods for inferring the DFE. We performed simulations using the 
-    human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
-
-    A challenge when comparing quantitative DFE estimates among tools and with simulations is differing conventions of factors of two in terms of how the selection coefficient ($s$) is defined.
+    from genetic data. The new framework of modeling selection in \stdpopsim is ideally designed to benchmark and compare
+    the performance of different methods for inferring the DFE. 
+    %IG: I don't think we need this intro.
+    %We performed simulations using the 
+    %human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
+    %
+    A challenge when comparing different inferred and simulated distributions of the selection coefficient ($s$) is differing convention regarding its definition.
     The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
-    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$. %https://academic.oup.com/genetics/article/207/3/1103/5930676
-    Moreover, all three inference tools directly infer the distribution of the selection coefficient scaled by an inferred ancestral effective population size $N_e$, but they assume different scaling.
+    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
+     %https://academic.oup.com/genetics/article/207/3/1103/5930676
+    Moreover, inferred distributions of selection coefficients are typically scaled by an inferred ancestral
+    effective population size $N_e$, and different methods assume different scaling.
     \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$.
@@
     For comparison, here we have normalized all inference results to the convention of SLiM.
+    \igcomment{Can we summarize this in a table or in a short methods section?
+    For each method, we could specify the scaling we apply to compare its inferred distribution to the one used in simulation.}
 
@@
         \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_OutOfAfricaArchaicAdmixture_5R19_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
-        \caption{A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
-        Simulations were performed using a human out-of-Africa model \citep{ragsdale2019models} with a gamma-distributed DFE
-        acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
-        parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
-        The human model has three extant populations (CEU, CHB, and YRI), and parameter estimates from each
-        sample are shown in different colors.
-        Estimates from three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
-        and 3) \dadicli \citep{Huang2023}.
+        \caption{Inferred versus simulated DFEs from simulated human genomes.
+        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
+        acting on exons.
+        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; left) and shape parameter (right).
+        The DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
+        by three different methods: \grapes, \polydfe , and \dadicli. The dashed horizontal lines represents the simulated values.
+        \igcomment{Does the boxplot represent the three simulation reps? Is each method run several times?}
 }
@@
 
-    In Figures \ref{fig:homsap-dfe.constant} and \ref{fig:homsap-dfe.ooa} we show a comparison of estimates
-    from three software packages for inferring the DFE from genetic data:
-    \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
+    \igcomment{I start with the OOA model and then cite the constant-size model as supplementary.}
+    We start by examining DFE inference from human genomes simulated under an out-of-Africa demographic model
+    with selection acting on exons (same data used for Figure \ref{fig:1pop-human-demography}; right).
+    In this setting, DFE is inferred separately for each of the three simulated populations
+    using three different DFE inference methods: \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
     and \grapes \citep{galtier2016adaptive}.
     In each case we note that the DFE is estimated from segregating sites only,
-    without the use of substitution data, as a current limitation of \stdpopsim
-    is that we are not able to simulate an outgroup sequence to estimate
-    divergence. 
-    Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
-    while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
-    In a constant size population, all methods perform reasonably well, although \dadicli seems to
-    perform slightly worse than the other methods examined, at least for the small number of replicates we simulated here.
-    Notably only \polydfe overlapped the true mean selection coefficient, whereas 
+    without the use of substitution data, as \stdpopsim's models currently do not
+    contain outgroup populations or species that can be used to estimate divergence.
+    %IG: no need to map figures. Just describe results and cite figs
+    %Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
+    %while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
+    %In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
+    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
+    The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
+    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters.
+    Population history appears to influence different methods in slightly different ways.
+    For instance, \dadicli and \grapes produce more accurate estimates from the YRI genomes 
+    (relative to the other two populations), whereas \polydfe
+    produces the most accurate estimate from the CHB genomes.
+    To factor out the influence of demographic changes, we also analyzed genomes generated using similar
+    simulations, but with a simpler demographic model with no population size changes (Figure \ref{fig:homsap-dfe.constant}).
+    In this simpler setting, \polydfe overlapped the true mean selection coefficient and shape parameter, whereas 
     the other methods do not. 
-    For the shape parameter in constant size simulations \grapes and \polydfe are roughly
-    equivalent in their performance whereas \dadicli seems to yield overestimates.
-
-    In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
-    In particular, variation in accuracy of the parameters estimated in each
-    sample population depends on the methods used, and each method seems to 
-    have its own biases with respect to the way in which population history 
-    affects parameter estimates. For instance, for \dadicli we see that estimates of 
-    the shape parameter are most accurate in the YRI sample, while for \polydfe
-    estimates are least accurate in the YRI sample. 
-    With respect to the strength of selection, as in constant size populations,
-    all methods are performing reasonably well, although \polydfe seems to
-    lead in terms of accuracy.
+    \grapes exhibits similar biases here as in the simulations with more complex demography,
+    but with milder under-estimation of the shape parameter, but \dadicli shows an inverse pattern
+    with overestimated shape parameter and slightly underestimated mean. 
 \kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things.
@@
 I would also do this for the demogrpahic models as well as the DFE models}
+\igcomment{I suspect that this way of comparing inferred distributions to simulated ones might be a bit misleading. Wouldn't it be better to use something like what we do in fig 5C?
+I predict that if we do this, we'll get a reasonable match for \dadicli}
 
@@
         \label{fig:vaquita-dfe}
-        A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
-        Simulations were performed using a model of vaquita porpoise demography \citep{robinson2022critically} with a gamma-distributed DFE
-        acting on exons parameterized by a mean selection coefficient, a shape parameter, and a relationship between selection and dominance. Estimates of the 
-        parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
-        This vaquita model has a single extant population, and parameter estimates from each
-        of the three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
-        and 3) \dadicli \citep{Huang2023,kim2017inference}.
-        C) Shown is the simulated distribution of $2 h s$, where $h$ is the dominance coefficient, along with the simulated distribution of $s$ and the median inferred DFEs.
-The distribution of $2 h s$ is multimodal because the simulated relationship between $h$ and $s$ is not continuous.
+        Inferred versus simulated DFEs from simulated vaquita genomes.
+        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
+        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$). 
+        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; panel A) and shape parameter (panel B).
+        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
+        \grapes, \polydfe, and \dadicli.
+        Panel C shows the simulated distribution of $2 h s$,
+        along with the simulated distribution of $s$ and the median inferred DFEs. \igcomment{What is median inferred DFE?}
+        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
         }
@@
 
-    In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
-    from the same three DFE-inference packages on simulations of the
-    vaquita porpoise genome.
-Importantly, these simulations use the vaquita demogrpahic model shown in Fig.
-\ref{fig:1pop-vaquita-demography} and a DFE that includes recessive deleterious mutations, as suggested in \cite{robinson2022critically}.
-All methods
-    perform uniformly worse that in the human genome simulations, with consistent underestimation of the mean selection coefficient    
-    and overestimation of the shape parameter.
-This is not surprising as a mis-specified model is being fit to the data.
-All DFE inference methods used here assume an addive model (no dominance, $h = 1/2$) while the simulated data from the vaquita  was generated from a DFE in which more deleterious mutations are more recessive ($h < 1/2$)
-
-    Reassuringly, for all three software packages, the inferred mean and shape of the distribution of selection coefficients 
-    is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient
-    and the selection coefficient (Fig.~\ref{fig:vaquita-dfe}C).
+    %IG: describe experiment and cite figure later
+    %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
+    Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome,
+    as generated for the analysis shown in Fig. \ref{fig:1pop-vaquita-demography} (right).
+    Recall that these simulations apply a gamma-dsitributed selection coefficient to mutations in exons
+    and a relationship between the selection and dominance coefficients.
+    In this setting, all methods perform uniformly worse than in the human genome simulations,
+    with consistent underestimation of the mean selection coefficient    
+    and overestimation of the shape parameter (Figure \ref{fig:vaquita-dfe}A-B). 
+    Similar patterns are observed when re-running the simulations with a simpler demographic model
+    without population size changes (Figure \ref{fig:vaquita-dfe.constant}).
+    We suspect that this is due to the fact that all DFE inference methods used here assume an additive model ($h = \frac{1}{2}$),
+    while mutations in the simulated genomes were more recessive ($h < \frac{1}{2}$), and their dominance decreased with the strength of selection (see above). 
+    To examine the influence of this model misspecification, we compared the inferred distributions of $s$ to the 
+    simulated distribution of $2hs$, which would have been equal to $s$ under an additive model (Figure \ref{fig:vaquita-dfe}C).
+    Reassuringly, all three methods infer distributions that appear to fit the main mode of the simulated distribution of $2hs$.
     This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
     where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
-   \kelcomment{I think we should move this next sentence to the Discussion and expand on it a bit....update on 1/10--I moved this to the Discussion and fleshed it out}
+   \kelcomment{I moved the last sentence from here to the Discussion and fleshed it out}
 
@@
 Indeed, the distribution of $hs$ is much more accurately estimated (Fig 4b).
-Overall, these results are consistent with prior work showing that it is challenging to separately infer the DFE of $s$ along with $h$ (cite PMID: 24830675; PMID: 39302992; PMID: 34951958 ) and that a large proportion (>5%) of recessive deleterious mutations can confound inferences of $s$ (PMID: 37074880).
+Overall, these results are consistent with prior work showing that it is challenging to separately infer the DFE of $s$ along with $h$ (cite PMID: 24830675; PMID: 39302992; PMID: 34951958 ) and that a large proportion (>5\%) of recessive deleterious mutations can confound inferences of $s$ (PMID: 37074880).
 Given the plethora of evidence for recessive deleterious mutations (cite PMID: 4630587, PMID: 39114967 , PMID: 30013096 , PMID: 21098719 ), these results suggest the need for care in DFE inference.
-Kyriazis & Lohmueller (PMID: 39302992 ) inferred the DFE of $s$ for nonsynonymous mutations in humans when considering different values of $h$.
+Kyriazis and Lohmueller (PMID: 39302992 ) inferred the DFE of $s$ for nonsynonymous mutations in humans when considering different values of $h$.
 Intriguingly, additive models show a satisfactory fit to the SFS of nonsynonymous variants in humans (Kyriazis et al) and more complex models of dominance do not substantially improve the fit.
@@
     under constant population size  where all sites along the simulated chromosome can experience selected and neutral mutation
-    as opposed to exonic regions only.
+    as opposed to exonic regions only.\igcomment{This figure is not cited anywhere. Do we really need it?}
     }

### 2025-02-13 analysis2_manuscript.tex

@@
     to benchmark and compare the performance of different methods for
-    inferring demographic history, the targets of recent positive selection,
-    and the distribution of fitness effects from genetic data.
+    inferring demographic history, the distribution of fitness effects,
+    and the targets of recent positive selection, from genetic data.
 
@@
 
-\section*{Example Applications with Selection}
-    \label{applications}
-    In this section, we present a series of examples of how the new models of selection
-    in \stdpopsim can be used to benchmark different
-    methods for population genetic inference.
-    We focus on three main areas: demographic inference in the 
-    single and multi-population settings, inference of the DFE,
-    and the detection of selective sweeps, particularly in the context of realistic recombination maps and background selection.
+%IG: commented out this intro paragraph. Better to dive right in.
+%\section*{Example Applications with Selection}
+%    \label{applications}
+%    In this section, we present a series of examples of how the new models of selection
+%    in \stdpopsim can be used to benchmark different
+%    methods for population genetic inference.
+%    We focus on three main areas: demographic inference in the 
+%    single and multi-population settings, inference of the DFE,
+%    and the detection of selective sweeps, particularly in the context of realistic recombination maps and background selection.
 
@@
     the effective population size over time, $N_e(t)$, from genetic data. This can be done
-    using a variety of methods, including the pairwise sequentially Markovian coalescent
-    (PSMC) \citep{li2011inference}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
+    using a variety of methods, including the sequentially Markovian coalescent
+    \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
     \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
@@
     as well as through identity by descent information \citep{santiago2020recent}.
-    While these methods perform well under neutral evolution, because of its effects on
-    patterns of genetic variation, natural selection on linked sites can bias estimates of $N_e(t)$
-    away from true census population sizes simply because selection increases the 
+    Since all of these methods assume neutral evolution, they are typically applied to genomic
+    segments that are not expected to be directly affected by selection (e.g., by masking exons
+    and conserved elements).
+    However, natural selection acting on linked sites (that are masked out of the analysis)
+    can bias estimates of $N_e(t)$
+    away from true census population sizes by increasing the 
     rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
-
-    \igcomment{Questions about methods here:
-    (1) Which DFE model was used here? The previous version of the text cited Huber, but the production config says Kim;
-    (2) Was DFE applied to all exonic sites or nonsynonymous? (I think that all exonic sites, but if non-syn, we need to clarify);
-    (3) We mention 3 reps, but this is not used in the results (see also comment in next para)
-    }
-    Using \stdpopsim, we can easily examine the effect of selection on the effective
-    population sizes inferred by different methods.
+    Using \stdpopsim, we can easily examine the effect of selection on linked sites,
+    which is commonly referred to as \emph{background selection}, 
+    on different methods for inferring $N_e(t)$.
+
     For this purpose, we simulated human genomes with and without natural selection.
     In both scenarios, we ran three replicate simulations
-    under the out of Africa (OOA) demographic model
-    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second}.
-    Simulations with selection were implemented by modeling purifying selection on exons,
-    defined from the HAVANA group release 104 for the human genome,
+    under the out of Africa (OOA) demographic model of
+    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
+    Simulations with selection were implemented by modeling purifying selection on exons
+    % IG: moved details to methods
+    %, defined from the HAVANA group release 104 for the human genome,
     using a DFE inferred by \cite{kim2017inference}.
-    Each of the six simulated datasets contained 300 genomes, 100 for each geographic group: CEU, CHB, and YRI.
-    All simulations were implemented using a relatively short python script (as illustrated in Figure 1B).
-    % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
-    % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
-    % IG: added a note about this above.
     Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
@@
 
-    \igcomment{Questions about inference and summary:
-    (1) In simulation with selection, do we infer Ne from all sites (exons and non-exons)?
-        This is what I understand from the code, but if not, we should clarify.
-    (2) The effect of selection on inference is difficult to glean from Fig 2.
-        The main thing I see there is that the influence is very small.
-        Maybe we should have a separate plot showing each method in a diff panel with inferences from all 6 datasets?
-        This might allow us to see differences between the two scenarios (with/without selection).  
-    }
-    %IG: I commented this out because it repeats some of the info above and we now start by directly
-    %discussing the results rather than describing the figure
-    %In Figure \ref{fig:1pop-human-demography}, we show six panels, each of which shows the estimated $N_e(t)$
-    %from data simulated under the five-population model of human out-of-Africa demography
-    %from Ragsdale \textit{et al.} (2019)
-    %with samples taken from three modern geographic groups (CEU, CHB, and YRI from top to bottom).
-    %The left column shows the results of $N_e(t)$ estimates made from simulations without selection, 
-    %while the right column shows the results
-    %from simulations with deleterious mutations occurring in exons.
-     % IG: commented this out, because these details are better left for the fig caption
-    %In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
-    %in various colors for the four methods. Dotted lines on the top row are to visually reinforce that inference is done
-    %from a model without selection.
-    We see that all methods generally perform reasonably well in the absence of selection (Figure \ref{fig:1pop-human-demography}; left).
-    The only exception to this is that GONE seems to overestimate $N_e(t)$ from the CHB genomes,
-    %IG: I commented this out because this seems like a very minor blip
-    %and to a lesser extent in the CEU sample.
-    presumably due to patterns of linkage disequilibrium that result from the combination
-    of bottleneck and growth in the populations ancestral to these samples.
-    %
-    \igcomment{I don't really see evidence for the observations listed below in Fig 2. See my suggestion above to enhance the comparison.}
-    In the presence of selection, different methods behave differently. 
-    For instance stairwayplot, which uses the SFS to infer $N_e(t)$, 
-    is reasonably close to the true $N_e(t)$ both with and without selection,
-    although it seems to infer significant population size changes that are not part of the true history.
-    Likewise \smcpp seems to be robust to the presence of selection,
-    but the timing of population size changes is slightly inaccurate. 
-    \msmc seems to be the most affected by the presence of selection, 
-    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$ particularly in the very recent past,
-    although estimates from \msmc have quite a bit of added noise even in the absence of selection.
-
-    % TODO: legend here needs to be changed. The dotted line indicates no selection, the solid selection
-
-    \begin{figure}[t]
+    \begin{figure}[b!]
         \centering
-        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
+        %IG: NEED TO LINK NEW FIG HERE
+        %\includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
+        \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
         \caption{
-        \label{fig:1pop-human-demography}
-        Inference of $N_e(t)$ from a human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
-        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ obtained from genomes simulated
-        without selection, while the right column shows estimates of $N_e(t)$ from genomes simulated with a gamma-distributed   
-        DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
+        \label{fig:human-demography}
+        Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
+        with and without purifying selection on exons. 
+        Rows correspond to the three extant populations in the simulation: CHB, CEU, and YRI.
+        Columns correspond to the four inference methods:    
         \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
-        \igcomment{This is a nice figure but it's difficult to see if there is any subtle influence of selection from it.
-        We should think of one multi-panel main figure to summarize different aspects of the $N_e(t)$ analysis, and push the rest to the supplement.
-        Specifically for this figure, I'm not sure we need to use different line types for left and right since the plots are separated.}  
+        Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
+        and the three datasets simulated with the background effects of purifying selection on exons (orange),
+        alongside the true values of $N_e$ used in simulation (black).
         }
     \end{figure}
-    %DRS-fig: reminder to update the legend to just say ``exons'' instead of ensembl_havana_whatever...
-    %IG: this seems to have been taken care of
-
-    We note that while selection in these simulations directly acts on a small fraction of sites in the analyzed genomes,
-    it has a widespread influence on genetic diversity through background selection \citep[e.g.][]{charlesworth1993effect}.
-    Our results here suggest that the combined effect of direct selection and background selection does not considerably
-    influence infrerence of $N_e(t)$ for human populations under reasonable assumptions about their evolutionary history.
+
+    \igcomment{Rewrote this paragraph based on new figs.}
     %
+    All four inference methods appear to infer similar $N_e(t)$ trajectories from
+    genomes simulated without selection (Figure \ref{fig:human-demography}; blue) and
+    genomes simulated with selection (Figure \ref{fig:human-demography}; orange).
+    Thus, background selection does not appear to notably bias the estimates
+    produced by these methods.
+    As noted in previous studies that compared demography inference methods
+    on neutral simulations \citep{adrion2020community},
+    we see that each method has its strengths and weaknesses.
+    Methods based on the sequenial Markovian coalescent (\smcpp and \msmc)
+    appear to overal produce the most accurate trajectories of $N_e(t)$
+    in this setting, although they tend to under-estimate the population
+    bottleneck associated with the population split.
+    On the other hand, \stairway, which utilizes the site frequency spectrum,
+    produces more noisy estimates,
+    but it does not appear to under-estimate the population bottleneck.
+    The inference of \gone, which utilizes identity by descent,
+    is targeted toward demographic changes in more recent time frames ($\sim 200$ generations),
+    explaining its overall noisy estimates.
+
+%IG: commented old version of figure here
+%    \begin{figure}[t]
+%        \centering
+%        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
+%        \caption{
+%        \label{fig:1pop-human-demography}
+%        Inference of $N_e(t)$ from a human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
+%        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ obtained from genomes simulated
+%        without selection, while the right column shows estimates of $N_e(t)$ from genomes simulated with a gamma-distributed   
+%        DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
+%        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+%        }
+%    \end{figure}
+
     To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
@@
     We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
-    and since no genetic map is currently available, we assumed a constant recombination rate across the genome.
+    and a constant recombination rate across the genome (see \textbf{Methods}).
     In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
-    annotated from the vaquita genome assembly \igcomment{Maybe cite Morin et al. 2010 for assembly and annotations?}.
-    \igcomment{I added this about the h-s relationship. Please check!}
-    In this DFE model, we implemented a relationship between the dominance coefficient ($h$) and the selection coefficient ($s$)
-    as follows: $s<-0.1 \Rightarrow h=0$; $s\in [-0.1,-0.01) \Rightarrow h=0.01$; $s\in [-0.01,-0.001) \Rightarrow h=0.1$;
-    and $s\geq -0.001 \Rightarrow h=0.4$.
-    This approach is similar to what was used in the simulations cunducted by \textcite{robinson2022critically} and is easily
-    implemented in \stdpopsim's \texttt{DFE} class.
-    Using these simulations, we see that all methods infer $N_e(t)$ reasonably well with or without selection
-    (Figure \ref{fig:1pop-vaquita-demography}), although they do not accurately identify the boundary time
-    between the two epochs.
+    annotated from the vaquita genome assembly.
+    This DFE model implements a relationship between selection and dominance 
+    in which very deleterious mutations (with selection coefficient $s<-0.1$) are fully recessive
+    (dominance coefficient $h=0$) and weakly deleterious mutations ($s\geq -0.001$) are nearly
+    additive ($h=0.4$; see \textbf{Methods}).
+    Here, we see that the site-frequency-based method (\stairway) produces more accurate
+    inference than the two methods based on the sequenctioal Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
+    Reagrdless of these difference, we see that background selection does not
+    considerably influence the inference of any of the four methods also for this species,
+    which has a very different evolutionary history than humans.
+
     % MFR: Should we say somewhere that people use to expect from BGS?
 That is, that it can be seen as a reduction in local rates of coalescences and thus that we would expect these methods to be somewhat downwardly biased in the presence of BGS?
-    % IG: we can address this in the discussion.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final}
-    \caption{
-    \label{fig:1pop-vaquita-demography}
-    Performance of methods to infer $N_e(t)$ from simulations of the vaquita porpoise genome under a single population
-    model of declining population size \citep{robinson2022critically} with and without background selection on exons. 
-    The left panel shows estimates of $N_e(t)$ from simulations
-    without selection, while the right panel shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
-    DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
-    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
-    \igcomment{see my comment on fig 2. We should decide what part of this stays in the main figure and what ends up in the supp.}  
-    }
-\end{figure}
+    % IG: we can address this in the discussion, or in the revised text above.
+
+%\begin{figure}[t]
+%    \centering
+%    \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final}
+%    \caption{
+%    \label{fig:1pop-vaquita-demography}
+%    Performance of methods to infer $N_e(t)$ from simulations of the vaquita porpoise genome under a single population
+%    model of declining population size \citep{robinson2022critically} with and without background selection on exons. 
+%    The left panel shows estimates of $N_e(t)$ from simulations
+%    without selection, while the right panel shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
+%    DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
+%    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+%
+%    }
+%\end{figure}
 
@@
     For comparison, here we have normalized all inference results to the convention of SLiM.
-    \igcomment{Can we summarize this in a table or in a short methods section?
-    For each method, we could specify the scaling we apply to compare its inferred distribution to the one used in simulation.}
+    \igcomment{We should add a detailed description in the Methods of what we do to scale the output of each DFE-inference method.}
 
@@
 
-    \begin{figure}[htbp]
+    \begin{figure}[b!]
         \centering
         \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_OutOfAfricaArchaicAdmixture_5R19_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
+        \igcomment{NEED TO UPDATE FIG WITH NEW VERSION}
         \caption{Inferred versus simulated DFEs from simulated human genomes.
         Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
-        acting on exons.
-        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; left) and shape parameter (right).
-        The DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
-        by three different methods: \grapes, \polydfe , and \dadicli. The dashed horizontal lines represents the simulated values.
-        \igcomment{Does the boxplot represent the three simulation reps? Is each method run several times?}
+        acting on exons (see \textbf{Methods}).
+        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
+        by three different methods: \grapes, \polydfe , and \dadicli.
+        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
+        shown for each DFE inferred from all three simulated datasets,
+        with median values marked by horizontal bars
+        and simulated values represented by dashed horizontal lines.
+        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets;
+        white bars represent the distribution used in the simulation.
         }
@@
     We start by examining DFE inference from human genomes simulated under an out-of-Africa demographic model
-    with selection acting on exons (same data used for Figure \ref{fig:1pop-human-demography}; right).
+    with selection acting on exons (see \textbf{Methods}).
     In this setting, DFE is inferred separately for each of the three simulated populations
@@
     The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
-    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters.
+    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters (Figure \ref{fig:homsap-dfe.ooa}A-B).
+    As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$; Figure \ref{fig:homsap-dfe.ooa}C),
+    while \grapes under-estimates the fraction of mutations with $s<0.01$ and considerably over-estimates the fraction of
+    mutations under weak selection ($s\geq-0.001$).
     Population history appears to influence different methods in slightly different ways.
@@
 I would also do this for the demogrpahic models as well as the DFE models}
-\igcomment{I suspect that this way of comparing inferred distributions to simulated ones might be a bit misleading. Wouldn't it be better to use something like what we do in fig 5C?
-I predict that if we do this, we'll get a reasonable match for \dadicli}
+\igcomment{I think that this is better conveyed now in the added histograms for $s$ in panel C}
 
@@
     %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
-    Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome,
-    as generated for the analysis shown in Fig. \ref{fig:1pop-vaquita-demography} (right).
+    Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome.
     Recall that these simulations apply a gamma-dsitributed selection coefficient to mutations in exons
-    and a relationship between the selection and dominance coefficients.
+    and a relationship between the selection and dominance coefficients (see \textbf{Methods}).
     In this setting, all methods perform uniformly worse than in the human genome simulations,
@@
     We suspect that this is due to the fact that all DFE inference methods used here assume an additive model ($h = \frac{1}{2}$),
-    while mutations in the simulated genomes were more recessive ($h < \frac{1}{2}$), and their dominance decreased with the strength of selection (see above). 
+    while mutations in the simulated genomes were more recessive ($h < \frac{1}{2}$), and their dominance decreased with the strength of selection. 
     To examine the influence of this model misspecification, we compared the inferred distributions of $s$ to the 
@@
     In this study we performed simulations of the human and vaquita porpoise genomes
-    using the \stdpopsim{} framework. As stated above, simulations did or did not include a model 
-    of (negative) selection acting on exonic regions, matched to the human and vaquita porpoise genomes.
-    Three replicates of whole-genome simulations were performed for each species, with 
-    a sample size of 100 diploid individuals. Demographic models are as stated above, each
-    using the exon annotations from the human and vaquita porpoise genomes respectively.
-    For simulations with selection, we sampled 10 diploid individuals from each of two populations (YRI and CEU). 
-    To decrease runtimes, we used a SLiM scaling factor of 2 with a burn-in of 2N generations.
-\kelcomment{I'm confused about the 100 diploid individuals samples vs.
-the 10 diploid individuals sampled.
-Did some analyses use 100 and others use 10?
-That's fine, but maybe we can specify that more clearly...}
-
-    \subsection*{Demographic inference}
-    We assessed the accuracy of demographic inferences from
-    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
-    Each of these methods was run using the same set of simulations in replicate.
-    In each case, the non-recombining portion
-    of the genome was masked to remove any bias due to recombination rate % DRS: might need more detail here
-    variation. Comparisons between inferences made with and without selection were 
-    of particular interest. For \msmc, we used a random sample of $N=6$ genomes with 20
-    iterations of the EM algorithm. For \gone, we set max\_snps to 500000,
-    the number of generations to 2000, and the number of bins to 400, all representing default settings.
-
-    \subsection*{Distribution of fitness effects inference}
-    We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
-    \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
-    Each software package was run using the same set of simulations.
-    For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
-    A key distinction between our analyses and those done in practice
-    is that we have complete knowledge of which sites are under selection in the simulation,
-    thus our inferences is a best case scenario. Empirical inference of the DFE is a much
-    harder problem as some mutations assumed to be free of selection
-    may in reality be under selection. As above, all methods were run using default settings.
-    \kelcomment{I think we need a lot more detail here...how big was the coding sequence length?
-How was demogrpahic inference done for the different models via the different programs?
-How was the unscaling of E[s] done?, etc}
+    using the \stdpopsim{} framework.
+    All simulations were executed using the SLiM engine with a burn-in of $2N$ generations
+    (where $N$ is the initial population size), and a scaling factor of 2 (to reduce running time).
+    Each simulation was implemented using a relatively short python script, similar to the one shown in Figure 1B
+    (see \textbf{Data availablitiy} below).
+    \igcomment{We should add a link to the analysis2 repo in data/code availablity section}
+    % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
+    % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
+    % IG: added a note about this above.
+
+    \subsubsection*{Simulations of complete human genomes}
+    Simulation of human genomes used a genetic map from the HapMap Project
+    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
+    the out of Africa (OOA) demographic model with archaic admixture from
+    \cite{ragsdale2019models} (\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}).
+    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
+    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
+    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
+    Each simulation generate 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
+    with each genome consisting of 22 autosomes.
+    Using these settings, we generated three replicate datasets without selection
+    and three replicate datasets with selection.
+    Three additional datasets, used in Figure \ref{fig:homsap-dfe.constant},
+    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $10,000$.
+    Each of these simulations generated 100 diploid genomes.
+
+    \subsubsection*{Simulations of complete vaquita porpoise genomes}
+    Simulation of vaquita porpoise genomes used 
+    the two-epoch demographic model from \cite{robinson2022critically} (\stdpopsim label \texttt{Vaquita2Epoch\_1R22}).
+    Since no genetic map is currently available for this species, we assumed a constant recombination rate of  $10^{-8}$
+    across all chromosomes \citep{morin2021}.
+    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
+    \cite{robinson2022critically} (\stdpopsim label \texttt{Gamma\_R22}) to exons annotated from the vaquita genome assembly
+    (\cite{morin2021}; \stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}).
+    This DFE model implements a relationship between selection and dominance as follows:
+    very deleterious mutations (with $s<-0.1$) were set to be recessive ($h=0$),
+    mutations with $s\in [-0.1,-0.01)$ were set to have a dominance coefficient of $h=0.01$,
+    mutations with $s\in [-0.01,-0.001)$ were set to have a dominance coefficient of $h=0.1$,
+    and very mild mutations (with $s\in [-0.001,0)$) were set to have a nearly additive effect ($h=0.4$).
+    This approach is similar to what was used in the simulations cunducted by \textcite{robinson2022critically}
+    and is easily implemented in \stdpopsim's \texttt{DFE} class.
+    Each simulation generate 100 diploid genomes,
+    with each genome consisting of 21 autosomes.
+    Using these settings, we generated three replicate datasets without selection
+    and three replicate datasets with selection.
+    Three additional datasets, used in Figure \ref{fig:vaquita-dfe.constant},
+    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $3,500$.
+    \igcomment{If we end up using Fig. S3, then we should add a line describing it's simulation here.}
+
+    \subsubsection*{Simulations with selective sweeps}
+
+    To test methods for detecting selective sweeps, we simulated 100 evenly distributed 5Mb
+    segments of human chromosome 1.
+    Each segment was simulated in each of the following three settings:
+    (1) neutral, (2) background selection (BGS), and (3) hard sweep with BGS.
+    In all three settings, we applied a genetic map from the HapMap Project
+    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
+    the three-population out of Africa (OOA) demographic model from
+    \cite{gutenkunst2009inferring} (\stdpopsim label \texttt{OutOfAfrica\_3G09}).
+    In the two settings with BGS, we applied the DFE model with gamma-distributed selection coefficients inferred by
+    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
+    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
+    Hard sweeps were modeled by generating a single beneficial mutation 
+    with selection coefficient of $s = 0.03$ in the center of the 5Mb genomic segment
+    and discarding simulations in which the beneficial mutation did not reach (or exceed) a terminal frequency of $95\%$.
+    For each of the 100 genomic segments of chromosome 1, we generated 1,000 replicate datasets under each of the three settings.
+    Each simulation generated 10 diploid genomes for each of the two sampled populations: YRI and CEU.
+    The genomes simulated with a hard sweep and BGS were used for sweep detection,
+    and genomes simulated under the other two settings were used to calibrate the thresholds of the test statistics
+    used by the three sweep-detection methods (see below).
+
+    \subsection*{Inference of $N_e(t)$}
+    %We assessed the accuracy of demographic inferences from
+    %\msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+    Inference of $N_e(t)$ was conducted on simulated datasets after masking non-recombining portion
+    of the genome (to avoid bias caused by extreme recombination rate variation)
+    and exons (to avoid the direct effect of natural selection).
+    Thus, inference from datasets simulated with selection was not influenced by direct selection
+    acting on exons, but it was influenced by selection acting on linked sites (background selection).    
+    %IG: I commented because it doesn't seem relevant for methods
+    %Comparisons between inferences made with and without selection were 
+    %of particular interest.
+    For \msmc \citep{Schiffels2020}, we used a random sample of $N=6$ simulated genomes with 20
+    iterations of the EM algorithm.
+    For \gone \citep{santiago2020recent}, we set \texttt{max\_snps} to 500,000,
+    the number of generations to 2,000, and the number of bins to 400, all representing default settings.
+    For \stairway \citep{liu2020stairway} and \smcpp \citep{terhorst2017robust} we used the default settings.
+    \igcomment{Nothing more to say about \stairway plot and \smcpp?}
+
+    \subsection*{DFE inference}
+    \igcomment{I commented all the text that was here because it wasn't relevant to the methods. We might want to add it to the discussion.
+    Kirk also noted that a lot of methods-level details are missing:
+    how big was the coding sequence length?
+    How was demographic inference done for the different models via the different programs?
+    How was the unscaling of E[s] done?, etc.
+    Can someone write this up?}
+    % We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
+    % \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
+    % Each software package was run using the same set of simulations.
+    % For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
+    % A key distinction between our analyses and those done in practice
+    % is that we have complete knowledge of which sites are under selection in the simulation,
+    % thus our inferences is a best case scenario. Empirical inference of the DFE is a much
+    % harder problem as some mutations assumed to be free of selection
+    % may in reality be under selection. As above, all methods were run using default settings.
 
     \subsection*{Selective sweeps detection}
-    We ran simulations of human chromosome 1 using the three population out-of-Africa model
-    (identified as OutOfAfrica\_3G09 in the stdpopsim library; \cite{gutenkunst2009inferring})
-    and the genetic map estimated in \citet{international2007second} (HapMapII\_GRCh38).
-    For computational efficiency, we simulated a 5Mb region flanking 100 evenly distributed points along the chromosome 1.
-    We generated 1000 replicates of each of three classes of simulations:
-    (i) neutral (with the specified demography and genetic map),
-    (ii) background selection, in which we applied the DFE estimated in \citet{kim2017inference} (Gamma\_K17) to exon annotations (taken from Ensembl; ensembl\_havana\_104\_exons),
-    and (iii) sweep with background selection, in which we simulated the hard sweep as described with the addition of constraint in exonic regions (same as for \emph{ii}).
-    In the sweep simulations, we placed a single beneficial mutation in the center of the simulated region with a selection coefficient of $s = 0.03$, and only sampled if the beneficial mutation reached terminal frequency of 0.95 or greater.
-
-    We then benchmarked the performance of three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
-    2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff for nucleotide diversity ($\pi$) to find regions of reduced diversity.
-    \diploshic was trained on data simulated using \texttt{discoal} \citep{kern2016discoal}.
-    Nucleotide diversity was computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently} over 10 equally sized and non-overlapping subwindows.
-    The composite likelihood ratio (output by \sweepfinder) was computed at 21 uniformly distributed points along the simulated region.
-    To train the \diploshic classifier, 
-    For training simulations, we simulated 20 haploid genomic regions of 1.21Mbp
-    with a constant population size of 10,000 individuals, a mutation rate of 2e-8, and a recombination rate of 2e-8.
+
+    In this analysis, we compared three different criteria for sweep detection:
+    (1) reduced nucleotide diversity ($\pi$),
+    (2) the composite likelihood ratio (CLR) computed by \sweepfinder \citep{degiorgio2016sweepfinder2},
+    and (3) a probability for sweeps computed using the supervised learning approach of \diploshic \citep{kern2018diplos}.
+    Nucleotide diversity was computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently}
+    over 10 equally sized and non-overlapping subwindows, and 
+    the CLR was computed at 21 uniformly distributed positions along the simulated region.
+    To train the \diploshic classifier, we used \texttt{discoal} \citep{kern2016discoal} to simulate 20 haploid sequences of length 1.21Mbp
+    with a constant population size of $N_e=10,000$, a mutation rate of $2\times 10^{-8}$, and a recombination rate of $2\times 10^{-8}$.
     For \texttt{discoal} simulations that included selective sweeps,
-    selection coefficients were drawn from a uniform distribution between 0.001 and 0.10.
-    The time of the sweep was chosen from a uniform distribution between 0 and 0.01 Ne generations in the past.
-    For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between 1e-4 and 0.1.
-    All \texttt{discoal} training simulations were done under a model of equilibrium population size, 
-    so we expected it to perform suboptimally given the variation in population size in the out-of-Africa model.
-
-    To call sweeps, we empirically determined the distribution of each statistic (nucleotide diversity, \sweepfinder's CLR, and \diploshic's probability of sweep) under a null model and defined the top $5\%$ quantile as a cutoff to call a sweep (though regions with extremely low recombination rate were excluded from the computation).
-    We considered two null models: neutral and background selection (as explained before).
-    Power (or true positive rate) to detect a sweep is the proportion of simulations with a sweep that were correctly classified as a sweep for each method and at each of the 100 locations along chromosome 1.
+    selection coefficients were drawn from a uniform distribution between 0.001 and 0.10,
+    and the time of the sweep was chosen from a uniform distribution between 0 and $0.01N_e$ generations in the past.
+    For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between $0.0001$ and $0.1$.
+    We note that since the training set of \diploshic was generated under a simple model with
+    constant mutation and recombination rates and constant effective population size,
+    we expect it to perform suboptimally on the simulated segments of human chromosome 1,
+    which were generated using mroe relaistic simulations (see \textbf{Simulations with selective sweeps} above).
+
+    Each of the three sweep-detection methods was associated with a threshold for its test statistic ($\pi$, \sweepfinder's CLR, or \diploshic's probability)
+    that was determined using an empirical distribution representing a null model without sweeps.
+    We considered two null models: a neutral evolution model (neutral),
+    and a model with background selection from deleterious mutations in exons (BGS).
+    Each null model was represented by a different set of $100,000$ simulations generated according to the appropriate model
+    (see \textbf{Simulations with selective sweeps} above).
+    Thus, for each test statistic we determined two thresholds corresponding to these two different null models (neutral and BGS)
+    using the extreme $5\%$ quantile in each empirical distribution (and excluding regions with extremely low recombination rate).
+    \igcomment{Can we say what the rec rate cutoff was and how many regions were excluded here?
+    Shouldn't we have excluded these regions from the simulations altogether?}.
+    Then, power (or true positive rate) to detect a sweep by a given method was determined for
+    100 evenly-spaced positions on chromosome 1 as follows.
+    The sweep-detection method was applied to the 1,000 
+    simulated datasets generated with a hard sweep (and BGS) for that position,
+    and the power was set to be the proportion of simulations (out of 1,000)
+    in which the test statistic exceeded the threshold determined by each of the two null models (neutral and BGS).
 
@@
 
+\begin{figure}[b!]
+    \centering
+    %IG: NEED TO LINK NEW FIG HERE
+    %\includegraphics[width=\textwidth]
+    \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
+    \caption{
+    \label{fig:vaquita-demography}
+    Inference of $N_e(t)$ from vaquita porpoise genomes simulated under a single population 
+    model of declining population size \citep{robinson2022critically}
+    with and without purifying selection on exons. 
+    Inference is conducted using four methods:    
+    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+    Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
+    and the three datasets simulated with the background effects of purifying selection on exons (orange),
+    alongside the true values of $N_e$ used in simulation (black).
+}
+\end{figure}
+
 % constant size DFE figure
@@
     \caption{
+    \igcomment{Update fig and caption after we finalize Fig. 3}
     Performance of methods to infer the distribution of fitness effects (DFE) from genetic data.

### 2025-02-18 analysis2_manuscript.tex
Ilan's edits and comments (not including sweeps and the dicsussion) (#45)
* Ilan's edits and comments (not including sweeps and the dicsussion)

* More edits to manuscript: Methods, Ne(t), DFE inference

* more edits, especially in the results section on selective sweeps

---------

Co-authored-by: Ilan Gronau <gronau@gmail.com>

@@
 
-    %add PhoSin DFE figure here
-    \begin{figure}
-        \centering
-        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
-        \caption{
-        \label{fig:vaquita-dfe}
-        Inferred versus simulated DFEs from simulated vaquita genomes.
-        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
-        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$). 
-        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; panel A) and shape parameter (panel B).
-        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
-        \grapes, \polydfe, and \dadicli.
-        Panel C shows the simulated distribution of $2 h s$,
-        along with the simulated distribution of $s$ and the median inferred DFEs. \igcomment{What is median inferred DFE?}
-        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
-        }
-    \end{figure}
-
     %IG: describe experiment and cite figure later
@@
 
+    %add PhoSin DFE figure here
+    \begin{figure}[t!]
+        \centering
+        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
+        \caption{
+        \label{fig:vaquita-dfe}
+        Inferred versus simulated DFEs from simulated vaquita genomes.
+        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
+        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$). 
+        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; panel A) and shape parameter (panel B).
+        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
+        \grapes, \polydfe, and \dadicli.
+        Panel C shows the simulated distribution of $2 h s$,
+        along with the simulated distribution of $s$ and the median inferred DFEs. \igcomment{What is median inferred DFE?}
+        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
+        }
+    \end{figure}
+
 \section*{Detection of Selective Sweeps}
     \label{sweeps}
-    We next highlight how the new models of selection in \stdpopsim can be used to benchmark and compare the performance of different methods for detecting selective sweeps from genetic data.
-    Here, we performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring} with purifying selection acting on exons under the gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second}, but also include sweeps.
-    In particular, we introduced a beneficial mutation at a given location (that was varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
-    For each sweep location, we performed 1000 replicate simulations, sampling from them only if the selected allele reached a frequency of 0.95 or greater.
-    Rather than simulate the entire chromosome arm, replicate simulations were performed on 5cM regions of the chromosome at each of 100 different locations.
-    As recombination rate varies across the genome, the size of the region in base pairs simulated varied, however the recombination distance, which is critical for the linked effects of a sweep, was held constant.
+    We next highlight how \stdpopsim can be used to benchmark and compare
+    the performance of different methods for detecting selective sweeps from genetic data.
+    Because simulation of selective sweeps is very time consuming and evaluation of statistical power requires
+    many replicate simulations, we focused on human chromosome 1 and generated replicate simulations on 5cM segments
+    of the chromosome in 100 different locations. % (see \textbf{Methods}).
+    Selective sweeps were simulated by introducing a beneficial mutation in the center of the 5cM segment
+    with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$; see \textbf{Methods}).
+    Each segment was simulated 1000 times, while keeping only replicates in which the selected allele reached a frequency of 0.95 or greater.
+    These sweeps were modeled in the context of the three population out-of-Africa model from \cite{gutenkunst2009inferring}
+    and background selection from deleterious mutations in exons.
+    Sweep-detection methods typically base their inference on checking whether a test statistic
+    exceeds a certain threshold, which is determined according to some empirical null distribution.
+    In our simulations, we considered two different null distributions:
+    one based on neutral simulations, and another based on simulations with background selection
+    from deleterious mutations in exons (see \textbf{Methods}).
 
      %sweeep power figure
-    \begin{figure}
+    \begin{figure}[t!]
         \centering
@@
         Power to detect selective sweeps at 100 locations along human
-        chromosome 1 under a three population out-of-Africa model
-        \citep{gutenkunst2009inferring} when using null models with and without
-        purifying selection acting on exons under the gamma-distributed DFE
-        described above.  The HapMap II genetic map
-        \citep{international2007second} was used in each simulation.  Single
-        beneficial mutations were introduced at each location with a selection
-        coefficient of $s = 0.03$, and only sampled if they reached a terminal
-        frequency of 0.95 or greater.  Three methods for detecting sweeps were
-        used: 1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row), 2)
-        \diploshic \citep{kern2018diplos} (middle row), and 3) an empirical
-        cutoff of $\pi$ to find regions of reduced diversity (bottom row).
+        chromosome 1. Genomic segments were simulated with sweeps under a three population out-of-Africa model
+        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
+        %IG: seems like methods-level detail
+        %Single beneficial mutations were introduced at each location with a selection
+        % coefficient of $s = 0.03$, and only sampled if they reached a terminal
+        % frequency of 0.95 or greater.
+        Three methods for detecting sweeps were applied to simulated data:
+        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
+        2) \diploshic \citep{kern2018diplos} (middle row),
+        and 3) reduced diversity ($\pi$) (bottom row).
         Power (true positive rate) is shown for these methods for the CEU and YRI
-        samples (left and right respectively).  Under each method we show
-        heatmaps of exon density and local recombination rate along the
-        chromosome.  Power is shown as the proportion of replicates in which a
-        sweep was detected at each location.  To define
-        thresholds for calling a sweep, two different null models were used: a
-        neutral model (blue lines), and a model with purifying selection acting
-        on exons (red lines).
+        samples (left and right respectively).
+        The thresholds of the test statistics were set to control for
+        $5\%$ false positive rate under a neutral null model (blue)
+        and a null model with background selection from deleterious mutations in exons (red).
+        Also shown are heatmaps of exon density and local recombination rate along the
+        chromosome (bottom).
+        \igcomment{In fig, replace CLR with \sweepfinder; also in figs 6 S5 S6}
         }
@@
 
-    In Figure \ref{fig:chr1_power}, we show the power to detect selective sweeps at 100 locations along human chromosome 1 from
-    the simulations described above. We tested three methods for detecting sweeps: 1) \sweepfinder \citep{degiorgio2016sweepfinder2},
-    2) \diploshic \citep{kern2018diplos}, and 3) an empirical cutoff of $\pi$ to find regions of reduced diversity.
-    A few things are immediately apparent from this figure. First, the power to detect sweeps is generally higher in the YRI sample
-    than in the CEU sample, which is consistent with the fact that the YRI sample has a larger effective population size and thus both stronger selection and
-    more variation to detect a sweep. Further, the population size of YRI has been relatively constant over the last 100,000 years,
-    thus the out-of-Africa bottleneck has does not mute the signal of a sweep as much as in the CEU sample. This is consistent with
-    what is known about the effects of demography on the power to detect sweeps \citep[e.g.,][]{simonsen1995properties}.
-    Second, the power to detect sweeps is slightly lower when we use a null model with purifying selection acting on exons (consistent with \cite{schrider2020background}) but the effect is not
-    as strong as one might expect. 
-    \rngcomment{I'm confused by the terminology ``null model'' throughout this section.
-Are we always testing on simulated data that includes background selection, and ``null model'' refers to how the methods are tuned?
-Or are we comparing power on data simulated with and without background selection?} This is likely because the strength of selection acting on exons is relatively weak, and the
+    We evaluated the statistical power to detect selective sweeps for three methods:
+    1) \sweepfinder \citep{degiorgio2016sweepfinder2},
+    2) \diploshic \citep{kern2018diplos},
+    and 3) reduced diversity ($\pi$).
+    For each method, we determined a threshold for its test-statistic 
+    using each of the two null models separately (neutral and BGS) and controlling 
+    for $5\%$ false positive rate.
+    We then applied each method to data simulated with sweeps for two sampled populations (YRI and CEU)
+    in each of the 100 locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
+     First, we notice that power to detect sweeps is slightly lower when the test statistic threshold
+    is determined using simulated data with background selection (BGS null model).
+    This observation is consistent with what was previously shown by \cite{schrider2020background},
+    but the effect is relatively minor.
+    This is likely because the strength of selection acting on exons is relatively weak, and the
     sweep detection methods we have used are relatively robust to the effects of background selection. 
-    Finally and most importantly, the power to detect sweeps varies considerably among regions of the genome. 
-    Indeed variation in power along the chromosome is much greater than the variation in power between the CEU and YRI samples or between
-    models with or without purifying selection. This is likely due to the effects of recombination rate heterogeniety as well
-    as the density of functional elements along the chromosome.
-
-    Using this set of simulations, we can also compare the power of different methods to detect sweeps.
-    In general, the power of the methods varies with respect to the population history of the sample.
-    For instance \sweepfinder has marginally higher power to detect sweeps in the CEU sample than \diploshic,
-    likely as a results of the fact that we here trained \diploshic on simulations under a constant size population,
-    while the population size history of CEU is characterized by a strong recent bottleneck.
-    Conversely, \diploshic has higher power to detect sweeps in the YRI sample than \sweepfinder,
-    likely because that population size history more closely matches the population size history used to train \diploshic.
-    \sweepfinder power remains more constant along the chromosome, while the $\pi$ cutoff method varies wildly.
-
-    We further explored the performance of the methods by plotting their receiver operating characteristic (ROC) curves.
-    In Figure \ref{fig:roc-curves}, we show the ROC curves for the three methods for detecting sweeps.
-    False positive rates were established using neutral simulations that contained no sweeps
-    but were conditioned on the same population history as the CEU and YRI samples. \rngcomment{Did these simulations contain background selection?}
-    These suggest that the $\pi$ cutoff is a surprisingly effective method for detecting sweeps,
-    in these genomic regions at least, but that it lags behind \diploshic and \sweepfinder in terms of accuracy in the YRI sample.
-    %While this is not exhaustive benchmarking, we aim to give a sense to the reader
-    %of how \stdpopsim can be used to benchmark the performance of different methods easily and efficiently.
-    These ROC curves, however, hide the phenomenon of high variation in power (see Figure \ref{fig:chr1_power}) to detect sweeps along a realistic looking chromosome.
-    That is, although in aggregate the $\pi$ cutoff may seem effective, its power is highly heterogeneous across the chromosome.
-
-    \begin{figure}
+    Second, the power to detect sweeps is generally higher in the YRI sample
+    than in the CEU sample, which is consistent with the fact that YRI has a larger ancestral effective population size and
+    thus both stronger selection and more variation to detect a sweep \citep[e.g.,][]{simonsen1995properties}.
+    Further, the CEU population experienced the out-of-Africa bottleneck
+    that may mute the signal of a sweep.  
+    Finally and most importantly, the power to detect sweeps varies considerably along the chromosome,
+    and this variation is much greater than the variation in power between CEU and YRI or between
+    null models. This is likely due to the effects of recombination rate heterogeniety as well
+    as the density of functional elements along the chromosome  (Figure \ref{fig:chr1_power}; bottom).
+
+    We compared the overall performance of the three sweep-detection methods using
+    receiver operating characteristic (ROC) curves (Figure \ref{fig:roc-curves}).
+    Different methods perform better for each of the sampled populations,
+    with \sweepfinder performing betteer than the other two on the CEU sample,
+    and \diploshic performing better on the YRI sample.
+    This is likely a results of the fact that we trained \diploshic on simulations under a constant size population,
+    which more closely matches the demographic history of YRI, than that of CEU, which 
+    is characterized by a strong recent bottleneck followed by population expansion.
+    Interestingly, the simple method that uses reduced diversity ($\pi$) as an indicator
+    for a selective sweep performs reasonably well overall (especially on the CEU sample).
+    However, its accuracy is much more variable along the chromosome compared to the other two method (see Figure \ref{fig:chr1_power}),
+    and for quite a few locations along the chromosome it has power below $50\%$
+    when applied to both sample sets.
+    In that sense, \sweepfinder appears to have the least variation in power along the chromosome or across populations.
+
+    \begin{figure}[b!]
         \centering
@@
         \caption{
-        Power to detect selective sweeps at 100 locations along human
-        chromosome 1 under a three population out-of-Africa model
-        \citep{gutenkunst2009inferring}  when using null models with and
-        without purifying selection as a function of local recombination rate.
-        Simulations are as described in Figure \ref{fig:chr1_power}.  True
-        positive rate is shown for these methods for samples from the CEU and
-        YRI samples (left and right respectively).  To define thresholds for
-        calling a sweep, two different null models were used: a neutral model
-        (blue lines), and a model with purifying selection acting on exons (red
-        lines).  Points represent the power at each location and trend lines
-        along with 95\% confidence intervals are shown fit to those points.
+        Power to detect selective sweeps 
+        %IG: shortened the title line
+        %at 100 locations along human
+        %chromosome 1 under a three population out-of-Africa model
+        %\citep{gutenkunst2009inferring}  when using null models with and
+        %without purifying selection
+        as a function of local recombination rate.
+        This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
+        but with the 100 genomic segments sorted according to their
+        average recombination rates instead of position along chromosome 1.
+        \igcomment{Specify how the curves were computed}
+        %IG: copoed from the caption of fig 5
+	Genomic segments were simulated with sweeps under a three population out-of-Africa model
+        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
+        Three methods for detecting sweeps were applied to simulated data:
+        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
+        2) \diploshic \citep{kern2018diplos} (middle row),
+        and 3) reduced diversity ($\pi$) (bottom row).
+        Power (true positive rate) is shown for these methods for the CEU and YRI
+        samples (left and right respectively).
+        The thresholds of the test statistics were set to control for
+        $5\%$ false positive rate under a neutral null model (blue)
+        and a null model with background selection from deleterious mutations in exons (red).
         }
@@
     To further dissect the reasons for high variance (along the chromosome) in power to detect sweeps,
-    we show in Figure \ref{fig:power-recomb} the relationship between power and local recombination rate for our sweep simulations.
+    we examined the relationship between power and local recombination rate for our sweep simulations ( Figure \ref{fig:power-recomb}).
     We see that power to detect sweeps is
-    a decreasing function of local recombination rate for \sweepfinder and $\pi$ outliers,
+    a decreasing function of local recombination rate for \sweepfinder and reduced diversity,
     which is consistent with sweeps having a larger genomic footprint in regions of low
@@
     probably because it is designed to distinguish regions where a sweep happened from regions linked to a sweep.
-    We also see that the presence of purifying selection acting on exons has a relatively small effect on the power to detect sweeps,    
-    but one that is present across the entire range of recombinations rates observed in human chromosome 1. 
-    Unsurprisingly, the empirical cutoff of $\pi$ to detect sweeps is the less robust to the effects of recombination rate than the other two methods,
-    which are specifically designed to find selective sweeps.
-
-    In Figure \ref{fig:power-exon}, we show the relationship between power and local exon density for our sweep simulations.
-    There is little relationship between power (across all methods) and exon density,
-    which is somewhat surprising given that background selection should somewhat interfere with the sweep signature.
-    It is unclear why that is so, but the effects of realistic background selection may not be as similar to sweeps as previously thought \citep{schrider2020background}.
+    These results confirm that reduced genetic diversity performs well as an indicator for selective sweeps only in regions with
+    low recombination rate.
+    The other two methods, which were specifically designed for sweep detection, are considerably more robust 
+    to the effects of recombination rate than the simple indicator of reduced genetic diversity.
+    These methods are particularly robust to recombination rate variation when applied to 
+    the YRI sample, especially when using a neutral null model to calibrate the test statistics.
+    The other null model, which assumes background selection from exons,
+    appears to consistently lead to reduced power for detecting sweeps regardless of the recombination rate.
+    Interestingly, we do not see a clear correlation between power to detect sweeps and
+    local exon density (Figure \ref{fig:power-exon}).
+    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature.
+    This might suggest that the effects of realistic background selection on genome variation
+    may not be as similar to those of sweeps as previously thought \citep{schrider2020background}.
 
@@
 
-    To test methods for detecting selective sweeps, we simulated 100 evenly distributed 5Mb
+    To test methods for detecting selective sweeps, we simulated 100 evenly distributed 5cM
     segments of human chromosome 1.
+    As recombination rate varies across the genome, the size of the region in base pairs simulated varied, however the recombination distance,
+    which is critical for the linked effects of a sweep, was held constant.
     Each segment was simulated in each of the following three settings:
@@
     Hard sweeps were modeled by generating a single beneficial mutation 
-    with selection coefficient of $s = 0.03$ in the center of the 5Mb genomic segment
+    with selection coefficient of $s = 0.03$ in the center of the 5cM genomic segment
     and discarding simulations in which the beneficial mutation did not reach (or exceed) a terminal frequency of $95\%$.
@@
 
-    In this analysis, we compared three different criteria for sweep detection:
-    (1) reduced nucleotide diversity ($\pi$),
-    (2) the composite likelihood ratio (CLR) computed by \sweepfinder \citep{degiorgio2016sweepfinder2},
-    and (3) a probability for sweeps computed using the supervised learning approach of \diploshic \citep{kern2018diplos}.
-    Nucleotide diversity was computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently}
-    over 10 equally sized and non-overlapping subwindows, and 
-    the CLR was computed at 21 uniformly distributed positions along the simulated region.
-    To train the \diploshic classifier, we used \texttt{discoal} \citep{kern2016discoal} to simulate 20 haploid sequences of length 1.21Mbp
+    Each of the three sweep-detection methods is based on a test statistic.
+    Reduced nucleotide diversity is based on nucleotide diversity  ($\pi$) 
+    computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently}
+    over 10 equally sized and non-overlapping subwindows.
+    \sweepfinder \citep{degiorgio2016sweepfinder2} is based a composite likelihood ratio (CLR) 
+    computed at 21 uniformly distributed positions along the simulated region.
+    \diploshic \citep{kern2018diplos} uses supervised learning to train a sweep classifier.
+    Its training phase used \texttt{discoal} \citep{kern2016discoal} to simulate 20 haploid sequences of length 1.21Mbp
     with a constant population size of $N_e=10,000$, a mutation rate of $2\times 10^{-8}$, and a recombination rate of $2\times 10^{-8}$.
@@
 
-    Each of the three sweep-detection methods was associated with a threshold for its test statistic ($\pi$, \sweepfinder's CLR, or \diploshic's probability)
-    that was determined using an empirical distribution representing a null model without sweeps.
+    The three sweep-detection methods were associated with thresholds for their respective test statistics
+    ($\pi$, \sweepfinder's CLR, or \diploshic's probability),
+    which were determined using an empirical distribution representing a null model without sweeps.
     We considered two null models: a neutral evolution model (neutral),
@@
     Receiver operating characteristic (ROC) curves for the three methods for detecting sweeps.
-    False positive rates were established using neutral simulations that contained no sweeps
-    but were conditioned on the same population history as the CEU and YRI samples.
+    True positive rates (TPR) were computed across all 100,000 datasets simulated with selective sweeps,
+    and false positive rates (FPR) were established by applying the methods to genomes simulated
+    under neutral evolution with the same demographic history and genetic map
+    used in the simulations that included sweeps (see \textbf{Methods}).
+    \igcomment{Add AUCs?}
     }
@@
     \caption{
-    Power to detect selective sweeps at 100 locations along human chromosome 1 under a three-population out-of-Africa model \citep{gutenkunst2009inferring}  
-    with and without purifying selection as a function of local exon density.
-    Simulations are as described in Figure \ref{fig:chr1_power}.
-    Each panel shows the results of simulations with purifying selection acting on exons (red; labelled BGS)
-    and without purifying selection (blue; labelled Neutral).
-    True positive rate is shown for these methods for the CEU and YRI samples (left and right respectively).
-    Points represent the power at each location and trend lines along with 95\% confidence intervals are shown fit to those points.
+    %IG: copied new caption from fig 6 and adjusted.
+        Power to detect selective sweeps as a function of local exon density.
+        This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
+        but with the 100 genomic segments sorted according to their
+        exon density instead of position along chromosome 1.
+        \igcomment{Specify exactly how exon density is computed (\% of bases in exons?) and how the curves were computed}
+        %IG: copoed from the caption of fig 5
+	Genomic segments were simulated with sweeps under a three population out-of-Africa model
+        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
+        Three methods for detecting sweeps were applied to simulated data:
+        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
+        2) \diploshic \citep{kern2018diplos} (middle row),
+        and 3) reduced diversity ($\pi$) (bottom row).
+        Power (true positive rate) is shown for these methods for the CEU and YRI
+        samples (left and right respectively).
+        The thresholds of the test statistics were set to control for
+        $5\%$ false positive rate under a neutral null model (blue)
+        and a null model with background selection from deleterious mutations in exons (red).
     }

### 2025-02-18 analysis2_manuscript.tex
figure updates after Ilan's review

@@
         \centering
-        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_OutOfAfricaArchaicAdmixture_5R19_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
+        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
         \igcomment{NEED TO UPDATE FIG WITH NEW VERSION}
@@
     %IG: NEED TO LINK NEW FIG HERE
-    %\includegraphics[width=\textwidth]
+    \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final.pdf}
     \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}

### 2025-02-27 analysis2_manuscript.tex

@@
     Furthermore, different simulation and inference tools may use different structures
-    and parameter scalings for their models.
+    and parameter scalings for their models,
+    making it difficult to directly compare across these tools and the studies that use them.
 
@@
     both accessible to a wide range of researchers and highly reproducible. 
-    Further the framework should include the complex realities of 
+    Further, the framework should include the complex realities of 
     genomes, including heterogenous recombination rates, 
     variation in the size and density of functional elements, and
-    non-equilibrium demographic histories. Ideally, one should be able
-    to model all of these features using estimates/annotations obtained
+    non-equilibrium demographic histories, and where possible, modeling these features using estimates/annotations obtained
     from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.
@@
     %    
-    For example, say that we wish to implement a model of selection consistent with the
+    For example, say we wanted to implement a model of selection consistent with the
     DFE inferred by \citep{kim2017inference} for exons in the human genome.
     The associated \texttt{DFE} object is defined with two mutation types,
-    neutral and deleterious, corresponding to synonymous and non-synonimous sites.
-    Their proportions are set to $0.3$ and $0.7$ respectively, following their expected
-    prevelanece in exons.
+    neutral and deleterious, corresponding to synonymous and non-synonimous sites,
+    with their proportions set to following their expected
+    prevalence in exons.
     The neutral mutation type is associted with a fixed selection coefficient (0),
@@
     \igcomment{modified this paragraph to focus on the catalog and moved simulation details to next paragraph}
-    As with other \stdpopsim models, when incorporating selection we aim to hew as closely as possible
-    to realistic genomes.
-    Thus, the \stdpopsim catalog provides \texttt{Annotation} object based on public annotations
-    of specices' functional genomic elements 
+    As with other aspects of \stdpopsim, we intend the selection models to be as biologicaly accurate as possible.
+    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publically available functional genomic elements 
     and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
-    Since the vast majority of published DFE estimates are based on a limited number of species,
-    \stdpopsim currently has implemented DFEs for four species:
+    There are relatively few species with published DFE estimates in the literature.
+    \stdpopsim currently has implemented DFEs for four species in the catalogue:
     \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
@@
     While the DFEs are from a limited number of sepecies, these DFEs can be applied to other species
-    in the catalog. For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
+    in the catalog if and when it is suitable. 
+    For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
     estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
@@
     %and can be combined with other \texttt{DFE} objects 
-    Simulations with selection can currently be generated only 
+    Simulations with selection are generated 
     using the SLiM simulation engine, which is
     %is perhaps
-    the most flexible simulation engine for modeling selection available to-date
-    (msprime currently cannot be used to simulate models of selection).
+    the most flexible simulation engine for modeling selection available to-date.
     %IG: this is a nice statement, but I couldn't figure where to plug it after my changes so I'm leaving it commented
@@
     and conserved elements).
-    However, natural selection acting on linked sites (that are masked out of the analysis)
+    However, natural selection acting on linked sites (even when masked out of the analysis)
     can bias estimates of $N_e(t)$
@@
         \centering
-        %IG: NEED TO LINK NEW FIG HERE
-        %\includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
-        \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
+        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
         \caption{
         \label{fig:human-demography}
+        %citations aren't always allowed in captions. We should check with MBE.
         Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
@@
     inference than the two methods based on the sequenctioal Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
-    Reagrdless of these difference, we see that background selection does not
+    Reagrdless of these difference, we again see that background selection does not
     considerably influence the inference of any of the four methods also for this species,
@@
     %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
-    For comparison, here we have normalized all inference results to the convention of SLiM.
-    \igcomment{We should add a detailed description in the Methods of what we do to scale the output of each DFE-inference method.}
-
+    For comparison, here we have normalized all inference results to the convention of SLiM (see \textbf{Methods}).
     % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?
@@
         \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
-        \igcomment{NEED TO UPDATE FIG WITH NEW VERSION}
         \caption{Inferred versus simulated DFEs from simulated human genomes.
@@
     Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome.
-    Recall that these simulations apply a gamma-dsitributed selection coefficient to mutations in exons
+    Recall that these simulations apply a gamma-distributed selection coefficient to mutations in exons
     and a relationship between the selection and dominance coefficients (see \textbf{Methods}).
@@
         \grapes, \polydfe, and \dadicli.
-        Panel C shows the simulated distribution of $2 h s$,
-        along with the simulated distribution of $s$ and the median inferred DFEs. \igcomment{What is median inferred DFE?}
+        Panel C shows the simulated distribution of $2 h s$ (gamma parameters averaged across replicates),
+        along with the simulated distribution of $s$.
         The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
@@
     Different methods perform better for each of the sampled populations,
-    with \sweepfinder performing betteer than the other two on the CEU sample,
+    with \sweepfinder performing better than the other two on the CEU sample,
     and \diploshic performing better on the YRI sample.
@@
     We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Fig 4).
-This performance is not surprising as decreasing h for a given distribution of s will lead to a decrease in hs, or an apparent decrease in the amount of selection.
+    This performance is not surprising as decreasing h for a given distribution of s will lead to a decrease in hs,
+    or an apparent decrease in the amount of selection.
     Fitting a mis-specified DFE that assumes $h=0.5$ will then lead to underestimates of $E[s]$.
     Indeed, the distribution of $hs$ is much more accurately estimated (Fig 4b).
-Overall, these results are consistent with prior work showing that it is challenging to separately infer the DFE of $s$ along with $h$ (cite PMID: 24830675; PMID: 39302992; PMID: 34951958 ) and that a large proportion (>5\%) of recessive deleterious mutations can confound inferences of $s$ (PMID: 37074880).
-Given the plethora of evidence for recessive deleterious mutations (cite PMID: 4630587, PMID: 39114967 , PMID: 30013096 , PMID: 21098719 ), these results suggest the need for care in DFE inference.
+    Overall, these results are consistent with prior work showing that it is challenging to separately infer the DFE of $s$ along with $h$ (cite PMID: 24830675; PMID: 39302992; PMID: 34951958 )
+    and that a large proportion (>5\%) of recessive deleterious mutations can confound inferences of $s$ (PMID: 37074880).
+    Given the plethora of evidence for recessive deleterious mutations (cite PMID: 4630587, PMID: 39114967 , PMID: 30013096 , PMID: 21098719 ),
+    these results suggest the need for care in DFE inference.
     Kyriazis and Lohmueller (PMID: 39302992 ) inferred the DFE of $s$ for nonsynonymous mutations in humans when considering different values of $h$.
     Intriguingly, additive models show a satisfactory fit to the SFS of nonsynonymous variants in humans (Kyriazis et al) and more complex models of dominance do not substantially improve the fit.
-Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome, show a similar fit to the SFS and may be more biologically realistic than fully additive models, given the evidence for the presence of recessive deleterious mutations.
+    Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome, show a similar fit to the SFS and may be more biologically realistic than fully additive models,
+    given the evidence for the presence of recessive deleterious mutations.
     More broadly, new computational approaches are needed to precisely quantify the DFE and the covariance of $h$ and $s$.
@@
     Each simulation was implemented using a relatively short python script, similar to the one shown in Figure 1B
-    (see \textbf{Data availablitiy} below).
+    (see \textbf{Data availablity} below).
     \igcomment{We should add a link to the analysis2 repo in data/code availablity section}
+    %https://github.com/popsim-consortium/analysis2/tree/main
     % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
@@
     % may in reality be under selection. As above, all methods were run using default settings.
+    As described above, each method scales DFE slightly differently. 
+    To make direct comparisons among them,
+    we rescaled the selection coefficent for \grapes  as $s \mu / \theta$,
+    \polydfe as $2 \mu s / \theta$, and
+    \dadicli $\alpha * \nu / (\theta / (4 \mu L p))$.
+    In each case, $\mu$ is the mutation rate,
+    $\theta$ is each method's estimated value of the population scaled mutation rate ($\theta = 4N_{e}\mu$),
+    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution, $L$ is the total number of sites available for the analysis,
+    $p$ is the proportion of sites under selection.
 

### 2025-03-04 analysis2_manuscript.tex
A few more edits after Silas' + Dan's passes

@@
 \newcommand{\kelcomment}[1]{\textcolor{blue}{KEL: #1}}
+
+\newcommand{\drscomment}[1]{\textcolor{purple}{DRS: #1}}
 
@@
     DFE estimation, and selective sweep detection across different species and scenarios. 
-    Our results highlight how selection can bias demographic inference, 
+    Our results demonstrate the robustness of demographic inference methods to selection on linked sites, 
     reveal the sensitivity of DFE inference methods to model assumptions, 
@@
     understand the effects of complex evolutionary processes on genetic
-    variation \citep[e.g.,][]{galloway2020few},\kelcomment{also cite: PMID: 32487519 } to provide a null model for hypothesis testing
+    variation \citep[e.g.,][]{galloway2020few},\kelcomment{also cite: PMID: 32487519 } \drscomment{and maybe a few others?
+PMIDs: 24102784 and thornton2019polygenic} to provide a null model for hypothesis testing
     \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
@@
     While that is so, the original version of \stdpopsim did not include
-    models of selection, which is a major limitation for more empirical
+    models of selection, which is a major limitation for more empirically motivated
     applications of population genetic simulation. In particular, modeling selection
     through simulation is critical for understanding processes such
-    as adaptation, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
+    as adaptation\citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
     background selection on genetic diversity
@@
     both accessible to a wide range of researchers and highly reproducible. 
-    Further, the framework should include the complex realities of 
+    Furthermore the framework should include the complex realities of 
     genomes, including heterogenous recombination rates, 
     variation in the size and density of functional elements, and
-    non-equilibrium demographic histories, and where possible, modeling these features using estimates/annotations obtained
+    non-equilibrium demographic histories. Ideally, one should be able
+    to model all of these features simultaneously using estimates and annotations obtained
     from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.
@@
     to benchmark and compare the performance of different methods for
-    inferring demographic history, the distribution of fitness effects,
-    and the targets of recent positive selection, from genetic data.
+    using genetic data to infer demographic history, the distribution of fitness effects,
+    and the targets of recent positive selection.
 
@@
     This was accomplished through the use of a standardized interface
-    and data structure that allows for curated addition of new
+    and data structure that allows for the curated addition of new
     species and models, and can use either
@@
     as the backend engine for efficient population genetic simulation.
-    \igcomment{edited here to add details on the spec of the DFE and Annotation classes
-    moved below the information on how they are used in simulation. Please check!}
+    %IG: edited here to add details on the spec of the DFE and Annotation classes
+    %moved below the information on how they are used in simulation. Please check!
     To extend \stdpopsim's functionality to include models of selection, 
@@
     their proportions, where 
-    each mutation type is associated with a dominance coefficient (in $[0,1]$),
+    each mutation type is associated with a dominance coefficient
+    (currently limited to values between 0 and 1, inclusive),
     the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
@@
     Each \texttt{Annotation} object is typically associated with a collection of functional
-    genomic elements specified using a GFF3 file. %
+    genomic elements whose coordinates are specified using a GFF3 file. %
     %    
-    For example, say we wanted to implement a model of selection consistent with the
-    DFE inferred by \citep{kim2017inference} for exons in the human genome.
+    For example, say that we wish to implement a model of selection consistent with the
+    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
     The associated \texttt{DFE} object is defined with two mutation types,
-    neutral and deleterious, corresponding to synonymous and non-synonimous sites,
-    with their proportions set to following their expected
-    prevalence in exons.
+    neutral and deleterious, corresponding to synonymous and non-synonimous sites.
+    with their proportions set to follow their expected
+    prevelanece in exons. 
     The neutral mutation type is associted with a fixed selection coefficient (0),
@@
     a gamma-distributed selection coefficient
-    with mean and shape as inferred by \citep{kim2017inference}.
-    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients
+    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).
+    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep{huber2018gene}
     by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below).
@@
         distribution of fitness effects (DFE) and a set of functional
-        annotations. \rngcomment{I find the images too small to read the text comfortably.}
-        \igcomment{I changed this to full width. We might want to reduce font size in the panel titles.}}
+        annotations. }
     \end{figure}
 
-    \igcomment{modified this paragraph to focus on the catalog and moved simulation details to next paragraph}
-    As with other aspects of \stdpopsim, we intend the selection models to be as biologicaly accurate as possible.
+    As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
     Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publically available functional genomic elements 
     and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
+    By using annotation objects, one can, for example, simulate populations where selected mutations (whose selection
+    coefficients are drawn from a specified DFE) occur within annotation regions (e.g. exons), while all mutations
+    occurring outside of these regions do not affect fitness.
     There are relatively few species with published DFE estimates in the literature.
@@
     a genetic map if available, a model of demography, and pairings of DFEs with genomic segments
-    using either the python API or the command line interface (Figure \ref{fig:schematic}B-C).
+    using either the python API or the command-line interface (Figure \ref{fig:schematic}B-C).
     %IG: commented this out because I think it's covered by the text I wrote above
@@
     %is perhaps
-    the most flexible simulation engine for modeling selection available to-date.
+    the most flexible simulation engine for modeling selection available to date
+    (msprime's selection models are relatively limited and thus not currently used by \stdpopsim).
     %IG: this is a nice statement, but I couldn't figure where to plug it after my changes so I'm leaving it commented
@@
     At this release the \stdpopsim catalog comprises 24 species (i.e. genome representations)
-    for which we have implemented 28 demographic models, have 37 genetic maps, and 7 DFEs. %someone should double check the numbers
-    \igcomment{do we want to also mention the number of annotation objects?}
+    for which we have included 28 demographic models, 37 genetic maps, and 7 DFEs. %someone should double check the numbers
     Additions to the catalog are ongoing---we point the reader to our previously published 
@@
     or contact one of the PopSim Consortium members for guidance.
+    \igcomment{Mention releas number above, update the num of species and models and maybe also mention the number of annotation objects.}
 
@@
     the effective population size over time, $N_e(t)$, from genetic data. This can be done
-    using a variety of methods, including the sequentially Markovian coalescent
+    using a variety of methods, including the sequential Markovian coalescent
     \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
     \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
+    \drscomment{One could use dadi for that, but then we should also cite things like fastsimcoal and ABC, while explaining
+    that users can use those methods to infer piecewise-constant N(t) histories as well with a prespecified number of epochs
+    and priors around their params, but maybe that is too much of a digression here? I'm fine either way}
     through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
@@
     and conserved elements).
-    However, natural selection acting on linked sites (even when masked out of the analysis)
-    can bias estimates of $N_e(t)$
+    However, natural selection acting on linked sites---even those that are masked out of the
+    analysis---can bias estimates of $N_e(t)$
     away from true census population sizes by increasing the 
     rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
-    Using \stdpopsim, we can easily examine the effect of selection on linked sites,
-    which is commonly referred to as \emph{background selection}, 
-    on different methods for inferring $N_e(t)$.
-
-    For this purpose, we simulated human genomes with and without natural selection.
-    In both scenarios, we ran three replicate simulations
-    under the out of Africa (OOA) demographic model of
-    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
-    Simulations with selection were implemented by modeling purifying selection on exons
-    % IG: moved details to methods
-    %, defined from the HAVANA group release 104 for the human genome,
-    using a DFE inferred by \cite{kim2017inference}.
-    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
-    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
-    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
-    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.
+    Using \stdpopsim, we can easily examine how different methods for inferring $N_e(t)$
+    are influenced by selection on linked sites,
+    which in the case of negative selection is commonly referred to as \emph{background selection}
+    \citep{charlesworth1993effect,hudson1995deleterious}.
 
@@
 
-    \igcomment{Rewrote this paragraph based on new figs.}
-    %
-    All four inference methods appear to infer similar $N_e(t)$ trajectories from
-    genomes simulated without selection (Figure \ref{fig:human-demography}; blue) and
-    genomes simulated with selection (Figure \ref{fig:human-demography}; orange).
-    Thus, background selection does not appear to notably bias the estimates
-    produced by these methods.
+    For this purpose, we simulated human genomes with and without natural selection.
+    In both scenarios, we ran three replicate simulations
+    under the out of Africa (OOA) demographic model of
+    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
+    Simulations with selection were implemented by modeling purifying selection on exons
+    % IG: moved details to methods
+    %, defined from the HAVANA group release 104 for the human genome,
+    using a DFE inferred by \cite{kim2017inference}.
+    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
+    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
+    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
+    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.
+
+    Although the estimates produced by the four methods differ from one another, 
+    for each method the estimates produced from data with selection (Figure
+    \ref{fig:human-demography}; orange) and without selection (Figure
+    \ref{fig:human-demography}; blue) are fairly similar.
+    Thus, this particular parameterization of background selection does not
+    appear to notably bias the estimates produced by these methods.
     As noted in previous studies that compared demography inference methods
@@
     we see that each method has its strengths and weaknesses.
-    Methods based on the sequenial Markovian coalescent (\smcpp and \msmc)
+    Methods based on the sequential Markovian coalescent (\smcpp and \msmc)
     appear to overal produce the most accurate trajectories of $N_e(t)$
@@
     is targeted toward demographic changes in more recent time frames ($\sim 200$ generations),
-    explaining its overall noisy estimates.
+    explaining its noisy estimates for all but the most recent time periods.
 
@@
     Here, we see that the site-frequency-based method (\stairway) produces more accurate
-    inference than the two methods based on the sequenctioal Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
-    Reagrdless of these difference, we again see that background selection does not
-    considerably influence the inference of any of the four methods also for this species,
+    inference than the two methods based on the sequential Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
+    Reagrdless of these differences, we see that background selection again does not
+    appear to considerably influence the inference of any of the four methods for this species,
     which has a very different evolutionary history than humans.
@@
     Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE) of new mutations
-    from genetic data. The new framework of modeling selection in \stdpopsim is ideally designed to benchmark and compare
+    from genetic data. The new framework of modeling selection in \stdpopsim makes it an ideal tool for easily benchmarking and comparing
     the performance of different methods for inferring the DFE. 
@@
         \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
+        \drscomment{I think it would be good to increase the font size for the A,B,C panel legends and also
+        draw a horizontal line above panel C so it is clear that all three plots at the bottom are part of C.}
         \caption{Inferred versus simulated DFEs from simulated human genomes.
@@
         and simulated values represented by dashed horizontal lines.
-        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets;
+        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
         white bars represent the distribution used in the simulation.
@@
 
-    \igcomment{I start with the OOA model and then cite the constant-size model as supplementary.}
     We start by examining DFE inference from human genomes simulated under an out-of-Africa demographic model
     with selection acting on exons (see \textbf{Methods}).
-    In this setting, DFE is inferred separately for each of the three simulated populations
+    In this setting, the DFE is inferred separately for each of the three simulated populations
     using three different DFE inference methods: \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
@@
     Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
+    Note that while the estimation accuracy for the DFE parameters themselves (Figure \ref{fig:homsap-dfe.ooa}A-B) is useful,
+    examining the difference between the actual and inferred distributions can be more helpful for showing how well a method
+    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C).
     The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
-    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters (Figure \ref{fig:homsap-dfe.ooa}A-B).
-    As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$; Figure \ref{fig:homsap-dfe.ooa}C),
-    while \grapes under-estimates the fraction of mutations with $s<0.01$ and considerably over-estimates the fraction of
-    mutations under weak selection ($s\geq-0.001$).
+    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters.
+    As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$),
+    while \grapes considerably over-estimates the fraction of mutations under weak selection ($s\geq-0.0001$).
+    % IG: commented this out to simplify the basic observation
+    % and under-estimates the fraction of mutations with $s<-0.01$.
+
     Population history appears to influence different methods in slightly different ways.
@@
     (relative to the other two populations), whereas \polydfe
-    produces the most accurate estimate from the CHB genomes.
+    produces its most accurate estimate from the CHB genomes.
     To factor out the influence of demographic changes, we also analyzed genomes generated using similar
-    simulations, but with a simpler demographic model with no population size changes (Figure \ref{fig:homsap-dfe.constant}).
+    simulations but with no population size changes (Figure \ref{fig:homsap-dfe.constant}).
     In this simpler setting, \polydfe overlapped the true mean selection coefficient and shape parameter, whereas 
     the other methods do not. 
-    \grapes exhibits similar biases here as in the simulations with more complex demography,
+    \grapes exhibits similar biases without population size changes as in the simulations with more complex demography,
     but with milder under-estimation of the shape parameter, but \dadicli shows an inverse pattern
     with overestimated shape parameter and slightly underestimated mean.
-\kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things.
-That said, I think it would be good to plot the empirical (in the case simulated) SFS along with the SFSs from the MLE models.
-I would also do this for the demogrpahic models as well as the DFE models}
-\igcomment{I think that this is better conveyed now in the added histograms for $s$ in panel C}
+
+    \begin{figure}[b!]
+        \centering
+        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
+        \caption{
+        \label{fig:vaquita-dfe}
+        Inferred versus simulated DFEs from simulated vaquita genomes.
+        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
+        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$) (see \textbf{Methods}).
+        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
+        \grapes, \polydfe, and \dadicli.
+        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
+        shown for each DFE inferred from all three simulated datasets,
+        with median values marked by horizontal bars
+        and simulated values represented by dashed horizontal lines. 
+        (C) Binned distribution of $2hs$ implied by the simulated DFE compared with
+        the distribution of $s$ implied by the average DFE inferred for each method from the three simulated datasets (averaging the inferred gamma parameters).
+        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
+        }
+    \end{figure}
 
@@
     %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
-    Next, we compared these three DFE-inference methods on simulations of the vaquita porpoise genome.
+    Next, we benchmarked these three DFE-inference methods on simulations of the vaquita porpoise genome.
     Recall that these simulations apply a gamma-distributed selection coefficient to mutations in exons
@@
     This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
-    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
-   \kelcomment{I moved the last sentence from here to the Discussion and fleshed it out}
-
-    %add PhoSin DFE figure here
-    \begin{figure}[t!]
-        \centering
-        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
-        \caption{
-        \label{fig:vaquita-dfe}
-        Inferred versus simulated DFEs from simulated vaquita genomes.
-        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
-        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$). 
-        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; panel A) and shape parameter (panel B).
-        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
-        \grapes, \polydfe, and \dadicli.
-        Panel C shows the simulated distribution of $2 h s$ (gamma parameters averaged across replicates),
-        along with the simulated distribution of $s$.
-        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
-        }
-    \end{figure}
+    where their selective effect is $h s$ rather than $s$.
+    %KEL: I moved the last sentence from here to the Discussion and fleshed it out
 
@@
     is determined using simulated data with background selection (BGS null model).
-    This observation is consistent with what was previously shown by \cite{schrider2020background},
-    but the effect is relatively minor.
+    %This observation is consistent with what was previously shown by \cite{schrider2020background},
+    %but the effect is relatively minor. DRS: cutting this because my paper showed that power decreases when our test set has BGS and our training set does not, so not quite the same.
     This is likely because the strength of selection acting on exons is relatively weak, and the
@@
     and this variation is much greater than the variation in power between CEU and YRI or between
-    null models. This is likely due to the effects of recombination rate heterogeniety as well
-    as the density of functional elements along the chromosome  (Figure \ref{fig:chr1_power}; bottom).
+    null models. This is likely due to the effects of recombination rate heterogeniety (Figure \ref{fig:chr1_power}).% as well
+    %as the density of functional elements along the chromosome.
+    %DRS: cut the density of func elements because that stuff doesn't seem to matter much here
 
@@
     and \diploshic performing better on the YRI sample.
-    This is likely a results of the fact that we trained \diploshic on simulations under a constant size population,
+    This is may result from the fact that we trained \diploshic on simulations under a constant size population,
     which more closely matches the demographic history of YRI, than that of CEU, which 
     is characterized by a strong recent bottleneck followed by population expansion.
+    \drscomment{note: I am couching the language above a bit because we have previously shown that shic doesn't suffer in the sweep vs.
+nonsweep question when doing other train-on-constant-test-on-nonequilib benchmarks.
+Could be something else going on here (or some interaction between demog misspec and rec rate variation), but in any case I am not certain of the cause so I am toning down our certainty slightly.}
     Interestingly, the simple method that uses reduced diversity ($\pi$) as an indicator
@@
 
-    To further dissect the reasons for high variance (along the chromosome) in power to detect sweeps,
+    To further dissect the reasons for high variance along the chromosome in power to detect sweeps,
     we examined the relationship between power and local recombination rate for our sweep simulations ( Figure \ref{fig:power-recomb}).
@@
     \diploshic on the other hand shows a slight increase in power with recombination rates,
-    probably because it is designed to distinguish regions where a sweep happened from regions linked to a sweep.
+    probably because it is designed to distinguish regions where a sweep happened from regions linked to a sweep,
+    and if the recombination rate is too low (or selection too strong) then the window examined by \diploshic may
+    not be large enough to exhibit enough of a recovery in patterns of diversity flanking the sweep.
     These results confirm that reduced genetic diversity performs well as an indicator for selective sweeps only in regions with
@@
     We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Fig 4).
-    This performance is not surprising as decreasing h for a given distribution of s will lead to a decrease in hs,
+    This performance is not surprising as decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
     or an apparent decrease in the amount of selection.
@@
     Overall, these results are consistent with prior work showing that it is challenging to separately infer the DFE of $s$ along with $h$ (cite PMID: 24830675; PMID: 39302992; PMID: 34951958 )
-    and that a large proportion (>5\%) of recessive deleterious mutations can confound inferences of $s$ (PMID: 37074880).
+    and that a large proportion ($>5\%$) of recessive deleterious mutations can confound inferences of $s$ (PMID: 37074880).
     Given the plethora of evidence for recessive deleterious mutations (cite PMID: 4630587, PMID: 39114967 , PMID: 30013096 , PMID: 21098719 ),
@@
 
-    Although we have shown how the new API can be used to benchmark the performance of different methods,
+    Regarding sweep detection, the inclusion of selection models into \stdpopsim allowed us to easily compare the
+    effectiveness of different methods on data reflecting the demographic history, recombination rate variation,
+    and genomic locations of negatively selected sites present in human populations. These results clearly demonstrate
+    the impact of recombination rate variation, and underscore the importance of training methods on data with variable
+    recombination rates (e.g. following Schrider and Kern \citeyear{schrider2017soft}). In addition, our results
+    illustrate a strength of likelihood methods like \sweepfinder, which do not require a predefined window size and
+    thus will perform better than \diploshic---which only examines relative summarie of diversity across a window instead
+    of the actual values of these diversity measures---on very strong sweeps that effect diversity across a wide stretch of
+    chromosomes that extends far beyond \diploshic's window size. \drscomment{Maybe this isn't the coolest stuff ever but
+    I thought it might be worth saying something about our sweep detection stuff in the Discussion.}
+
+    Although we have shown how the new \stdpopsim API can be used to benchmark the performance of different methods,
     the benchmarks shown here merely scratch the surface of what can be done with the new API.
     We envision that, moving forward, many researchers will use
-    the new API to simulate data under a wide range of models of selection, and to develop
+    the new API to simulate data under a wide range of models of selection, demography, and genome architecture, and to develop
     new methods for the analysis of genetic data in the presence of selection. 
@@
 
-    \subsection*{Simulations}
+    \subsection*{Whole genome simulations}
     In this study we performed simulations of the human and vaquita porpoise genomes
@@
 
-    \subsubsection*{Simulations with selective sweeps}
+    \subsection*{Inference of $N_e(t)$}
+    %We assessed the accuracy of demographic inferences from
+    %\msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+    Inference of $N_e(t)$ was conducted on simulated datasets after masking non-recombining portion
+    of the genome (to avoid bias caused by extreme recombination rate variation)
+    and exons (to avoid the direct effect of natural selection).
+    Thus, inference from datasets simulated with selection was not influenced by direct selection
+    acting on exons, but it was influenced by selection acting on linked sites (background selection).    
+    %IG: I commented because it doesn't seem relevant for methods
+    %Comparisons between inferences made with and without selection were 
+    %of particular interest.
+    For \msmc \citep{Schiffels2020}, we used a random sample of $N=6$ simulated genomes with 20
+    iterations of the EM algorithm.
+    For \gone \citep{santiago2020recent}, we set \texttt{max\_snps} to 500,000,
+    the number of generations to 2,000, and the number of bins to 400, all representing default settings.
+    For \stairway \citep{liu2020stairway} and \smcpp \citep{terhorst2017robust} we used the default settings.
+    \igcomment{Nothing more to say about \stairway plot and \smcpp?}
+
+    \subsection*{DFE inference}
+    \igcomment{I commented all the text that was here because it wasn't relevant to the methods. We might want to add it to the discussion.
+    Kirk also noted that a lot of methods-level details are missing:
+    how big was the coding sequence length?
+    How was demographic inference done for the different models via the different programs?,
+    etc.
+    Can someone write this up?}
+
+    % We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
+    % \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
+    % Each software package was run using the same set of simulations.
+    % For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
+    % A key distinction between our analyses and those done in practice
+    % is that we have complete knowledge of which sites are under selection in the simulation,
+    % thus our inferences is a best case scenario. Empirical inference of the DFE is a much
+    % harder problem as some mutations assumed to be free of selection
+    % may in reality be under selection. As above, all methods were run using default settings.
+    \igcomment{I edited the description below written by Silas. Please check to make sure it's still accurate.}
+    As described above, each DFE-inference method scales selection coefficients slightly differently.
+    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by SLiM.
+    The output of \grapes  was scaled by a factor of  $\mu / \theta$,
+    and the output of \polydfe was saled by $2 \mu / \theta$, 
+    where $\mu$ is the assumed mutation rate and 
+    $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
+    \igcomment{Do we really multiple $\alpha$ by $\nu$ in rescaling for \dadicli?}
+    For \dadicli, we report $\alpha * \nu / (\theta / (4 \mu L p))$, where
+    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution,
+    $L$ is the total number of sites available for the analysis,
+    $p$ is the proportion of sites under selection.
+
+    \subsection*{Simulations with selective sweeps}
 
@@
     For each of the 100 genomic segments of chromosome 1, we generated 1,000 replicate datasets under each of the three settings.
-    Each simulation generated 10 diploid genomes for each of the two sampled populations: YRI and CEU.
-    The genomes simulated with a hard sweep and BGS were used for sweep detection,
-    and genomes simulated under the other two settings were used to calibrate the thresholds of the test statistics
+    As with the whole-genome simulations (see above), we used the SLiM engine with a burn-in of $2N$ generations
+    (where $N$ is the initial population size), and a scaling factor of 2.
+    Each simulation generated 10 diploid genomic segments for each of the two sampled populations: YRI and CEU.
+    The sequences simulated with a hard sweep and BGS were used for sweep detection,
+    and sequences simulated under the other two settings were used to calibrate the thresholds of the test statistics
     used by the three sweep-detection methods (see below).
 
-    \subsection*{Inference of $N_e(t)$}
-    %We assessed the accuracy of demographic inferences from
-    %\msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
-    Inference of $N_e(t)$ was conducted on simulated datasets after masking non-recombining portion
-    of the genome (to avoid bias caused by extreme recombination rate variation)
-    and exons (to avoid the direct effect of natural selection).
-    Thus, inference from datasets simulated with selection was not influenced by direct selection
-    acting on exons, but it was influenced by selection acting on linked sites (background selection).    
-    %IG: I commented because it doesn't seem relevant for methods
-    %Comparisons between inferences made with and without selection were 
-    %of particular interest.
-    For \msmc \citep{Schiffels2020}, we used a random sample of $N=6$ simulated genomes with 20
-    iterations of the EM algorithm.
-    For \gone \citep{santiago2020recent}, we set \texttt{max\_snps} to 500,000,
-    the number of generations to 2,000, and the number of bins to 400, all representing default settings.
-    For \stairway \citep{liu2020stairway} and \smcpp \citep{terhorst2017robust} we used the default settings.
-    \igcomment{Nothing more to say about \stairway plot and \smcpp?}
-
-    \subsection*{DFE inference}
-    \igcomment{I commented all the text that was here because it wasn't relevant to the methods. We might want to add it to the discussion.
-    Kirk also noted that a lot of methods-level details are missing:
-    how big was the coding sequence length?
-    How was demographic inference done for the different models via the different programs?
-    How was the unscaling of E[s] done?, etc.
-    Can someone write this up?}
-    % We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
-    % \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
-    % Each software package was run using the same set of simulations.
-    % For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
-    % A key distinction between our analyses and those done in practice
-    % is that we have complete knowledge of which sites are under selection in the simulation,
-    % thus our inferences is a best case scenario. Empirical inference of the DFE is a much
-    % harder problem as some mutations assumed to be free of selection
-    % may in reality be under selection. As above, all methods were run using default settings.
-    As described above, each method scales DFE slightly differently. 
-    To make direct comparisons among them,
-    we rescaled the selection coefficent for \grapes  as $s \mu / \theta$,
-    \polydfe as $2 \mu s / \theta$, and
-    \dadicli $\alpha * \nu / (\theta / (4 \mu L p))$.
-    In each case, $\mu$ is the mutation rate,
-    $\theta$ is each method's estimated value of the population scaled mutation rate ($\theta = 4N_{e}\mu$),
-    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution, $L$ is the total number of sites available for the analysis,
-    $p$ is the proportion of sites under selection.
-
     \subsection*{Selective sweeps detection}
+    %\drscomment{Maybe it makes sense to move this up to right after the sweep simulation section. Should make it easier for readers to follow.}
+    %IG: I moved that section down here to bring them together
 
@@
     we expect it to perform suboptimally on the simulated segments of human chromosome 1,
-    which were generated using mroe relaistic simulations (see \textbf{Simulations with selective sweeps} above).
+    which were generated using more relaistic simulations (see \textbf{Simulations with selective sweeps} above).
 
@@
     The sweep-detection method was applied to the 1,000 
-    simulated datasets generated with a hard sweep (and BGS) for that position,
+    simulated datasets generated with a hard sweep (neutral and BGS) for that position,
     and the power was set to be the proportion of simulations (out of 1,000)
@@
 \section*{Acknowledgments}\label{acknowledgements}
+We thank Tom Booker for discussions and encouragement to incorporate realistic BGS models into \stdpopsim.
 
@@
     \label{funding}
-    Multiple funding sources supported this work.
     MFR, ST, NP, and ADK were supported in part by NIH award R35GM148253.
     RNG and LNT were supported in part by NIH award R35GM149235.
+    DRS was supported in part by NIH award R35GM138286.
 
@@
     \centering
-    %IG: NEED TO LINK NEW FIG HERE
     \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final.pdf}
-    \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
     \caption{
@@
     \caption{
-    \igcomment{Update fig and caption after we finalize Fig. 3}
-    Performance of methods to infer the distribution of fitness effects (DFE) from genetic data.
-    Simulations were performed using a human constant size model with a gamma-distributed DFE
-    acting on exons parameterized by a mean selection coefficient and a shape parameter. On each row we show the true parameter 
-    as a black line, and the estimates as boxplots from replicate simulations. The human model has three extant populations (CEU, CHB, and YRI).
-    %DRS-fig: only one population is shown here, but 3 referenced in the legend. Wrong legend or wrong fig?
-    %DRS-fig: If it is the right fig, get rid of the pop0 legend inset
-    Each row of the figure shows estimates from a different method: 1) GRAPES \citep{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe}, and 3) \dadicli \citep{Huang2023}.
+       Inferred versus simulated DFEs from human genomes simulated with
+       a simple demographic model with one population with a constant size of $10,000$
+       and a gamma-distributed DFE acting on exons (see \textbf{Methods}).
+        DFE is inferred by three different methods: \grapes, \polydfe , and \dadicli.
+        Mean absolute value of selection coefficient ($\lvert E(s) \rvert$, panel A) and shape parameter (panel B)
+        are shown for each DFE inferred from all three simulated datase
+        with median values marked by horizontal bars
+        and simulated values represented by dashed horizontal lines.
     }
@@
     \label{fig:vaquita-dfe.constant}
-    A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
-    Simulations were performed using a model of vaquita porpoise genome under constant population size with a gamma-distributed DFE
-    acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
-    parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
-    This vaquita model has a single extant population, and parameter estimates from each
-    of the three different methods are shown: 1) GRAPES \citep{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe},
-    and 3) \dadicli \citep{Huang2023}.}
-    %DRS-fig: remove the pop 0 legend again
+        Inferred versus simulated DFEs from vaquita genomes simulated with
+        a simple demographic model with one population with a constant size of $3,500$
+        and a gamma-distributed DFE acting on exons 
+        with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$) (see \textbf{Methods}).
+        DFE is inferred by three different methods: \grapes, \polydfe , and \dadicli.
+        Mean absolute value of selection coefficient ($\lvert E(s) \rvert$, panel A) and shape parameter (panel B)
+        are shown for each DFE inferred from all three simulated datase
+        with median values marked by horizontal bars
+        and simulated values represented by dashed horizontal lines.
+    }
 \end{figure}
-% DRS-fig: ditch the unnecessary legends from panels A and B. Also, are we deciding which version of panel C to include here? I guess I prefer
-% to bottom one but with the legend of the top one (``Original s DFE'' sounds a little clunky).
-
-% constant all site size DFE figure
+
 \begin{figure}[h]
@@
     under constant population size  where all sites along the simulated chromosome can experience selected and neutral mutation
-    as opposed to exonic regions only.\igcomment{This figure is not cited anywhere. Do we really need it?}
-    }
-    %DRS-fig: human? vaquita? need more info in this figure legend, and also get rid of the pop 0 thing again
+    as opposed to exonic regions only.\\
+    \igcomment{This fig is not cited anywhere. Do we really need it?}
+    }
 \end{figure}
-
-\begin{table}[ht]
-\centering
-\small
-\caption{\bf{Performance of DFE methods across the simulation scenarios}. 
-Bold rows show the lowest mean absolute error (MAE) for the two DFE parameters
-for each species.}
-\begin{tabular}{lllrrrrrr}
-\toprule
-species ID & demography & annotation & \makecell{MAE \\ $E|s|$ \\ dadi-cli} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi-cli} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
-\midrule
-HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
-HomSap & Constant & exons & \bf{0.0012} & 0.0101 & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
-HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
-HomSap & OOA 5R19 & exons & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
-PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
-PhoSin & Constant & exons & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
-PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
-PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
-\bottomrule
-\end{tabular}
-\label{tab:dfe_table}
-\end{table}
 
@@
 
+\begin{table}[ht]
+\centering
+\small
+
+\caption{\bf{Performance of DFE methods across the simulation scenarios}. 
+Bold rows show the lowest mean absolute error (MAE) for the two DFE parameters
+for each species.\\
+\igcomment{This table is not cited anywhere. The analysis of ``all sites" is also not mentioned anywhere. Do we want to add it?}
+}
+\begin{tabular}{lllrrrrrr}
+\toprule
+species ID & demography & annotation & \makecell{MAE \\ $E|s|$ \\ dadi-cli} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi-cli} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
+\midrule
+HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
+HomSap & Constant & exons & \bf{0.0012} & 0.0101 & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
+HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
+HomSap & OOA 5R19 & exons & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
+PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
+PhoSin & Constant & exons & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
+PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
+PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
+\bottomrule
+\end{tabular}
+\label{tab:dfe_table}
+\end{table}
+
 \stopsupplement

### 2025-03-04 analysis2_manuscript.tex

@@
     DFE estimation, and selective sweep detection across different species and scenarios. 
-    Our results demonstrate the robustness of demographic inference methods to selection on linked sites, 
+    Our results highlight how selection can bias demographic inference, 
     reveal the sensitivity of DFE inference methods to model assumptions, 
@@
     non-equilibrium demographic histories. Ideally, one should be able
-    to model all of these features simultaneously using estimates and annotations obtained
+    to model all of these features simultaneously using estimates/annotations obtained
     from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.
@@
     as the backend engine for efficient population genetic simulation.
-    %IG: edited here to add details on the spec of the DFE and Annotation classes
-    %moved below the information on how they are used in simulation. Please check!
+    \igcomment{edited here to add details on the spec of the DFE and Annotation classes
+    moved below the information on how they are used in simulation. Please check!}
     To extend \stdpopsim's functionality to include models of selection, 
@@
         distribution of fitness effects (DFE) and a set of functional
-        annotations. }
+        annotations. \rngcomment{I find the images too small to read the text comfortably.}
+        \igcomment{I changed this to full width. We might want to reduce font size in the panel titles.}}
     \end{figure}
 
+    \igcomment{modified this paragraph to focus on the catalog and moved simulation details to next paragraph}
     As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
@@
     coefficients are drawn from a specified DFE) occur within annotation regions (e.g. exons), while all mutations
-    occurring outside of these regions do not affect fitness.
+    occurring outside of these regions are fitness neutral.
     There are relatively few species with published DFE estimates in the literature.
@@
     for which we have included 28 demographic models, 37 genetic maps, and 7 DFEs. %someone should double check the numbers
+    \igcomment{do we want to also mention the number of annotation objects?}
     Additions to the catalog are ongoing---we point the reader to our previously published 
@@
     or contact one of the PopSim Consortium members for guidance.
-    \igcomment{Mention releas number above, update the num of species and models and maybe also mention the number of annotation objects.}
 
@@
     the effective population size over time, $N_e(t)$, from genetic data. This can be done
-    using a variety of methods, including the sequential Markovian coalescent
+    using a variety of methods, including the sequentially Markovian coalescent
     \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
@@
     rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
-    Using \stdpopsim, we can easily examine how different methods for inferring $N_e(t)$
-    are influenced by selection on linked sites,
-    which in the case of negative selection is commonly referred to as \emph{background selection}
-    \citep{charlesworth1993effect,hudson1995deleterious}.
+    Using \stdpopsim, we can easily examine the effect of selection on linked sites,
+    which in the case of negative selection is commonly referred to as \emph{background selection} \citep{charlesworth1993effect,hudson1995deleterious}, 
+    on different methods for inferring $N_e(t)$.
+
+    For this purpose, we simulated human genomes with and without natural selection.
+    In both scenarios, we ran three replicate simulations
+    under the out of Africa (OOA) demographic model of
+    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
+    Simulations with selection were implemented by modeling purifying selection on exons
+    % IG: moved details to methods
+    %, defined from the HAVANA group release 104 for the human genome,
+    using a DFE inferred by \cite{kim2017inference}.
+    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
+    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
+    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
+    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
+        %IG: NEED TO LINK NEW FIG HERE
+        %\includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
+        \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
         \caption{
@@
 
-    For this purpose, we simulated human genomes with and without natural selection.
-    In both scenarios, we ran three replicate simulations
-    under the out of Africa (OOA) demographic model of
-    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
-    Simulations with selection were implemented by modeling purifying selection on exons
-    % IG: moved details to methods
-    %, defined from the HAVANA group release 104 for the human genome,
-    using a DFE inferred by \cite{kim2017inference}.
-    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
-    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
-    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
-    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.
-
+    \igcomment{Rewrote this paragraph based on new figs.}
+    %
     Although the estimates produced by the four methods differ from one another, 
@@
     we see that each method has its strengths and weaknesses.
-    Methods based on the sequential Markovian coalescent (\smcpp and \msmc)
+    Methods based on the sequenial Markovian coalescent (\smcpp and \msmc)
     appear to overal produce the most accurate trajectories of $N_e(t)$
@@
     Here, we see that the site-frequency-based method (\stairway) produces more accurate
-    inference than the two methods based on the sequential Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
+    inference than the two methods based on the sequenctioal Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
     Reagrdless of these differences, we see that background selection again does not
@@
         and simulated values represented by dashed horizontal lines.
-        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
+        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets;
         white bars represent the distribution used in the simulation.
@@
 
+    \igcomment{I start with the OOA model and then cite the constant-size model as supplementary.}
     We start by examining DFE inference from human genomes simulated under an out-of-Africa demographic model
@@
     Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
-    Note that while the estimation accuracy for the DFE parameters themselves (Figure \ref{fig:homsap-dfe.ooa}A-B) is useful,
+    Note that while the estimation accuracy for the DFE parameters themselves (Figure \ref{fig:homsap-dfe.ooa}A,B) is useful,
     examining the difference between the actual and inferred distributions can be more helpful for showing how well a method
-    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C).
+    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C)
     The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
-    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters.
-    As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$),
-    while \grapes considerably over-estimates the fraction of mutations under weak selection ($s\geq-0.0001$).
-    % IG: commented this out to simplify the basic observation
-    % and under-estimates the fraction of mutations with $s<-0.01$.
-
+    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters (Figure \ref{fig:homsap-dfe.ooa}A-B).
+    As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$; Figure \ref{fig:homsap-dfe.ooa}C),
+    while \grapes under-estimates the fraction of mutations with $-1<s<-0.01$ and considerably over-estimates the fraction of
+    mutations under weak selection ($s\geq-0.001$).
     Population history appears to influence different methods in slightly different ways.
@@
     with overestimated shape parameter and slightly underestimated mean. 
-
-    \begin{figure}[b!]
-        \centering
-        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
-        \caption{
-        \label{fig:vaquita-dfe}
-        Inferred versus simulated DFEs from simulated vaquita genomes.
-        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
-        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$) (see \textbf{Methods}).
-        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
-        \grapes, \polydfe, and \dadicli.
-        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
-        shown for each DFE inferred from all three simulated datasets,
-        with median values marked by horizontal bars
-        and simulated values represented by dashed horizontal lines. 
-        (C) Binned distribution of $2hs$ implied by the simulated DFE compared with
-        the distribution of $s$ implied by the average DFE inferred for each method from the three simulated datasets (averaging the inferred gamma parameters).
-        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
-        }
-    \end{figure}
+\kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things.
+That said, I think it would be good to plot the empirical (in the case simulated) SFS along with the SFSs from the MLE models.
+I would also do this for the demogrpahic models as well as the DFE models}
+\igcomment{I think that this is better conveyed now in the added histograms for $s$ in panel C}
 
@@
     where their selective effect is $h s$ rather than $s$.
-    %KEL: I moved the last sentence from here to the Discussion and fleshed it out
+   \kelcomment{I moved the last sentence from here to the Discussion and fleshed it out}
+
+    %add PhoSin DFE figure here
+    \begin{figure}[t!]
+        \centering
+        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
+        \caption{
+        \label{fig:vaquita-dfe}
+        Inferred versus simulated DFEs from simulated vaquita genomes.
+        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
+        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$). 
+        Distributions are compared via their mean absolute value ($\lvert E(s) \rvert $; panel A) and shape parameter (panel B).
+        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
+        \grapes, \polydfe, and \dadicli.
+        Panel C shows the simulated distribution of $2 h s$ (gamma parameters averaged across replicates),
+        along with the simulated distribution of $s$.
+        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
+        }
+    \end{figure}
 
@@
 
-    \subsection*{Whole genome simulations}
+    \subsection*{Simulations}
     In this study we performed simulations of the human and vaquita porpoise genomes
@@
 
+    \subsubsection*{Simulations with selective sweeps}
+
+    To test methods for detecting selective sweeps, we simulated 100 evenly distributed 5cM
+    segments of human chromosome 1.
+    As recombination rate varies across the genome, the size of the region in base pairs simulated varied, however the recombination distance,
+    which is critical for the linked effects of a sweep, was held constant.
+    Each segment was simulated in each of the following three settings:
+    (1) neutral, (2) background selection (BGS), and (3) hard sweep with BGS.
+    In all three settings, we applied a genetic map from the HapMap Project
+    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
+    the three-population out of Africa (OOA) demographic model from
+    \cite{gutenkunst2009inferring} (\stdpopsim label \texttt{OutOfAfrica\_3G09}).
+    In the two settings with BGS, we applied the DFE model with gamma-distributed selection coefficients inferred by
+    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
+    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
+    Hard sweeps were modeled by generating a single beneficial mutation 
+    with selection coefficient of $s = 0.03$ in the center of the 5cM genomic segment
+    and discarding simulations in which the beneficial mutation did not reach (or exceed) a terminal frequency of $95\%$.
+    For each of the 100 genomic segments of chromosome 1, we generated 1,000 replicate datasets under each of the three settings.
+    Each simulation generated 10 diploid genomes for each of the two sampled populations: YRI and CEU.
+    The genomes simulated with a hard sweep and BGS were used for sweep detection,
+    and genomes simulated under the other two settings were used to calibrate the thresholds of the test statistics
+    used by the three sweep-detection methods (see below).
+
     \subsection*{Inference of $N_e(t)$}
@@
     how big was the coding sequence length?
-    How was demographic inference done for the different models via the different programs?,
-    etc.
+    How was demographic inference done for the different models via the different programs?
+    How was the unscaling of E[s] done?, etc.
     Can someone write this up?}
-
     % We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
@@
     % may in reality be under selection. As above, all methods were run using default settings.
-    \igcomment{I edited the description below written by Silas. Please check to make sure it's still accurate.}
-    As described above, each DFE-inference method scales selection coefficients slightly differently.
-    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by SLiM.
-    The output of \grapes  was scaled by a factor of  $\mu / \theta$,
-    and the output of \polydfe was saled by $2 \mu / \theta$, 
-    where $\mu$ is the assumed mutation rate and 
-    $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
-    \igcomment{Do we really multiple $\alpha$ by $\nu$ in rescaling for \dadicli?}
-    For \dadicli, we report $\alpha * \nu / (\theta / (4 \mu L p))$, where
-    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution,
-    $L$ is the total number of sites available for the analysis,
+    As described above, each method scales DFE slightly differently. 
+    To make direct comparisons among them,
+    we rescaled the selection coefficent for \grapes  as $s \mu / \theta$,
+    \polydfe as $2 \mu s / \theta$, and
+    \dadicli $\alpha * \nu / (\theta / (4 \mu L p))$.
+    In each case, $\mu$ is the mutation rate,
+    $\theta$ is each method's estimated value of the population scaled mutation rate ($\theta = 4N_{e}\mu$),
+    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution, $L$ is the total number of sites available for the analysis,
     $p$ is the proportion of sites under selection.
 
-    \subsection*{Simulations with selective sweeps}
-
-    To test methods for detecting selective sweeps, we simulated 100 evenly distributed 5cM
-    segments of human chromosome 1.
-    As recombination rate varies across the genome, the size of the region in base pairs simulated varied, however the recombination distance,
-    which is critical for the linked effects of a sweep, was held constant.
-    Each segment was simulated in each of the following three settings:
-    (1) neutral, (2) background selection (BGS), and (3) hard sweep with BGS.
-    In all three settings, we applied a genetic map from the HapMap Project
-    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
-    the three-population out of Africa (OOA) demographic model from
-    \cite{gutenkunst2009inferring} (\stdpopsim label \texttt{OutOfAfrica\_3G09}).
-    In the two settings with BGS, we applied the DFE model with gamma-distributed selection coefficients inferred by
-    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
-    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
-    Hard sweeps were modeled by generating a single beneficial mutation 
-    with selection coefficient of $s = 0.03$ in the center of the 5cM genomic segment
-    and discarding simulations in which the beneficial mutation did not reach (or exceed) a terminal frequency of $95\%$.
-    For each of the 100 genomic segments of chromosome 1, we generated 1,000 replicate datasets under each of the three settings.
-    As with the whole-genome simulations (see above), we used the SLiM engine with a burn-in of $2N$ generations
-    (where $N$ is the initial population size), and a scaling factor of 2.
-    Each simulation generated 10 diploid genomic segments for each of the two sampled populations: YRI and CEU.
-    The sequences simulated with a hard sweep and BGS were used for sweep detection,
-    and sequences simulated under the other two settings were used to calibrate the thresholds of the test statistics
-    used by the three sweep-detection methods (see below).
-
     \subsection*{Selective sweeps detection}
-    %\drscomment{Maybe it makes sense to move this up to right after the sweep simulation section. Should make it easier for readers to follow.}
-    %IG: I moved that section down here to bring them together
+    \drscomment{Maybe it makes sense to move this up to right after the sweep simulation section. Should make it easier for readers to follow.}
 
@@
     we expect it to perform suboptimally on the simulated segments of human chromosome 1,
-    which were generated using more relaistic simulations (see \textbf{Simulations with selective sweeps} above).
+    which were generated using mroe relaistic simulations (see \textbf{Simulations with selective sweeps} above).
 
@@
     \centering
+    %IG: NEED TO LINK NEW FIG HERE
     \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final.pdf}
+    \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
     \caption{
@@
     \caption{
-       Inferred versus simulated DFEs from human genomes simulated with
-       a simple demographic model with one population with a constant size of $10,000$
-       and a gamma-distributed DFE acting on exons (see \textbf{Methods}).
-        DFE is inferred by three different methods: \grapes, \polydfe , and \dadicli.
-        Mean absolute value of selection coefficient ($\lvert E(s) \rvert$, panel A) and shape parameter (panel B)
-        are shown for each DFE inferred from all three simulated datase
-        with median values marked by horizontal bars
-        and simulated values represented by dashed horizontal lines.
+    \igcomment{Update fig and caption after we finalize Fig. 3}
+    Performance of methods to infer the distribution of fitness effects (DFE) from genetic data.
+    Simulations were performed using a human constant size model with a gamma-distributed DFE
+    acting on exons parameterized by a mean selection coefficient and a shape parameter. On each row we show the true parameter 
+    as a black line, and the estimates as boxplots from replicate simulations. The human model has three extant populations (CEU, CHB, and YRI).
+    %DRS-fig: only one population is shown here, but 3 referenced in the legend. Wrong legend or wrong fig?
+    %DRS-fig: If it is the right fig, get rid of the pop0 legend inset
+    Each row of the figure shows estimates from a different method: 1) GRAPES \citep{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe}, and 3) \dadicli \citep{Huang2023}.
     }
@@
     \label{fig:vaquita-dfe.constant}
-        Inferred versus simulated DFEs from vaquita genomes simulated with
-        a simple demographic model with one population with a constant size of $3,500$
-        and a gamma-distributed DFE acting on exons 
-        with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$) (see \textbf{Methods}).
-        DFE is inferred by three different methods: \grapes, \polydfe , and \dadicli.
-        Mean absolute value of selection coefficient ($\lvert E(s) \rvert$, panel A) and shape parameter (panel B)
-        are shown for each DFE inferred from all three simulated datase
-        with median values marked by horizontal bars
-        and simulated values represented by dashed horizontal lines.
-    }
+    A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
+    Simulations were performed using a model of vaquita porpoise genome under constant population size with a gamma-distributed DFE
+    acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
+    parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
+    This vaquita model has a single extant population, and parameter estimates from each
+    of the three different methods are shown: 1) GRAPES \citep{galtier2016adaptive}, 2) polyDFE \citep{tataru2020polydfe},
+    and 3) \dadicli \citep{Huang2023}.}
+    %DRS-fig: remove the pop 0 legend again
 \end{figure}
-
+% DRS-fig: ditch the unnecessary legends from panels A and B. Also, are we deciding which version of panel C to include here? I guess I prefer
+% to bottom one but with the legend of the top one (``Original s DFE'' sounds a little clunky).
+
+% constant all site size DFE figure
 \begin{figure}[h]
@@
     under constant population size  where all sites along the simulated chromosome can experience selected and neutral mutation
-    as opposed to exonic regions only.\\
-    \igcomment{This fig is not cited anywhere. Do we really need it?}
+    as opposed to exonic regions only.\igcomment{This figure is not cited anywhere. Do we really need it?}
     }
+    %DRS-fig: human? vaquita? need more info in this figure legend, and also get rid of the pop 0 thing again
 \end{figure}
+
+\begin{table}[ht]
+\centering
+\small
+\caption{\bf{Performance of DFE methods across the simulation scenarios}. 
+Bold rows show the lowest mean absolute error (MAE) for the two DFE parameters
+for each species.}
+\begin{tabular}{lllrrrrrr}
+\toprule
+species ID & demography & annotation & \makecell{MAE \\ $E|s|$ \\ dadi-cli} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi-cli} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
+\midrule
+HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
+HomSap & Constant & exons & \bf{0.0012} & 0.0101 & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
+HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
+HomSap & OOA 5R19 & exons & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
+PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
+PhoSin & Constant & exons & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
+PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
+PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
+\bottomrule
+\end{tabular}
+\label{tab:dfe_table}
+\end{table}
 
@@
 
-\begin{table}[ht]
-\centering
-\small
-
-\caption{\bf{Performance of DFE methods across the simulation scenarios}. 
-Bold rows show the lowest mean absolute error (MAE) for the two DFE parameters
-for each species.\\
-\igcomment{This table is not cited anywhere. The analysis of ``all sites" is also not mentioned anywhere. Do we want to add it?}
-}
-\begin{tabular}{lllrrrrrr}
-\toprule
-species ID & demography & annotation & \makecell{MAE \\ $E|s|$ \\ dadi-cli} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi-cli} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
-\midrule
-HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
-HomSap & Constant & exons & \bf{0.0012} & 0.0101 & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
-HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
-HomSap & OOA 5R19 & exons & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
-PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
-PhoSin & Constant & exons & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
-PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
-PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
-\bottomrule
-\end{tabular}
-\label{tab:dfe_table}
-\end{table}
-
 \stopsupplement

### 2025-03-04 analysis2_manuscript.tex
tweaks to sentence that described getting rescaled s for dadi

--- 2025-03-04 analysis2_manuscript.tex
@@
     $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
-    \igcomment{Do we really multiple $\alpha$ by $\nu$ in rescaling for \dadicli?}
-    For \dadicli, we report $\alpha * \nu / (\theta / (4 \mu L p))$, where
-    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution,
-    $L$ is the total number of sites available for the analysis,
+    For \dadicli, the gamma distribution is parameterized using the shape and rate ($\alpha$ and $\nu$, respectively). 
+    The mean value of the distribution, which is $s$ is transformed from $\alpha$ and $\nu$ and rescaled as $\alpha * \nu / (\theta / (4 \mu L p))$,
+    where $L$ is the total number of sites available for the analysis,
     $p$ is the proportion of sites under selection.

### 2025-03-04 analysis2_manuscript.tex
added missing and

@@
     where $L$ is the total number of sites available for the analysis,
-    $p$ is the proportion of sites under selection.
+    and $p$ is the proportion of sites under selection.
 

### 2025-03-06 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
 \kelcomment{also cite: PMID: 36617238 }
-    While this is so, joint simulation of complex demography and selection
-    is challenging,  requires a deep understanding of the underlying
-    evolutionary processes, and necessitates a number of parameter choices including
-    the strength of selection and the
-    recombination rate, as well as a parameterized demographic model.
+    However, joint simulation of complex demography and selection
+    is challenging, requiring a deep understanding of the underlying
+    evolutionary processes, and necessitating a number of parameter choices including
+    the strength of selection, the
+    recombination rate, and a parameterized demographic model.
     This is a daunting task for many researchers, and can be a barrier to
@@
     population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
-    While that is so, the original version of \stdpopsim did not include
+    Yet, the original version of \stdpopsim did not include
     models of selection, which is a major limitation for more empirically motivated
@@
     through simulation is critical for understanding processes such
-    as adaptation\citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
+    as adaptation \citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
     background selection on genetic diversity
@@
     to have a single, unified framework for simulating both neutral and
-    non-neutral evolutionary processes, and to be able to compare the
+    non-neutral evolutionary processes and to be able to compare the
     results of these simulations to empirical data in a manner that is
     both accessible to a wide range of researchers and highly reproducible. 
-    Furthermore the framework should include the complex realities of 
+    Furthermore, the framework should include the complex realities of 
     genomes, including heterogenous recombination rates, 
@@
     % goals
-    In this manuscript, we provide an overview of a major new addition
+    Here, we provide an overview of a major new addition
     to \stdpopsim---the inclusion of models of selection.
@@
     The \texttt{DFE} class provides an interface for specifying
-    a distribution of the fitness effects of new mutations.
+    a distribution of the fitness effects (DFE) of new mutations.
     Each \texttt{DFE} object specifies a list of mutation types and
@@
     Each \texttt{Annotation} object is typically associated with a collection of functional
-    genomic elements whose coordinates are specified using a GFF3 file. %
+    genomic elements whose coordinates are specified using a GFF3 annotation file. %
     %    
@@
     neutral and deleterious, corresponding to synonymous and non-synonimous sites.
-    with their proportions set to follow their expected
+    with their proportions set to their expected
     prevelanece in exons. 
@@
     \end{figure}
-
+    \kelcomment{For Figure 1 at the bottom, it says "Kyriazis et al. 2013". This should be "2023" instead}
     As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
@@
     which conditions on the introduction of a selected mutation at a given time and position
-    and with a given selection coefficient. Further the user can specify the minimum frequency
+    and with a given selection coefficient. Further, the user can specify the minimum frequency
     of the selected allele at the time of sampling. As these extended events are implemented
@@
     using a variety of methods, including the sequential Markovian coalescent
-    \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
-    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
-    \drscomment{One could use dadi for that, but then we should also cite things like fastsimcoal and ABC, while explaining
-    that users can use those methods to infer piecewise-constant N(t) histories as well with a prespecified number of epochs
-    and priors around their params, but maybe that is too much of a digression here? I'm fine either way}
+    \citep{li2011inference,Schiffels2020,terhorst2017robust},
+
+    %\kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
+    %\igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
+    %\drscomment{One could use dadi for that, but then we should also cite things like fastsimcoal and ABC, while explaining
+    %that users can use those methods to infer piecewise-constant N(t) histories as well with a prespecified number of epochs
+    %and priors around their params, but maybe that is too much of a digression here? I'm fine either way}
     through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
@@
     \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
-    Simulations with selection were implemented by modeling purifying selection on exons
+    Simulations with selection were implemented by modeling purifying selection on nonsynonymous mutations
     % IG: moved details to methods
@@
     Methods based on the sequential Markovian coalescent (\smcpp and \msmc)
-    appear to overal produce the most accurate trajectories of $N_e(t)$
+    appear to produce the most accurate trajectories of $N_e(t)$ overall
     in this setting, although they tend to under-estimate the population
@@
     Note that while the estimation accuracy for the DFE parameters themselves (Figure \ref{fig:homsap-dfe.ooa}A-B) is useful,
-    examining the difference between the actual and inferred distributions can be more helpful for showing how well a method
+    examining the difference between the actual and inferred distributions of $s$ can be more helpful for showing how well a method
     has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C).
@@
     the other methods do not. 
+    \kelcomment{From looking at the SI Figure, these differences seem really slight.
+Also, with only 3 replicates, I think it's hard to conclusively say that some methods overlap the true values while others do not.
+I might instead just say here that all methods are close, but deviate slightly in different ways.
     \grapes exhibits similar biases without population size changes as in the simulations with more complex demography,
@@
         Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
-        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$) (see \textbf{Methods}).
+        acting on nonsynonymous mutations with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$) (see \textbf{Methods}).
         The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
@@
     %IG: describe experiment and cite figure later
+    \kelcomment{Throughout this section, we talk about "mutations in exons".
+Is it more accurate to say "nonsynonymous mutations"?
+In other words, the simulations of exons include by synonymous and nonsynonymous mutations, but only the nonsynonymous ones experience selection and only nonsynonymous variants are used in DFE inference.
+Did I get that right?
+Maybe saying that clearly here would be good}
     %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
     Next, we benchmarked these three DFE-inference methods on simulations of the vaquita porpoise genome.
-    Recall that these simulations apply a gamma-distributed selection coefficient to mutations in exons
+    Recall that these simulations apply a gamma distribution of selection coefficients to nonsynonymous mutations 
     and a relationship between the selection and dominance coefficients (see \textbf{Methods}).

### 2025-03-06 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
     using each of the two null models separately (neutral and BGS) and controlling 
-    for $5\%$ false positive rate.
+    for a $5\%$ false positive rate.
     We then applied each method to data simulated with sweeps for two sampled populations (YRI and CEU)
@@
     One important finding we made using the new stdpopsim selection API is that inference of the DFE is sensitive to the dominance of deleterious mutations.
-    All three of the inference methods of the DFE applied here assume that deleterious mutations act in an additive fashion (h=0.5) by default.
+    All three of the inference methods of the DFE applied here assume that deleterious mutations act in an additive fashion ($h=0.5$) by default.
     However, the simulated DFE for the vaquita genome included highly recessive mutations.
-    Specifically, these simulations used a DFE where mutations with $s>0.01$ are quite recessive, with $h<=0.01$ and where mutations with $0.001<s<0.01$ have $h=0.1$.
+    Specifically, these simulations used a DFE where mutations with $s<-0.01$ are quite recessive, with $h<=0.01$ and where mutations with $-0.01<s<-0.001$ have $h=0.1$.
     We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Fig 4).
@@
     DRS was supported in part by NIH award R35GM138286.
+    KEL was supported by NIH award R35GM119856.
 

### 2025-03-10 analysis2_manuscript.tex
abstract

--- 2025-03-04 analysis2_manuscript.tex
@@
 \author[14,15]{Xin Huang}
+\author[5]{Kevin Korfmann}
 % \author[16]{Leonardo Nicola Martin Iasi}
@@
     However, simulating realistic models that incorporate both selection and complex demographic histories
-    is challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and explore theoretical predictions.
-    \stdpopsim is a tool for facilitating the simulation of a contiually expanding catalog of population genetic models.
+    is challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and to explore theoretical predictions.
+    \stdpopsim, a simulation library that provides an extensive catalog of population genetic models,
+    aims to provide a standardized interface for simulating realistic models that incorporate both selection and complex demographic histories.
     Here we present a major extension to the \stdpopsim{} framework that enables simulation of various modes
@@
     and show how genomic features like recombination rate and the density of functional sequence influence power to detect selective sweeps.
-    This extension to \stdpopsim{} provides a powerful new resource for the population genetics community
+    This extension to \stdpopsim{} provides a powerful, new resource for the population genetics community
     to explore the interplay between selection and other evolutionary forces in a reproducible framework.
@@
     variation in the size and density of functional elements, and
-    non-equilibrium demographic histories. Ideally, one should be able
+    non-equilibrium demographic histories. Moreover, one should be able
     to model all of these features simultaneously using estimates and annotations obtained
@@
     $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
-    For \dadicli, the gamma distribution is parameterized using the shape and rate ($\alpha$ and $\nu$, respectively). 
-    The mean value of the distribution, which is $s$ is transformed from $\alpha$ and $\nu$ and rescaled as $\alpha * \nu / (\theta / (4 \mu L p))$,
-    where $L$ is the total number of sites available for the analysis,
-    and $p$ is the proportion of sites under selection.
+    \igcomment{Do we really multiple $\alpha$ by $\nu$ in rescaling for \dadicli?}
+    For \dadicli, we report $\alpha * \nu / (\theta / (4 \mu L p))$, where
+    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution,
+    $L$ is the total number of sites available for the analysis,
+    $p$ is the proportion of sites under selection.
 

### 2025-03-10 analysis2_manuscript.tex
intro

@@
     style=authoryear,
+    uniquename=false,
     ]{biblatex}
@@
     Thus, it is imperative to jointly account for the effects of selection
-    and demography when inferring evolutionary history from genetic data \citep{sheehan2016deep,johri2020toward}.
-    However, this is a challenging task from a modeling perspective \citep{johri2022prospect}. % MFR this is a bit unclear
+    and demography when inferring evolutionary history from genetic data \citep[e.g.,][]{sheehan2016deep,johri2020toward},
+    yet standardized tools for doing so remain limited.
 
@@
     and forward-in-time simulations of complex demography and selection
-    \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.  \kelcomment{maybe also cite https://doi.org/10.1093/bioinformatics/btn522}
+    \citep[e.g.,][]{gillespie1984molecular, hernandez2008flexible, thornton2014c++, haller2019slim}. 
     The development of simulation tools has been driven by the need to
     understand the effects of complex evolutionary processes on genetic
-    variation \citep[e.g.,][]{galloway2020few},\kelcomment{also cite: PMID: 32487519 } \drscomment{and maybe a few others?
-PMIDs: 24102784 and thornton2019polygenic} to provide a null model for hypothesis testing
+    variation \citep[e.g.,][]{peischl2013accumulation, thornton2019polygenic, galloway2020few, zhang2020impact}, 
+     to provide a null model for hypothesis testing
     \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
@@
     and increasingly to provide a basis for machine learning and other
-    simulation-based inference methods \citep[e.g.,][]{beaumont2002approximate,pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}.
-\kelcomment{also cite: PMID: 36617238 }
+    simulation-based inference methods 
+    \citep[e.g.,][]{beaumont2002approximate,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic, zhang2023maladapt, smith2023dispersenn2}.
     While this is so, joint simulation of complex demography and selection
@@
     provides a standardized interface for accessing a wide range of
-    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
-    While that is so, the original version of \stdpopsim did not include
+    population genetic models, and has begun to be be widely adopted by the community 
+    \citep[e.g.,][]{speidel2021inferring, wang2021automatic, yang2022trustgwas, degiorgio2022spatially, browning2023fast, schweiger2023ultrafast, temple2024modeling, haag2025pandora}.
+    While that is so, previous releases of \stdpopsim did not include
     models of selection, which is a major limitation for more empirically motivated
@@
     through simulation is critical for understanding processes such
-    as adaptation\citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
-    background selection on genetic diversity
+    as adaptation\citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps 
+    \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, 
+    and the impact of background selection on genetic diversity
     \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.

### 2025-03-10 analysis2_manuscript.tex
up to Inference of Ne

@@
     species and models, and can use either
-    msprime \citep{Baumdicker2022} or SLiM \citep{haller2019slim}
+    \texttt{msprime} \citep{Baumdicker2022} or \texttt{SLiM} \citep{haller2019slim}
     as the backend engine for efficient population genetic simulation.
-    %IG: edited here to add details on the spec of the DFE and Annotation classes
-    %moved below the information on how they are used in simulation. Please check!
     To extend \stdpopsim's functionality to include models of selection, 
     we introduce two new classes
-    %an important new class representing a specfic distribution of fitness effects
     to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
@@
     and the parameters of this distribution.
-    Each \texttt{Annotation} object is typically associated with a collection of functional
-    genomic elements whose coordinates are specified using a GFF3 file. %
-    %    
-    For example, say that we wish to implement a model of selection consistent with the
+    Each \texttt{Annotation} object represents a collection of functional
+    genomic elements (e.g., exons, introns, intergenic regions, etc.)
+    whose coordinates are specified using a GFF3 file.      
+    More concretely, consider implementing a model of selection consistent with the
     DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
     The associated \texttt{DFE} object is defined with two mutation types,
-    neutral and deleterious, corresponding to synonymous and non-synonimous sites.
-    with their proportions set to follow their expected
-    prevelanece in exons. 
-    The neutral mutation type is associted with a fixed selection coefficient (0),
+    neutral and deleterious, corresponding to synonymous and non-synonymous sites.
+    with their proportions set to follow their expected prevalence in exons. 
+    The neutral mutation type is associated with a fixed selection coefficient (0),
     and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
@@
     As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
-    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publically available functional genomic elements 
+    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publicly available functional genomic elements 
     and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
@@
     humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
-    While the DFEs are from a limited number of sepecies, these DFEs can be applied to other species
+    While the DFEs are from a limited number of species, these DFEs can be applied to other species
     in the catalog if and when it is suitable. 
@@
     a published DFE included in the catalog. This flexibility allows for a wide range of
-    models of selection to be simulated. %, and for the results of these simulations to be
-    %compared to empirical data or used as a null model for hypothesis testing.
-    % DRS: commenting that stuff out because there are obviously other use cases that could be
-    % listed here (e.g. ML/ABC, and if we don't know enough about a species to do that then
-    % we probably have no business doing hypothesis testing either, which could be just as
-    % misleading or maybe even more so), so the simplest solution is to cut this as it is
-    % not needed here.
+    models of selection to be simulated.
 
@@
     using either the python API or the command-line interface (Figure \ref{fig:schematic}B-C).
-    %IG: commented this out because I think it's covered by the text I wrote above
-    %A \texttt{DFE} object can apply to specified intervals within a chromosome,
-    %and can be combined with other \texttt{DFE} objects 
     Simulations with selection are generated
-    using the SLiM simulation engine, which is
-    %is perhaps
+    using the \texttt{SLiM} simulation engine, which is perhaps
     the most flexible simulation engine for modeling selection available to date
-    (msprime's selection models are relatively limited and thus not currently used by \stdpopsim).
-    %IG: this is a nice statement, but I couldn't figure where to plug it after my changes so I'm leaving it commented
-    %all in an automated and user-friendly fashion.
+    (\texttt{msprime}'s selection models are relatively limited and thus not currently used by \stdpopsim).
 
@@
     on top of the existing \stdpopsim API, they can be combined with other models of selection
-    and demography to provide a rich model of the combined effects of multiple disparate evolutionary processes
+    and demography to provide a model of the combined effects of multiple, disparate evolutionary processes
     on genetic variation in a population.
@@
     % current numbers of DFEs / species / etc
-    % DRS: are we giving this a version number?
-    At this release the \stdpopsim catalog comprises 24 species (i.e. genome representations)
-    for which we have included 28 demographic models, 37 genetic maps, and 7 DFEs. %someone should double check the numbers
+    % ADK: check version number
+    At this release (version 0.3XX) the \stdpopsim catalog comprises 27 species (i.e. genome representations)
+    for which we have included 35 demographic models, 38 genetic maps, 12 DFEs, and 8 annotation objects. %ADK: checked this on 3/10/25
     Additions to the catalog are ongoing---we point the reader to our previously published 

### 2025-03-10 analysis2_manuscript.tex
through Ne

@@
     \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
-    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
-    \drscomment{One could use dadi for that, but then we should also cite things like fastsimcoal and ABC, while explaining
-    that users can use those methods to infer piecewise-constant N(t) histories as well with a prespecified number of epochs
-    and priors around their params, but maybe that is too much of a digression here? I'm fine either way}
+    %\igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
+    %\drscomment{One could use dadi for that, but then we should also cite things like fastsimcoal and ABC, while explaining
+    % that users can use those methods to infer piecewise-constant N(t) histories as well with a prespecified number of epochs
+    % and priors around their params, but maybe that is too much of a digression here? I'm fine either way}
+    % ADK: I think we are good as is
     through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
     as well as through identity by descent information \citep{santiago2020recent}.
-    Since all of these methods assume neutral evolution, they are typically applied to genomic
+    Since all of these methods assume neutrality, they are typically applied to genomic
     segments that are not expected to be directly affected by selection (e.g., by masking exons
@@
     analysis---can bias estimates of $N_e(t)$
-    away from true census population sizes by increasing the 
+    away from census population sizes by increasing the 
     rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
-    Using \stdpopsim, we can easily examine how different methods for inferring $N_e(t)$
+    Using \stdpopsim, we can readily examine how different methods for inferring $N_e(t)$
     are influenced by selection on linked sites,
@@
 
-    For this purpose, we simulated human genomes with and without natural selection.
+    To do this we simulated human genomes with and without natural selection.
     In both scenarios, we ran three replicate simulations
@@
     \ref{fig:human-demography}; blue) are fairly similar.
-    Thus, this particular parameterization of background selection does not
+    Thus, this particular parameterization of background selection
+    in combination with selection only at exonic sites in the out-of-Africa demographic model does not
     appear to notably bias the estimates produced by these methods.

### 2025-03-10 analysis2_manuscript.tex
up to discussion

@@
     the performance of different methods for inferring the DFE.
-    %IG: I don't think we need this intro.
-    %We performed simulations using the 
-    %human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.
-    %
-    A challenge when comparing different inferred and simulated distributions of the selection coefficient ($s$) is differing convention regarding its definition.
-    The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
+
+    A challenge when comparing different inferred and simulated distributions of the selection coefficient ($s$)
+    is differing conventions in the literature regarding its definition.
+    The simulator \texttt{SLiM} and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
     On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
@@
     %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
-    For comparison, here we have normalized all inference results to the convention of SLiM (see \textbf{Methods}).
+    For comparison, here we have normalized all inference results to the convention of \texttt{SLiM} (see \textbf{Methods}).
     % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?
@@
     contain outgroup populations or species that can be used to estimate divergence.
-    %IG: no need to map figures. Just describe results and cite figs
-    %Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
-    %while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
-    %In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
+
     Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
@@
 
-    %IG: describe experiment and cite figure later
-    %In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
     Next, we benchmarked these three DFE-inference methods on simulations of the vaquita porpoise genome.
     Recall that these simulations apply a gamma-distributed selection coefficient to mutations in exons
-    and a relationship between the selection and dominance coefficients (see \textbf{Methods}).
+    with a relationship between the selection and dominance coefficients (see \textbf{Methods}).
     In this setting, all methods perform uniformly worse than in the human genome simulations,
@@
     where their selective effect is $h s$ rather than $s$.
-    %KEL: I moved the last sentence from here to the Discussion and fleshed it out
 
@@
     using each of the two null models separately (neutral and BGS) and controlling 
-    for $5\%$ false positive rate.
-    We then applied each method to data simulated with sweeps for two sampled populations (YRI and CEU)
-    in each of the 100 locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
+    for a $5\%$ false positive rate.
+    We then applied each method to data simulated with sweeps for two samples (YRI and CEU)
+    in each of 100 locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
     First, we notice that power to detect sweeps is slightly lower when the test statistic threshold
@@
     thus both stronger selection and more variation to detect a sweep \citep[e.g.,][]{simonsen1995properties}.
-    Further, the CEU population experienced the out-of-Africa bottleneck
-    that may mute the signal of a sweep.  
+    Conversely, the CEU population experienced the out-of-Africa bottleneck that may mute the signal of a sweep.  
     Finally and most importantly, the power to detect sweeps varies considerably along the chromosome,
@@
     is characterized by a strong recent bottleneck followed by population expansion.
-    \drscomment{note: I am couching the language above a bit because we have previously shown that shic doesn't suffer in the sweep vs.
+    %\drscomment{note: I am couching the language above a bit because we have previously shown that shic doesn't suffer in the sweep vs.
 nonsweep question when doing other train-on-constant-test-on-nonequilib benchmarks.
@@
     \diploshic on the other hand shows a slight increase in power with recombination rates,
-    probably because it is designed to distinguish regions where a sweep happened from regions linked to a sweep,
-    and if the recombination rate is too low (or selection too strong) then the window examined by \diploshic may
+    presumably because it is designed to distinguish regions where a sweep happened from regions linked to a sweep,
+    and if the recombination rate is too low (or selection too strong) then the genomic window examined by \diploshic may
     not be large enough to exhibit enough of a recovery in patterns of diversity flanking the sweep.

### 2025-03-11 analysis2_manuscript.tex
through discussion

@@
     However, the simulated DFE for the vaquita genome included highly recessive mutations.
-    Specifically, these simulations used a DFE where mutations with $s>0.01$ are quite recessive, with $h<=0.01$ and where mutations with $0.001<s<0.01$ have $h=0.1$.
-    We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Fig 4).
+    Specifically, these simulations used a DFE where mutations with $s>0.01$ are quite recessive, 
+    with $h<=0.01$, and where mutations with $0.001<s<0.01$ have $h=0.1$.
+    We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}).
     This performance is not surprising as decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
@@
     Fitting a mis-specified DFE that assumes $h=0.5$ will then lead to underestimates of $E[s]$.
-    Indeed, the distribution of $hs$ is much more accurately estimated (Fig 4b).
-    Overall, these results are consistent with prior work showing that it is challenging to separately infer the DFE of $s$ along with $h$ (cite PMID: 24830675; PMID: 39302992; PMID: 34951958 )
-    and that a large proportion ($>5\%$) of recessive deleterious mutations can confound inferences of $s$ (PMID: 37074880).
-    Given the plethora of evidence for recessive deleterious mutations (cite PMID: 4630587, PMID: 39114967 , PMID: 30013096 , PMID: 21098719 ),
+    Indeed, the distribution of $hs$ is much more accurately estimated (Figure \ref{fig:vaquita-dfe}b).
+    Overall, these results are consistent with prior work showing that it is challenging to separately
+    infer the DFE of $s$ along with $h$ \citep{veeramah2014evidence, kyriazis2024constraining, balick2022overcoming}
+    and that a large proportion ($>5\%$) of recessive deleterious mutations can confound inferences of $s$ \citep{wade2023quantifying}.
+    Given the plethora of evidence for recessive deleterious mutations 
+    \citep{mukai1972mutation, agrawal2011inferences, huber2018gene, di2024revisiting,},
     these results suggest the need for care in DFE inference.
-    Kyriazis and Lohmueller (PMID: 39302992 ) inferred the DFE of $s$ for nonsynonymous mutations in humans when considering different values of $h$.
-    Intriguingly, additive models show a satisfactory fit to the SFS of nonsynonymous variants in humans (Kyriazis et al) and more complex models of dominance do not substantially improve the fit.
-    Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome, show a similar fit to the SFS and may be more biologically realistic than fully additive models,
+    Kyriazis and Lohmueller \citeyear{kyriazis2024constraining} inferred the DFE of $s$ for nonsynonymous mutations
+    in humans when considering different values of $h$.
+    Intriguingly, additive models showed a satisfactory fit to the SFS of nonsynonymous variants in humans
+    and more complex models of dominance did not substantially improve the fit.
+    Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome,
+    show a similar fit to the SFS and may be more biologically realistic than fully additive models,
     given the evidence for the presence of recessive deleterious mutations.
@@
 
-    Regarding sweep detection, the inclusion of selection models into \stdpopsim allowed us to easily compare the
+    Regarding sweep detection, the inclusion of selection models into \stdpopsim allowed us to readily compare the
     effectiveness of different methods on data reflecting the demographic history, recombination rate variation,
@@
     the impact of recombination rate variation, and underscore the importance of training methods on data with variable
-    recombination rates (e.g. following Schrider and Kern \citeyear{schrider2017soft}). In addition, our results
+    recombination rates (e.g., following Schrider and Kern \citeyear{schrider2017soft}). In addition, our results
     illustrate a strength of likelihood methods like \sweepfinder, which do not require a predefined window size and
@@
     of the actual values of these diversity measures---on very strong sweeps that effect diversity across a wide stretch of
-    chromosomes that extends far beyond \diploshic's window size. \drscomment{Maybe this isn't the coolest stuff ever but
-    I thought it might be worth saying something about our sweep detection stuff in the Discussion.}
+    chromosomes that extends far beyond \diploshic's window size.
 
@@
     models, and for benchmarking methods for the analysis of genetic data
-    on common ground
-    %, using the same simulations and parameterizations of models of selection
-    % DRS: this sentence got pretty bloated with my edits so I commented part of it out
-    in a way that was not possible before.
+    on common ground in a way that was not possible before.
 

### 2025-03-11 analysis2_manuscript.tex
data availability; authors con't

@@
 \newcommand{\tskit}{\texttt{tskit}\xspace}
+\newcommand{\slim}{\texttt{SLiM}\xspace}
+\newcommand{\msprime}{\texttt{msprime}\xspace}
 
@@
 % \author[9]{Jean Cury}
-% \author[10]{Ignacio Echevarria}
-% \author[11]{Benjamin C. Haller}
-% \author[12,13]{Ahmed R. Hasan}
+\author[10]{Peter D. Fields}
+\author[11]{Benjamin C. Haller}
 \author[14,15]{Xin Huang}
+\author[30]{Ben Jeffrey}
 \author[5]{Kevin Korfmann}
+\author[5]{Jiseon Min}
 % \author[16]{Leonardo Nicola Martin Iasi}
@@
 % \author[29]{Anthony Wilder Wohns}
-% \author[30]{Yan Wong}
+\author[30]{Yan Wong}
 % \author[31]{Franz Baumdicker}
-% \author[32]{Reed A. Cartwright}
-% \author[33]{Gregor Gorjanc}
-\author[4]{Kirk E. Lohmueller}
+%\author[32]{Reed A. Cartwright}
+\author[33]{Gregor Gorjanc}
+\author[28]{Christian D. Huber}
+
 \author[34]{Ryan N. Gutenkunst}
 \author[30]{Jerome Kelleher}
+\author[4]{Kirk E. Lohmueller}
 \author[35]{Aaron P. Ragsdale}
@@
 % \affil[9]{\small{Université Paris-Saclay, CNRS, INRIA, Laboratoire Interdisciplinaire des Sciences du Numérique, UMR 9015 Orsay, France}}
-% \affil[10]{\small{School of Life Sciences, University of Glasgow, Glasgow, UK}}
-% \affil[11]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
+\affil[10]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
+\affil[11]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
 % \affil[12]{\small{Department of Cell and Systems Biology, University of Toronto, Toronto ON, Canada}}
@@
 % \affil[27]{\small{Department of Integrative Biology, University of California, Berkeley, Berkeley CA, USA}}
-% \affil[28]{\small{Department of Biostatistics, University of Washington, Seattle WA, USA}}
+\affil[28]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
 % \affil[29]{\small{Broad Institute of MIT and Harvard, Cambridge MA 02142, USA}}
@@
 % \affil[32]{\small{School of Life Sciences and The Biodesign Institute, Arizona State University, Tempe AZ, USA}}
-% \affil[33]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
-\affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson AZ 85721, USA}}
+\affil[33]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
+\affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
 \affil[35]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
@@
     species and models, and can use either
-    \texttt{msprime} \citep{Baumdicker2022} or \texttt{SLiM} \citep{haller2019slim}
+    \msprime \citep{Baumdicker2022} or \slim \citep{haller2019slim}
     as the backend engine for efficient population genetic simulation.
@@
     Simulations with selection are generated
-    using the \texttt{SLiM} simulation engine, which is perhaps
+    using the \slim simulation engine, which is perhaps
     the most flexible simulation engine for modeling selection available to date
-    (\texttt{msprime}'s selection models are relatively limited and thus not currently used by \stdpopsim).
+    (\msprime's selection models are relatively limited and thus not currently used by \stdpopsim).
 
@@
     is differing conventions in the literature regarding its definition.
-    The simulator \texttt{SLiM} and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
+    The simulator \slim and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
     On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
@@
     %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
-    For comparison, here we have normalized all inference results to the convention of \texttt{SLiM} (see \textbf{Methods}).
+    For comparison, here we have normalized all inference results to the convention of \slim (see \textbf{Methods}).
     % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?
@@
     using the \stdpopsim{} framework.
-    All simulations were executed using the SLiM engine with a burn-in of $2N$ generations
+    All simulations were executed using the \slim engine with a burn-in of $2N$ generations
     (where $N$ is the initial population size), and a scaling factor of 2 (to reduce running time).
@@
     (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
-    Each simulation generate 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
+    Each simulation generated 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
     with each genome consisting of 22 autosomes.
@@
     As described above, each DFE-inference method scales selection coefficients slightly differently.
-    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by SLiM.
+    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
     The output of \grapes  was scaled by a factor of  $\mu / \theta$,
@@
     For each of the 100 genomic segments of chromosome 1, we generated 1,000 replicate datasets under each of the three settings.
-    As with the whole-genome simulations (see above), we used the SLiM engine with a burn-in of $2N$ generations
+    As with the whole-genome simulations (see above), we used the \slim engine with a burn-in of $2N$ generations
     (where $N$ is the initial population size), and a scaling factor of 2.
@@
 \section*{Data availability}\label{data_availability}
+The release of \stdpopsim is available at \url{https://github.com/popsim-consortium/stdpopsim}.
+This paper coincides with version 0.3.XX of \stdpopsim. %be sure to check this
+The entire analysis pipeline associated with this paper is available
+as a snakemake workflow at \url{https://github.com/popsim-consortium/analysis2}.
 

### 2025-03-12 analysis2_manuscript.tex
added DFE method details

@@
     Can someone write this up?}
+    % ADK: I've added a section below that describes the DFE inference methods used in this study.
 
@@
     % may in reality be under selection. As above, all methods were run using default settings.
-    \igcomment{I edited the description below written by Silas. Please check to make sure it's still accurate.}
+    The DFE inference using \dadicli was conducted by first inferring a demographic model from neutral sites,
+    using a two epoch demographic model with 100 repeated runs of the optimization procedure.
+    \dadicli then proceeds by using a cached strategy wherein we used 2000 points in the $gamma$ distribution.
+    Finally, the DFE was inferred using a ratio of non-synonymous to synonymous mutations of 2.31, again with 
+    100 repeated runs of the optimization procedure.
+    For \polydfe, we used model 'C' which uses a mixture of a gamma distribution and an exponential distribution,
+    where the gamma component is used for negatively selected mutations. For optimization we used
+    the default BFGS optimizer with the '-e' flag to automatically estimate the initial parameters values.
+    For \grapes, we inferred the DFE under the `GammaZero' model, which assumes that selection coefficients
+    are drawn from a single gamma distribution. We used the `-no-div-param' flag to disable the use
+    of divergence-based parameter estimation. All other settings were left at their default values.
+    All methods used the same exon annotations as generated in the simulations. The total number of coding sites
+    in the human and vaquita genomes used were 61,644,801 and 31,039513 respectively.
+
+    % ADK: I've added a section below that describes the DFE inference methods used in this study.
+    For comparison among methods, estimates were rescaled. 
     As described above, each DFE-inference method scales selection coefficients slightly differently.
@@
     $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
-    \igcomment{Do we really multiple $\alpha$ by $\nu$ in rescaling for \dadicli?}
     For \dadicli, we report $\alpha * \nu / (\theta / (4 \mu L p))$, where

### 2025-03-12 analysis2_manuscript.tex
one more DFE method edit
more methods

@@
     In this study we performed simulations of the human and vaquita porpoise genomes
-    using the \stdpopsim{} framework.
-    All simulations were executed using the \slim engine with a burn-in of $2N$ generations
-    (where $N$ is the initial population size), and a scaling factor of 2 (to reduce running time).
-    Each simulation was implemented using a relatively short python script, similar to the one shown in Figure 1B
+    using the \stdpopsim{} framework. We ran replicate datasets for each genome/parameter combination,
+    in which chromosomes were simulated separately from one another.
+    All simulations were executed using the \slim engine with a burn-in of $10N$ generations
+    (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
+    All simulations were implemented using as part of a snakemake workflow,
+    where the code utilized is similar to the one shown in Figure 1B.
+    The complete simulation workflow is available at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/simulation.snake}
     (see \textbf{Data availablity} below).
-    \igcomment{We should add a link to the analysis2 repo in data/code availablity section}
-    %https://github.com/popsim-consortium/analysis2/tree/main
-    % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
-    % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
-    % IG: added a note about this above.
 
@@
     Each of these simulations generated 100 diploid genomes.
+    The configuration file used in combination with our snakemake workflow
+    for these simulations can be found at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_HomSap.yml}.
 
@@
     \igcomment{If we end up using Fig. S3, then we should add a line describing it's simulation here.}
+    The configuration file used in combination with our snakemake workflow
+    for these simulations can be found at 
+    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_PhoSin.yml}.
 
     \subsection*{Inference of $N_e(t)$}
-    %We assessed the accuracy of demographic inferences from
-    %\msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
-    Inference of $N_e(t)$ was conducted on simulated datasets after masking non-recombining portion
+    Inference of $N_e(t)$ was conducted on simulated datasets after masking non-recombining portions
     of the genome (to avoid bias caused by extreme recombination rate variation)
@@
     acting on exons, but it was influenced by selection acting on linked sites (background selection).
-    %IG: I commented because it doesn't seem relevant for methods
-    %Comparisons between inferences made with and without selection were 
-    %of particular interest.
+
+    We assessed the accuracy of demographic inferences from three software packages:
+    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent},
+     and \smcpp \citep{terhorst2017robust}.
     For \msmc \citep{Schiffels2020}, we used a random sample of $N=6$ simulated genomes with 20
@@
     the number of generations to 2,000, and the number of bins to 400, all representing default settings.
-    For \stairway \citep{liu2020stairway} and \smcpp \citep{terhorst2017robust} we used the default settings.
-    \igcomment{Nothing more to say about \stairway plot and \smcpp?}
+    For \stairway \citep{liu2020stairway} we did one run of theta estimation for each replicate dataset,
+    and used a `dimFactor' of 5000, both of which represent default settings.
+    For \smcpp \citep{terhorst2017robust} we used the unfolded SFS (`--unfold'), 
+    along with the true mutation rate under which the simulations were generated. 
+    The complete snakemake workflow for demographic inference can be found at
+    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/n_t.snake}.
 
     \subsection*{DFE inference}
-    \igcomment{I commented all the text that was here because it wasn't relevant to the methods. We might want to add it to the discussion.
-    Kirk also noted that a lot of methods-level details are missing:
-    how big was the coding sequence length?
-    How was demographic inference done for the different models via the different programs?,
-    etc.
-    Can someone write this up?}
+    %\igcomment{I commented all the text that was here because it wasn't relevant to the methods. We might want to add it to the discussion.
+    %Kirk also noted that a lot of methods-level details are missing:
+    %how big was the coding sequence length?
+    %How was demographic inference done for the different models via the different programs?,
+    %etc.
+    %Can someone write this up?
     % ADK: I've added a section below that describes the DFE inference methods used in this study.
@@
     $p$ is the proportion of sites under selection. 
+    Complete details for the DFE inference analysis can be found in the
+    snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.
 
@@
     \subsection*{Selective sweeps detection}
-    %\drscomment{Maybe it makes sense to move this up to right after the sweep simulation section. Should make it easier for readers to follow.}
-    %IG: I moved that section down here to bring them together
 
@@
     over 10 equally sized and non-overlapping subwindows.
-    \sweepfinder \citep{degiorgio2016sweepfinder2} is based a composite likelihood ratio (CLR) 
+    \sweepfinder \citep{degiorgio2016sweepfinder2} is based on a composite likelihood ratio (CLR) 
     computed at 21 uniformly distributed positions along the simulated region.
@@
     For \texttt{discoal} simulations that included selective sweeps,
-    selection coefficients were drawn from a uniform distribution between 0.001 and 0.10,
-    and the time of the sweep was chosen from a uniform distribution between 0 and $0.01N_e$ generations in the past.
+    selection coefficients were drawn from a uniform distribution between $0.001$ and $0.10$,
+    and the time of the sweep was chosen from a uniform distribution between $0$ and $0.01N_e$ generations in the past.
     For soft sweeps, the frequency at which selection was introduced was drawn from a uniform distribution between $0.0001$ and $0.1$.
@@
     Then, power (or true positive rate) to detect a sweep by a given method was determined for
-    100 evenly-spaced positions on chromosome 1 as follows.
-    The sweep-detection method was applied to the 1,000 
+    100 evenly-spaced positions on chromosome 1 as follows:
+    each sweep-detection method was applied to the 1,000 
     simulated datasets generated with a hard sweep (neutral and BGS) for that position,
@@
     in which the test statistic exceeded the threshold determined by each of the two null models (neutral and BGS).
+    We note that using an empirical distribution to determine the threshold for calling a selective
+    sweep with \diploshic deviates from the original implementation of the method,
+    however for consistency among methods we chose to take this approach.
+    The complete snakemake workflow for sweep detection can be found at
+    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/sweep_simulate.snake}.
 

### 2025-03-12 analysis2_manuscript.tex
small stuff

@@
 
+    Inference of the DFE was done here using exon annotations of the genome in question.
+    A key distinction between our analyses and those done in practice
+    is that we have complete knowledge of which sites are under selection in the simulation,
+    thus our inferences represent a best case scenario. Empirical inference of the DFE is a much
+    harder problem as some mutations assumed to be free of selection
+    may in reality be under selection
+
     One important finding we made using the new stdpopsim selection API is that inference of the DFE is sensitive to the dominance of deleterious mutations.
@@
     \subsection*{DFE inference}
-    %\igcomment{I commented all the text that was here because it wasn't relevant to the methods. We might want to add it to the discussion.
-    %Kirk also noted that a lot of methods-level details are missing:
-    %how big was the coding sequence length?
-    %How was demographic inference done for the different models via the different programs?,
-    %etc.
-    %Can someone write this up?
-    % ADK: I've added a section below that describes the DFE inference methods used in this study.
-
-    % We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
-    % \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
-    % Each software package was run using the same set of simulations.
-    % For these analyses we used the exon annotations from the human and vaquita porpoise genomes respectively.
-    % A key distinction between our analyses and those done in practice
-    % is that we have complete knowledge of which sites are under selection in the simulation,
-    % thus our inferences is a best case scenario. Empirical inference of the DFE is a much
-    % harder problem as some mutations assumed to be free of selection
-    % may in reality be under selection. As above, all methods were run using default settings.
+    % ADK: I've tried to deal with all the comments here    
+    We used three software packages to infer the distribution of fitness effects (DFE) from genetic data:
+    \dadicli \citep{Huang2023,kim2017inference}, \polydfe \citep{tataru2020polydfe}, and \grapes \citep{galtier2016adaptive}.
+    Each software package was run using the same set of simulations.
     The DFE inference using \dadicli was conducted by first inferring a demographic model from neutral sites,
@@
     Thus, for each test statistic we determined two thresholds corresponding to these two different null models (neutral and BGS)
-    using the extreme $5\%$ quantile in each empirical distribution (and excluding regions with extremely low recombination rate).
-    \igcomment{Can we say what the rec rate cutoff was and how many regions were excluded here?
-    Shouldn't we have excluded these regions from the simulations altogether?}.
+    using the extreme $5\%$ quantile in each empirical distribution.
+    Simulated regions where the central 1Mb had a recombination rate of less than $10^{-3}cM$ were excluded from the analysis.
+    This excluded 7 regions from the analysis. 
     Then, power (or true positive rate) to detect a sweep by a given method was determined for

### 2025-03-12 analysis2_manuscript.tex
added pi sanity checks

@@
     using a variety of methods, including the sequential Markovian coalescent
-    \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
+    \citep{li2011inference,Schiffels2020,terhorst2017robust}, 
     %\igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
@@
     using a DFE inferred by \cite{kim2017inference}.
+    A basic summary of nucleotide diversity from these simulations is shown in Figure \ref{fig:homsap-summary-stats},
+    which shows that the presence of selection decreases polymorphism in simulations with selection only slightly
+    at the genome-wide level, but more substantially within exonic regions under both a constant population size
+    and the out-of-Africa demographic model.
     Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
@@
     additive ($h=0.4$; see \textbf{Methods}).
-    Here, we see that the site-frequency-based method (\stairway) produces more accurate
+    As above, a basic summary of nucleotide diversity from these simulations is shown in Figure \ref{fig:vaquita-summary-stats},
+    which again shows that the presence of selection decreases polymorphism in simulations with selection, but that the effect
+    is only readily apparent within exonic regions. Further in this case,
+    we see that the two-epoch demographic model used in the simulations actually leads to a slight increase
+    in diversity versus the constant population size model.
+
+    In comparing demography inference methods, we see that the site-frequency-based method (\stairway) produces more accurate
     inference than the two methods based on the sequential Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
@@
 
-    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
+    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}, Table \ref{tab:dfe_table}).
     Note that while the estimation accuracy for the DFE parameters themselves (Figure \ref{fig:homsap-dfe.ooa}A-B) is useful,
@@
     with overestimated shape parameter and slightly underestimated mean. 
+    As a sanity check, we also inferred the DFE from human genome simulations with a constant population size 
+    (Figure \ref{fig:homsap-dfe.constant}, Table \ref{tab:dfe_table}), but largely the results are similar.
 
@@
     with consistent underestimation of the mean selection coefficient    
-    and overestimation of the shape parameter (Figure \ref{fig:vaquita-dfe}A-B). 
+    and overestimation of the shape parameter (Figure \ref{fig:vaquita-dfe}A-B, Table \ref{tab:dfe_table}). 
     Similar patterns are observed when re-running the simulations with a simpler demographic model
-    without population size changes (Figure \ref{fig:vaquita-dfe.constant}).
+    without population size changes (Figure \ref{fig:vaquita-dfe.constant}, Table \ref{tab:dfe_table}).
     We suspect that this is due to the fact that all DFE inference methods used here assume an additive model ($h = \frac{1}{2}$),
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figures/HomSap/summary_stats.png}
+    \caption{
+    \label{fig:homsap-summary-stats}
+    Summaries of diversity for human simulations. The left and right panels show
+    nucleotide diversity ($\pi$) calculated over all three replicate datasets
+    for the Constant size and Out-of-Africa demographic models respectively.
+    Each panel shows four violins representing the distribution of $\pi$ values
+    under a model without selection (blue), a model with background selection (orange),
+    calculated from either the entire genome (labeled neutral or DFE)
+    or from exonic regions only (labeled exons-only).
+    }
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figures/PhoSin/summary_stats.png}
+    \caption{
+    \label{fig:vaquita-summary-stats}
+    Summaries of diversity for vaquita simulations. The left and right panels show
+    nucleotide diversity ($\pi$) calculated over all three replicate datasets
+    for the Constant size and two-epoch demographic models respectively.
+    Each panel shows four violins representing the distribution of $\pi$ values
+    under a model without selection (blue), a model with background selection (orange),
+    calculated from either the entire genome (labeled neutral or DFE)
+    or from exonic regions only (labeled exons-only).
+    }
+\end{figure}
+
 \begin{figure}[b!]
@@
     as opposed to exonic regions only.\\
-    \igcomment{This fig is not cited anywhere. Do we really need it?}
     }
@@
 Bold rows show the lowest mean absolute error (MAE) for the two DFE parameters
-for each species.\\
-\igcomment{This table is not cited anywhere. The analysis of ``all sites" is also not mentioned anywhere. Do we want to add it?}
-}
+for each species.}
+
 \begin{tabular}{lllrrrrrr}
@@
 \midrule
-HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
+%HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
 HomSap & Constant & exons & \bf{0.0012} & 0.0101 & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
-HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
+%HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
 HomSap & OOA 5R19 & exons & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
-PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
+%PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
 PhoSin & Constant & exons & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
-PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
+%PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
 PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\

### 2025-03-12 analysis2_manuscript.tex
added discussion stuff

@@
 \newcommand{\msprime}{\texttt{msprime}\xspace}
+\newcommand{\snakemake}{\texttt{snakemake}\xspace}
 
@@
     % ADK: check version number
-    At this release (version 0.3XX) the \stdpopsim catalog comprises 27 species (i.e. genome representations)
+    At this release (version 0.3) the \stdpopsim catalog comprises 27 species (i.e. genome representations)
     for which we have included 35 demographic models, 38 genetic maps, 12 DFEs, and 8 annotation objects. %ADK: checked this on 3/10/25
@@
     Simulations with selection were implemented by modeling purifying selection on exons
-    % IG: moved details to methods
-    %, defined from the HAVANA group release 104 for the human genome,
     using a DFE inferred by \cite{kim2017inference}.
@@
     appear to notably bias the estimates produced by these methods.
+    In light of Figure \ref{fig:human-demography}, this makes sense-- 
+    the effect of selection is only apparent within exonic regions and presumably sites
+    very closely linked, thus masking of exons should be sufficient to remove the effects of selection
+    from our simulations. 
     As noted in previous studies that compared demography inference methods
@@
     Methods based on the sequential Markovian coalescent (\smcpp and \msmc)
-    appear to overal produce the most accurate trajectories of $N_e(t)$
+    appear overall to produce the most accurate trajectories of $N_e(t)$
     in this setting, although they tend to under-estimate the population
@@
 
-    In comparing demography inference methods, we see that the site-frequency-based method (\stairway) produces more accurate
-    inference than the two methods based on the sequential Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
-    Reagrdless of these differences, we see that background selection again does not
+    In comparing demography inference methods, we see that the SFS-based method (\stairway) produces more accurate
+    inference than the two methods based on the sequential Markovian coalescent 
+    (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
+    Regardless of these differences, we see that background selection again does not
     appear to considerably influence the inference of any of the four methods for this species,
-    which has a very different evolutionary history than humans.
+    at least for the parameterization used here, 
+    even though vaquita has a very different evolutionary history than humans.
 
@@
         \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
-        \drscomment{I think it would be good to increase the font size for the A,B,C panel legends and also
-        draw a horizontal line above panel C so it is clear that all three plots at the bottom are part of C.}
         \caption{Inferred versus simulated DFEs from simulated human genomes.
@@
         average recombination rates instead of position along chromosome 1.
-        \igcomment{Specify how the curves were computed}
-        %IG: copoed from the caption of fig 5
     	Genomic segments were simulated with sweeps under a three population out-of-Africa model
@@
         and a null model with background selection from deleterious mutations in exons (red).
+        Fitted lines represent loess smoothed regressions.
         }
@@
 
-    Inference of the DFE was done here using exon annotations of the genome in question.
-    A key distinction between our analyses and those done in practice
-    is that we have complete knowledge of which sites are under selection in the simulation,
-    thus our inferences represent a best case scenario. Empirical inference of the DFE is a much
-    harder problem as some mutations assumed to be free of selection
-    may in reality be under selection
-
-    One important finding we made using the new stdpopsim selection API is that inference of the DFE is sensitive to the dominance of deleterious mutations.
-    All three of the inference methods of the DFE applied here assume that deleterious mutations act in an additive fashion (h=0.5) by default.
+    The exploration of performance among methods done here is not meant to be thorough 
+    benchmarking, but instead meant to highlight the utility of the \stdpopsim resource. 
+    Having said that, there are a number of interesting results that bear discussion. 
+    The first pertains to estimation of demographic history ($N_e(t)$) in the context of
+    linked purifying selection. Our genomic simulations of humans and vaquita that incorporate
+    selection model published DFEs that operate at exonic positions of the genome only. 
+    As shown in Figures \ref{fig:homsap-summary-stats} and \ref{fig:vaquita-summary-stats},
+    this combination only produces sizeable decreases in polymorphism at exonic positions, 
+    which represent a small minority of sites within the genome, and does not lead to large-scale
+    perturbations in genome-wide diversity. Thus it is not surprising that estimates of $N_e(t)$
+    from genome-wide data from which exons have been masked should be accurate 
+    (e.g., Figures \ref{fig:human-demography} and \ref{fig:vaquita-demography}). 
+    However the question remains to what extent this would be true in empirical data where
+    selection is not constrained to coding positions. For instance, we know
+    that many genomes harbor large numbers
+    of conserved non-coding elements \citep[e.g.,]{siepel2005evolutionarily} with extremely
+    strong selection acting on them \citep{katzman2007human,mclean2008dispensability},
+    thus we caution extrapolating from our simulation results. Future studies might build upon
+    the annotations we used here to include selection at other portions of the genome as well
+    as mixtures of DFEs across sites.
+
+    A second important finding we made using the new stdpopsim selection API is that inference of the DFE
+    is sensitive to the dominance of deleterious mutations.
+    All three of the inference methods of the DFE applied here assume that deleterious mutations act
+    in an additive fashion (h=0.5) by default.
     However, the simulated DFE for the vaquita genome included highly recessive mutations.
@@
 
+    Further, inference of the DFE was done here using exon annotations of the genome in question.
+    A key distinction between our analyses and those done in practice
+    is that we have complete knowledge of which sites are under selection in the simulation,
+    thus our inferences represent a best case scenario. Empirical inference of the DFE is a much
+    harder problem as some mutations assumed to be free of selection
+    may in reality be under selection.
+
     Regarding sweep detection, the inclusion of selection models into \stdpopsim allowed us to readily compare the
@@
     the impact of recombination rate variation, and underscore the importance of training methods on data with variable
-    recombination rates (e.g., following Schrider and Kern \citeyear{schrider2017soft}). In addition, our results
-    illustrate a strength of likelihood methods like \sweepfinder, which do not require a predefined window size and
-    thus will perform better than \diploshic---which only examines relative summarie of diversity across a window instead
-    of the actual values of these diversity measures---on very strong sweeps that effect diversity across a wide stretch of
-    chromosomes that extends far beyond \diploshic's window size.
+    recombination rates \citep[e.g.,][]{schrider2017soft}. In addition, our results
+    illustrate a strength of likelihood methods like \sweepfinder, that do not require a predefined window size.
+    Window-based methods, such as \diploshic, tend to lag in performance when the perturbation due to a sweep is at 
+    a much greater or smaller scale than the chosen window size, as can occur with very strong sweeps or low levels of 
+    recombination. Moreover while all the methods examined here retain adequate power for sweep identification across
+    scenarios, including the inclusion of background selection, our results underscore the importance of developing 
+    methods that are well powered in the face of unknown demographic histories. Indeed, the largest differences
+    in power we observed across simulated scenarios were between demographic histories and not the inclusion of background
+    selection.
 
@@
     (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
-    All simulations were implemented using as part of a snakemake workflow,
+    All simulations were implemented using as part of a \snakemake workflow,
     where the code utilized is similar to the one shown in Figure 1B.
@@
     Each of these simulations generated 100 diploid genomes.
-    The configuration file used in combination with our snakemake workflow
+    The configuration file used in combination with our \snakemake workflow
     for these simulations can be found at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_HomSap.yml}.
@@
     \igcomment{If we end up using Fig. S3, then we should add a line describing it's simulation here.}
-    The configuration file used in combination with our snakemake workflow
+    The configuration file used in combination with our \snakemake workflow
     for these simulations can be found at 
@@
     along with the true mutation rate under which the simulations were generated. 
-    The complete snakemake workflow for demographic inference can be found at
+    The complete \snakemake workflow for demographic inference can be found at
     \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/n_t.snake}.
@@
     Complete details for the DFE inference analysis can be found in the
-    snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.
+    \snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.
 
@@
     however for consistency among methods we chose to take this approach.
-    The complete snakemake workflow for sweep detection can be found at
+    The complete \snakemake workflow for sweep detection can be found at
     \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/sweep_simulate.snake}.
@@
 The entire analysis pipeline associated with this paper is available
-as a snakemake workflow at \url{https://github.com/popsim-consortium/analysis2}.
+as a \snakemake workflow at \url{https://github.com/popsim-consortium/analysis2}.
 
@@
         but with the 100 genomic segments sorted according to their
-        exon density instead of position along chromosome 1.
-        \igcomment{Specify exactly how exon density is computed (\% of bases in exons?) and how the curves were computed}
-        %IG: copoed from the caption of fig 5
+        exon density (percentage of exonic bases per window)instead of position along chromosome 1.
 	    Genomic segments were simulated with sweeps under a three population out-of-Africa model
@@
         and a null model with background selection from deleterious mutations in exons (red).
+        Fitted lines represent loess smoothed regressions.
     }

### 2025-03-12 analysis2_manuscript.tex
author list clean up

@@
 
-% \author[1,+]{M. Elise Lauterbur}
-% \author[3,*]{Ariella L. Gladstein}
-\author[4,*]{Graham Gower}
-\author[5*]{Nathaniel S. Pope}
-\author[5*]{Murillo F. Rodrigues}
-\author[5*]{Silas Tittes}
-\author[34*]{Linh N. Tran}
-\author[6]{Georgia Tsambos}
-
-\author[2]{Maria Izabel A. Cavassim}
-% \author[5,7]{Jeff Adrion}
-% \author[5]{Saurabh Belsare}
-% \author[8]{Arjun Biddanda}
-% \author[5]{Victoria Caudill}
-% \author[9]{Jean Cury}
-\author[10]{Peter D. Fields}
-\author[11]{Benjamin C. Haller}
-\author[14,15]{Xin Huang}
-\author[30]{Ben Jeffrey}
-\author[5]{Kevin Korfmann}
-\author[5]{Jiseon Min}
-% \author[16]{Leonardo Nicola Martin Iasi}
-% \author[17]{Ekaterina Noskova}
-% \author[18]{Jana Obšteter}
-% \author[19]{Vitor Antonio Corrêa Pavinato}
-% \author[20,21]{Alice Pearson}
-% \author[22,23]{David Peede}
-% \author[24]{Manolo F. Perez}
-\author[5]{Chris C. R. Smith}
-% \author[25]{Jeffrey P. Spence}
-% \author[5]{Anastasia Teterina}
-
-\author[5]{Scott T. Small}
-\author[5]{Clara T. Rehmann}
-% \author[26]{Per Unneberg}
-% \author[27]{Juan Manuel Vazquez}
-% \author[28]{Ryan K. Waples}
-% \author[29]{Anthony Wilder Wohns}
-\author[30]{Yan Wong}
-% \author[31]{Franz Baumdicker}
-%\author[32]{Reed A. Cartwright}
-\author[33]{Gregor Gorjanc}
-\author[28]{Christian D. Huber}
-
-\author[34]{Ryan N. Gutenkunst}
-\author[30]{Jerome Kelleher}
-\author[4]{Kirk E. Lohmueller}
-\author[35]{Aaron P. Ragsdale}
-
-\author[37]{Daniel R. Schrider}
-\author[38]{Ilan Gronau}
-\author[5,36]{Peter L. Ralph}
-\author[5]{Andrew D. Kern}
+\author[1,*]{Graham Gower}
+\author[2*]{Nathaniel S. Pope}
+\author[2*]{Murillo F. Rodrigues}
+\author[2*]{Silas Tittes}
+\author[3*]{Linh N. Tran}
+
+\author[4]{Ornob Alam}
+\author[5]{Maria Izabel A. Cavassim}
+\author[6]{Peter D. Fields}
+\author[7]{Benjamin C. Haller}
+\author[8,9]{Xin Huang}
+\author[10]{Ben Jeffrey}
+\author[5]{Christopher C. Kyriazis}
+\author[2]{Kevin Korfmann}
+\author[2]{Jiseon Min}
+\author[2]{Chris C. R. Smith}
+
+\author[2]{Scott T. Small}
+\author[2]{Clara T. Rehmann}
+\author[11]{Inés Rebollo}
+\author[10]{Yan Wong}
+\author[12]{Gregor Gorjanc}
+\author[13]{Christian D. Huber}
+\author[]{Yu Zhang}
+
+\author[3]{Ryan N. Gutenkunst}
+\author[10]{Jerome Kelleher}
+\author[5]{Kirk E. Lohmueller}
+\author[14]{Aaron P. Ragsdale}
+\author[15]{Daniel R. Schrider}
+\author[16]{Ilan Gronau}
+\author[2,17]{Peter L. Ralph}
+\author[2]{Andrew D. Kern}
 
 \affil[*]{\small{These authors contributed equally to the paper.}}
-% \affil[+]{\small{Corresponding authors: lauterbur@gmail.com ; ilan.gronau@runi.ac.il.}}
-% \affil[1]{\small{Department of Ecology and Evolutionary Biology, University of Arizona, Tucson AZ 85719, USA}}
-% \affil[2]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
-% \affil[3]{\small{Embark Veterinary, Inc., Boston MA 02111, USA}}
-\affil[4]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
-\affil[5]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
-% \affil[6]{\small{School of Mathematics and Statistics, University of Melbourne, Australia}}
-% \affil[7]{\small{AncestryDNA, San Francisco CA 94107, USA}}
-% \affil[8]{\small{54Gene, Inc., Washington DC 20005, USA}}
-% \affil[9]{\small{Université Paris-Saclay, CNRS, INRIA, Laboratoire Interdisciplinaire des Sciences du Numérique, UMR 9015 Orsay, France}}
-\affil[10]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
-\affil[11]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
-% \affil[12]{\small{Department of Cell and Systems Biology, University of Toronto, Toronto ON, Canada}}
-% \affil[13]{\small{Department of Biology, University of Toronto Mississauga, Mississauga ON, Canada}}
-\affil[14]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
-\affil[15]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
-% \affil[16]{\small{Department of Evolutionary Genetics, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany}}
-% \affil[17]{\small{Computer Technologies Laboratory, ITMO University, St Petersburg, Russia}}
-% \affil[18]{\small{Agricultural Institute of Slovenia, Department of Animal Science, Ljubljana, Slovenia}}
-% \affil[19]{\small{Entomology Department, The Ohio State University, Wooster OH, USA}}
-% \affil[20]{\small{Department of Genetics, University of Cambridge, Cambridge, UK}}
-% \affil[21]{\small{Department of Zoology, University of Cambridge, Cambridge, UK}}
-% \affil[22]{\small{Department of Ecology, Evolution, and Organismal Biology, Brown University, Providence RI, USA}}
-% \affil[23]{\small{Center for Computational Molecular Biology, Brown University, Providence RI, USA}}
-% \affil[24]{\small{Department of Genetics and Evolution, Federal University of Sao Carlos, Sao Carlos 13565905, Brazil}}
-% \affil[25]{\small{Department of Genetics, Stanford University School of Medicine, Stanford CA 94305, USA}}
-% \affil[26]{\small{Department of Cell and Molecular Biology, National Bioinformatics Infrastructure Sweden, Science for Life Laboratory, Uppsala University, Husargatan 3, SE-752 37 Uppsala, Sweden}}
-% \affil[27]{\small{Department of Integrative Biology, University of California, Berkeley, Berkeley CA, USA}}
-\affil[28]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
-% \affil[29]{\small{Broad Institute of MIT and Harvard, Cambridge MA 02142, USA}}
-\affil[30]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
-% \affil[31]{\small{Cluster of Excellence - Controlling Microbes to Fight Infections, Eberhard Karls Universität Tübingen, Tübingen, Baden-Württemberg, Germany}}
-% \affil[32]{\small{School of Life Sciences and The Biodesign Institute, Arizona State University, Tempe AZ, USA}}
-\affil[33]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
-\affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
-\affil[35]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
-\affil[36]{\small{Department of Mathematics, University of Oregon, Eugene OR 97402, USA}}
-\affil[37]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
-\affil[38]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
+\affil[1]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
+\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
+\affil[3]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
+\affil[4]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
+\affil[5]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
+\affil[6]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
+\affil[7]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
+\affil[8]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
+\affil[9]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
+\affil[10]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
+\affil[11]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
+\affil[12]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
+\affil[13]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
+\affil[14]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
+\affil[15]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
+\affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
+\affil[17]{\small{Department of Mathematics, University of Oregon, Eugene OR 97402, USA}}
 

### 2025-03-12 analysis2_manuscript.tex
applied silas's diff

@@
     $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
-    For \dadicli, we report $\alpha * \nu / (\theta / (4 \mu L p))$, where
-    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution,
-    $L$ is the total number of sites available for the analysis,
-    $p$ is the proportion of sites under selection. 
+    For \dadicli, the gamma distribution is parameterized using the shape and rate ($\alpha$ and $\nu$, respectively). 
+    The mean value of the distribution, which is $s$ is transformed from $\alpha$ and $\nu$ and rescaled as $\alpha * \nu / (\theta / (4 \mu L p))$,
+    where $L$ is the total number of sites available for the analysis,
+    and $p$ is the proportion of sites under selection.
     Complete details for the DFE inference analysis can be found in the

### 2025-03-12 analysis2_manuscript.tex
applied kirk's changes

@@
     \citep[e.g.,][]{beaumont2002approximate,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic, zhang2023maladapt, smith2023dispersenn2}.
-    While this is so, joint simulation of complex demography and selection
-    is challenging,  requires a deep understanding of the underlying
-    evolutionary processes, and necessitates a number of parameter choices including
-    the strength of selection and the
-    recombination rate, as well as a parameterized demographic model.
+    However, joint simulation of complex demography and selection
+    is challenging, requiring a deep understanding of the underlying
+    evolutionary processes, and necessitating a number of parameter choices including
+    the strength of selection, the
+    recombination rate, and a parameterized demographic model.
     This is a daunting task for many researchers, and can be a barrier to
@@
     \citep[e.g.,][]{speidel2021inferring, wang2021automatic, yang2022trustgwas, degiorgio2022spatially, browning2023fast, schweiger2023ultrafast, temple2024modeling, haag2025pandora}.
-    While that is so, previous releases of \stdpopsim did not include
+    Yet, previous releases of \stdpopsim did not include
     models of selection, which is a major limitation for more empirically motivated
@@
     through simulation is critical for understanding processes such
-    as adaptation\citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps 
+    as adaptation \citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps 
     \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, 
@@
     The \texttt{DFE} class provides an interface for specifying
-    a distribution of the fitness effects of new mutations.
+    a distribution of the fitness effects (DFE)of new mutations.
     Each \texttt{DFE} object specifies a list of mutation types and
@@
     genomic elements (e.g., exons, introns, intergenic regions, etc.)
-    whose coordinates are specified using a GFF3 file.      
+    whose coordinates are specified using a GFF3 annotationfile.      
     More concretely, consider implementing a model of selection consistent with the
@@
     neutral and deleterious, corresponding to synonymous and non-synonymous sites.
-    with their proportions set to follow their expected prevalence in exons. 
+    with their proportions set to their expected prevalence in exons. 
     The neutral mutation type is associated with a fixed selection coefficient (0),
@@
     Methods based on the sequential Markovian coalescent (\smcpp and \msmc)
-    appear overall to produce the most accurate trajectories of $N_e(t)$
+    appear to produce the most accurate trajectories of $N_e(t)$ overall
     in this setting, although they tend to under-estimate the population
@@
     All three of the inference methods of the DFE applied here assume that deleterious mutations act
-    in an additive fashion (h=0.5) by default.
+    in an additive fashion ($h=0.5$) by default.
     However, the simulated DFE for the vaquita genome included highly recessive mutations.
-    Specifically, these simulations used a DFE where mutations with $s>0.01$ are quite recessive, 
-    with $h<=0.01$, and where mutations with $0.001<s<0.01$ have $h=0.1$.
+    Specifically, these simulations used a DFE where mutations with $s<-0.01$ are quite recessive, 
+    with $h<=0.01$, and where mutations with $-0.001<s<-0.01$ have $h=0.1$.
     We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}).
@@
     DRS was supported in part by NIH award R35GM138286.
-
+    KEL was supported by NIH award R35GM119856.
     %other funding sources?

### 2025-03-12 analysis2_manuscript.tex
Adk edit 3.10 (#56)
* abstract

* intro

* up to Inference of Ne

* through Ne

* up to discussion

* through discussion

* data availability; authors con't

* added DFE method details

* one more DFE method edit

* more methods

* small stuff

* added pi sanity checks

* fig edit from silas

* added discussion stuff

* author list clean up

* applied silas's diff

* applied kirk's changes

@@
     The \texttt{DFE} class provides an interface for specifying
-    a distribution of the fitness effects (DFE)of new mutations.
+    a distribution of the fitness effects (DFE) of new mutations.
     Each \texttt{DFE} object specifies a list of mutation types and
@@
     \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
-    Simulations with selection were implemented by modeling purifying selection on exons
+    Simulations with selection were implemented by modeling purifying selection on nonsynonymous mutations
     using a DFE inferred by \cite{kim2017inference}.
@@
     for a $5\%$ false positive rate.
+
     We then applied each method to data simulated with sweeps for two samples (YRI and CEU)
@@
     KEL was supported by NIH award R35GM119856.
+
     %other funding sources?

### 2025-03-12 analysis2_manuscript.tex
edited schematic per Kirk's comment

--- 2025-03-12 analysis2_manuscript.tex
@@
     To factor out the influence of demographic changes, we also analyzed genomes generated using similar
-    simulations but with no population size changes (Figure \ref{fig:homsap-dfe.constant}).
+    simulations but with no population size changes (Figure \ref{fig:homsap-dfe.constant}, Table \ref{tab:dfe_table}).
     In this simpler setting, \polydfe overlapped the true mean selection coefficient and shape parameter, whereas 
-    the other methods do not. 
+    the other methods do not show as close a correspondence.
     \grapes exhibits similar biases without population size changes as in the simulations with more complex demography,
@@
     with overestimated shape parameter and slightly underestimated mean. 
-    As a sanity check, we also inferred the DFE from human genome simulations with a constant population size 
-    (Figure \ref{fig:homsap-dfe.constant}, Table \ref{tab:dfe_table}), but largely the results are similar.
+    While that is so, it is hard to be definitive as our sample size is small with 3 simulated replicates.
 

### 2025-03-16 analysis2_manuscript.tex
edits through disco

@@
 
-\title{Accesible, realistic genome simulation with selection using \stdpopsim}
+\title{Accessible, realistic genome simulation with selection using \stdpopsim}
 
@@
 \affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
-\affil[17]{\small{Department of Mathematics, University of Oregon, Eugene OR 97402, USA}}
+\affil[17]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
 
@@
     Natural selection is a fundamental evolutionary force that shapes patterns of genetic variation across species. 
-    However, simulating realistic models that incorporate both selection and complex demographic histories
+    However, simulations incorporating realistic selection along heterogeneous genomes in complex demographic histories
     is challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and to explore theoretical predictions.
-    \stdpopsim, a simulation library that provides an extensive catalog of population genetic models,
-    aims to provide a standardized interface for simulating realistic models that incorporate both selection and complex demographic histories.
+    \stdpopsim is a community-maintained simulation library that already provides an extensive catalog of
+    species-specific population genetic models,
     Here we present a major extension to the \stdpopsim{} framework that enables simulation of various modes
-    of natural selection, including background selection, selective sweeps, and arbitrary distributions of fitness effects (DFE).
-    This extension maintains \stdpopsim's core principles of reproducibility and standardization while adding support
+    of natural selection, including background selection, selective sweeps,
+    and arbitrary distributions of fitness effects (DFE) acting on annotated subsets of the genome (for instance, exons).
+    This extension maintains \stdpopsim's core principles of reproducibility and accessibility while adding support
     for species-specific genomic annotations and published DFE estimates. 
     We demonstrate the utility of this framework by benchmarking methods for demographic inference,
-    DFE estimation, and selective sweep detection across different species and scenarios. 
+    DFE estimation, and selective sweep detection across several species and scenarios. 
     Our results demonstrate the robustness of demographic inference methods to selection on linked sites, 
@@
     This extension to \stdpopsim{} provides a powerful, new resource for the population genetics community
-    to explore the interplay between selection and other evolutionary forces in a reproducible framework.
+    to explore the interplay between selection and other evolutionary forces in a reproducible, low-barrier framework.
 
@@
     and demography when inferring evolutionary history from genetic data \citep[e.g.,][]{sheehan2016deep,johri2020toward},
-    yet standardized tools for doing so remain limited.
+    yet tools for doing so remain limited.
 
@@
     \citep[e.g.,][]{beaumont2002approximate,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic, zhang2023maladapt, smith2023dispersenn2}.
-    However, joint simulation of complex demography and selection
-    is challenging, requiring a deep understanding of the underlying
-    evolutionary processes, and necessitating a number of parameter choices including
-    the strength of selection, the
-    recombination rate, and a parameterized demographic model.
-    This is a daunting task for many researchers, and can be a barrier to
+    However, joint simulation of complex demography and selection on realistic genomes is challenging,
+    requiring many choices including the strength and rate of selection,
+    and a parameterized demographic model,
+    as well as implementation of genetic maps and other genomic features related to the strength of selection.
+    This is a daunting task, and can be a barrier to
     the adoption of simulation-based methods in population genetics.
@@
     Ideally, one would like
-    to have a single, unified framework for simulating both neutral and
+    to have a simple, unified framework for simulating both neutral and
     non-neutral evolutionary processes and to be able to compare the
@@
 
+    \stdpopsim{} aims to provide such a tool,
+    thus lowering the barrier to production of complex, multifaceted simulations,
+    and allowing a wider range of researchers to make use of high-quality simulations
+    in developing or testing their work.
+    The project also has the goal of doing open, collaborative, and inclusive work,
+    viewing capacity-building across the community as equal in importance to software development.
+
     % goals
@@
     models of the distribution of fitness effects (DFE).
-    Further we describe how these models can be combined with genomic
-    annotations that are specific to a given species available
+    Further we describe how these can be applied not only genome-wide but to
+    genomic regions with a particular annotation, specific to a given species available
     through the \stdpopsim API, and combined with a parameterized model of
@@
     their proportions, where 
-    each mutation type is associated with a dominance coefficient
+    each mutation type is associated with a dominance coefficient,
     (currently limited to values between 0 and 1, inclusive),
@@
     There are relatively few species with published DFE estimates in the literature.
-    \stdpopsim currently has implemented DFEs for four species in the catalogue:
-    \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
-    humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
+    \stdpopsim currently has implemented DFEs for five species in the catalogue:
+    %% MISSING SOME CITATIONS IF WE"RE AIMING TO HAVE ALL OF THEM IN HERE, I THINK?
+    \textit{Arabidopsis thaleana} \citep{huber2018gene},
+    \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
+    humans \citep{huber2017determining,kim2017inference},
+    \textit{Mus musculus} \citep{booker2021selective},
+    and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
     While the DFEs are from a limited number of species, these DFEs can be applied to other species
@@
     estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
+    Indeed, the DFE labeled \texttt{HomSap/Mixed_K23} is recommended by \citet{kyriazis2022using}
+    for ``general'' use across species (lacking more specific information).
     Furthermore, the user can specify a custom DFE and provide their own annotations
     of functional elements to simulate selection in a species for which we do not yet have 
-    a published DFE included in the catalog. This flexibility allows for a wide range of
-    models of selection to be simulated.
+    a published DFE included in the catalog.
 
@@
     % sweep interface
-    In addition to the \texttt{DFE} class, we have also implemented a class (\texttt{stdpopsim.ext.selective\_sweep})
+    In addition to the \texttt{DFE} class, we have also implemented a class (\texttt{stdpopsim.selective\_sweep})
     that enables selective sweeps to be simulated in \stdpopsim.
-    This class augments a demographic model with an ``extended event''
+    This class augments a demographic model with an event
     which conditions on the introduction of a selected mutation at a given time and position
     and with a given selection coefficient. Further, the user can specify the minimum frequency
-    of the selected allele at the time of sampling. As these extended events are implemented
+    of the selected allele at the time of sampling. As this functionality is implemented
     on top of the existing \stdpopsim API, they can be combined with other models of selection
@@
     Simulations with selection were implemented by modeling purifying selection on nonsynonymous mutations
-    using a DFE inferred by \cite{kim2017inference}.
+    using a DFE inferred by \citet{kim2017inference}.
     A basic summary of nucleotide diversity from these simulations is shown in Figure \ref{fig:homsap-summary-stats},
@@
     from our simulations. 
-    As noted in previous studies that compared demography inference methods
+    As noted in previous studies that compared demographic inference methods
     on neutral simulations \citep{adrion2020community},
@@
     annotated from the vaquita genome assembly.
-    This DFE model implements a relationship between selection and dominance 
+    This DFE implements a relationship between selection and dominance 
     in which very deleterious mutations (with selection coefficient $s<-0.1$) are fully recessive
@@
     which again shows that the presence of selection decreases polymorphism in simulations with selection, but that the effect
-    is only readily apparent within exonic regions. Further in this case,
-    we see that the two-epoch demographic model used in the simulations actually leads to a slight increase
+    is only readily apparent within exonic regions.
+    We also see that the two-epoch demographic model used in the simulations actually leads to a slight increase
     in diversity versus the constant population size model.
 
-    In comparing demography inference methods, we see that the SFS-based method (\stairway) produces more accurate
+    For these simulations, the SFS-based method (\stairway) produces more accurate
     inference than the two methods based on the sequential Markovian coalescent 
@@
     %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
-    For comparison, here we have normalized all inference results to the convention of \slim (see \textbf{Methods}).
+    For comparison, here we have normalized all inference results to the convention of \slim
+    (also used by \stdpopsim; see \textbf{Methods}).
     % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?
@@
     (relative to the other two populations), whereas \polydfe
-    produces its most accurate estimate from the CHB genomes.
+    shows very similar accuracy across all populations.
+    % produces its most accurate estimate from the CHB genomes.
     To factor out the influence of demographic changes, we also analyzed genomes generated using similar
@@
     Each segment was simulated 1000 times, while keeping only replicates in which the selected allele reached a frequency of 0.95 or greater.
-    These sweeps were modeled in the context of the three population out-of-Africa model from \cite{gutenkunst2009inferring}
+    These sweeps were modeled in the context of the three population out-of-Africa model from \citet{gutenkunst2009inferring}
     and background selection from deleterious mutations in exons.
@@
     and 3) reduced diversity ($\pi$).
-    For each method, we determined a threshold for its test-statistic 
-    using each of the two null models separately (neutral and BGS) and controlling 
-    for a $5\%$ false positive rate.
+    For each method, we determined a threshold for its test statistic 
+    using each of the two null models separately -- neutral and with background selection (BGS)
+    -- and controlling for a $5\%$ false positive rate.
 
@@
     in each of 100 locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
-    First, we notice that power to detect sweeps is slightly lower when the test statistic threshold
+    First and most strikingly, the power to detect sweeps varies considerably along the chromosome,
+    and this variation is much greater than the variation in power between CEU and YRI or between
+    null models. This is likely due to the effects of recombination rate heterogeniety (Figure \ref{fig:chr1_power}).
+    Next, we notice that power to detect sweeps is slightly lower when the test statistic threshold
     is determined using simulated data with background selection (BGS null model).
@@
     sweep detection methods we have used are relatively robust to the effects of background selection. 
-    Second, the power to detect sweeps is generally higher in the YRI sample
+    Finally, the power to detect sweeps is generally higher in the YRI sample
     than in the CEU sample, which is consistent with the fact that YRI has a larger ancestral effective population size and
@@
     Conversely, the CEU population experienced the out-of-Africa bottleneck that may mute the signal of a sweep.
-    Finally and most importantly, the power to detect sweeps varies considerably along the chromosome,
-    and this variation is much greater than the variation in power between CEU and YRI or between
-    null models. This is likely due to the effects of recombination rate heterogeniety (Figure \ref{fig:chr1_power}).% as well
-    %as the density of functional elements along the chromosome.
-    %DRS: cut the density of func elements because that stuff doesn't seem to matter much here
 
@@
     when applied to both sample sets.
-    In that sense, \sweepfinder appears to have the least variation in power along the chromosome or across populations.
+    \sweepfinder appears to have the least variation in power along the chromosome or across populations.
 
@@
     To further dissect the reasons for high variance along the chromosome in power to detect sweeps,
-    we examined the relationship between power and local recombination rate for our sweep simulations ( Figure \ref{fig:power-recomb}).
+    we examined the relationship between power and local recombination rate (Figure \ref{fig:power-recomb}).
     We see that power to detect sweeps is
@@
     of fitness effects of new mutations, and the detection of selective sweeps.
-    In general, this is an important step forward for the field of population genetics,
-    as it allows for a wide range of models of selection to be easily simulated and compared to
+    This allows for a wide range of models of selection to be easily simulated and compared to
     empirical data, with reproducibility, computational efficiency, and rigor.
@@
     The exploration of performance among methods done here is not meant to be thorough 
-    benchmarking, but instead meant to highlight the utility of the \stdpopsim resource. 
+    benchmarking, but rather to highlight the utility of the \stdpopsim resource. 
     Having said that, there are a number of interesting results that bear discussion. 
@@
     that many genomes harbor large numbers
-    of conserved non-coding elements \citep[e.g.,]{siepel2005evolutionarily} with extremely
+    of conserved non-coding elements \citep[e.g.,][]{siepel2005evolutionarily} with extremely
     strong selection acting on them \citep{katzman2007human,mclean2008dispensability},
@@
 
-    A second important finding we made using the new stdpopsim selection API is that inference of the DFE
+    A second important finding is that inference of the DFE
     is sensitive to the dominance of deleterious mutations.
@@
     \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
-    the out of Africa (OOA) demographic model with archaic admixture from
-    \cite{ragsdale2019models} (\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}).
+    the out of Africa demographic model with archaic admixture from
+    \citet{ragsdale2019models} (\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}).
     Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
-    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
-    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
+    \citet{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
+    (\citet{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
     Each simulation generated 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
@@
     Simulation of vaquita porpoise genomes used 
-    the two-epoch demographic model from \cite{robinson2022critically} (\stdpopsim label \texttt{Vaquita2Epoch\_1R22}).
+    the two-epoch demographic model from \citet{robinson2022critically} (\stdpopsim label \texttt{Vaquita2Epoch\_1R22}).
     Since no genetic map is currently available for this species, we assumed a constant recombination rate of  $10^{-8}$
@@
     Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
-    \cite{robinson2022critically} (\stdpopsim label \texttt{Gamma\_R22}) to exons annotated from the vaquita genome assembly
-    (\cite{morin2021}; \stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}).
+    \citet{robinson2022critically} (\stdpopsim label \texttt{Gamma\_R22}) to exons annotated from the vaquita genome assembly
+    (\citet{morin2021}; \stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}).
     This DFE model implements a relationship between selection and dominance as follows:
@@
     and very mild mutations (with $s\in [-0.001,0)$) were set to have a nearly additive effect ($h=0.4$).
-    This approach is similar to what was used in the simulations cunducted by \textcite{robinson2022critically}
+    This approach is similar to what was used in the simulations conducted by \textcite{robinson2022critically}
     and is easily implemented in \stdpopsim's \texttt{DFE} class.

### 2025-03-16 analysis2_manuscript.tex
plr edits to rest of paper

@@
     Furthermore, the framework should include the complex realities of 
-    genomes, including heterogenous recombination rates, 
+    genomes, including heterogeneous recombination rates, 
     variation in the size and density of functional elements, and
@@
     As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
-    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publicly available functional genomic elements 
+    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on species' publicly available functional genomic elements 
     and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
@@
     \stdpopsim currently has implemented DFEs for five species in the catalogue:
-    %% MISSING SOME CITATIONS IF WE"RE AIMING TO HAVE ALL OF THEM IN HERE, I THINK?
-    \textit{Arabidopsis thaleana} \citep{huber2018gene},
-    \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
-    humans \citep{huber2017determining,kim2017inference},
+    \textit{Arabidopsis thaliana} \citep{huber2018gene},
+    \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining,zhen2021greater},
+    humans \citep{huber2017determining,kim2017inference,kyriazis2022using,castellano2019compmarison,castellano2019compmarison,rodrigues2024shared},
     \textit{Mus musculus} \citep{booker2021selective},
@@
     estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
-    Indeed, the DFE labeled \texttt{HomSap/Mixed_K23} is recommended by \citet{kyriazis2022using}
+    Indeed, the DFE labeled \texttt{HomSap/Mixed\_K23} is recommended by \citet{kyriazis2022using}
     for ``general'' use across species (lacking more specific information).
@@
     and this variation is much greater than the variation in power between CEU and YRI or between
-    null models. This is likely due to the effects of recombination rate heterogeniety (Figure \ref{fig:chr1_power}).
+    null models. This is likely due to the effects of recombination rate heterogeneity (Figure \ref{fig:chr1_power}).
     Next, we notice that power to detect sweeps is slightly lower when the test statistic threshold
@@
         This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
-        but with the 100 genomic segments sorted according to their
+        but with the 100 genomic segments plotted against their
         average recombination rates instead of position along chromosome 1.
@@
         samples (left and right respectively).
-        The thresholds of the test statistics were set to control for
+        The thresholds of the test statistics were set to obtain a
         $5\%$ false positive rate under a neutral null model (blue)
@@
 
+    \subsection*{Annotations}
+    Currently available annotation data in \stdpopsim were downloaded from Ensembl.
+    Coding sequence and exon annotations are available for
+    \textit{Arabidopsis thaliana} \citep{cheng2017araport},
+    \textit{Drosophila melanogaster} \citep{hoskins2015release},
+    \textit{Homo sapiens} \citep{hunt2018ensembl},
+    and \textit{Phocoena sinus} \citep[vaquita porpoise,][]{morin2021}.
+
+    \subsection*{Quality control and software development}
+    Following the protocol established previously for other aspects of \stdpopsim{} \citep{adrion2020community},
+    new DFEs are subjet to a rigorous quality control (QC) process,
+    by which a separate individual re-implements the same DFE, and the two implementations
+    are programatically tested for equality.
+    The QC process identified a substantial bug in one DFE:
+    the \textit{Drosophila melanogaster} DFE \texttt{LognormalPlusPositive\_R16} \citet{booker2021selective}.
+    Due to an error in coding the log-normal distribution, deleterious selection coefficients
+    were several orders of magnitude too large.
+    The bug affected results in a separate project from one of our groups
+    that made use of this DFE before QC had been done.
+    All changes and additions to the code must pass a comprehensive set of unit tests:
+    at time of writing, the library includes over 11,000 lines of testing code,
+    that run 2,862 unit tests, including comparisons to QC implementations
+    and cross-checks for agreement, including between \slim and \msprime implementations of the same models.
+    For instance, for testing we make use of \slim's scriptability to separately record in simulation metadata
+    information about current population sizes, DFEs, etcetera,
+    which are then programmatically compared to the models used to generate the \slim script
+    that ran the simulation.
+    We make use of automatic test ``code coverage'' tools to help ensure that all portions of the code base
+    are validated by these unit tests.
+
     \subsection*{Whole genome simulations}
     In this study we performed simulations of the human and vaquita porpoise genomes
-    using the \stdpopsim{} framework. We ran replicate datasets for each genome/parameter combination,
+    using \stdpopsim{}. We ran replicate datasets for each genome/parameter combination,
     in which chromosomes were simulated separately from one another.
@@
     (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
-    All simulations were implemented using as part of a \snakemake workflow,
-    where the code utilized is similar to the one shown in Figure 1B.
+    All simulations were implemented using a \snakemake workflow,
+    where the code utilized is similar to that shown in Figure~\ref{fig:schematic}B.
     The complete simulation workflow is available at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/simulation.snake}
-    (see \textbf{Data availablity} below).
+    (see \textbf{Data availability} below).
 
@@
     Simulation of human genomes used a genetic map from the HapMap Project
-    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
+    \citep[\stdpopsim label \texttt{HapMapII\_GRCh38}]{international2007second}  and
     the out of Africa demographic model with archaic admixture from
-    \citet{ragsdale2019models} (\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}).
+    \citet[\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}]{ragsdale2019models}.
     Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
-    \citet{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
-    (\citet{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
-    Each simulation generated 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
+    \citet[\stdpopsim label \texttt{Gamma\_K17}]{kim2017inference}  to exons annotated by the HAVANA group release 104
+    \citep[\stdpopsim label \texttt{ensembl\_havana\_104\_exons}]{ensembl2018}.
+    Each simulation generated 100 diploid genomes from each of three extant human populations (labeled YRI, CEU, and CHB),
     with each genome consisting of 22 autosomes.
@@
     Simulation of vaquita porpoise genomes used 
-    the two-epoch demographic model from \citet{robinson2022critically} (\stdpopsim label \texttt{Vaquita2Epoch\_1R22}).
+    the two-epoch demographic model from \citet[\stdpopsim label \texttt{Vaquita2Epoch\_1R22}]{robinson2022critically}.
     Since no genetic map is currently available for this species, we assumed a constant recombination rate of  $10^{-8}$
@@
     Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
-    \citet{robinson2022critically} (\stdpopsim label \texttt{Gamma\_R22}) to exons annotated from the vaquita genome assembly
-    (\citet{morin2021}; \stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}).
+    \citet[\stdpopsim label \texttt{Gamma\_R22}]{robinson2022critically}  to exons annotated from the vaquita genome assembly
+    \citet[\stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}]{morin2021}.
     This DFE model implements a relationship between selection and dominance as follows:
-    very deleterious mutations (with $s<-0.1$) were set to be recessive ($h=0$),
-    mutations with $s\in [-0.1,-0.01)$ were set to have a dominance coefficient of $h=0.01$,
-    mutations with $s\in [-0.01,-0.001)$ were set to have a dominance coefficient of $h=0.1$,
-    and very mild mutations (with $s\in [-0.001,0)$) were set to have a nearly additive effect ($h=0.4$).
-    This approach is similar to what was used in the simulations conducted by \textcite{robinson2022critically}
+    very deleterious mutations (with $s<-0.1$) are recessive ($h=0$),
+    mutations with $s\in [-0.1,-0.01)$ have a dominance coefficient of $h=0.01$,
+    mutations with $s\in [-0.01,-0.001)$ have a dominance coefficient of $h=0.1$,
+    and very mild mutations (with $s\in [-0.001,0)$) are nearly additive ($h=0.4$).
+    This discretized relationship matches that used by \textcite{robinson2022critically},
     and is easily implemented in \stdpopsim's \texttt{DFE} class.
-    Each simulation generate 100 diploid genomes,
+    Each simulation generates 100 diploid genomes,
     with each genome consisting of 21 autosomes.
@@
     The output of \grapes  was scaled by a factor of  $\mu / \theta$,
-    and the output of \polydfe was saled by $2 \mu / \theta$, 
+    and the output of \polydfe was scaled by $2 \mu / \theta$, 
     where $\mu$ is the assumed mutation rate and 
@@
     we expect it to perform suboptimally on the simulated segments of human chromosome 1,
-    which were generated using more relaistic simulations (see \textbf{Simulations with selective sweeps} above).
+    which were generated using more realistic simulations (see \textbf{Simulations with selective sweeps} above).
 
@@
     using the extreme $5\%$ quantile in each empirical distribution.
-    Simulated regions where the central 1Mb had a recombination rate of less than $10^{-3}cM$ were excluded from the analysis.
+    Simulated regions where the central 1Mb had a recombination rate of less than $10^{-3}$cM were excluded from the analysis.
     This excluded 7 regions from the analysis. 
@@
     \label{funding}
-    MFR, ST, NP, and ADK were supported in part by NIH award R35GM148253.
-    RNG and LNT were supported in part by NIH award R35GM149235.
-    DRS was supported in part by NIH award R35GM138286.
-    KEL was supported by NIH award R35GM119856.
-
-    %other funding sources?
+    % we got too many people to do initials
+    M.~Rodgrigues, S.~Tittes, N.~Pope, and A.~Kern were supported in part by NIH award R35GM148253.
+    R.~Gutenkunst and L.~Tran were supported in part by NIH award R35GM149235.
+    D.~Schrider was supported in part by NIH award R35GM138286.
+    K.~Lohmueller was supported by NIH award R35GM119856.
+    P.~Ralph, B.~Haller, A.~Kern, and B.~Jefferey
+    were supported in part by NIH award R01HG012473.
+    J.~Kelleher was supported by EPSRC grant EP/X024881/1
+    and the Robertson Foundation.
+
 \printbibliography

### 2025-03-17 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
     All changes and additions to the code must pass a comprehensive set of unit tests:
-    at time of writing, the library includes over 11,000 lines of testing code,
+    at the time of writing, the library includes over 11,000 lines of testing code,
     that run 2,862 unit tests, including comparisons to QC implementations

### 2025-03-17 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
     and cross-checks for agreement, including between \slim and \msprime implementations of the same models.
-    For instance, for testing we make use of \slim's scriptability to separately record in simulation metadata
+    For instance, for testing we make use of \slim's scriptability to separately record in-simulation metadata
     information about current population sizes, DFEs, etcetera,

### 2025-03-17 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
     models of the distribution of fitness effects (DFE).
-    Further we describe how these can be applied not only genome-wide but to
+    Further, we describe how these can be applied not only genome-wide but to
     genomic regions with a particular annotation, specific to a given species available

### 2025-03-17 analysis2_manuscript.tex
edits to affiliations

@@
 \author[2]{Jiseon Min}
+\author[11]{Inés Rebollo}
+\author[2]{Clara T. Rehmann}
 \author[2]{Chris C. R. Smith}
-
 \author[2]{Scott T. Small}
-\author[2]{Clara T. Rehmann}
-\author[11]{Inés Rebollo}
+\author[12]{Georgia Tsambos}
+
 \author[10]{Yan Wong}
-\author[12]{Gregor Gorjanc}
-\author[13]{Christian D. Huber}
-\author[]{Yu Zhang}
+\author[13]{Gregor Gorjanc}
+\author[14]{Christian D. Huber}
+\author[13]{Yu Zhang}
 
@@
 \author[5]{Kirk E. Lohmueller}
-\author[14]{Aaron P. Ragsdale}
-\author[15]{Daniel R. Schrider}
-\author[16]{Ilan Gronau}
-\author[2,17]{Peter L. Ralph}
+\author[15]{Aaron P. Ragsdale}
+\author[16]{Daniel R. Schrider}
+\author[17]{Ilan Gronau}
+\author[2,18]{Peter L. Ralph}
 \author[2]{Andrew D. Kern}
@@
 \affil[11]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
-\affil[12]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
-\affil[13]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
-\affil[14]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
-\affil[15]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
-\affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
-\affil[17]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
+\affil[12]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
+\affil[13]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
+\affil[14]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
+\affil[15]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
+\affil[16]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
+\affil[17]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
+\affil[18]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
 

### 2025-03-17 analysis2_manuscript.tex
more edits

@@
 
+% author tranche explanation:
+% co-first authors
 \author[1,*]{Graham Gower}
-\author[2*]{Nathaniel S. Pope}
-\author[2*]{Murillo F. Rodrigues}
-\author[2*]{Silas Tittes}
-\author[3*]{Linh N. Tran}
-
+\author[2,*]{Nathaniel S. Pope}
+\author[2,*]{Murillo F. Rodrigues}
+\author[2,*]{Silas Tittes}
+\author[3,*]{Linh N. Tran}
+
+% additional authors
 \author[4]{Ornob Alam}
@@
 \author[10]{Ben Jeffrey}
+\author[2]{Kevin Korfmann}
 \author[5]{Christopher C. Kyriazis}
-\author[2]{Kevin Korfmann}
 \author[2]{Jiseon Min}
@@
 \author[2]{Clara T. Rehmann}
+\author[2]{Scott T. Small}
 \author[2]{Chris C. R. Smith}
-\author[2]{Scott T. Small}
 \author[12]{Georgia Tsambos}
-
 \author[10]{Yan Wong}
+\author[13]{Yu Zhang}
+
+% PI additional authors
+\author[14]{Christian D. Huber}
 \author[13]{Gregor Gorjanc}
-\author[14]{Christian D. Huber}
-\author[13]{Yu Zhang}
-
+\author[15]{Aaron P. Ragsdale}
+
+% PIs that were very involved in the project
+\author[16]{Ilan Gronau}
 \author[3]{Ryan N. Gutenkunst}
@@
 \author[5]{Kirk E. Lohmueller}
-\author[15]{Aaron P. Ragsdale}
-\author[16]{Daniel R. Schrider}
-\author[17]{Ilan Gronau}
+\author[17]{Daniel R. Schrider}
 \author[2,18]{Peter L. Ralph}
@@
 \affil[15]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
-\affil[16]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
-\affil[17]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
+\affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
+\affil[17]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
 \affil[18]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
@@
     and cross-checks for agreement, including between \slim and \msprime implementations of the same models.
-    For instance, for testing we make use of \slim's scriptability to separately record in-simulation metadata
-    information about current population sizes, DFEs, etcetera,
+    For instance, for testing we make use of \slim's scriptability to separately record in metadata
+    of the simulation current population sizes, DFEs, etcetera,
     which are then programmatically compared to the models used to generate the \slim script
@@
 We thank Tom Booker for discussions and encouragement to incorporate realistic BGS models into \stdpopsim.
+We thank the numerous users of the \stdpopsim package and well as
+workshop participants who provided feedback and suggestions for improvements.
 

### 2025-03-17 analysis2_manuscript.tex
discussion of AUC scores for sweeps

@@
         Three methods for detecting sweeps were applied to simulated data:
-        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
+        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row -- labelled CLR),
         2) \diploshic \citep{kern2018diplos} (middle row),
@@
         chromosome (bottom).
-        \igcomment{In fig, replace CLR with \sweepfinder; also in figs 6 S5 S6}
         }
@@
     receiver operating characteristic (ROC) curves (Figure \ref{fig:roc-curves}).
-    Different methods perform better for each of the sampled populations,
-    with \sweepfinder performing better than the other two on the CEU sample,
-    and \diploshic performing better on the YRI sample.
-    This is may result from the fact that we trained \diploshic on simulations under a constant size population,
+    Performance varies for each of the sampled populations.
+    In the CEU sample, \sweepfinder and \diploshic perform similarly with a high area under the curve (AUC) score,
+    (AUC = 0.906 and 0.903 respectively), but both underperform the simple method
+    that uses reduced diversity ($\pi$) as an indicator (AUC = 0.932).
+    In the YRI sample, \diploshic outperforms the other two methods,
+    with an AUC of 0.979, followed by using outliers of $\pi$ with an AUC of 0.957,
+    and then \sweepfinder with an AUC of 0.939.
+    We note that we trained \diploshic on simulations under a constant size population,
     which more closely matches the demographic history of YRI, than that of CEU, which 
@@
     Interestingly, the simple method that uses reduced diversity ($\pi$) as an indicator
-    for a selective sweep performs reasonably well overall (especially on the CEU sample).
+    for a selective sweep performs very well overall (especially on the CEU sample).
     However, its accuracy is much more variable along the chromosome compared to the other two method (see Figure \ref{fig:chr1_power}),
@@
         Three methods for detecting sweeps were applied to simulated data:
-        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
+        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row-- labelled CLR),
         2) \diploshic \citep{kern2018diplos} (middle row),
@@
     used in the simulations that included sweeps (see \textbf{Methods}).
-    \igcomment{Add AUCs?}
+    AUC scores are shown for each method. \sweepfinder is labelled CLR in the figure.
     }

### 2025-03-17 analysis2_manuscript.tex
dan found a typo

@@
     and cross-checks for agreement, including between \slim and \msprime implementations of the same models.
-    For instance, for testing we make use of \slim's scriptability to separately record in metadata
-    of the simulation current population sizes, DFEs, etcetera,
+    For instance, for testing we make use of \slim's scriptability to separately record as metadata
+    the simulation parameters (current population sizes, DFEs, etc.),
     which are then programmatically compared to the models used to generate the \slim script

### 2025-03-17 analysis2_manuscript.tex
updated my affiliation and added some explanation to ROC curves

@@
 \author[2,*]{Nathaniel S. Pope}
-\author[2,*]{Murillo F. Rodrigues}
+\author[2,19,*]{Murillo F. Rodrigues}
 \author[2,*]{Silas Tittes}
@@
 \affil[18]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
+\affil[19]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton OR 97006, USA}}
 
@@
     and then \sweepfinder with an AUC of 0.939.
+    Note, however, that despite the apparent good performance of the $\pi$ statistic,
+    it showed extreme heterogeneity across the genome. % MFR: I think this is an important point. Basically there is a Simpson's paradox here.
     We note that we trained \diploshic on simulations under a constant size population,
@@
     and false positive rates (FPR) were established by applying the methods to genomes simulated
-    under neutral evolution with the same demographic history and genetic map
+    under BGS applied on exons with the same demographic history and genetic map
     used in the simulations that included sweeps (see \textbf{Methods}).

### 2025-03-19 analysis2_manuscript.tex
Minor edits 3.17 (#59)
* edits to affiliations

* more edits

* added updated ROC plot

* discussion of AUC scores for sweeps

* dan found a typo

* updated my affiliation and added some explanation to ROC curves

* oops

---------

Co-authored-by: Murillo F. Rodrigues <murillofer.rodrigues@gmail.com>
Ilan's edits: - marked with %IG non-trivial changes - removed previous resolved comments

--- 2025-03-17 analysis2_manuscript.tex
@@
 \affil[15]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
-\affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya, Israel}}
+\affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya 4610101, Israel}}
 \affil[17]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
 \affil[18]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
+\affil[19]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton OR 97006, USA}}
 
@@
     \stdpopsim is a community-maintained simulation library that already provides an extensive catalog of
-    species-specific population genetic models,
+    species-specific population genetic models.
     Here we present a major extension to the \stdpopsim{} framework that enables simulation of various modes
@@
     Our results demonstrate the robustness of demographic inference methods to selection on linked sites, 
-    reveal the sensitivity of DFE inference methods to model assumptions, 
+    reveal the sensitivity of DFE-inference methods to model assumptions, 
     and show how genomic features like recombination rate and the density of functional sequence influence power to detect selective sweeps.
@@
     To meet the growing need for the interpretation,
-    analysis, and exploration of realistic and complex evolutionary models
+    analysis, and exploration of realistic and complex evolutionary models,
     the field of population genetics has increasingly turned to simulation.
@@
     from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.
-
-    \stdpopsim{} aims to provide such a tool,
+    %
+    \stdpopsim{} provides a framework for such models,
     thus lowering the barrier to production of complex, multifaceted simulations,
@@
     in developing or testing their work.
-    The project also has the goal of doing open, collaborative, and inclusive work,
+    The project also promotes open, collaborative, and inclusive work,
     viewing capacity-building across the community as equal in importance to software development.
@@
     genomic elements (e.g., exons, introns, intergenic regions, etc.)
-    whose coordinates are specified using a GFF3 annotationfile.      
+    whose coordinates are specified using a GFF3 annotation file.      
     More concretely, consider implementing a model of selection consistent with the
@@
     The associated \texttt{DFE} object is defined with two mutation types,
-    neutral and deleterious, corresponding to synonymous and non-synonymous sites.
+    neutral and deleterious, corresponding to synonymous and non-synonymous sites,
     with their proportions set to their expected prevalence in exons. 
@@
     of the selected allele at the time of sampling. As this functionality is implemented
-    on top of the existing \stdpopsim API, they can be combined with other models of selection
+    on top of the existing \stdpopsim API, it can be combined with other models of selection
     and demography to provide a model of the combined effects of multiple, disparate evolutionary processes
@@
     % current numbers of DFEs / species / etc
-    % ADK: check version number
-    At this release (version 0.3) the \stdpopsim catalog comprises 27 species (i.e. genome representations)
+    At this release (version 0.3.0) the \stdpopsim catalog comprises 27 species (i.e. genome representations)
     for which we have included 35 demographic models, 38 genetic maps, 12 DFEs, and 8 annotation objects. %ADK: checked this on 3/10/25
@@
 
-%IG: commented out this intro paragraph. Better to dive right in.
-%\section*{Example Applications with Selection}
-%    \label{applications}
-%    In this section, we present a series of examples of how the new models of selection
-%    in \stdpopsim can be used to benchmark different
-%    methods for population genetic inference.
-%    We focus on three main areas: demographic inference in the 
-%    single and multi-population settings, inference of the DFE,
-%    and the detection of selective sweeps, particularly in the context of realistic recombination maps and background selection.
-
     \section*{Inference of $N_e(t)$ in the context of selection}
@@
     \citep{li2011inference,Schiffels2020,terhorst2017robust}, 
-    %\igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
-    %\drscomment{One could use dadi for that, but then we should also cite things like fastsimcoal and ABC, while explaining
-    % that users can use those methods to infer piecewise-constant N(t) histories as well with a prespecified number of epochs
-    % and priors around their params, but maybe that is too much of a digression here? I'm fine either way}
-    % ADK: I think we are good as is
     through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
@@
         %citations aren't always allowed in captions. We should check with MBE.
-        Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
+        %IG:I removed citations from all fig captions. They are prominently specified in the main text
+        Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model
         with and without purifying selection on exons. 
@@
         Columns correspond to the four inference methods:    
-        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+        \stairway, \msmc, \smcpp and \gone.
         Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
@@
     under the out of Africa (OOA) demographic model of
-    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
+    \citet{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
     Simulations with selection were implemented by modeling purifying selection on nonsynonymous mutations
     using a DFE inferred by \citet{kim2017inference}.
-    A basic summary of nucleotide diversity from these simulations is shown in Figure \ref{fig:homsap-summary-stats},
-    which shows that the presence of selection decreases polymorphism in simulations with selection only slightly
-    at the genome-wide level, but more substantially within exonic regions under both a constant population size
-    and the out-of-Africa demographic model.
-    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
-    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
-    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
-    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.
+    A basic summary of nucleotide diversity from these simulations
+    shows that the presence of selection decreases polymorphism only slightly
+    at the genome-wide level, but more substantially within exonic regions (Figure \ref{fig:homsap-summary-stats}).
+    %IG: no need to specify the two scenarios shown in fig
+    %under both a constant population size and the out-of-Africa demographic model
+    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 
+    %IG: removed numbering and reordered to match fig 2
+    \stairway \citep{liu2020stairway}, \msmc \citep{Schiffels2020}, \smcpp \citep{terhorst2017robust} and \gone \citep{santiago2020recent}.
 
@@
     \ref{fig:human-demography}; blue) are fairly similar.
-    Thus, this particular parameterization of background selection
-    in combination with selection only at exonic sites in the out-of-Africa demographic model does not
-    appear to notably bias the estimates produced by these methods.
-    In light of Figure \ref{fig:human-demography}, this makes sense-- 
-    the effect of selection is only apparent within exonic regions and presumably sites
-    very closely linked, thus masking of exons should be sufficient to remove the effects of selection
-    from our simulations. 
+    Thus, this particular parameterization of background selection and demographic model,
+    in combination with selection only at exonic sites, %in the out-of-Africa demographic model IG: it's probably not just this specific demo model
+    does not appear to notably bias the estimates produced by these methods.
+    This makes sense in light of our observation that genetic diversity is mostly affected
+    within exonic regions and presumably sites very closely linked (Figure \ref{fig:homsap-summary-stats}),
+    thus masking of exons should be sufficient to remove the major effects of selection. 
     As noted in previous studies that compared demographic inference methods
@@
 
-%IG: commented old version of figure here
-%    \begin{figure}[t]
-%        \centering
-%        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
-%        \caption{
-%        \label{fig:1pop-human-demography}
-%        Inference of $N_e(t)$ from a human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
-%        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ obtained from genomes simulated
-%        without selection, while the right column shows estimates of $N_e(t)$ from genomes simulated with a gamma-distributed   
-%        DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
-%        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
-%        }
-%    \end{figure}
-
     To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
     we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
-    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
+    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita by \textcite{robinson2022critically},
     and a constant recombination rate across the genome (see \textbf{Methods}).
@@
     additive ($h=0.4$; see \textbf{Methods}).
-    As above, a basic summary of nucleotide diversity from these simulations is shown in Figure \ref{fig:vaquita-summary-stats},
-    which again shows that the presence of selection decreases polymorphism in simulations with selection, but that the effect
-    is only readily apparent within exonic regions.
-    We also see that the two-epoch demographic model used in the simulations actually leads to a slight increase
-    in diversity versus the constant population size model.
-
+    As in the human simulations, we see that the presence of selection
+    decreases polymorphism in genomes simulated with selection, but that the effect
+    is only readily apparent within exonic regions (Figure \ref{fig:vaquita-summary-stats}).
+    %IG: last sentence seems like an unnecessary distraction. The constant-size simulations are only used to make a minor point in the next section
+    %We also see that the two-epoch demographic model used in the simulations actually leads to a slight increase
+    %in diversity versus the constant population size model.
+    %IG: merged paragraphs
     For these simulations, the SFS-based method (\stairway) produces more accurate
@@
 
-    % MFR: Should we say somewhere that people use to expect from BGS?
-That is, that it can be seen as a reduction in local rates of coalescences and thus that we would expect these methods to be somewhat downwardly biased in the presence of BGS?
-    % IG: we can address this in the discussion, or in the revised text above.
-
-%\begin{figure}[t]
-%    \centering
-%    \includegraphics[width=\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/estimated_Ne_t_final}
-%    \caption{
-%    \label{fig:1pop-vaquita-demography}
-%    Performance of methods to infer $N_e(t)$ from simulations of the vaquita porpoise genome under a single population
-%    model of declining population size \citep{robinson2022critically} with and without background selection on exons. 
-%    The left panel shows estimates of $N_e(t)$ from simulations
-%    without selection, while the right panel shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
-%    DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
-%    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
-%
-%    }
-%\end{figure}
+    %MFR: Should we say somewhere that people use to expect from BGS? 
+    %That is, that it can be seen as a reduction in local rates of coalescences
+    %and thus that we would expect these methods to be somewhat downwardly biased in the presence of BGS?
+    %IG: we can address this in the discussion, or in the revised text above.
 
@@
     the performance of different methods for inferring the DFE.
+    %IG: Moved up the methods here to provide context for their mention in the next para
+    Here, we considered three DFE-inference methods:
+    \grapes \citep{galtier2016adaptive}, \polydfe \citep{tataru2020polydfe}, and \dadicli \citep{Huang2023}.
 
@@
     On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
-     %https://academic.oup.com/genetics/article/207/3/1103/5930676
     Moreover, inferred distributions of selection coefficients are typically scaled by an inferred ancestral
@@
     \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$.
-    %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
-    For comparison, here we have normalized all inference results to the convention of \slim
+    To allow direct comparison, here we have normalized all inference results to the convention of \slim
     (also used by \stdpopsim; see \textbf{Methods}).
-    % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?
+    %IG - citations suggested in earlier versions to be added (I'm not sure they're necessary)
+    % precursor of polyDFE  from 2017:     https://academic.oup.com/genetics/article/207/3/1103/5930676
+    % fastDFE - the successor to polyDFE:  https://academic.oup.com/mbe/article/41/5/msae070/7641109
+    % paper that uses GRAPES from 2020:    https://academic.oup.com/genetics/article/216/2/559/6066183
 
@@
     In this setting, the DFE is inferred separately for each of the three simulated populations
-    using three different DFE inference methods: \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
-    and \grapes \citep{galtier2016adaptive}.
+    using each of the three DFE-inference methods.
     In each case we note that the DFE is estimated from segregating sites only,
@@
 
-    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}, Table \ref{tab:dfe_table}).
-    Note that while the estimation accuracy for the DFE parameters themselves (Figure \ref{fig:homsap-dfe.ooa}A-B) is useful,
+    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}). %, Table \ref{tab:dfe_table}). IG: Table seems less relevant here
+    Note that while the estimation accuracy for the DFE parameters themselves is informative (Figure \ref{fig:homsap-dfe.ooa}A-B),
     examining the difference between the actual and inferred distributions of $s$ can be more helpful for showing how well a method
@@
     while \grapes considerably over-estimates the fraction of mutations under weak selection ($s\geq-0.0001$).
-    % IG: commented this out to simplify the basic observation
-    % and under-estimates the fraction of mutations with $s<-0.01$.
 
@@
     shows very similar accuracy across all populations.
-    % produces its most accurate estimate from the CHB genomes.
     To factor out the influence of demographic changes, we also analyzed genomes generated using similar
@@
     the other methods do not show as close a correspondence.
-    \grapes exhibits similar biases without population size changes as in the simulations with more complex demography,
-    but with milder under-estimation of the shape parameter, but \dadicli shows an inverse pattern
+    \grapes exhibits similar biases as observed in the simulations with more complex demography,
+    but with milder under-estimation of the shape parameter,
+    whereas \dadicli shows an inverse pattern
     with overestimated shape parameter and slightly underestimated mean.
-    While that is so, it is hard to be definitive as our sample size is small with 3 simulated replicates.
+    While providing possible insights into the effects of the demographic history
+    on accuracy of different DFE-inference methods, substantiating these insights
+    will require more than three replicate simulations per scenario.
 
@@
     without population size changes (Figure \ref{fig:vaquita-dfe.constant}, Table \ref{tab:dfe_table}).
-    We suspect that this is due to the fact that all DFE inference methods used here assume an additive model ($h = \frac{1}{2}$),
+    We suspect that this is due to the fact that all DFE-inference methods used here assume an additive model ($h = \frac{1}{2}$),
     while mutations in the simulated genomes were more recessive ($h < \frac{1}{2}$), and their dominance decreased with the strength of selection. 
@@
     many replicate simulations, we focused on human chromosome 1 and generated replicate simulations on 5cM segments
-    of the chromosome in 100 different locations. % (see \textbf{Methods}).
+    %IG: took out the 7 segments with low rec rate
+    of the chromosome in 93 different locations. % (see \textbf{Methods}).
     Selective sweeps were simulated by introducing a beneficial mutation in the center of the 5cM segment
-    with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$; see \textbf{Methods}).
-    Each segment was simulated 1000 times, while keeping only replicates in which the selected allele reached a frequency of 0.95 or greater.
+    with a moderately strong selection coefficient ($s = 0.03$; $2N_es \sim 600$; see \textbf{Methods}).
+    Each segment was simulated 1000 times, while keeping only replicates in which the selected allele reached a frequency of 95\% or greater.
     These sweeps were modeled in the context of the three population out-of-Africa model from \citet{gutenkunst2009inferring}
@@
      %sweeep power figure
-    \begin{figure}[t!]
+    \begin{figure}[b!]
         \centering
@@
         \caption{
-        Power to detect selective sweeps at 100 locations along human
+        %IG: better not mention 100 if we're filtering out 7 segments
+        %IG: removed citations from fig captions; they are prominently specified in main text
+        Power to detect selective sweeps along human
         chromosome 1. Genomic segments were simulated with sweeps under a three population out-of-Africa model
-        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
-        %IG: seems like methods-level detail
-        %Single beneficial mutations were introduced at each location with a selection
-        % coefficient of $s = 0.03$, and only sampled if they reached a terminal
-        % frequency of 0.95 or greater.
+        %\citep{gutenkunst2009inferring}
+        and with background selection from deleterious mutations in exons.
         Three methods for detecting sweeps were applied to simulated data:
-        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row -- labelled CLR),
-        2) \diploshic \citep{kern2018diplos} (middle row),
-        and 3) reduced diversity ($\pi$) (bottom row).
+        \sweepfinder (top row -- labelled CLR),
+        \diploshic (middle row),
+        and reduced diversity ($\pi$) (bottom row).
         Power (true positive rate) is shown for these methods for the CEU and YRI
@@
         and a null model with background selection from deleterious mutations in exons (red).
-        Also shown are heatmaps of exon density and local recombination rate along the
+        Also shown are heatmaps of local recombination rate and exon density along the
         chromosome (bottom).
@@
     We evaluated the statistical power to detect selective sweeps for three methods:
-    1) \sweepfinder \citep{degiorgio2016sweepfinder2},
-    2) \diploshic \citep{kern2018diplos},
-    and 3) reduced diversity ($\pi$).
+    \sweepfinder \citep{degiorgio2016sweepfinder2},
+    \diploshic \citep{kern2018diplos},
+    and reduced diversity ($\pi$).
     For each method, we determined a threshold for its test statistic 
@@
     -- and controlling for a $5\%$ false positive rate.
-
+    %
     We then applied each method to data simulated with sweeps for two samples (YRI and CEU)
-    in each of 100 locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
+    in each of the simulated locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
     First and most strikingly, the power to detect sweeps varies considerably along the chromosome,
@@
     is determined using simulated data with background selection (BGS null model).
-    %This observation is consistent with what was previously shown by \cite{schrider2020background},
-    %but the effect is relatively minor. DRS: cutting this because my paper showed that power decreases when our test set has BGS and our training set does not, so not quite the same.
     This is likely because the strength of selection acting on exons is relatively weak, and the
@@
     is characterized by a strong recent bottleneck followed by population expansion.
-    %\drscomment{note: I am couching the language above a bit because we have previously shown that shic doesn't suffer in the sweep vs.
-nonsweep question when doing other train-on-constant-test-on-nonequilib benchmarks.
-Could be something else going on here (or some interaction between demog misspec and rec rate variation), but in any case I am not certain of the cause so I am toning down our certainty slightly.}
     Interestingly, the simple method that uses reduced diversity ($\pi$) as an indicator
@@
         Power to detect selective sweeps 
-        %IG: shortened the title line
-        %at 100 locations along human
-        %chromosome 1 under a three population out-of-Africa model
-        %\citep{gutenkunst2009inferring}  when using null models with and
-        %without purifying selection
         as a function of local recombination rate.
         This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
-        but with the 100 genomic segments plotted against their
+        but with the genomic segments plotted against their
         average recombination rates instead of position along chromosome 1.
     	Genomic segments were simulated with sweeps under a three population out-of-Africa model
-        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
+        and with background selection from deleterious mutations in exons.
         Three methods for detecting sweeps were applied to simulated data:
-        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row-- labelled CLR),
-        2) \diploshic \citep{kern2018diplos} (middle row),
-        and 3) reduced diversity ($\pi$) (bottom row).
+        \sweepfinder (top row-- labelled CLR),
+        \diploshic (middle row),
+        and reduced diversity ($\pi$) (bottom row).
         Power (true positive rate) is shown for these methods for the CEU and YRI
@@
     The first pertains to estimation of demographic history ($N_e(t)$) in the context of
-    linked purifying selection. Our genomic simulations of humans and vaquita that incorporate
-    selection model published DFEs that operate at exonic positions of the genome only. 
+    linked purifying selection. Our simulations of humans and vaquita genomes
+    model deleterious mutations only at exonic positions of the genome. 
     As shown in Figures \ref{fig:homsap-summary-stats} and \ref{fig:vaquita-summary-stats},
@@
     which represent a small minority of sites within the genome, and does not lead to large-scale
-    perturbations in genome-wide diversity. Thus it is not surprising that estimates of $N_e(t)$
+    perturbations in genome-wide diversity. Thus, it is not surprising that estimates of $N_e(t)$
     from genome-wide data from which exons have been masked should be accurate 
     (e.g., Figures \ref{fig:human-demography} and \ref{fig:vaquita-demography}). 
-    However the question remains to what extent this would be true in empirical data where
+    However, the question remains to what extent this would be true in empirical data where
     selection is not constrained to coding positions. For instance, we know
@@
     of conserved non-coding elements \citep[e.g.,][]{siepel2005evolutionarily} with extremely
-    strong selection acting on them \citep{katzman2007human,mclean2008dispensability},
-    thus we caution extrapolating from our simulation results. Future studies might build upon
+    strong selection acting on them \citep{katzman2007human,mclean2008dispensability}.
+    We thus caution extrapolating from our results. Future studies might build upon
     the annotations we used here to include selection at other portions of the genome as well
@@
     is sensitive to the dominance of deleterious mutations.
-    All three of the inference methods of the DFE applied here assume that deleterious mutations act
+    All three of the DFE-inference methods applied here assume that deleterious mutations act
     in an additive fashion ($h=0.5$) by default.
@@
     Specifically, these simulations used a DFE where mutations with $s<-0.01$ are quite recessive, 
-    with $h<=0.01$, and where mutations with $-0.001<s<-0.01$ have $h=0.1$.
-    We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}).
+    with $h<=0.01$, and where mutations with $-0.01<s<-0.001$ have $h=0.1$.
+    We found that the three DFE-inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}A-B).
     This performance is not surprising as decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
@@
     Fitting a mis-specified DFE that assumes $h=0.5$ will then lead to underestimates of $E[s]$.
-    Indeed, the distribution of $hs$ is much more accurately estimated (Figure \ref{fig:vaquita-dfe}b).
+    Indeed, the distribution of $hs$ is much more accurately estimated (Figure \ref{fig:vaquita-dfe}C).
     Overall, these results are consistent with prior work showing that it is challenging to separately
@@
     illustrate a strength of likelihood methods like \sweepfinder, that do not require a predefined window size.
-    Window-based methods, such as \diploshic, tend to lag in performance when the perturbation due to a sweep is at 
-    a much greater or smaller scale than the chosen window size, as can occur with very strong sweeps or low levels of 
-    recombination. Moreover while all the methods examined here retain adequate power for sweep identification across
+    Window-based methods, such as \diploshic, tend to lag in performance when the genomic region affected by the sweep is
+    larger than the chosen window size, as can occur with very strong sweeps or low levels of 
+    recombination. Moreover, while all the methods examined here retain adequate power for sweep identification across
     scenarios, including the inclusion of background selection, our results underscore the importance of developing 
-    methods that are well powered in the face of unknown demographic histories. Indeed, the largest differences
-    in power we observed across simulated scenarios were between demographic histories and not the inclusion of background
-    selection.
+    methods that are well powered in the face of unknown demographic histories. Indeed, larger differences
+    in power were observed when applying a method to genomes simulated with different demographic histories (YRI versus CEU)
+    than when calibrating a method using a null model with and without background selection.
 
@@
     new methods for the analysis of genetic data in the presence of selection. 
-    The additions to \stdpopsim described here further expand this tool's utility for both 
+    The additions described here further expand the utility of \stdpopsim for both 
     characterizing the patterns of diversity expected under a wide range of population genetic
@@
     are programatically tested for equality.
-    The QC process identified a substantial bug in one DFE:
-    the \textit{Drosophila melanogaster} DFE \texttt{LognormalPlusPositive\_R16} \citet{booker2021selective}.
-    Due to an error in coding the log-normal distribution, deleterious selection coefficients
-    were several orders of magnitude too large.
-    The bug affected results in a separate project from one of our groups
-    that made use of this DFE before QC had been done.
-    All changes and additions to the code must pass a comprehensive set of unit tests:
+    %IG: tried to put a more positive spin on the DFE bug example
+    This QC process has proven its effectiveness in numerous occasions in detecting implementation bugs
+    and resolving modeling ambiguities.
+    For example, this process detected  an error in coding the log-normal distribution of
+    the \textit{Drosophila melanogaster} DFE inferred by \citet[\stdpopsim label  \texttt{LognormalPlusPositive\_R16}]{booker2021selective}.
+    This faulty DFE model was used in a study conducted by one of our groups before it underwent the QC process,
+    where we noticed that deleterious selection coefficients
+    were several orders of magnitude larger than expected.
+    Subsequently, the QC process was able to find and fix this bug.
+    In addition to QC, all changes and additions to the code must pass a comprehensive set of unit tests:
     at the time of writing, the library includes over 11,000 lines of testing code,
@@
     In this study we performed simulations of the human and vaquita porpoise genomes
-    using \stdpopsim{}. We ran replicate datasets for each genome/parameter combination,
+    using \stdpopsim{}. We ran replicate simulations for each genome/parameter combination,
     in which chromosomes were simulated separately from one another.
@@
     were simulated with the same model of selection and a simple demographic model with one population with a constant size of $3,500$.
-    \igcomment{If we end up using Fig. S3, then we should add a line describing it's simulation here.}
     The configuration file used in combination with our \snakemake workflow
@@
     acting on exons, but it was influenced by selection acting on linked sites (background selection).    
-
+    %
     We assessed the accuracy of demographic inferences from three software packages:
@@
     and \smcpp \citep{terhorst2017robust}.
-    For \msmc \citep{Schiffels2020}, we used a random sample of $N=6$ simulated genomes with 20
+    For \msmc %\citep{Schiffels2020}, IG: no need to double-cite
+    we used a random sample of $N=6$ simulated genomes with 20
     iterations of the EM algorithm.
-    For \gone \citep{santiago2020recent}, we set \texttt{max\_snps} to 500,000,
+    For \gone we set \texttt{max\_snps} to 500,000,
     the number of generations to 2,000, and the number of bins to 400, all representing default settings.
-    For \stairway \citep{liu2020stairway} we did one run of theta estimation for each replicate dataset,
-    and used a `dimFactor' of 5000, both of which represent default settings.
-    For \smcpp \citep{terhorst2017robust} we used the unfolded SFS (`--unfold'), 
+    For \stairway we did one run of $\theta$ estimation for each replicate dataset,
+    and used a \texttt{dimFactor} of 5000, both of which represent default settings.
+    For \smcpp we used the unfolded SFS (\texttt{--unfold}), 
     along with the true mutation rate under which the simulations were generated. 
@@
     Each software package was run using the same set of simulations.
-    The DFE inference using \dadicli was conducted by first inferring a demographic model from neutral sites,
+    DFE inference using \dadicli was conducted by first inferring a demographic model from neutral sites,
     using a two epoch demographic model with 100 repeated runs of the optimization procedure.
-    \dadicli then proceeds by using a cached strategy wherein we used 2000 points in the $gamma$ distribution.
+    \dadicli then proceeds by using a cached strategy wherein we used 2000 points in the gamma distribution.
     Finally, the DFE was inferred using a ratio of non-synonymous to synonymous mutations of 2.31, again with 
     100 repeated runs of the optimization procedure.
-    For \polydfe, we used model 'C' which uses a mixture of a gamma distribution and an exponential distribution,
+    For \polydfe, we used model `C' which uses a mixture of a gamma distribution and an exponential distribution,
     where the gamma component is used for negatively selected mutations. For optimization we used
-    the default BFGS optimizer with the '-e' flag to automatically estimate the initial parameters values.
+    the default BFGS optimizer with the \texttt{-e} flag to automatically estimate the initial parameters values.
     For \grapes, we inferred the DFE under the `GammaZero' model, which assumes that selection coefficients
-    are drawn from a single gamma distribution. We used the `-no-div-param' flag to disable the use
+    are drawn from a single gamma distribution. We used the \texttt{-no-div-param} flag to disable the use
     of divergence-based parameter estimation. All other settings were left at their default values.
@@
     % ADK: I've added a section below that describes the DFE inference methods used in this study.
-    For comparison among methods, estimates were rescaled. 
-    As described above, each DFE-inference method scales selection coefficients slightly differently.
-    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
-    The output of \grapes  was scaled by a factor of  $\mu / \theta$,
-    and the output of \polydfe was scaled by $2 \mu / \theta$, 
+    To facilitate direct comparison between inference methods, 
+    we scale each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
+    %IG: expanded a bit to justify each scaling decision
+    \grapes outputs the mean and shape of the inferred distribution of $4N_es$.
+    We thus scale the mean by a factor of  $\frac{\mu}{\theta}$, 
     where $\mu$ is the assumed mutation rate and
-    $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
-    For \dadicli, the gamma distribution is parameterized using the shape and rate ($\alpha$ and $\nu$, respectively). 
-    The mean value of the distribution, which is $s$ is transformed from $\alpha$ and $\nu$ and rescaled as $\alpha * \nu / (\theta / (4 \mu L p))$,
+    $\theta=4N_{e}\mu$ is the population-scaled mutation rate, as estimated by \grapes.
+    \polydfe is similarly scaled by $\frac{2\mu}{\theta}$,
+    reflecting the fact that it defines the selection coefficient such that a homozygote has fitness $1+ 2s$.
+    \dadicli infers a gamma distribution for $N_es$ and parameterizes this distribution using the shape ($\alpha$) and rate ($\nu$).
+    %IG: we're still missing justification for the scaling by L and p, but we can leave this for revision
+    Thus, the mean value of $s$ is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$,
     where $L$ is the total number of sites available for the analysis,
@@
 
-    To test methods for detecting selective sweeps, we simulated 100 evenly distributed 5cM
+    To test methods for detecting selective sweeps, we considered 100 evenly distributed 5cM
     segments of human chromosome 1.
@@
     which is critical for the linked effects of a sweep, was held constant.
+    %IG: moved this up here since it makes mroe sense to describe it as if these segments were never simulated
+    % also converted to rec rate from cM. Please check!
+    Seven segments in which the central 1Mb had an average recombination rate below $10^{-11}$ were excluded from the analysis.
     Each segment was simulated in each of the following three settings:
@@
     In all three settings, we applied a genetic map from the HapMap Project
-    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
+    \citep[\stdpopsim label \texttt{HapMapII\_GRCh38}]{international2007second} and
     the three-population out of Africa (OOA) demographic model from
-    \cite{gutenkunst2009inferring} (\stdpopsim label \texttt{OutOfAfrica\_3G09}).
+    \citet[\stdpopsim label \texttt{OutOfAfrica\_3G09}]{gutenkunst2009inferring}.
     In the two settings with BGS, we applied the DFE model with gamma-distributed selection coefficients inferred by
-    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
-    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
+    \citet[\stdpopsim label \texttt{Gamma\_K17}]{kim2017inference} to exons annotated by the HAVANA group release 104
+    \citep[\stdpopsim label \texttt{ensembl\_havana\_104\_exons}]{ensembl2018}.
     Hard sweeps were modeled by generating a single beneficial mutation 
@@
     For each of the 100 genomic segments of chromosome 1, we generated 1,000 replicate datasets under each of the three settings.
-    As with the whole-genome simulations (see above), we used the \slim engine with a burn-in of $2N$ generations
+    We used the \slim engine with a burn-in of $2N$ generations
     (where $N$ is the initial population size), and a scaling factor of 2.
-    Each simulation generated 10 diploid genomic segments for each of the two sampled populations: YRI and CEU.
+    Each simulation generated ten diploid genomic segments for each of the two sampled populations: YRI and CEU.
     The sequences simulated with a hard sweep and BGS were used for sweep detection,
@@
     computed directly from the underlying tree sequences using \tskit \citep{ralph2020efficiently}
-    over 10 equally sized and non-overlapping subwindows.
+    over ten equally sized and non-overlapping subwindows.
     \sweepfinder \citep{degiorgio2016sweepfinder2} is based on a composite likelihood ratio (CLR) 
@@
     and a model with background selection from deleterious mutations in exons (BGS).
-    Each null model was represented by a different set of $100,000$ simulations generated according to the appropriate model
+    %IG: deducted 7000 datasets for the low-rec regions
+    Each null model was represented by a different set of $93,000$ simulations generated according to the appropriate model
     (see \textbf{Simulations with selective sweeps} above).
@@
     using the extreme $5\%$ quantile in each empirical distribution.
-    Simulated regions where the central 1Mb had a recombination rate of less than $10^{-3}$cM were excluded from the analysis.
-    This excluded 7 regions from the analysis. 
     Then, power (or true positive rate) to detect a sweep by a given method was determined for
-    100 evenly-spaced positions on chromosome 1 as follows:
+    positions on chromosome 1 as follows:
+    %IG: the 1000 simulations are with hard sweep + BGS
     each sweep-detection method was applied to the 1,000 
-    simulated datasets generated with a hard sweep (neutral and BGS) for that position,
+    simulated datasets generated with a hard sweep and BGS for that position,
     and the power was set to be the proportion of simulations (out of 1,000)
@@
     however for consistency among methods we chose to take this approach.
-    The complete \snakemake workflow for sweep detection can be found at
+    The complete \snakemake workflow for sweep detection (including simulations) can be found at
     \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/sweep_simulate.snake}.
@@
 The release of \stdpopsim is available at \url{https://github.com/popsim-consortium/stdpopsim}.
-This paper coincides with version 0.3.XX of \stdpopsim. %be sure to check this
+This paper coincides with version 0.3.0 of \stdpopsim.
 The entire analysis pipeline associated with this paper is available
@@
 
+%IG: moved supp table here, because otherwise it appeared between the supp figs
+%IG: edited caption and removed annotations column, since they're all exons now.
+%IG: moved commented "all sites" rows to bottom
+\begin{table}[h]
+\centering
+\small
+
+\caption{\textbf{Performance of DFE-inference methods in four simulation scenarios}.
+For each of the three methods, we report the mean absolute error (MAE) of the mean selection coefficient and shape of its inferred DFE.
+In each simulation scenario, the lowest MAEs are highlighted in bold.}
+
+\begin{tabular}{lllrrrrrr}
+\toprule
+species ID & demography & \makecell{MAE \\ $E|s|$ \\ dadi-cli} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi-cli} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
+\midrule
+HomSap & Constant  & \bf{0.0012} & 0.0101 & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
+HomSap & OOA 5R19  & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
+PhoSin & Constant  & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
+PhoSin & Vaquita2Epoch 1R22  & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
+%HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
+%HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
+%PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
+%PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
+\bottomrule
+\end{tabular}
+\label{tab:dfe_table}
+\end{table}
+
 \begin{figure}[h]
@@
     \label{fig:homsap-summary-stats}
-    Summaries of diversity for human simulations. The left and right panels show
-    nucleotide diversity ($\pi$) calculated over all three replicate datasets
-    for the Constant size and Out-of-Africa demographic models respectively.
-    Each panel shows four violins representing the distribution of $\pi$ values
-    under a model without selection (blue), a model with background selection (orange),
-    calculated from either the entire genome (labeled neutral or DFE)
+    Genetic diversity in human simulations.
+    Nucleotide diversity ($\pi$) was calculated for genomes simulated
+    under constant population size (left panel) and under the out-of-Africa demographic model (right panel).
+    Each panel shows four violin plots representing the distribution of $\pi$
+    in genomes simulated without selection (blue) and genomes simulated with background selection (orange),
+    calculated from either the entire genome (labeled neutral or DFE),
     or from exonic regions only (labeled exons-only).
+    Each violin plot summarizes three replicate simulations generated under the relevant scenario.
     }
@@
     \label{fig:vaquita-summary-stats}
-    Summaries of diversity for vaquita simulations. The left and right panels show
-    nucleotide diversity ($\pi$) calculated over all three replicate datasets
-    for the Constant size and two-epoch demographic models respectively.
-    Each panel shows four violins representing the distribution of $\pi$ values
-    under a model without selection (blue), a model with background selection (orange),
-    calculated from either the entire genome (labeled neutral or DFE)
+    Genetic diversity in vaquite simulations.
+    Nucleotide diversity ($\pi$) was calculated for genomes simulated
+    under constant population size (left panel) and under the two-epoch demographic model (right panel).
+    Each panel shows four violin plots representing the distribution of $\pi$
+    in genomes simulated without selection (blue) and genomes simulated with background selection (orange),
+    calculated from either the entire genome (labeled neutral or DFE),
     or from exonic regions only (labeled exons-only).
+    Each violin plot summarizes three replicate simulations generated under the relevant scenario.
      }
@@
 
-\begin{figure}[b!]
+\begin{figure}[h]
     \centering
@@
     Inference of $N_e(t)$ from vaquita porpoise genomes simulated under a single population 
-    model of declining population size \citep{robinson2022critically}
-    with and without purifying selection on exons. 
-    Inference is conducted using four methods:    
-    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
+    model of declining population size with and without purifying selection on exons. 
+    Inference is conducted using four methods: \stairway, \msmc, \smcpp and \gone.
     Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
@@
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 
+%IG: this fig is not cited anywhere and does not appear to provide crucial insights
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=\textwidth]{figures/HomSap/Constant/HomSap_Constant_Gamma_K17_all_sites_DFE_plot}
+%    \caption{
+%    \label{fig:homsap-dfe.constant.all_sites}
+%    Performance of methods to infer the distribution of fitness effects (DFE) from genetic data
+%    under constant population size  where all sites along the simulated chromosome can experience selected and neutral mutation
+%    as opposed to exonic regions only.\\
+%    }
+%\end{figure}
+
+% roc curves for sweeps
 \begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{figures/HomSap/Constant/HomSap_Constant_Gamma_K17_all_sites_DFE_plot}
-    \caption{
-    \label{fig:homsap-dfe.constant.all_sites}
-    Performance of methods to infer the distribution of fitness effects (DFE) from genetic data
-    under constant population size  where all sites along the simulated chromosome can experience selected and neutral mutation
-    as opposed to exonic regions only.\\
-    }
-\end{figure}
-
-% roc curves for sweeps
-\begin{figure}
     \centering
@@
     \label{fig:roc-curves}
-    Receiver operating characteristic (ROC) curves for the three methods for detecting sweeps.
-    True positive rates (TPR) were computed across all 100,000 datasets simulated with selective sweeps,
+    Receiver operating characteristic (ROC) curves for the three methods for detecting sweeps:
+    \sweepfinder (labeled CLR), \diploshic, and reduced diversity ($\pi$).
+    True positive rates (TPR) were computed across all datasets simulated with selective sweeps,
     and false positive rates (FPR) were established by applying the methods to genomes simulated
-    under neutral evolution with the same demographic history and genetic map
-    used in the simulations that included sweeps (see \textbf{Methods}).
-    AUC scores are shown for each method. \sweepfinder is labelled CLR in the figure.
-    }
-    %DRS-fig: can we include AUCs here?
+    under the same settings but without sweeps (the BGS model; see \textbf{Methods}).
+    Area under the curve (AUC) scores are shown for each method.
+    }
 \end{figure}
@@
 %sweep power vs exon density
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
     \caption{
-    %IG: copied new caption from fig 6 and adjusted.
-        Power to detect selective sweeps as a function of local exon density.
+        Power to detect selective sweeps 
+        as a function of local exon density.
         This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
-        but with the 100 genomic segments sorted according to their
-        exon density (percentage of exonic bases per window)instead of position along chromosome 1.
+        but with the genomic segments plotted against their
+        exon density (percentage of exonic bases per window) instead of position along chromosome 1.
     	Genomic segments were simulated with sweeps under a three population out-of-Africa model
-        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
+        and with background selection from deleterious mutations in exons.
         Three methods for detecting sweeps were applied to simulated data:
-        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row -- labelled CLR),
-        2) \diploshic \citep{kern2018diplos} (middle row),
-        and 3) reduced diversity ($\pi$) (bottom row).
+        \sweepfinder (top row-- labelled CLR),
+        \diploshic (middle row),
+        and reduced diversity ($\pi$) (bottom row).
         Power (true positive rate) is shown for these methods for the CEU and YRI
@@
 
-\begin{table}[ht]
-\centering
-\small
-
-\caption{\bf{Performance of DFE methods across the simulation scenarios}. 
-Bold rows show the lowest mean absolute error (MAE) for the two DFE parameters
-for each species.}
-
-\begin{tabular}{lllrrrrrr}
-\toprule
-species ID & demography & annotation & \makecell{MAE \\ $E|s|$ \\ dadi-cli} & \makecell{MAE \\ $E|s|$ \\ grapes} & \makecell{MAE \\ $E|s|$ \\ polyDFE} & \makecell{MAE \\ shape \\ dadi-cli} & \makecell{MAE \\ shape \\ grapes} & \makecell{MAE \\ shape \\ polyDFE} \\
-\midrule
-%HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
-HomSap & Constant & exons & \bf{0.0012} & 0.0101 & 0.0023 & 0.030 & 0.0086 & \bf{0.0068} \\
-%HomSap & OOA 5R19 & all sites & 0.012 & 0.012 & \bf{0.0074} & \bf{0.027} & 0.055 & 0.035 \\
-HomSap & OOA 5R19 & exons & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
-%PhoSin & Constant & all sites & \bf{0.024} & 0.025 & 0.024 & 0.21 & 0.24 & \bf{0.19} \\
-PhoSin & Constant & exons & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
-%PhoSin & Vaquita2Epoch 1R22 & all sites & \bf{0.024} & 0.025 & 0.024 & 0.20 & 0.23 & \bf{0.18} \\
-PhoSin & Vaquita2Epoch 1R22 & exons & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
-\bottomrule
-\end{tabular}
-\label{tab:dfe_table}
-\end{table}
-
 \stopsupplement

### 2025-03-19 analysis2_manuscript.tex
one more tweak
Update analysis2_manuscript.tex
Co-authored-by: Andrew Kern <adkern@uoregon.edu>

@@
     in developing or testing their work.
-    The project also promotes open, collaborative, and inclusive work,
+    The project also holds as a central goal open, collaborative, and inclusive work,
     viewing capacity-building across the community as equal in importance to software development.
@@
     and discarding simulations in which the beneficial mutation did not reach (or exceed) a terminal frequency of $95\%$.
-    For each of the 100 genomic segments of chromosome 1, we generated 1,000 replicate datasets under each of the three settings.
+    For each of the 93 genomic segments of chromosome 1, we generated 1,000 replicate datasets under each of the three settings.
     We used the \slim engine with a burn-in of $2N$ generations

### 2025-03-19 analysis2_manuscript.tex
Update analysis2_manuscript.tex
Co-authored-by: Andrew Kern <adkern@uoregon.edu>

@@
     \ref{fig:human-demography}; blue) are fairly similar.
-    Thus, this particular parameterization of background selection and demographic model,
+    Thus, this particular parameterization of background selection and demography,
     in combination with selection only at exonic sites, %in the out-of-Africa demographic model IG: it's probably not just this specific demo model

### 2025-03-19 analysis2_manuscript.tex
Update analysis2_manuscript.tex
Co-authored-by: Andrew Kern <adkern@uoregon.edu>

@@
     \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$.
-    To allow direct comparison, here we have normalized all inference results to the convention of \slim
+    To allow direct comparison, here we have normalized all inference results to the convention used by \slim
     (also used by \stdpopsim; see \textbf{Methods}).

### 2025-03-19 analysis2_manuscript.tex
Update analysis2_manuscript.tex
Co-authored-by: Andrew Kern <adkern@uoregon.edu>

@@
 
-    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}). %, Table \ref{tab:dfe_table}). IG: Table seems less relevant here
+    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}, Table \ref{tab:dfe_table}). 
     Note that while the estimation accuracy for the DFE parameters themselves is informative (Figure \ref{fig:homsap-dfe.ooa}A-B),

### 2025-03-19 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
     methods that are well powered in the face of unknown demographic histories. Indeed, larger differences
-    in power were observed when applying a method to genomes simulated with different demographic histories (YRI versus CEU)
+    in power were observed when applying a method to genomes simulated with different demographic histories (i.e., YRI versus CEU)
     than when calibrating a method using a null model with and without background selection.

### 2025-03-19 analysis2_manuscript.tex
Ilan's final round of edits (#60)
* Ilan's edits:
- marked with %IG non-trivial changes
- removed previous resolved comments

* one more tweak

* Update analysis2_manuscript.tex

Co-authored-by: Andrew Kern <adkern@uoregon.edu>

* Update analysis2_manuscript.tex

Co-authored-by: Andrew Kern <adkern@uoregon.edu>

* Update analysis2_manuscript.tex

Co-authored-by: Andrew Kern <adkern@uoregon.edu>

* Update analysis2_manuscript.tex

Co-authored-by: Andrew Kern <adkern@uoregon.edu>

* Update analysis2_manuscript.tex

* Update analysis2_manuscript.tex

---------

Co-authored-by: Ilan Gronau <gronau

@@
 \author[2,*]{Nathaniel S. Pope}
-\author[2,19,*]{Murillo F. Rodrigues}
+\author[2,3,*]{Murillo F. Rodrigues}
 \author[2,*]{Silas Tittes}
-\author[3,*]{Linh N. Tran}
+\author[4,*]{Linh N. Tran}
 
 % additional authors
-\author[4]{Ornob Alam}
-\author[5]{Maria Izabel A. Cavassim}
-\author[6]{Peter D. Fields}
-\author[7]{Benjamin C. Haller}
-\author[8,9]{Xin Huang}
-\author[10]{Ben Jeffrey}
+\author[5]{Ornob Alam}
+\author[6]{Maria Izabel A. Cavassim}
+\author[7]{Peter D. Fields}
+\author[8]{Benjamin C. Haller}
+\author[9,10]{Xin Huang}
+\author[11]{Ben Jeffrey}
 \author[2]{Kevin Korfmann}
-\author[5]{Christopher C. Kyriazis}
+\author[6]{Christopher C. Kyriazis}
 \author[2]{Jiseon Min}
-\author[11]{Inés Rebollo}
+\author[12]{Inés Rebollo}
 \author[2]{Clara T. Rehmann}
@@
 \author[2]{Chris C. R. Smith}
-\author[12]{Georgia Tsambos}
-\author[10]{Yan Wong}
-\author[13]{Yu Zhang}
+\author[13]{Georgia Tsambos}
+\author[11]{Yan Wong}
+\author[14]{Yu Zhang}
 
 % PI additional authors
-\author[14]{Christian D. Huber}
-\author[13]{Gregor Gorjanc}
-\author[15]{Aaron P. Ragsdale}
+\author[15]{Christian D. Huber}
+\author[14]{Gregor Gorjanc}
+\author[16]{Aaron P. Ragsdale}
 
 % PIs that were very involved in the project
-\author[16]{Ilan Gronau}
-\author[3]{Ryan N. Gutenkunst}
-\author[10]{Jerome Kelleher}
-\author[5]{Kirk E. Lohmueller}
-\author[17]{Daniel R. Schrider}
-\author[2,18]{Peter L. Ralph}
+\author[17]{Ilan Gronau}
+\author[4]{Ryan N. Gutenkunst}
+\author[11]{Jerome Kelleher}
+\author[6]{Kirk E. Lohmueller}
+\author[18]{Daniel R. Schrider}
+\author[2,19]{Peter L. Ralph}
 \author[2]{Andrew D. Kern}
@@
 \affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
-\affil[3]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
-\affil[4]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
-\affil[5]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
-\affil[6]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
-\affil[7]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
-\affil[8]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
-\affil[9]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
-\affil[10]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
-\affil[11]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
-\affil[12]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
-\affil[13]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
-\affil[14]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
-\affil[15]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
-\affil[16]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya 4610101, Israel}}
-\affil[17]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
-\affil[18]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
-\affil[19]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton OR 97006, USA}}
+\affil[3]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton OR 97006, USA}}
+\affil[4]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
+\affil[5]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
+\affil[6]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
+\affil[7]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
+\affil[8]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
+\affil[9]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
+\affil[10]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}
+\affil[11]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
+\affil[12]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
+\affil[13]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
+\affil[14]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
+\affil[15]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
+\affil[16]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
+\affil[17]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya 4610101, Israel}}
+\affil[18]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
+\affil[19]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
 
@@
     and \smcpp \citep{terhorst2017robust}.
-    For \msmc %\citep{Schiffels2020}, IG: no need to double-cite
+    For \msmc 
     we used a random sample of $N=6$ simulated genomes with 20

### 2025-03-19 analysis2_manuscript.tex
first edit
commas

minor edits

@@
 \author[2,19]{Peter L. Ralph}
-\author[2]{Andrew D. Kern}
+\author[2,$\heartsuit$]{Andrew D. Kern}
 
 \affil[*]{\small{These authors contributed equally to the paper.}}
+\affil[$\heartsuit$]{\small{To whom correspondence should be addressed.}} 
 \affil[1]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
-\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
-\affil[3]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton OR 97006, USA}}
-\affil[4]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
-\affil[5]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
-\affil[6]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
-\affil[7]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
-\affil[8]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
+\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene, OR, 97402, USA}}
+\affil[3]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton, OR, 97006, USA}}
+\affil[4]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ, 85721, USA}}
+\affil[5]{\small{Center for Genomics \& Systems Biology, New York University, New York, NY, 10003, USA}}
+\affil[6]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, CA, USA}}
+\affil[7]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609, USA}}
+\affil[8]{\small{Department of Computational Biology, Cornell University, Ithaca, NY, USA}}
 \affil[9]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
@@
 \affil[12]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
-\affil[13]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
+\affil[13]{\small{Department of Genome Sciences, University of Washington, Seattle, WA, 98195, USA}}
 \affil[14]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
-\affil[15]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
-\affil[16]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
+\affil[15]{\small{Department of Biology, Pennsylvania State University, University Park, PA, 16802, USA}}
+\affil[16]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison, WI, USA}}
 \affil[17]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya 4610101, Israel}}
-\affil[18]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
-\affil[19]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
+\affil[18]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill, NC, 27599, USA}}
+\affil[19]{\small{Department of Data Science, University of Oregon, Eugene, OR, 97402, USA}}
 
@@
     \label{abstract}
-    Natural selection is a fundamental evolutionary force that shapes patterns of genetic variation across species. 
+    Selection is a fundamental evolutionary force that shapes patterns of genetic variation across species.
     However, simulations incorporating realistic selection along heterogeneous genomes in complex demographic histories
@@
     Here we present a major extension to the \stdpopsim{} framework that enables simulation of various modes
-    of natural selection, including background selection, selective sweeps,
+    of selection, including background selection, selective sweeps,
     and arbitrary distributions of fitness effects (DFE) acting on annotated subsets of the genome (for instance, exons).
@@
     reveal the sensitivity of DFE-inference methods to model assumptions,
-    and show how genomic features like recombination rate and the density of functional sequence influence power to detect selective sweeps.
+    and show how genomic features, like recombination rate and the density of functional sequence, influence power to detect selective sweeps.
     This extension to \stdpopsim{} provides a powerful, new resource for the population genetics community
@@
     \label{introduction}
-    % natural selection
-    Natural selection is a fundamental force in evolution, shaping the
+    % selection
+    Selection is a fundamental force in evolution, shaping the
     genetic diversity of populations and driving the adaptation of
-    species to their environments. The effects of natural selection
+    species to their environments. The effects of selection
     on genetic variation are complex, and can be difficult to disentangle
@@
     and demography when inferring evolutionary history from genetic data \citep[e.g.,][]{sheehan2016deep,johri2020toward},
-    yet tools for doing so remain limited.
+    yet the number of tools for doing so remain limited.
 
@@
     and forward-in-time simulations of complex demography and selection
-    \citep[e.g.,][]{gillespie1984molecular, hernandez2008flexible, thornton2014c++, haller2019slim}. 
+    \citep[e.g.,][]{gillespie1984molecular, hernandez2008flexible, thornton2014c++, haller2019slim, gaynor2021AlphaSimR}.
     The development of simulation tools has been driven by the need to
@@
     \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
-    to explore the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature},
+    to explore the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature, macleod2014effects},
     and increasingly to provide a basis for machine learning and other
@@
     provides a standardized interface for accessing a wide range of
-    population genetic models, and has begun to be be widely adopted by the community 
+    population genetic models, and has been widely adopted by the community
     \citep[e.g.,][]{speidel2021inferring, wang2021automatic, yang2022trustgwas, degiorgio2022spatially, browning2023fast, schweiger2023ultrafast, temple2024modeling, haag2025pandora}.
@@
     There are relatively few species with published DFE estimates in the literature.
-    \stdpopsim currently has implemented DFEs for five species in the catalogue:
+    \stdpopsim currently has implemented DFEs for five species in the catalog:
     \textit{Arabidopsis thaliana} \citep{huber2018gene},
@@
     report detailing this effort \citep{lauterbur2023expanding}. We welcome contributions from the
-    community and encourage those interested in contributing to visit https://popsim-consortium.github.io/stdpopsim-docs
+    community and encourage those interested in contributing to visit \url{https://popsim-consortium.github.io/stdpopsim-docs}
     or contact one of the PopSim Consortium members for guidance.
@@
     and conserved elements).
-    However, natural selection acting on linked sites---even those that are masked out of the
+    However, selection acting on linked sites---even those that are masked out of the
     analysis---can bias estimates of $N_e(t)$
@@
 
-    To do this we simulated human genomes with and without natural selection.
+    To do this we simulated human genomes with and without selection.
     In both scenarios, we ran three replicate simulations
@@
     but it does not appear to under-estimate the population bottleneck.
-    The inference of \gone, which utilizes identity by descent,
+    The inference of \gone, which utilizes linkage-disequilibrium,
     is targeted toward demographic changes in more recent time frames ($\sim 200$ generations),
@@
     The simulator \slim and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
-    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$.
+    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+2s$.
     Moreover, inferred distributions of selection coefficients are typically scaled by an inferred ancestral
@@
         Three methods for detecting sweeps were applied to simulated data:
-        \sweepfinder (top row -- labelled CLR),
+        \sweepfinder (top row -- labeled CLR),
         \diploshic (middle row),
@@
     for a selective sweep performs very well overall (especially on the CEU sample).
-    However, its accuracy is much more variable along the chromosome compared to the other two method (see Figure \ref{fig:chr1_power}),
+    However, its accuracy is much more variable along the chromosome compared to the other two methods (see Figure \ref{fig:chr1_power}),
     and for quite a few locations along the chromosome it has power below $50\%$
@@
         Three methods for detecting sweeps were applied to simulated data:
-        \sweepfinder (top row-- labelled CLR),
+        \sweepfinder (top row-- labeled CLR),
         \diploshic (middle row),
@@
     As shown in Figures \ref{fig:homsap-summary-stats} and \ref{fig:vaquita-summary-stats},
-    this combination only produces sizeable decreases in polymorphism at exonic positions, 
+    this combination only produces sizable decreases in polymorphism at exonic positions,
     which represent a small minority of sites within the genome, and does not lead to large-scale
@@
     Given the plethora of evidence for recessive deleterious mutations
-    \citep{mukai1972mutation, agrawal2011inferences, huber2018gene, di2024revisiting,},
+    \citep{mukai1972mutation, agrawal2011inferences, huber2018gene, di2024revisiting},
     these results suggest the need for care in DFE inference.
@@
     harder problem as some mutations assumed to be free of selection
-    may in reality be under selection.
+    may in reality be under selection and strength of selection can change over time or with environment.
 
@@
     Following the protocol established previously for other aspects of \stdpopsim{} \citep{adrion2020community},
-    new DFEs are subjet to a rigorous quality control (QC) process,
+    new DFEs are subject to a rigorous quality control (QC) process,
     by which a separate individual re-implements the same DFE, and the two implementations
-    are programatically tested for equality.
+    are programmatically tested for equality.
     %IG: tried to put a more positive spin on the DFE bug example
@@
     Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
-    \citet[\stdpopsim label \texttt{Gamma\_R22}]{robinson2022critically}  to exons annotated from the vaquita genome assembly
+    \citet[\stdpopsim label \texttt{Gamma\_R22}]{robinson2022critically} to exons annotated from the vaquita porpoise genome assembly
     \citet[\stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}]{morin2021}.
@@
     %
-    We assessed the accuracy of demographic inferences from three software packages:
+    We assessed the accuracy of demographic inferences from four software packages:
     \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent},
@@
     All methods used the same exon annotations as generated in the simulations. The total number of coding sites
-    in the human and vaquita genomes used were 61,644,801 and 31,039513 respectively.
+    in the human and vaquita genomes used were 61,644,801 and 31,039,513 respectively.
 
@@
     and the Robertson Foundation.
+    G.~Gorjanc and Y.~Zhang were supported in part by BBSRC (BBS/E/RL/230001A, BBS/E/RL/230001C, BB/T014067/1),
+    Chinese Scolarship Council (CSC) and Genus.
 
@@
     \label{fig:vaquita-summary-stats}
-    Genetic diversity in vaquite simulations.
+    Genetic diversity in vaquita simulations.
     Nucleotide diversity ($\pi$) was calculated for genomes simulated
@@
         Mean absolute value of selection coefficient ($\lvert E(s) \rvert$, panel A) and shape parameter (panel B)
-        are shown for each DFE inferred from all three simulated datase
+        are shown for each DFE inferred from all three simulated datasets
         with median values marked by horizontal bars

### 2025-03-20 analysis2_manuscript.tex
Minor edits

--- 2025-03-19 analysis2_manuscript.tex
@@
     Following the protocol established previously for other aspects of \stdpopsim{} \citep{adrion2020community},
-    new DFEs are subjet to a rigorous quality control (QC) process,
+    new DFEs are subject to a rigorous quality control (QC) process,
     by which a separate individual re-implements the same DFE, and the two implementations

### 2025-03-20 analysis2_manuscript.tex
Minor edits (#64)
Co-authored-by: Andrew Kern <adkern@uoregon.edu>

--- 2025-03-19 analysis2_manuscript.tex
@@
 \author[2,19]{Peter L. Ralph}
-\author[2,$\heartsuit$]{Andrew D. Kern}
+\author[2]{Andrew D. Kern}
 
 \affil[*]{\small{These authors contributed equally to the paper.}}
-\affil[$\heartsuit$]{\small{To whom correspondence should be addressed.}} 
 \affil[1]{\small{Section for Molecular Ecology and Evolution, Globe Institute, University of Copenhagen, Denmark}}
-\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene, OR, 97402, USA}}
-\affil[3]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton, OR, 97006, USA}}
-\affil[4]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ, 85721, USA}}
-\affil[5]{\small{Center for Genomics \& Systems Biology, New York University, New York, NY, 10003, USA}}
-\affil[6]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, CA, USA}}
-\affil[7]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609, USA}}
-\affil[8]{\small{Department of Computational Biology, Cornell University, Ithaca, NY, USA}}
+\affil[2]{\small{Institute of Ecology and Evolution, University of Oregon, Eugene OR 97402, USA}}
+\affil[3]{\small{Division of Genetics, Oregon National Primate Center, Oregon Health and Science University, Beaverton OR 97006, USA}}
+\affil[4]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson, AZ 85721, USA}}
+\affil[5]{\small{Center for Systems Biology, New York University, New York NY, 10012 USA}}
+\affil[6]{\small{Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles CA, USA}}
+\affil[7]{\small{The Jackson Laboratory, Bar Harbor, ME, 04609 USA}}
+\affil[8]{\small{Department of Computational Biology, Cornell University, Ithaca NY, USA}}
 \affil[9]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
@@
 \affil[12]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
-\affil[13]{\small{Department of Genome Sciences, University of Washington, Seattle, WA, 98195, USA}}
+\affil[13]{\small{Department of Genome Sciences, University of Washington, Seattle WA, USA}}
 \affil[14]{\small{The Roslin Institute and Royal (Dick) School of Veterinary Studies, University of Edinburgh, Edinburgh EH25 9RG, UK}}
-\affil[15]{\small{Department of Biology, Pennsylvania State University, University Park, PA, 16802, USA}}
-\affil[16]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison, WI, USA}}
+\affil[15]{\small{Department of Biology, Pennsylvania State University, University Park, PA 16802, USA}}
+\affil[16]{\small{Department of Integrative Biology, University of Wisconsin-Madison, Madison WI, USA}}
 \affil[17]{\small{Efi Arazi School of Computer Science, Reichman University, Herzliya 4610101, Israel}}
-\affil[18]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill, NC, 27599, USA}}
-\affil[19]{\small{Department of Data Science, University of Oregon, Eugene, OR, 97402, USA}}
+\affil[18]{\small{Department of Genetics, University of North Carolina at Chapel Hill, Chapel Hill NC 27599, USA}}
+\affil[19]{\small{Department of Data Science, University of Oregon, Eugene OR 97402, USA}}
 

### 2025-03-20 analysis2_manuscript.tex
changed funding (#66)

@@
     K.~Lohmueller was supported by NIH award R35GM119856.
-    P.~Ralph, B.~Haller, A.~Kern, and B.~Jefferey
+    P.~Ralph, B.~Haller, J.~Kelleher, A.~Kern, and B.~Jefferey
     were supported in part by NIH award R01HG012473.
-    J.~Kelleher was supported by EPSRC grant EP/X024881/1
-    and the Robertson Foundation.
     G.~Gorjanc and Y.~Zhang were supported in part by BBSRC (BBS/E/RL/230001A, BBS/E/RL/230001C, BB/T014067/1),

### 2025-03-20 analysis2_manuscript.tex
removing Ben from NIH funding
Huber funding
chris' edits

@@
 \usepackage{makecell}
+\usepackage[left]{lineno}
 
@@
 \newcommand{\igcomment}[1]{\textcolor{orange}{IG: #1}}
-
+\linenumbers
 \title{Accessible, realistic genome simulation with selection using \stdpopsim}
@@
     \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining,zhen2021greater},
-    humans \citep{huber2017determining,kim2017inference,kyriazis2022using,castellano2019compmarison,castellano2019compmarison,rodrigues2024shared},
+    humans \citep{huber2017determining,kim2017inference,kyriazis2023using,castellano2019compmarison,castellano2019compmarison,rodrigues2024shared},
     \textit{Mus musculus} \citep{booker2021selective},
@@
     estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
-    Indeed, the DFE labeled \texttt{HomSap/Mixed\_K23} is recommended by \citet{kyriazis2022using}
-    for ``general'' use across species (lacking more specific information).
+    Indeed, the DFE labeled \texttt{HomSap/Mixed\_K23} is recommended by \citet{kyriazis2023using}
+    for ``general'' use across mammals (lacking more specific information).
     Furthermore, the user can specify a custom DFE and provide their own annotations
@@
     through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
-    as well as through identity by descent information \citep{santiago2020recent}.
+    as well as through linkage-disequilibrium information \citep{santiago2020recent}.
     Since all of these methods assume neutrality, they are typically applied to genomic
@@
     K.~Lohmueller was supported by NIH award R35GM119856.
-    P.~Ralph, B.~Haller, J.~Kelleher, A.~Kern, and B.~Jefferey
+    P.~Ralph, B.~Haller, J.~Kelleher, and A.~Kern
     were supported in part by NIH award R01HG012473.
@@
     Chinese Scolarship Council (CSC) and Genus.
+    C ~Huber was supported by NIH award R35GM146886.
 
@@
 \beginsupplement
+\resetlinenumber
 

### 2025-03-20 analysis2_manuscript.tex
changes and comments from bhaller

--- 2025-03-20 analysis2_manuscript.tex
@@
     However, simulations incorporating realistic selection along heterogeneous genomes in complex demographic histories
-    is challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and to explore theoretical predictions.
+    are challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and to explore theoretical predictions.
+    % BCH: "simulations... are challenging".  I would also reverse the order: "limiting our ability to explore... and to benchmark", because as
+    % it is written not, "...and to explore" is very far from its first half "limiting our ability to...", making it hard for the reader to parse.
     \stdpopsim is a community-maintained simulation library that already provides an extensive catalog of
@@
     and arbitrary distributions of fitness effects (DFE) acting on annotated subsets of the genome (for instance, exons).
+    % BCH: I'd put "DFEs", since it is plural, although I recognize that others might disagree...
     This extension maintains \stdpopsim's core principles of reproducibility and accessibility while adding support
@@
     reveal the sensitivity of DFE-inference methods to model assumptions,
-    and show how genomic features, like recombination rate and the density of functional sequence, influence power to detect selective sweeps.
-    This extension to \stdpopsim{} provides a powerful, new resource for the population genetics community
+    and show how genomic features, like recombination rate and functional sequence density, influence power to detect selective sweeps.
+    % BCH: "the density of functional sequence" does not seem grammatical to me; either "sequences" plural, or the fix I've recommended, which I think is better.
+    This extension to \stdpopsim{} provides a powerful new resource for the population genetics community
     to explore the interplay between selection and other evolutionary forces in a reproducible, low-barrier framework.
@@
     that can mimic the effects of selection \citep{simonsen1995properties},
-    and lead to spurious inferences about the strength and targets of genetic adaptation
+    leading to spurious inferences about the strength and targets of genetic adaptation
     \citep{simonsen1995properties,akey2004population,nielsen2005genomic}.
@@
     \citep[e.g.,][]{gillespie1984molecular, hernandez2008flexible, thornton2014c++, haller2019slim, gaynor2021AlphaSimR}.
-    The development of simulation tools has been driven by the need to
-    understand the effects of complex evolutionary processes on genetic
+    The development of simulation tools has been driven by several needs:
+    understanding the effects of complex evolutionary processes on genetic
     variation \citep[e.g.,][]{peischl2013accumulation, thornton2019polygenic, galloway2020few, zhang2020impact},
-     to provide a null model for hypothesis testing
+    obtaining a null model for hypothesis testing
     \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
-    to explore the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature, macleod2014effects},
-    and increasingly to provide a basis for machine learning and other
+    exploring the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature, macleod2014effects},
+    and increasingly, providing a basis for machine learning and other
     simulation-based inference methods
@@
     However, joint simulation of complex demography and selection on realistic genomes is challenging,
-    requiring many choices including the strength and rate of selection,
-    and a parameterized demographic model,
-    as well as implementation of genetic maps and other genomic features related to the strength of selection.
+    requiring a parameterized demographic model,
+    many choices including the strength and rate of selection,
+    and implementation of genetic maps and other genomic features related to the strength of selection.
+    % BCH: this reordering reads better to me since it was hard to tell whether "many choices" encompassed the parameterized demographic model or not;
+    % perhaps a different way of restructuring it would be better, if I've distorted the intended meaning, such as "many choices regarding parameters
+    % controlling selection and demography", e.g.
     This is a daunting task, and can be a barrier to
     the adoption of simulation-based methods in population genetics.
-    Furthermore, different simulation and inference tools may use different structures
+    Furthermore, different simulation and inference tools often use different structures % BCH: considered may -> might, and then decided "often" is more accurate!
     and parameter scalings for their models,
@@
     to have a simple, unified framework for simulating both neutral and
-    non-neutral evolutionary processes and to be able to compare the
+    non-neutral evolutionary processes, and to be able to compare the
     results of these simulations to empirical data in a manner that is
@@
     in developing or testing their work.
-    The project also holds as a central goal open, collaborative, and inclusive work,
-    viewing capacity-building across the community as equal in importance to software development.
+    A central goal of the project is open, collaborative, and inclusive work;
+    we view capacity-building across the community as equal in importance to software development.
 
@@
 \section*{Implementing Selection in \stdpopsim}
+    % BCH: I mostly skimmed/skipped this section; if you'd find it useful for me to review it, let me know, I'm just a bit swamped.  :->
     \label{selection}
@@
         annotations. }
+        % BCH: I notice that a rescaling factor of 20 is used here, which gives me pause.  I think this is my only major
+        % comment on this manuscript: I wonder whether we ought to add a brief discussion of the dangers of rescaling,
+        % specifically with respect to simulations involving selection since that is the topic at hand?  Dabi and Schrider
+        % discuss this at some length in their recent paper.  I'm not suggesting rehashing everything they say, but rather,
+        % adding one paragraph talking about the fact that stdpopsim offers a rescling parameter (as seen in this figure),
+        % but that the use of it might be particularly problematic/unpredictable in simulations involving selection,
+        % especially for large values of Q, and thus it ought to be used with caution, and with an effort to confirm that
+        % the rescaling factor chosen is not substantially distorting the key results from the simulations, with a cite
+        % to Dabi and Schrider.  (And probably also a cite to the other recent rescaling paper, which I forget the authors of.)
+        % My impression from D&S is that Q=20 is quite high, and might well distort things significantly; so I wonder if we
+        % also might want to show a smaller Q, like 2 or 5, in this figure so as not to give our imprimatur to people using Q=20!
     \end{figure}
@@
     In this paper, we have presented an important new addition to the \stdpopsim{} library
-    that allows for simulating genetic variation in the presence of selection.
+    that enables the simulation of genetic variation in the presence of selection.
     We have demonstrated the utility of this new API by showing how it can be used to benchmark
@@
     benchmarking, but rather to highlight the utility of the \stdpopsim resource.
-    Having said that, there are a number of interesting results that bear discussion.
+    Having said that, there are a number of interesting results that merit discussion.
     The first pertains to estimation of demographic history ($N_e(t)$) in the context of
-    linked purifying selection. Our simulations of humans and vaquita genomes
+    linked purifying selection. Our simulations of human and vaquita genomes
     model deleterious mutations only at exonic positions of the genome.
@@
     strong selection acting on them \citep{katzman2007human,mclean2008dispensability}.
-    We thus caution extrapolating from our results. Future studies might build upon
+    We thus caution regarding extrapolation from our results. Future studies might build upon
     the annotations we used here to include selection at other portions of the genome as well
@@
     We found that the three DFE-inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}A-B).
-    This performance is not surprising as decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
+    This performance is not surprising since decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$, % "as" vs. "since" is a pet peeve :->
     or an apparent decrease in the amount of selection.
@@
     in humans when considering different values of $h$.
-    Intriguingly, additive models showed a satisfactory fit to the SFS of nonsynonymous variants in humans
+    Intriguingly, additive models showed a satisfactory fit to the SFS of nonsynonymous variants in humans,
     and more complex models of dominance did not substantially improve the fit.
@@
     Further, inference of the DFE was done here using exon annotations of the genome in question.
-    A key distinction between our analyses and those done in practice
-    is that we have complete knowledge of which sites are under selection in the simulation,
-    thus our inferences represent a best case scenario. Empirical inference of the DFE is a much
-    harder problem as some mutations assumed to be free of selection
-    may in reality be under selection and strength of selection can change over time or with environment.
+    A key distinction between our analyses and those done in practice % BCH: I don't like "in practice"; "those done in empirical work"?
+    is that we have complete knowledge of which sites are under selection in the simulation.
+    Our inferences therefore represent a best case scenario. Empirical inference of the DFE is a much
+    harder problem, since some mutations assumed to be free of selection
+    might in reality be under selection, and the strength of selection can change over time or across environments.
 
     Regarding sweep detection, the inclusion of selection models into \stdpopsim allowed us to readily compare the
-    effectiveness of different methods on data reflecting the demographic history, recombination rate variation,
+    effectiveness of different detection methods using data that reflects the demographic history, recombination rate variation,
     and genomic locations of negatively selected sites present in human populations. These results clearly demonstrate
@@
     recombination rates \citep[e.g.,][]{schrider2017soft}. In addition, our results
-    illustrate a strength of likelihood methods like \sweepfinder, that do not require a predefined window size.
+    illustrate a strength of likelihood methods like \sweepfinder that do not require a predefined window size.
     Window-based methods, such as \diploshic, tend to lag in performance when the genomic region affected by the sweep is
@@
     scenarios, including the inclusion of background selection, our results underscore the importance of developing
-    methods that are well powered in the face of unknown demographic histories. Indeed, larger differences
+    methods that are well powered in the face of unknown demographic histories.
+Indeed, larger differences % BCH: not crazy about "well powered"; "with adequate statistical power even in the face of..."?
     in power were observed when applying a method to genomes simulated with different demographic histories (i.e., YRI versus CEU)
@@
     models, and for benchmarking methods for the analysis of genetic data
-    on common ground in a way that was not possible before.
+    on common ground in a way that was not possible before. % BCH: "on common ground" seems like an odd phrasing to me? "on a level playing field", perhaps?
 
@@
     new DFEs are subject to a rigorous quality control (QC) process,
-    by which a separate individual re-implements the same DFE, and the two implementations
+    by which a different individual re-implements the same DFE, and the two implementations
     are programmatically tested for equality.
     %IG: tried to put a more positive spin on the DFE bug example
-    This QC process has proven its effectiveness in numerous occasions in detecting implementation bugs
+    This QC process has proven its effectiveness on numerous occasions in detecting implementation bugs
     and resolving modeling ambiguities.
@@
     at the time of writing, the library includes over 11,000 lines of testing code,
-    that run 2,862 unit tests, including comparisons to QC implementations
+    comprising 2,862 unit tests, including comparisons to QC implementations
     and cross-checks for agreement, including between \slim and \msprime implementations of the same models.
@@
 We thank Tom Booker for discussions and encouragement to incorporate realistic BGS models into \stdpopsim.
-We thank the numerous users of the \stdpopsim package and well as
+We thank the numerous users of the \stdpopsim package, as well as
 workshop participants who provided feedback and suggestions for improvements.

### 2025-03-21 analysis2_manuscript.tex
Python with capital P

--- 2025-03-20 analysis2_manuscript.tex
@@
         can be used in simulations.
-        \textbf{(B)} The python API and \textbf{(C)} the command line interface.
+        \textbf{(B)} The Python API and \textbf{(C)} the command line interface.
         The user specifies a species, a portion of
@@
     a genetic map if available, a model of demography, and pairings of DFEs with genomic segments
-    using either the python API or the command-line interface (Figure \ref{fig:schematic}B-C).
+    using either the Python API or the command-line interface (Figure \ref{fig:schematic}B-C).
     Simulations with selection are generated

### 2025-03-21 analysis2_manuscript.tex
Edits to selective_sweep API description

@@
         can be used in simulations.
-        \textbf{(B)} The Python API and \textbf{(C)} the command line interface.
+        \textbf{(B)} The python API and \textbf{(C)} the command line interface.
         The user specifies a species, a portion of
@@
     a genetic map if available, a model of demography, and pairings of DFEs with genomic segments
-    using either the Python API or the command-line interface (Figure \ref{fig:schematic}B-C).
+    using either the python API or the command-line interface (Figure \ref{fig:schematic}B-C).
     Simulations with selection are generated
@@
     % sweep interface
-    In addition to the \texttt{DFE} class, we have also implemented a class (\texttt{stdpopsim.selective\_sweep})
-    that enables selective sweeps to be simulated in \stdpopsim.
-    This class augments a demographic model with an event
-    which conditions on the introduction of a selected mutation at a given time and position
-    and with a given selection coefficient. Further, the user can specify the minimum frequency
-    of the selected allele at the time of sampling. As this functionality is implemented
-    on top of the existing \stdpopsim API, it can be combined with other models of selection
-    and demography to provide a model of the combined effects of multiple, disparate evolutionary processes
+    While the \texttt{DFE} class facilitates the simulation of non-neutral variation across a sequence,
+    it does not provide a means of specifying the origin and fate of particular selected alleles.
+    Thus, we have also implemented an API in \stdpopsim (\texttt{stdpopsim.selective\_sweep}) that enables 
+    selective sweeps to be simulated with a high degree of control.
+    This function augments a demographic model by introducing an advantageous mutation over a given time interval,
+    at a given position along the sequence; and conditions the simulation such that the mutation is above a particular 
+    frequency at the beginning and end of the sweep period.
+    As this functionality is implemented on top of the existing \stdpopsim API, it may be combined with other 
+    models of selection and demography to simulate the combined effects of multiple, disparate evolutionary processes
     on genetic variation in a population.

### 2025-03-21 analysis2_manuscript.tex
clean up of Ben's edits

--- 2025-03-21 analysis2_manuscript.tex
@@
     However, simulations incorporating realistic selection along heterogeneous genomes in complex demographic histories
-    is challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and to explore theoretical predictions.
+    are challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and to explore theoretical predictions.
     \stdpopsim is a community-maintained simulation library that already provides an extensive catalog of
@@
     reveal the sensitivity of DFE-inference methods to model assumptions,
-    and show how genomic features, like recombination rate and the density of functional sequence, influence power to detect selective sweeps.
-    This extension to \stdpopsim{} provides a powerful, new resource for the population genetics community
+    and show how genomic features, like recombination rate and functional sequence density, influence power to detect selective sweeps.
+    This extension to \stdpopsim{} provides a powerful new resource for the population genetics community
     to explore the interplay between selection and other evolutionary forces in a reproducible, low-barrier framework.
@@
     the adoption of simulation-based methods in population genetics.
-    Furthermore, different simulation and inference tools may use different structures
+    Furthermore, different simulation and inference tools often use different structures
     and parameter scalings for their models,
@@
     We found that the three DFE-inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}A-B).
-    This performance is not surprising as decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
+    This performance is not surprising since decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
     or an apparent decrease in the amount of selection.
@@
     Further, inference of the DFE was done here using exon annotations of the genome in question.
-    A key distinction between our analyses and those done in practice
-    is that we have complete knowledge of which sites are under selection in the simulation,
-    thus our inferences represent a best case scenario. Empirical inference of the DFE is a much
-    harder problem as some mutations assumed to be free of selection
-    may in reality be under selection and strength of selection can change over time or with environment.
+    A key distinction between our analyses and those done in empirical work 
+    is that we have complete knowledge of which sites are under selection in the simulation.
+    Our inferences therefore represent a best case scenario. Empirical inference of the DFE is a much
+    harder problem, since some mutations assumed to be free of selection
+    might in reality be under selection, and the strength of selection can change over time or across environments.
 
     Regarding sweep detection, the inclusion of selection models into \stdpopsim allowed us to readily compare the
-    effectiveness of different methods on data reflecting the demographic history, recombination rate variation,
+    effectiveness of different methods using data that reflects the demographic history, recombination rate variation,
     and genomic locations of negatively selected sites present in human populations. These results clearly demonstrate
@@
     scenarios, including the inclusion of background selection, our results underscore the importance of developing
-    methods that are well powered in the face of unknown demographic histories. Indeed, larger differences
+    methods with adequate statistical power even in the face of unknown demographic histories. Indeed, larger differences
     in power were observed when applying a method to genomes simulated with different demographic histories (i.e., YRI versus CEU)

### 2025-03-21 analysis2_manuscript.tex
minor edits (#73)

@@
 \affil[11]{\small{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford OX3 7LF, UK}}
-\affil[12]{\small{Instituto Nacional de Investigación Agropecuaria (INIA), Estación Experimental Las Brujas, Ruta 48 km 10, Canelones, Uruguay}}
+\affil[12]{\small{Department of Agronomy and Plant Genetics, University of Minnesota, St. Paul, MN, 55108, USA}}
 \affil[13]{\small{Department of Genome Sciences, University of Washington, Seattle, WA, 98195, USA}}
@@
     in developing or testing their work.
-    A central goal of the project is open, collaborative, and inclusive work;
+    A central goal of the \stdpopsim project is open, collaborative, and inclusive work;
     we view capacity-building across the community as equal in importance to software development.
@@
     genomic regions with a particular annotation, specific to a given species available
-    through the \stdpopsim API, and combined with a parameterized model of
+    through the \stdpopsim application programming interface (API), and combined with a parameterized model of
     demography and recombination to provide a realistic simulation of
@@
     % current numbers of DFEs / species / etc
-    At this release (version 0.3.0) the \stdpopsim catalog comprises 27 species (i.e. genome representations)
+    At this release (version 0.3.0) the \stdpopsim catalog comprises 27 species (i.e. genome representations),
+    including vertebrates, invertebrates, plants, and bacteria,
     for which we have included 35 demographic models, 38 genetic maps, 12 DFEs, and 8 annotation objects. %ADK: checked this on 3/10/25

### 2025-03-22 analysis2_manuscript.tex

@@
     (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
+    We note in passing that aggressive scaling of forward-in-time simulations
+    can lead to quite different outcomes when selection in involved \cite{dabi2025population},
+    and we caution users against using high degrees of scaling for drawing
+    final inferences, although we recommend it during development to save
+    on computational costs. 
     All simulations were implemented using a \snakemake workflow,

### 2025-03-22 analysis2_manuscript.tex
Changed scaling factor in schematic (#75)
* altered shematic

* added text on scaling to methods

* fixed curly quotes

* refs from Dan

--- 2025-03-21 analysis2_manuscript.tex
@@
     (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
+    We note that aggressive scaling of forward-in-time simulations
+    can lead to quite different outcomes when selection in involved
+    \cite{uricchio2014robust,ferrari2024towards, dabi2025population},
+    and we caution users against using high degrees of scaling for drawing
+    final inferences, although we recommend it during development to save
+    on computational costs. 
     All simulations were implemented using a \snakemake workflow,

### 2025-05-21 review-responses.tex
resubmission stuff (#80)
* adding original submission text file

* added review response stuff

\begin{minipage}[b]{2.5in}
    Resubmission Cover Letter \\
    {\it Molecular Biology and Evolution}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{2.5in}
      Andrew Kern (on behalf of all authors)  \\
    \today
  \end{minipage}

  \vskip 2em

  \noindent
  {\bf To the Editor(s) -- }

  \vskip 1em

  We are writing to submit a revised version of our manuscript,
  MBE-25-0312
  titled
  ``Accessible, realistic genome simulation with selection using stdpopsim''.
  We appreciate the thoughtful comments by the reviewers.
  We've done our best to address all the comments,
  as detailed in our Response to Reviewers.

  In addition to a Response to Reviewers (in which page/line numbers refer to the revised manuscript file),
  we also provide a pdf with differences to the previously submitted version highlighted.

  \vskip 2em

  \noindent \hspace{4em}
  \begin{minipage}{3in}
  \noindent
  {\bf Sincerely,}

  \vskip 2em

  {\bf
     Andrew Kern (on behalf of all authors)
  }\\
  \end{minipage}

  \vskip 4em

  \pagebreak

  %%%%%%%%%%%%%%
  \reviewersection{AE}
  \begin{quote}
    Two experts in population genetics examined this manuscript and gave Very High and Medium marks in the Impact of the resource reported. 
    Despite this mixed evaluations, I think this extension of stdpopsim contributed by many authors should be very useful in the community of population genomics. 
    However, referees found many places in the manuscript that need clarification. 
    These comments should be addressed in revision. 
  \end{quote}

  Thanks for your comments on our paper.

  \begin{point}{\revref} % 
    In addition, regarding the point discussed in lines 368-373, a recent paper Marsh \& Johri 2024 MBE 41(7):msae118 
    might be acknowledged because it also developed simulations including the annonation for D. mel and humans and also mutation and recombination rate heterogeneity. 
  \end{point}

  \reply{
    TODO. \revref
  }

  \reviewersection{1}
  \begin{quote}
    In this manuscript, a large team of authors (Gower, Pope, Rodrigues, Tittes et al) present the next 
    set of developments for the stdpopsim software, which facilitates "standard" population-genetic 
    simulations that can be set to follow estimated demographic and genetic (mutation rates, recombination maps) 
    parameters in a number of species and populations, as well as a framework for entering new species and populations 
    into the collection. stdpopsim has become an important tool in population-genetic research. 
    The authors here represent a large community of people invested in the maintenance and improvement of the software, 
    which is open and undergoes thorough unit testing. 
    The addition of selection to stdpopsim's capabilities is a substantial advance warranting a new publication. 
    MBE also seems to me an appropriate outlet.  
\end{quote}

  \begin{quote}
    In addition to announcing the new features available in stdpopsim, the paper describes three applications of the software: 
    one on demographic inference in the presence of background selection, another on DFE estimation, and a third on detection of selective sweeps. 
    Two of these contrast the results obtained using human vs. vaquita porpoise parameters.  
\end{quote}

  \begin{quote}
    I am generally supportive of the paper and of the broader stdpopsim project. 
    That said, I do have some constructive thoughts and questions:  \end{quote}

  Thanks -- we're glad you appreciate both the project and the paper.

  \begin{point}{}
    Perhaps my main constructive reaction is that this paper is a bit light on information about the ways in which the new selection features are implemented. 
    There is a section "implementing selection in stdpopsim", but it is not always easy to figure out from the description here how things work under the hood. 
    Instead, the section is more like a short tour of the new features aimed at users. 
    Moreover, there is no section of the methods that really goes over the details of implementation. 
    The ms says that SLiM is the backend, but there aren't many details about how information passed into stdpopsim is translated into SLiM. 
    It's also not clear to me how the selective sweep mode is implemented, given that it gives users the option to condition on some range of starting and ending allele frequencies. 
    I assume that the approach is to run full SLiM iterations until the sweep ends up as desired (though it may not be)? 
    But one might also simulate the allele-frequency trajectory for the sweeping allele alone until one "works", 
    and then do coalescent simulations to fill in flanking variation conditional on the allele-frequency trajectory, 
    a la Hudson \& Kaplan '88 (though I don't quite see how one would do this simultaneously with background selection, 
    which the sweep simulations include). In any case, I'm not sure how the conditions are being achieved from reading the paper. 
    (I've refrained from looking at the online documentation for stdpopsim 0.3.0, where there may be more answers, in order to review the paper as an independent product.) 
    I don't think the paper necessarily needs loads more text on this, but I think there should be more information hinting at implementation in main text and probably a new methods subsection with further details.  
\end{point}

  \reply{
  We have added further descriptions of the implementation to the TODO
  }

  \begin{point}{\revref} %line 212
    Relatedly, why is the dominance coefficient currently limited to between 0 and 1? It's fine that it is for now, but I was curious what makes it more difficult to include other values of h.
  \end{point}

  \reply{
      TODO. \revref
  }

  \begin{point}{\revref} 
    I think it would be helpful to say a bit more about the vaquita earlier in the manuscript. Although some of the authors have worked on the vaquita recently, it is still a fairly obscure organism, 
    so it would be worth describing its unfortunate situation and why it is an interesting comparison with humans throughout these simulations.  
\end{point}

  \reply{
     TODO. \revref
  }

  \begin{point}{\revref} % 
    It seems as though everything on the backend is being done by SLiM in the demonstrations here 
    (unless I'm missing something about how the sweeps are done), so the runtimes are not really the product of the effort of the authors here. 
    Still, I think it would be good to provide a bit more information on the runtimes for prospective users to get a sense of what doing selection in stdpopsim entails. 
    (It's mentioned that the sweeps were time-consuming, but details aren't given as far as I can tell.)  
\end{point}

  \reply{
      TODO. \revref
  }

  \begin{point}{\revref} % line 577
    Point of clarification about the flexibility of DFE specification: Is the DFE allowed to be a mixture distribution in a single genomic region? 
    Generally, the paper mentions users can provide a custom DFE, but what does that mean in practice? Does it have to be of a particular distribution family or form? 
    (i.e. the paper mentions "fixed value, normally-distributed, gamma-distributed"). 
    The vaquita DFE seems to be a mixture of several nonoverlapping DFEs with different h values, which suggests to me substantial flexibility, 
    but I'm not quite clear on the limits. Most of the figures feature fairly few replicates, which makes me think that these simulations are necessarily time/compute intensive.
  \end{point}

  \reply{
 TODO
  }

  \begin{point}{\revref} % line 577
    Sentence starting on line 383 about difficulty identifying h and s separately seems to be missing some important citations, 
    e.g. Fuller et al 2019 (pmid 30962618) and citations therein, including Simons et al 2014; Wright 1937.
  \end{point}

  \reply{
    TODO. \revref
  }

  \begin{point}{\revref} % line 577
    It might bear further comment that the gamma distribution is not flexible enough to accommodate the distribution of hs in the vaquita simulation 
    (which is what the DFE inference methods can really see), even if the s values are gamma distributed (bc of the big jump caused in hs by the switch in dominance). 
    This is interesting given how often gammas are used in DFE inference. The paper does mention that using h=1/2 as an assumption can lead to bad estimation of s, 
    but the point that a relationship between s and h can cause hs to be non-gamma even when s is gamma is perhaps a little more subtle. 
    Generally, I thought the lessons learned from the DFE exercise here could be put more clearly.
  \end{point}

  \reply{
    TODO. \revref
  }

  \begin{point}{\revref} % line 577
    Related to the previous point, I found it a bit odd to that both the style of the DFE (and in particular the h/s relationship) 
    and the demography switched when comparing vaquita and human DFE. 
    It seemed like a missed opportunity to do something hinted at earlier in the manuscript: 
    namely, to try pairing demographies and DFEs from different species. 
    I.e. it would be interesting in this context to look at simulations that have the human demography and vaquita DFE and vice versa, 
    even if only in a supplementary figure.
  \end{point}

  \reply{
    TODO. \revref
  }

  \begin{point}{\revref} % line 577
    Figure S1, right panel. Is pi being computed from all the genomes (CEU/CHB/YRI)? 
    I think it's okay given the point that's being made, 
    but it might strike some readers as odd given how much is often made of the pi differences 
    in these groups under the estimated demography, so maybe it should be said explicitly what's being done.
  \end{point}

  \reply{
    TODO. \revref
  }

  \begin{point}{\revref} % line 577
    Last comment, and this may be an idiosyncratic opinion on my part, so I do not expect the authors to implement, 
    respond to, or even quote this point in the response letter. I'm merely offering a reaction. 
    In my opinion, three motifs are at a slightly awkward level of detail/coverage. 
    On one hand, they're not enough to please partisans or enthusiasts of any of these three areas. 
    (The paper is clear about the fact that they're not meant to be thorough benchmarks.) 
    On the other hand, they are a big part of the paper, substantial on their own, 
    and the bulk of the discussion is given over to their substantive results rather than to stdpopsim per se. 
    I personally find many of the results interesting, but I also wondered whether the paper might be stronger 
    if these segments were streamlined. One way to do this might be to focus less on the three topics themselves, 
    but to change the emphasis to how selection results change when the species changes, 
    i.e. to focus the thread of the results primarily on a quick inspection of the human/vaquita differences
    (one might imagine adding in Drosophila too in such a framing) rather than on separate questions about the effect of 
    BGS on demographic estimation and sweep inference, and on the effect of h-s relationship on DFE inference. 
    Again, this is merely offered in the spirit of a friendly comment, and I expect no response.
  \end{point}

  \reply{
    Don't respond to this?. \revref
  }

  %%%%%%%%%%%%%%
  \reviewersection{2}
  \begin{quote}
    In this manuscript, Gower et al. extend the functionality of stdpopsim to include selection. 
    They test the performance of a few different methods that infer historical population sizes, 
    the distribution of fitness effects of new deleterious mutations, and those that detect selective sweeps using the simulations in SLiM. 
    The added functionality will be useful to the users of stdpopsim.
\end{quote}

  Thanks for your constructive comments on our paper.

  \begin{point}{\revref} % \&75
    How do you deal with overlapping annotations in a gff?
  \end{point}

  \reply{
    TODO. \revref
  }

  \begin{point}{}
    The inference of the deleterious DFE with GRAPES is quite inaccurate. 
    Is it possible that there is a factor of 2 error involved?
  \end{point}

  \reply{
TODO. \revref
  }

  \begin{point}{\revref} % 21:
    Figure 3c does not have any error bars, it’s difficult to conclude if the inference is actually significantly different from the truth. 
    Which “N” was used to obtain estimates of “s” in Figure 3a? 
    This should be mentioned in the legend.  \end{point}

  \reply{
  TODO.
  }

  % INTRO

  \begin{point}{\revref} % 24:
    You cite only Gillespie 1991 on line 56. 
    Requires many more citations, with at least one to Kimura.  
\end{point}
  \reply{
  TODO. \revref
  }

  \begin{point}{\revref} % 27:
    What fraction of the genome is under selection in the human genome you simulated?
  \end{point}

  \reply{
 TODO. \revref
  }

  \begin{point}{\revref} % 35:
   Bayesian $\to$ probabilistic (I think this is more general and would encompas any model-based approach, say max likelihood etc.; alternatively consider using model-based, but probabilistic/statistical assumes we have an underlying model)
  \end{point}

  \reply{
  Done.
  }

  \begin{point}{\revref} % 47:
    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature.” (line 349-350) -> I'm not sure what this means. 
    It's unclear to me why background selection will interfere with a sweep signature. 
    Do you have a citation for this?  
\end{point}

  \reply{
  TODO.
  }

### 2025-05-22 review-responses.tex
working on revisions

@@
   \reply{
-    TODO. \revref
+    Thanks for the suggestion. We have added this reference to the discussion section. \revref
   }
@@
 \end{point}
+
   \reply{
-  TODO. \revref
+  We have elaborated the list of citations for this point and have now cited Kimura as well. \revref
   }
@@
   \begin{point}{\revref} % 47:
-    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature.” (line 349-350) -> I'm not sure what this means. 
+    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature.” (line 349-350) $\to$ I'm not sure what this means. 
     It's unclear to me why background selection will interfere with a sweep signature. 

### 2025-05-22 analysis2_manuscript.tex
Revision steps (#79)
* adding decision letter to repo

* changes to schematic (issue #77)

* adding OG submission pdf to the repo for posterity

* added citation requested by AE
added review response stuff
working on revisions

@@
 \newcommand{\igcomment}[1]{\textcolor{orange}{IG: #1}}
+% Responses to reviews
+\input{review-response-commands}
+% set this to show line numbers and include responses to reviews or not
+\newif\ifreviewresponses
+\reviewresponsestrue  % include them
+% \reviewresponsesfalse  % don't include them
+\newcommand{\responsefile}{review-responses}  % name of the review responses file
+
 \linenumbers
@@
     from other evolutionary processes such as mutation, recombination,
-    and genetic drift \citep[e.g.,][]{gillespie1991causes}.
+    and genetic drift \citep[e.g.,][]{ohta1973slightly,lewontin1973distribution, kimura1985neutral, tajima1989statistical,gillespie1991causes, kern2018neutral}. \revpoint{2}{4}
     For instance, changes in population size can lead to fluctuations
@@
     local exon density (Figure \ref{fig:power-exon}).
-    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature.
+    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature. \revpoint{2}{7}
     This might suggest that the effects of realistic background selection on genome variation
@@
     (e.g., Figures \ref{fig:human-demography} and \ref{fig:vaquita-demography}).
+    Indeed our results echo those of \citet{marsh2024biases} who found \revpoint{AE}{1}
+    similar results on realistic human simulations that included selection.
     However, the question remains to what extent this would be true in empirical data where
@@
 \stopsupplement
+
+\includereviews
 \end{document}

### 2025-05-22 review-responses.tex
working further on revisions

@@
 
-  \begin{point}{\revref} %line 212
+  \begin{point}{\revref} 
     Relatedly, why is the dominance coefficient currently limited to between 0 and 1? It's fine that it is for now, but I was curious what makes it more difficult to include other values of h.
@@
 
-  \begin{point}{\revref} 
+  \begin{point}{} 
     I think it would be helpful to say a bit more about the vaquita earlier in the manuscript. Although some of the authors have worked on the vaquita recently, it is still a fairly obscure organism, 
@@
   \reply{
-     TODO. \revref
+     TODO. 
   }
 
-  \begin{point}{\revref} % 
+  \begin{point}{} % 
     It seems as though everything on the backend is being done by SLiM in the demonstrations here 
@@
   \reply{
-      TODO. \revref
+      TODO. 
   }
@@
 
-  \begin{point}{\revref} % line 577
+  \begin{point}{} 
     Related to the previous point, I found it a bit odd to that both the style of the DFE (and in particular the h/s relationship) 
@@
   \reply{
-    TODO. \revref
+    TODO. 
   }
@@
 
-  \begin{point}{\revref} % line 577
+  \begin{point}{} % line 577
     Last comment, and this may be an idiosyncratic opinion on my part, so I do not expect the authors to implement, 
@@
   \reply{
-    Don't respond to this?. \revref
+    Don't respond to this?. 
   }
@@
 
-  \begin{point}{\revref} % 27:
+  \begin{point}{} % 27:
     What fraction of the genome is under selection in the human genome you simulated?
@@
   \reply{
- TODO. \revref
-  }
-
-  \begin{point}{\revref} % 35:
-   Bayesian $\to$ probabilistic (I think this is more general and would encompas any model-based approach, say max likelihood etc.; alternatively consider using model-based, but probabilistic/statistical assumes we have an underlying model)
-  \end{point}
-
-  \reply{
-  Done.
+ For simulations with selection we assumed that selected mutations could only fall within CDS regions,
+ thus approximately 1\% of the genome is under selection.
   }
@@
   \reply{
-  TODO.
+  We have added a citation to a review paper from Wolfgang Stephan that discusses this point from a number of different perspectives. \revref
   }

### 2025-05-22 analysis2_manuscript.tex
working further on revisions

@@
     each mutation type is associated with a dominance coefficient,
-    (currently limited to values between 0 and 1, inclusive),
+    (currently limited to values between 0 and 1, inclusive), \revpoint{1}{2}
     the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
@@
     genomic elements (e.g., exons, introns, intergenic regions, etc.)
-    whose coordinates are specified using a GFF3 annotation file.
+    whose coordinates are specified using a GFF3 annotation file. \revpoint{2}{1}
     More concretely, consider implementing a model of selection consistent with the
@@
     \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep{huber2018gene}
-    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below).
+    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below). \revpoint{1}{5}
 
@@
     shows that the presence of selection decreases polymorphism only slightly
-    at the genome-wide level, but more substantially within exonic regions (Figure \ref{fig:homsap-summary-stats}).
+    at the genome-wide level, but more substantially within exonic regions (Figure \ref{fig:homsap-summary-stats}). \revpoint{1}{9}
     %IG: no need to specify the two scenarios shown in fig
@@
     examining the difference between the actual and inferred distributions of $s$ can be more helpful for showing how well a method
-    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C).
+    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C). \revpoint{2}{3}
     The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
@@
     As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$),
-    while \grapes considerably over-estimates the fraction of mutations under weak selection ($s\geq-0.0001$).
+    while \grapes considerably over-estimates the fraction of mutations under weak selection ($s\geq-0.0001$). \revpoint{2}{2}
 
@@
     local exon density (Figure \ref{fig:power-exon}).
-    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature. \revpoint{2}{7}
+    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature (reviewed in \citet{stephan2010genetic}) \revpoint{2}{6}
     This might suggest that the effects of realistic background selection on genome variation
@@
     Overall, these results are consistent with prior work showing that it is challenging to separately
-    infer the DFE of $s$ along with $h$ \citep{veeramah2014evidence, kyriazis2024constraining, balick2022overcoming}
+    infer the DFE of $s$ along with $h$ \citep{veeramah2014evidence, kyriazis2024constraining, balick2022overcoming} \revpoint{1}{6}
     and that a large proportion ($>5\%$) of recessive deleterious mutations can confound inferences of $s$ \citep{wade2023quantifying}.
@@
     Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome,
-    show a similar fit to the SFS and may be more biologically realistic than fully additive models,
+    show a similar fit to the SFS and may be more biologically realistic than fully additive models, \revpoint{1}{7}
     given the evidence for the presence of recessive deleterious mutations.

### 2025-05-22 review-responses.tex
continued

@@
   \begin{quote}
+    \color{gray}
     Two experts in population genetics examined this manuscript and gave Very High and Medium marks in the Impact of the resource reported. 
@@
   \begin{quote}
+    \color{gray}
     In this manuscript, a large team of authors (Gower, Pope, Rodrigues, Tittes et al) present the next 
@@
   \begin{quote}
+    \color{gray}
     In addition to announcing the new features available in stdpopsim, the paper describes three applications of the software: 
@@
   \begin{quote}
+    \color{gray}
     I am generally supportive of the paper and of the broader stdpopsim project. 
@@
   \reply{
-    TODO. \revref
+    Thanks for the suggestion. $\pi$ is being computed across all three populations,
+    as the point we were trying to make was about the effect of selection locally.
+    We have updated the figure legend to clarify this. \revref
   }
@@
   \begin{quote}
+    \color{gray}
     In this manuscript, Gower et al. extend the functionality of stdpopsim to include selection. 
@@
   \reply{
- For simulations with selection we assumed that selected mutations could only fall within CDS regions,
- thus approximately 1\% of the genome is under selection.
+ For simulations with selection we assumed that selected mutations could only fall within exonic regions,
+ thus approximately 2.1\% of the genome is under selection in human simulations and 2.2\% in vaquita simulations.
   }

### 2025-05-23 analysis2_manuscript.tex
continued
rev1 point 2

@@
     each mutation type is associated with a dominance coefficient,
-    (currently limited to values between 0 and 1, inclusive), \revpoint{1}{2}
+    \revpoint{1}{2}
     the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
@@
     or from exonic regions only (labeled exons-only).
+    Here we show the results of $\pi$ computed across all three populations (CEU, CHB, YRI)
+    taken together (i.e., not separately for each population).
     Each violin plot summarizes three replicate simulations generated under the relevant scenario. 
@@
 \end{figure}
-
+\revpoint{1}{9}
 \begin{figure}[h]

### 2025-06-06 review-responses.tex
response to 1 8

@@
   \reply{
-      TODO. \revref
+      This was a mistake in the original submission. $h$ is not limited to values between 0 and 1, and the 
+      implementation allows for overdominance and underdominance. The clause has been removed. \revref
   }
@@
 
-  \begin{point}{} 
+  \begin{point}{\revref} 
     Related to the previous point, I found it a bit odd to that both the style of the DFE (and in particular the h/s relationship) 
@@
   \reply{
-    TODO. 
+    Thank you for the suggestion. 
+    While it is possible to mix demographies and DFEs in stdpopsim catalog and study the effects, 
+    this is outside of the motivation for this paper.
+    Varying both DFE and demography in a systematic way and studying the effects would be interesting and useful, 
+    but there is less to learn from only swapping the DFE and demography of two species in this context.
+    We have clarified the point in the manuscript that when a given species of interest does not have published demographic and DFE inferences, 
+    one can be borrowed from another species as a next best option.  
   }

### 2025-06-06 analysis2_manuscript.tex
response to 1 8

@@
     While the DFEs are from a limited number of species, these DFEs can be applied to other species
-    in the catalog if and when it is suitable.
+    in the catalog if and when it is suitable,
+    namely, when a species of interest lacks genomic resources \revpoint{1}{8}.
     For instance, one could simulate a butterfly species with a gamma-distributed DFE originally

### 2025-06-09 analysis2_manuscript.tex
- Cleaned up old comments - Set up response to 1.5
Wrote responses to some reviewer comments + made associated edits to manuscript

--- 2025-05-23 analysis2_manuscript.tex
@@
         %citations aren't always allowed in captions. We should check with MBE.
-        %IG:I removed citations from all fig captions. They are prominently specified in the main text
         Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model
@@
     at the genome-wide level, but more substantially within exonic regions (Figure \ref{fig:homsap-summary-stats}). \revpoint{1}{9}
-    %IG: no need to specify the two scenarios shown in fig
-    %under both a constant population size and the out-of-Africa demographic model
     Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods:
-    %IG: removed numbering and reordered to match fig 2
     \stairway \citep{liu2020stairway}, \msmc \citep{Schiffels2020}, \smcpp \citep{terhorst2017robust} and \gone \citep{santiago2020recent}.
@@
     is only readily apparent within exonic regions (Figure \ref{fig:vaquita-summary-stats}).
-    %IG: last sentence seems like an unnecessary distraction. The constant-size simulations are only used to make a minor point in the next section
-    %We also see that the two-epoch demographic model used in the simulations actually leads to a slight increase
-    %in diversity versus the constant population size model.
-    %IG: merged paragraphs
     For these simulations, the SFS-based method (\stairway) produces more accurate
@@
     the performance of different methods for inferring the DFE.
-    %IG: Moved up the methods here to provide context for their mention in the next para
     Here, we considered three DFE-inference methods:
@@
     (also used by \stdpopsim; see \textbf{Methods}).
-    %IG - citations suggested in earlier versions to be added (I'm not sure they're necessary)
-    % precursor of polyDFE  from 2017:     https://academic.oup.com/genetics/article/207/3/1103/5930676
-    % fastDFE - the successor to polyDFE:  https://academic.oup.com/mbe/article/41/5/msae070/7641109
-    % paper that uses GRAPES from 2020:    https://academic.oup.com/genetics/article/216/2/559/6066183
 
@@
     many replicate simulations, we focused on human chromosome 1 and generated replicate simulations on 5cM segments
-    %IG: took out the 7 segments with low rec rate
     of the chromosome in 93 different locations. % (see \textbf{Methods}).
@@
         \caption{
-        %IG: better not mention 100 if we're filtering out 7 segments
-        %IG: removed citations from fig captions; they are prominently specified in main text
         Power to detect selective sweeps along human
@@
 
+    \subsection*{The DFE Class}
+    \revpoint{1}{5}
+    The DFE class has the following attributes for modeling natural selection in a given set of annotated genomic segments.
+    TBA
+
     \subsection*{Quality control and software development}
@@
     are programmatically tested for equality.
-    %IG: tried to put a more positive spin on the DFE bug example
     This QC process has proven its effectiveness on numerous occasions in detecting implementation bugs
@@
 
-    % ADK: I've added a section below that describes the DFE inference methods used in this study.
     To facilitate direct comparison between inference methods,
     we scale each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
-    %IG: expanded a bit to justify each scaling decision
     \grapes outputs the mean and shape of the inferred distribution of $4N_es$.
@@
     which is critical for the linked effects of a sweep, was held constant.
-    %IG: moved this up here since it makes mroe sense to describe it as if these segments were never simulated
-    % also converted to rec rate from cM. Please check!
     Seven segments in which the central 1Mb had an average recombination rate below $10^{-11}$ were excluded from the analysis.
@@
     and a model with background selection from deleterious mutations in exons (BGS).
-    %IG: deducted 7000 datasets for the low-rec regions
     Each null model was represented by a different set of $93,000$ simulations generated according to the appropriate model
@@
     positions on chromosome 1 as follows:
-    %IG: the 1000 simulations are with hard sweep + BGS
     each sweep-detection method was applied to the 1,000
@@
 
-%IG: moved supp table here, because otherwise it appeared between the supp figs
-%IG: edited caption and removed annotations column, since they're all exons now.
-%IG: moved commented "all sites" rows to bottom
 \begin{table}[h]
@@
 \end{figure}
-
-%IG: this fig is not cited anywhere and does not appear to provide crucial insights
-%\begin{figure}[h]
-%    \centering
-%    \includegraphics[width=\textwidth]{figures/HomSap/Constant/HomSap_Constant_Gamma_K17_all_sites_DFE_plot}
-%    \caption{
-%    \label{fig:homsap-dfe.constant.all_sites}
-%    Performance of methods to infer the distribution of fitness effects (DFE) from genetic data
-%    under constant population size  where all sites along the simulated chromosome can experience selected and neutral mutation
-%    as opposed to exonic regions only.\\
-%    }
-%\end{figure}
 

### 2025-06-09 review-responses.tex
Added responses to several comments by rev 1 & 2 + added relevant text to manuscript

--- 2025-05-22 review-responses.tex
@@
   \reply{
-  We have added further descriptions of the implementation to the TODO
+     %IG: I added this because the section I added for comment 1.5 below seems relevant here as well.
+     %    I think that this addresses the first part of the rev's comment, but details need to also be added for selective sweeps
+     We added a new section to the methods that provides detials on how DFE models are implemented using DFE objects \revref.
+     The end of this section contains a brief description of how the simulation engine uses the information specified in a given DFE object when sampling a new mutation.
+     TBD -- DETAILS ABOUT SELECTIVE SWEEPS.
+
   }
@@
   \reply{
-      TODO. \revref
+      This was a mistake in the original submission. The dominance coefficient is not limited to values between 0 and 1, and the 
+      implementation allows for overdominance and underdominance. This clause has been removed \revref,
+      %IG: I added a note about this in the 2nd part of the newly added methods section
+      and in the newly added methods section about DFE models, we added an explicit note that dominance coefficients can be negative or larger than 1. 
   }
@@
 
-  \begin{point}{\revref} % line 577
+  \begin{point}{} % line 577
     Point of clarification about the flexibility of DFE specification: Is the DFE allowed to be a mixture distribution in a single genomic region? 
@@
   \reply{
- TODO
+     We added a section to the methods with more details on \stdpopsim's DFE class, which answers these questions \revref.
+     In particular, we clarified that each DFE model is a mixture of different mutation types,
+     and we specified the distribution families that are currently supported by \stdpopsim.
+     We also described in greater detail how the relationship between dominance and selection coefficients is implemented,
+     which as the reviewer guessed, is done separately for each mutation type in the mixture.
+     This framework was designed to be very flexible, while allowing the implementation of most DFE models with few lines of code.
   }
@@
 
-  \begin{point}{\revref} % line 577
+  \begin{point}{} % line 577
     It might bear further comment that the gamma distribution is not flexible enough to accommodate the distribution of hs in the vaquita simulation 
@@
   \reply{
-    TODO. \revref
+    Following this insightful comment, we added a few sentences to the discussion addressing this issue \revref:
+    %IG: I added this quote of the text I added to the discussion to clarify our response to the reviewer.
+    %    We'll need to update after we finalize the version in the revision.
+    \textit{``
+    Our analysis suggests that the distributions of $s$ estimated by these methods
+    provide a reasonable approximation for the true distribution of $hs$ (Figure \ref{fig:vaquita-dfe}C).
+    This points to a notable limitation of most existing DFE inference methods,
+    which assume that selection coefficients are gamma-distributed.
+    As shown in Figure \ref{fig:vaquita-dfe}C,
+    while a gamma distribution might fit the distribution of $s$, a relationship between $h$ and $s$ might result in
+    $hs$ which is not well approximated by a gamma distribution.
+    "}
   }
@@
   \reply{
-    Thanks for the suggestion. $\pi$ is being computed across all three populations,
-    as the point we were trying to make was about the effect of selection locally.
-    We have updated the figure legend to clarify this. \revref
+    Thanks for the suggestion. In Fig. S1, $\pi$ is computed across all simulated genomes from all three populations,
+    as we wish to focus on the comparison between diversity estimates in different genomic segments in different simulation scenarios.
+    As the results are qualitatively similar when examining each of the three populations separately,
+    we chose to summarize them in a single plot using a measures of $\pi$ across all three populations.
+    We have updated the figure legend to clarify this \revref.
   }
 
-  \begin{point}{} % line 577
+  \begin{point}{}
     Last comment, and this may be an idiosyncratic opinion on my part, so I do not expect the authors to implement, 
@@
   \reply{
-    Don't respond to this?. 
+    %IG: I don't think we need to respond to this, but I did try to draft a polite response that clarifies how we see the paper
+    %    Feel free to shorten/remove this, if you think this is calling for trouble
+    We appreciate this comment.
+    The structure of this manuscript is indeed something we put a lot of thought into because of its somewhat unorthodox nature.
+    The main objective of the manuscript is to report on new capabilities of \stdpopsim,
+    so that the broad scientific community could use them to produce novel scientific discoveries.
+    The analyses presented in the manuscript are meant to demonstrate how features added to \stdpopsim
+    could be used to evaluate inference methods and test hypotheses regarding the effect of natural selection on patterns of diversity.
+    Consequently, the results of these analyses largely lead to more open questions than resolved questions,
+    and this is reflected in the discussion section, as noted by the reviewer.
+    While this is a bit unorthodox, we found this type of structure to be very useful in highlighting
+    \stdpopsim's features in our two previous publications \citep{adrion2020community,lauterbur2023expanding}.
+    We thus structured this manuscript similarly.
+
   }
@@
   \reply{
-  TODO.
+  %IG: I don't think that error bars are appropriate here and I tried to explain why
+  The main purpose of the plots in Figure 3C is to show the actual differences between simulated
+  and inferred distributions of $s$, since differences in mean and shape (shown in panels A-B) are not
+  always very informative or insightful.
+  To do this, we considered the average inferred DFE across the three replicate simualtions.
+  We do not see much use in adding error bars to these plots reflecting the variation across the three replicates.
+  Because of the somewhat noisy nature of this comparison, we only draw conclusions regarding qualitative differences
+  between inferred and simulated distributions, namely,
+  that  \polydfe and \dadicli tend to over-estimate the fraction of mutations under strong selection
+  amd \grapes tends to over-estimate the fraction of mutations under weak selection.
+  %IG: please check that this is correct
+  Regarding the estimates of $\text{E}[s]$ shown in Figure 3A (and also Figure 4A and associated supplementary figures),
+  these are typically values produced directly by the inference methods.
+  In the last paragraph of the methods section on DFE inference, we provide a complete description of the scaling
+  we apply to the output of each method so that they are directly comparable to $s$ as simulated in \slim.
   }
@@
   \reply{
-  We have added a citation to a review paper from Wolfgang Stephan that discusses this point from a number of different perspectives. \revref
+  We have added a citation to a review paper from Wolfgang Stephan that discusses this point from a number of different perspectives \revref.
   }

### 2025-06-09 analysis2_manuscript.tex
Added responses to several comments by rev 1 & 2 + added relevant text to manuscript

@@
     and the parameters of this distribution.
+    %IG: I moved this mention of h-s relationship up to here with a forward ref to newly added methods section
+    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep[][see \textbf{Methods}]{huber2018gene}.
+    %by defining ranges of selection coefficients and associating each range with a different dominance coefficient. 
+    %IG: I moved this label to the newly added methods section --> \revpoint{1}{5}
     Each \texttt{Annotation} object represents a collection of functional
@@
     with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).
-    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep{huber2018gene}
-    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below). \revpoint{1}{5}
 
@@
     Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}, Table \ref{tab:dfe_table}).
-    Note that while the estimation accuracy for the DFE parameters themselves is informative (Figure \ref{fig:homsap-dfe.ooa}A-B),
+    Note that while the estimation accuracy for the DFE parameters themselves is informative (Figure \ref{fig:homsap-dfe.ooa}A-B),\revpoint{2}{3}
     examining the difference between the actual and inferred distributions of $s$ can be more helpful for showing how well a method
-    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C). \revpoint{2}{3}
+    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C). 
     The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
@@
     or an apparent decrease in the amount of selection.
-    Fitting a mis-specified DFE that assumes $h=0.5$ will then lead to underestimates of $E[s]$.
-    Indeed, the distribution of $hs$ is much more accurately estimated (Figure \ref{fig:vaquita-dfe}C).
+    Thus, fitting a mis-specified DFE that assumes $h=0.5$ leads to underestimates of $E[s]$. \revpoint{1}{7}
+    %IG: added this to respond to comment 1.7.
+    %IG: I quote this in the response, so if we edit this, we should copy modified version to the response.
+    Our analysis suggests that the distributions of $s$ estimated by these methods
+    provide a reasonable approximation for the true distribution of $hs$ (Figure \ref{fig:vaquita-dfe}C).
+    This points to a notable limitation of most existing DFE inference methods,
+    which assume that selection coefficients are gamma-distributed.
+    As shown in Figure \ref{fig:vaquita-dfe}C,
+    while a gamma distribution might fit the distribution of $s$, a relationship between $h$ and $s$ might result in
+    $hs$ which is not well approximated by a gamma distribution.
+
+    %IG: split this into a separate paragraph because this ended up being a bit long
     Overall, these results are consistent with prior work showing that it is challenging to separately
@@
     Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome,
-    show a similar fit to the SFS and may be more biologically realistic than fully additive models, \revpoint{1}{7}
+    show a similar fit to the SFS and may be more biologically realistic than fully additive models, 
     given the evidence for the presence of recessive deleterious mutations.
@@
 
-    \subsection*{The DFE Class}
-    \revpoint{1}{5}
-    The DFE class has the following attributes for modeling natural selection in a given set of annotated genomic segments.
-    TBA
+    %IG: added this section to address rev's comment about lacking details in implementation
+    \subsection*{Implementing a DFE model using \stdpopsim's \texttt{DFE} class}
+
+    The \texttt{DFE} class 
+    has the following attributes for modeling natural selection in a given set of annotated genomic segments.\revpoint{1}{1}\revpoint{1}{5}
+    As with demographic models, each \texttt{DFE} object is associated with a unique identifier, description strings for documentation,
+    and a citation object for the publication containing information on the inference on which the DFE model is based.
+    Each \texttt{DFE} object also specifies a mixture of mutation types,
+    each associated with a frequency and a distribution of the seletion coefficient ($s$).
+    Thus, each \texttt{DFE} object contains a list of \texttt{MutationType} objects
+    and a separate list (of identical length) of positive frequencies (that sum up to 1).
+    Each \texttt{MutationType} object is associated with a distribution type, distribution parameters, and dominance coefficients.
+    The distribution types that are currently supported by \stdpopsim are:
+    \begin{itemize}
+       \item Fixed       (\texttt{f}), with one parameter specifying the fixed selection coefficient.
+       \item Exponential (\texttt{e}), with one parameter specifying the mean of the exponential distribution.
+       \item Gamma       (\texttt{g}), with two parameters specifying the mean and shape of the gamma distribution.
+       \item Normal      (\texttt{n}), with two parameters specifying the mean and standard deviation of the normal distribution.
+       \item Weibull     (\texttt{w}), with two parameters specifying the scale and shape of the Weibull distribution.
+       \item Uniform     (\texttt{u}), with two parameters specifying the minimum and maximum of the interval.
+       \item Positive log-normal  (\texttt{lp}), with two parameters specifying the mean and standard deviation of the normal distribution in log scale.
+       \item Negative log-normal  (\texttt{ln}), with two parameters specifying the mean and standard deviation of the normal distribution in log scale.
+    \end{itemize}
+
+    Typically, one would associate a \texttt{MutationType} object with a single dominance coefficient (using the \texttt{dominance\_coeff} attribute):
+    $h=0$ for recessive mutations, $h=\frac 12$ corresponding to an additive model (default value), and $h=1$ for completely dominant mutations.
+    The dominance coefficient, $h$,  can take any real value, including negative values (for underdominance)
+    and values greater than $1$ (overdominance).
+    There is also an option to model a relationship between the selection and dominance coefficients
+    by associating different ranges of selection coefficients with different dominance coefficients.
+    This is done using two attributes of the \texttt{MutationType} class:
+    \texttt{dominance\_coeff\_list} is used to specify a list of $n$ dominance coefficients,
+    and \texttt{dominance\_coeff\_breaks} is used to specify $n-1$ selection coefficient break points.
+    The $i$th dominance coefficient in the list is associated with selection coefficients in the range
+    between the $i-1$th and the $i$th break points.
+    For an explicit example, see the description of the vaquita porpoise genome simulations below.
+
+    During simulation, when a mutation is generated in an annotated region associated with a given \texttt{DFE} object,
+    a mutation type is sampled according to the frequencies of the different types.
+    Then, a selection coefficient is sampled for that mutation based on the distribution associated with the sampled mutation type.
+    Finally, a dominance coefficient is associated with the mutation either based on the fixed dominance coefficient of the mutation type
+    (using attribute \texttt{dominance\_coeff}), or based on the sampled selection coefficient
+    (using attributes \texttt{dominance\_coeff\_list} and \texttt{dominance\_coeff\_breaks}).    
+    For additional implementation details on the \texttt{DFE} class, see the {\stdpopsim docs} at
+    \url{https://popsim-consortium.github.io/stdpopsim-docs/main/introduction.html}.
 
@@
     This discretized relationship matches that used by \textcite{robinson2022critically},
-    and is easily implemented in \stdpopsim's \texttt{DFE} class.
+    and is easily implemented in \stdpopsim's \texttt{DFE} class (see above). %IG: added ref to newly added methods section
     Each simulation generates 100 diploid genomes,
@@
     or from exonic regions only (labeled exons-only).
-    Here we show the results of $\pi$ computed across all three populations (CEU, CHB, YRI)
-    taken together (i.e., not separately for each population).
+    %IG: edited here for simplicity and clarity
+    Here, $\pi$ is computed across all simulated genomes (from the three simulated populations)
+    considered as a single population.
+    %three populations (CEU, CHB, YRI) taken together (i.e., not separately for each population).
     Each violin plot summarizes three replicate simulations generated under the relevant scenario. 

### 2025-06-12 review-responses.tex
Response to factor of two review comment. Also updates to methods to clarify scaling.

--- 2025-05-22 review-responses.tex
@@
   \reply{
-TODO. \revref
+As noted in the paper, conversion of selection inferences between tools is challenging.
+We thus consulted with the developers of \polydfe and \grapes to confirm our conversions.
+(The developer of \dadicli is an author on this paper.)
+The inaccuracy of \grapes is a surprise to us.
+Note that \grapes is also less accurate than \polydfe and \dadicli in inferring the shape of the distribution (Fig.~\ref{fig:homsap-dfe.ooa}B), which suggests the inaccuracy is not simply a scaling issue.
   }

### 2025-06-12 analysis2_manuscript.tex
Response to factor of two review comment. Also updates to methods to clarify scaling.

--- 2025-06-06 analysis2_manuscript.tex
@@
     While the DFEs are from a limited number of species, these DFEs can be applied to other species
-    in the catalog if and when it is suitable,
-    namely, when a species of interest lacks genomic resources \revpoint{1}{8}.
+    in the catalog if and when it is suitable.
     For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
@@
     reflecting the fact that it defines the selection coefficient such that a homozygote has fitness $1+ 2s$.
-    \dadicli infers a gamma distribution for $N_es$ and parameterizes this distribution using the shape ($\alpha$) and rate ($\nu$).
-    %IG: we're still missing justification for the scaling by L and p, but we can leave this for revision
-    Thus, the mean value of $s$ is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$,
+    \dadicli defines the selection coefficient similarly to \polydfe and infers a gamma distribution for $2 N_e s$, parameterized by the shape ($\alpha$) and rate ($\nu$).
+    It also infers the total population-scaled mutation rate for selected sites over the considered region $\theta = 4 N_e \mu\, L\, p$,
     where $L$ is the total number of sites available for the analysis,
     and $p$ is the proportion of sites under selection.
+    %IG: we're still missing justification for the scaling by L and p, but we can leave this for revision
+    Thus, the mean value of $s$ in \slim terms is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$.
     Complete details for the DFE inference analysis can be found in the

### 2025-06-13 analysis2_manuscript.tex
Sbt dfe response (#83)
* response to 1 8

* response to 1 8 minor edit

@@
     While the DFEs are from a limited number of species, these DFEs can be applied to other species
-    in the catalog if and when it is suitable.
+    if and when it is suitable,
+    namely, when a species of interest lacks a pubished DFE in the catalog \revpoint{1}{8}.
     For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
@@
     reflecting the fact that it defines the selection coefficient such that a homozygote has fitness $1+ 2s$.
-    \dadicli defines the selection coefficient similarly to \polydfe and infers a gamma distribution for $2 N_e s$, parameterized by the shape ($\alpha$) and rate ($\nu$).
-    It also infers the total population-scaled mutation rate for selected sites over the considered region $\theta = 4 N_e \mu\, L\, p$,
+    \dadicli infers a gamma distribution for $N_es$ and parameterizes this distribution using the shape ($\alpha$) and rate ($\nu$).
+    %IG: we're still missing justification for the scaling by L and p, but we can leave this for revision
+    Thus, the mean value of $s$ is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$,
     where $L$ is the total number of sites available for the analysis,
     and $p$ is the proportion of sites under selection.
-    %IG: we're still missing justification for the scaling by L and p, but we can leave this for revision
-    Thus, the mean value of $s$ in \slim terms is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$.
     Complete details for the DFE inference analysis can be found in the

### 2025-06-25 analysis2_manuscript.tex
Further description of sweeps for ref 1

@@
     frequency at the beginning and end of the sweep period.
+    The conditioning is implemented via rejection sampling; if the frequency of the advantageous allele 
+    drifts out of the prescribed bounds, the simulation is restarted from the introduction of 
+    the advantageous allele. If there are competing advantageous alleles (e.g. simultaneous sweeps at distinct loci) 
+    then the simulation is restarted from the introduction of the first of these.
     As this functionality is implemented on top of the existing \stdpopsim API, it may be combined with other 
@@
     \url{https://popsim-consortium.github.io/stdpopsim-docs/main/introduction.html}.
+
+    % NSP: added to address reviewer 1's comments asking for more detail. Not adding
+    % any detail regarding the lower-level interface (particular "extended events") as
+    % these may change.
+    \subsection*{Selective sweeps and rejection sampling}
+
+    Selective sweeps are implemented in the \slim engine by
+    introducing a mutation with a fitness effect of $h s$ where $h$ is the
+    dominance coefficient (as for the DFE model) and $s \geq 0$ is the
+    selection coefficient.  The allele is introduced in a given population at
+    a specified time and genomic position, and may be either globally or locally
+    adaptive (so that the fitness effect is confined to the population of
+    origin) over a fixed duration. To model sweeps from standing genetic
+    variation (so-called ``soft sweeps''), the allele may be introduced at a
+    given frequency, in which case carriers are selected at random at the
+    initiation of the sweep.
+
+    Crucially, the advantageous allele must remain extant over the course of
+    the sweep; this conditioning is accomplished via rejection sampling,
+    whereby the simulation is restarted from the onset of the sweep if the
+    allele drifts out of the population of origin (or alternatively, below a
+    user-specified frequency). If there are competing advantageous alleles
+    (simultaneous sweeps at distinct loci) and one of these drifts out of
+    bounds, then the simulation is restarted from the onset of the first sweep;
+    this avoids conditioning on particular (partial) allele frequency
+    trajectories for the remaining alleles.  A downside of this strategy is
+    that rejection becomes more likely with increasing numbers of simultaneous
+    sweeps or with hard-to-satisfy constraints; but it allows for greater
+    flexibility in sweep specification and for interoperability with the
+    \stdpopsim DFE model.
 

### 2025-07-06 analysis2_manuscript.tex
Response to reviewer factor of 2 comment, plus methods clarification.
Update analysis2_manuscript.tex
Here are my edits on the revision. I added some context on the vaquita in a couple places. I also switched out the word "benchmark" in a few places to make it seem less like we're claiming we did a proper benchmarking. This is in line with the spirit of the response to the last comment from the first reviewer.

@@
     for species-specific genomic annotations and published DFE estimates.
-    We demonstrate the utility of this framework by benchmarking methods for demographic inference,
+    We demonstrate the utility of this framework by comparing methods for demographic inference,
     DFE estimation, and selective sweep detection across several species and scenarios.
@@
     This extension to \stdpopsim{} provides a powerful new resource for the population genetics community
-    to explore the interplay between selection and other evolutionary forces in a reproducible, low-barrier framework.
+    to explore the interplay between selection and other evolutionary forces in a reproducible, user-friendly framework.
 
@@
     Indeed, the DFE labeled \texttt{HomSap/Mixed\_K23} is recommended by \citet{kyriazis2023using}
-    for ``general'' use across mammals (lacking more specific information).
+    for ``general'' use across mammals, when lacking a more specific DFE.
     Furthermore, the user can specify a custom DFE and provide their own annotations
@@
     To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
-    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
+    we performed a similar analysis using the vaquita porpoise \textit{Phocoena sinus}.
+The vaquita offers an intriguing contast to the demographic history of humans.
+While human poulations continue to expand, the vaquita is critically endangered, with only approxmiately 10 individauls remaining.
+Previous demogrpahic inference from genetic data \textcite{robinson2022critically} found a long-term effective popualtion size of 4,485 indiviuals which then decreased to 2,807 individuals 2,162 generations ago.
+The population then severley decreased even more to its current size in the last few generations.
 We simulated 100 genomes under a two-epoch demographic model inferred for vaquita by \textcite{robinson2022critically},
@@
 
-    Next, we benchmarked these three DFE-inference methods on simulations of the vaquita porpoise genome.
-    Recall that these simulations apply a gamma-distributed selection coefficient to mutations in exons
-    with a relationship between the selection and dominance coefficients (see \textbf{Methods}).
+    Next, we examined the performance of these three DFE-inference methods when assuming a different DFE.
+To do this, we again turned to the vaquita.
+Unlike the human simulations where all mutations were assumed to be additive (h=0.5), the vaquita DFE applied a gamma-distributed selection coefficient to mutations in exons
+    with a relationship between the selection and dominance coefficients (see \textbf{Methods}). Specifically, more deleterious mutations in the vaquita DFE are assumed to be more recessive. 
     In this setting, all methods perform uniformly worse than in the human genome simulations,
@@
     that enables the simulation of genetic variation in the presence of selection.
-    We have demonstrated the utility of this new API by showing how it can be used to benchmark
+    We have demonstrated the utility of this new API by showing several illustritive exampls of how it can be used to benchmark
     the performance of different methods for demographic history inference, the inference of the distribution
@@
     Although we have shown how the new \stdpopsim API can be used to benchmark the performance of different methods,
-    the benchmarks shown here merely scratch the surface of what can be done with the new API.
+    the comparisons shown here merely scratch the surface of what can be done with the new API.
     We envision that, moving forward, many researchers will use
@@
     and $p$ is the proportion of sites under selection.
-    %IG: we're still missing justification for the scaling by L and p, but we can leave this for revision
     Thus, the mean value of $s$ in \slim terms is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$.

### 2025-07-08 review-responses.tex
fig3C error bars and runtime fig.

@@
     which assume that selection coefficients are gamma-distributed.
+    % DRS: see comment in main text on whether the limitation is the gamma for s or the fact that h is fixed
     As shown in Figure \ref{fig:vaquita-dfe}C,
@@
   and inferred distributions of $s$, since differences in mean and shape (shown in panels A-B) are not
-  always very informative or insightful.
-  To do this, we considered the average inferred DFE across the three replicate simualtions.
-  We do not see much use in adding error bars to these plots reflecting the variation across the three replicates.
-  Because of the somewhat noisy nature of this comparison, we only draw conclusions regarding qualitative differences
-  between inferred and simulated distributions, namely,
-  that  \polydfe and \dadicli tend to over-estimate the fraction of mutations under strong selection
-  amd \grapes tends to over-estimate the fraction of mutations under weak selection.
-  %IG: please check that this is correct
+  easy to judge assess what practical impact the parameter differences have on the realized distributions. 
+  We have added standard error bars to the mean values in Figure 3C, but the variation among the two parameters themselves 
+  is better for directly assessing the differences among methods. 
   Regarding the estimates of $\text{E}[s]$ shown in Figure 3A (and also Figure 4A and associated supplementary figures),
-  these are typically values produced directly by the inference methods.
+  these values were produced directly by the inference methods, which output their own estimate of $\theta = 4N_{e}\mu$,
+  where $\mu$ is taken from as the value in the stdpopsim catalog for each species. 
+  We have clarified this in the main text.
   In the last paragraph of the methods section on DFE inference, we provide a complete description of the scaling

### 2025-07-08 analysis2_manuscript.tex
fig3C error bars and runtime fig.

--- 2025-06-27 analysis2_manuscript.tex
@@
     drifts out of the prescribed bounds, the simulation is restarted from the introduction of
+    % DRS: not editing this because I haven't looked at the implementation, but this doesn't seem right.
+    % Obviously the allele must be allowed to drift outside of the bounds during the simulation, as long as it finishes
+    % within those bounds. So if it is fixed during the sim when the bounds don't extend to 1.0 (or lost when the
+    % bounds don't extent to 0.0) then we restart, and we restart if at the end of the sweep period we are outside of
+    % the bounds. right? If what I am saying is correct then maybe this should be rephrased, and more importantly the
+    % corresponding part of the methods should be corrected.
     the advantageous allele. If there are competing advantageous alleles (e.g. simultaneous sweeps at distinct loci) 
@@
         (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
-        shown for each DFE inferred from all three simulated datasets,
+        shown for each DFE inferred from all three simulation replicates,
         with median values marked by horizontal bars
@@
         (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
-        white bars represent the distribution used in the simulation. 
+        white bars represent the distribution used in the simulation. Standard error of the mean for each bin is shown in black. 
         }
@@
         (C) Binned distribution of $2hs$ implied by the simulated DFE compared with
-        the distribution of $s$ implied by the average DFE inferred for each method from the three simulated datasets (averaging the inferred gamma parameters).
+        the distribution of $s$ implied by the average DFE inferred for each method from the three simulated replicates (averaging the inferred gamma parameters).
         The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
@@
     The exploration of performance among methods done here is not meant to be thorough
-    benchmarking, but rather to highlight the utility of the \stdpopsim resource.
+    benchmarking. Rather, we include these analyses as examples to highlight the utility of \stdpopsim
+    for both benchmarking and other tasks where researchers would benefit from the ability to perform
+    simulations using parts of published population genetic models. in a streamlined and quality-controlled fashion.
     Having said that, there are a number of interesting results that merit discussion.
@@
     which assume that selection coefficients are gamma-distributed.
+    %DRS: Is the limitatio here that selection coefficients are gamma-distributed, or rather that h is fixed?
     As shown in Figure \ref{fig:vaquita-dfe}C,
-    while a gamma distribution might fit the distribution of $s$, a relationship between $h$ and $s$ might result in
-    $hs$ which is not well approximated by a gamma distribution.
+    while a gamma distribution might fit the conditional distribution of $s$ given $h$, a relationship between $h$ and $s$ might result in a magrinal distribution for $hs$ which is not well approximated by a gamma distribution.
 
@@
     %IG: added this section to address rev's comment about lacking details in implementation
+    %DRS: I think this section looks great.
     \subsection*{Implementing a DFE model using \stdpopsim's \texttt{DFE} class}
@@
     user-specified frequency). If there are competing advantageous alleles
+    % DRS: see my earlier comment about how we implement the conditioning here
+    % and correct as needed
     (simultaneous sweeps at distinct loci) and one of these drifts out of
@@
         Mean absolute value of selection coefficient ($\lvert E(s) \rvert$, panel A) and shape parameter (panel B)
-        are shown for each DFE inferred from all three simulated datasets
+        are shown for each DFE inferred from all three simulation replicates
         with median values marked by horizontal bars
@@
 
+%sweep power vs exon density
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8 \textwidth]{figures/HomSap/sims_runtime_plot.pdf}
+    \caption{
+    Runtime of simulated human chromosomes with and without background selection and demography.
+    Each point represents a single simulation of a single chromosome under conditions labeled by facets.
+    }
+    \label{fig:sim-runtime}
+\end{figure}
+
 \stopsupplement

### 2025-07-16 review-responses.tex
some light edits/comments from dan

--- 2025-06-27 review-responses.tex
@@
   these are typically values produced directly by the inference methods.
-  In the last paragraph of the methods section on DFE inference, we provide a complete description of the scaling
+  In the last paragraph of the new methods section on DFE inference, we provide a complete description of the scaling
   we apply to the output of each method so that they are directly comparable to $s$ as simulated in \slim.

### 2025-07-16 analysis2_manuscript.tex
some light edits/comments from dan

@@
         and simulated values represented by dashed horizontal lines.
-        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
-        white bars represent the distribution used in the simulation. Standard error of the mean for each bin is shown in black. 
+        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated replicates (averaging the inferred gamma parameters);
+        white bars represent the distribution used in the simulation. 
         }
@@
     As shown in Figure \ref{fig:vaquita-dfe}C,
-    while a gamma distribution might fit the conditional distribution of $s$ given $h$, a relationship between $h$ and $s$ might result in a magrinal distribution for $hs$ which is not well approximated by a gamma distribution.
+    while a gamma distribution might fit the distribution of $s$, a relationship between $h$ and $s$ might result in
+    $hs$ which is not well approximated by a gamma distribution.
 
@@
 
-%sweep power vs exon density
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8 \textwidth]{figures/HomSap/sims_runtime_plot.pdf}
-    \caption{
-    Runtime of simulated human chromosomes with and without background selection and demography.
-    Each point represents a single simulation of a single chromosome under conditions labeled by facets.
-    }
-    \label{fig:sim-runtime}
-\end{figure}
-
 \stopsupplement

### 2025-07-22 analysis2_manuscript.tex
Update analysis2_manuscript.tex

@@
     As shown in Figure \ref{fig:vaquita-dfe}C,
-    while a gamma distribution might fit the distribution of $s$, a relationship between $h$ and $s$ might result in
-    $hs$ which is not well approximated by a gamma distribution.
+    while a gamma distribution might fit the conditional distribution of $s$ given $h$, a relationship between $h$ and $s$ might result in a magrinal distribution for $hs$ which is not well approximated by a gamma distribution.
 

### 2025-07-22 analysis2_manuscript.tex
Clarify sweep conditioning

@@
     % sweep interface
-    While the \texttt{DFE} class facilitates the simulation of non-neutral variation across a sequence,
-    it does not provide a means of specifying the origin and fate of particular selected alleles.
-    Thus, we have also implemented an API in \stdpopsim (\texttt{stdpopsim.selective\_sweep}) that enables 
-    selective sweeps to be simulated with a high degree of control.
-    This function augments a demographic model by introducing an advantageous mutation over a given time interval,
-    at a given position along the sequence; and conditions the simulation such that the mutation is above a particular 
-    frequency at the beginning and end of the sweep period.
-    The conditioning is implemented via rejection sampling; if the frequency of the advantageous allele 
-    drifts out of the prescribed bounds, the simulation is restarted from the introduction of
-    % DRS: not editing this because I haven't looked at the implementation, but this doesn't seem right.
-    % Obviously the allele must be allowed to drift outside of the bounds during the simulation, as long as it finishes
-    % within those bounds. So if it is fixed during the sim when the bounds don't extend to 1.0 (or lost when the
-    % bounds don't extent to 0.0) then we restart, and we restart if at the end of the sweep period we are outside of
-    % the bounds. right? If what I am saying is correct then maybe this should be rephrased, and more importantly the
-    % corresponding part of the methods should be corrected.
-    the advantageous allele. If there are competing advantageous alleles (e.g. simultaneous sweeps at distinct loci) 
-    then the simulation is restarted from the introduction of the first of these.
-    As this functionality is implemented on top of the existing \stdpopsim API, it may be combined with other 
-    models of selection and demography to simulate the combined effects of multiple, disparate evolutionary processes
-    on genetic variation in a population.
+    While the \texttt{DFE} class facilitates the simulation of non-neutral
+    variation across a sequence, it does not provide a means of specifying the
+    origin and fate of particular selected alleles.  Thus, we have also
+    implemented an API in \stdpopsim (\texttt{stdpopsim.selective\_sweep}) that
+    enables selective sweeps to be simulated with a high degree of control.
+    This function augments a demographic model by introducing a mutation at a
+    particular time and position along the sequence; sets the selection
+    coefficient of the mutation to a positive value over a given time interval;
+    and conditions the simulation such that the mutation is above a particular
+    frequency at the beginning and end of the sweep period.  The conditioning
+    is implemented via rejection sampling; if the advantageous allele is lost
+    or does not have a frequency exceeding a user-specified threshold at the
+    beginning or end of the sweep, the simulation is restarted from the
+    introduction of the advantageous allele.  If there are competing
+    advantageous alleles (e.g.  simultaneous sweeps at distinct loci) then the
+    simulation is restarted from the introduction of the first of these.  As
+    this functionality is implemented on top of the existing \stdpopsim API, it
+    may be combined with other models of selection and demography to simulate
+    the combined effects of multiple, disparate evolutionary processes on
+    genetic variation in a population.
 
@@
 
-    Selective sweeps are implemented in the \slim engine by
-    introducing a mutation with a fitness effect of $h s$ where $h$ is the
-    dominance coefficient (as for the DFE model) and $s \geq 0$ is the
-    selection coefficient.  The allele is introduced in a given population at
-    a specified time and genomic position, and may be either globally or locally
-    adaptive (so that the fitness effect is confined to the population of
-    origin) over a fixed duration. To model sweeps from standing genetic
-    variation (so-called ``soft sweeps''), the allele may be introduced at a
-    given frequency, in which case carriers are selected at random at the
-    initiation of the sweep.
+    Selective sweeps are implemented in the \slim engine by introducing a
+    mutation with a fitness effect of $h s$ where $h$ is the dominance
+    coefficient (as for the DFE model) and $s \geq 0$ is the selection
+    coefficient. The allele is introduced in a given population at a specified
+    time and genomic position, and may be either globally or locally adaptive
+    (so that the fitness effect is confined to the population of origin) over a
+    fixed duration. The start of the sweep may be offset from the introduction
+    of the allele, to model selection on standing genetic variation (so-called
+    ``soft sweeps'').
 
@@
     whereby the simulation is restarted from the onset of the sweep if the
-    allele drifts out of the population of origin (or alternatively, below a
-    user-specified frequency). If there are competing advantageous alleles
-    % DRS: see my earlier comment about how we implement the conditioning here
-    % and correct as needed
-    (simultaneous sweeps at distinct loci) and one of these drifts out of
-    bounds, then the simulation is restarted from the onset of the first sweep;
-    this avoids conditioning on particular (partial) allele frequency
-    trajectories for the remaining alleles.  A downside of this strategy is
-    that rejection becomes more likely with increasing numbers of simultaneous
-    sweeps or with hard-to-satisfy constraints; but it allows for greater
-    flexibility in sweep specification and for interoperability with the
-    \stdpopsim DFE model.
+    allele drifts out of the population of origin. The conditioning may
+    optionally be extended to reject trajectories where the frequency at the
+    beginning and end of the sweep period is below a specified threshold. If
+    there are competing advantageous alleles (simultaneous sweeps at distinct
+    loci) and one of these does not satisfy the acceptance conditions, then the
+    simulation is restarted from the onset of the first sweep; this avoids
+    conditioning on particular (partial) allele frequency trajectories for the
+    remaining alleles.  A downside of this strategy is that rejection becomes
+    more likely with increasing numbers of simultaneous sweeps or with
+    hard-to-satisfy constraints; but it allows for greater flexibility in sweep
+    specification and for interoperability with the \stdpopsim DFE model.
 

### 2025-07-22 analysis2_manuscript.tex
Sbt runtime and errorbars (#88)
* added updated fig

* fig3C error bars and runtime fig.

@@
         and simulated values represented by dashed horizontal lines.
-        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated replicates (averaging the inferred gamma parameters);
-        white bars represent the distribution used in the simulation. 
+        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
+        white bars represent the distribution used in the simulation. Standard error of the mean for each bin is shown in black. 
         }

### 2025-07-25 review-responses.tex
Update review-responses.tex

--- 2025-07-08 review-responses.tex
@@
      We added a new section to the methods that provides detials on how DFE models are implemented using DFE objects \revref.
-     The end of this section contains a brief description of how the simulation engine uses the information specified in a given DFE object when sampling a new mutation.
-     TBD -- DETAILS ABOUT SELECTIVE SWEEPS.
-
+     The end of this section contains a brief description of how the simulation engine uses the information specified in a given DFE object when sampling a new mutation,
+     and how the conditioning for selective sweeps is implemented
+     (indeed, by rejection sampling).
+     We've also added a small note earlier in the paper, at \llname{sweep_conditioning}.
   }
@@
   \reply{
-     TODO. 
+      Good idea; we have added a bit more context, at \revref and \llname{vaquita_note}.
   }
@@
 
-  \begin{point}{} % line 577
+  \begin{point}{}
     Point of clarification about the flexibility of DFE specification: Is the DFE allowed to be a mixture distribution in a single genomic region? 
@@
 
-  \begin{point}{\revref} % line 577
-    Sentence starting on line 383 about difficulty identifying h and s separately seems to be missing some important citations, 
-    e.g. Fuller et al 2019 (pmid 30962618) and citations therein, including Simons et al 2014; Wright 1937.
+  \begin{point}{\revref}
+      Sentence starting [here] % on line 383 
+      about difficulty identifying $h$ and $s$ separately seems to be missing some important citations, 
+      e.g. \textcite{fuller2019measuring}, % Fuller et al 2019 (pmid 30962618) and citations therein, 
+      including 
+      \textcite{simons2014deleterious}; % Simons et al 2014;
+      \textcite{wright1937distribution}. % Wright 1937.
   \end{point}
@@
   \reply{
-    TODO. \revref
+      Good point; done \revref; note that we didn't make the list of citations as long as we might have,
+      opting instead to direct readers to \textcite{fuller2019measuring} for the review.
   }
 
-  \begin{point}{} % line 577
+  \begin{point}{}
     It might bear further comment that the gamma distribution is not flexible enough to accommodate the distribution of hs in the vaquita simulation 
@@
   \reply{
-    Following this insightful comment, we added a few sentences to the discussion addressing this issue \revref:
-    %IG: I added this quote of the text I added to the discussion to clarify our response to the reviewer.
-    %    We'll need to update after we finalize the version in the revision.
-    \textit{``
-    Our analysis suggests that the distributions of $s$ estimated by these methods
-    provide a reasonable approximation for the true distribution of $hs$ (Figure \ref{fig:vaquita-dfe}C).
-    This points to a notable limitation of most existing DFE inference methods,
-    which assume that selection coefficients are gamma-distributed.
-    % DRS: see comment in main text on whether the limitation is the gamma for s or the fact that h is fixed
-    As shown in Figure \ref{fig:vaquita-dfe}C,
-    while a gamma distribution might fit the distribution of $s$, a relationship between $h$ and $s$ might result in
-    $hs$ which is not well approximated by a gamma distribution.
-    "}
+    Following this insightful comment, we added a few sentences to the discussion addressing this issue, at \revref.
+    % No need to provide the quote here as they can easily see the change directly using the link.
   }
 
-  \begin{point}{\revref} 
+  \begin{point}{}
     Related to the previous point, I found it a bit odd to that both the style of the DFE (and in particular the h/s relationship) 
@@
     We have clarified the point in the manuscript that when a given species of interest does not have published demographic and DFE inferences, 
-    one can be borrowed from another species as a next best option.  
+    one can be borrowed from another species as a next best option at \revref,
+    and added another note about the aims of the paper, at \llname{not_the_point}.
   }
 
-  \begin{point}{\revref} % line 577
+\begin{point}{\revref}
     Figure S1, right panel. Is pi being computed from all the genomes (CEU/CHB/YRI)? 
@@
 As noted in the paper, conversion of selection inferences between tools is challenging.
-We thus consulted with the developers of \polydfe and \grapes to confirm our conversions.
+We thus spent quite a bit of time on this exact point,
+and consulted with the developers of \polydfe and \grapes to confirm our conversions.
 (The developer of \dadicli is an author on this paper.)
@@
  thus approximately 2.1\% of the genome is under selection in human simulations and 2.2\% in vaquita simulations.
+ We've added notes to this effect, at \revref{} and \llname{vaquita_percent}.
   }
@@
   \begin{point}{\revref} % 47:
-    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature.” (line 349-350) $\to$ I'm not sure what this means. 
+    ``This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature.'' % (line 349-350)
+      $\to$ I'm not sure what this means. 
     It's unclear to me why background selection will interfere with a sweep signature. 

### 2025-07-25 analysis2_manuscript.tex

@@
 \newcommand{\dadicli}{\texttt{dadi-cli}\xspace}
+\newcommand{\dadi}{\texttt{dadi}\xspace}
 \newcommand{\grapes}{\texttt{GRAPES}\xspace}
@@
     from other evolutionary processes such as mutation, recombination,
-    and genetic drift \citep[e.g.,][]{ohta1973slightly,lewontin1973distribution, kimura1985neutral, tajima1989statistical,gillespie1991causes, kern2018neutral}. \revpoint{2}{4}
+    and genetic drift \citep[e.g.,][]{ohta1973slightly,lewontin1973distribution, kimura1985neutral, Tajima1989, gillespie1991causes, kern2018neutral}. \revpoint{2}{4}
     For instance, changes in population size can lead to fluctuations
@@
     and the targets of recent positive selection.
+    The examples use genomes of both humans (\textit{Homo sapiens})
+    and the critically endangered Vaquita porpoise (\textit{Phocoena sinus}),
+    which provides an interesting comparison due to its small recent population size. \revpoint{1}{3}
 
@@
     if and when it is suitable,
-    namely, when a species of interest lacks a pubished DFE in the catalog \revpoint{1}{8}.
+    namely, when a species of interest lacks a published DFE in the catalog \revpoint{1}{8}.
     For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
@@
     and conditions the simulation such that the mutation is above a particular
-    frequency at the beginning and end of the sweep period.  The conditioning
+    frequency at the beginning and end of the sweep period.  The conditioning \llabel{sweep_conditioning}
     is implemented via rejection sampling; if the advantageous allele is lost
@@
     and a constant recombination rate across the genome (see \textbf{Methods}).
+    This is particularly interesting because the model was inferred assuming neutrality (with \dadi)
+    and used to inform conservation efforts for this critically endangered species. \llabel{vaquita_note}
     In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
@@
     The exploration of performance among methods done here is not meant to be thorough
-    benchmarking. Rather, we include these analyses as examples to highlight the utility of \stdpopsim
+    benchmarking.
+    Rather, we include these analyses as examples to highlight the utility of \stdpopsim \llabel{not_the_point}
     for both benchmarking and other tasks where researchers would benefit from the ability to perform
-    simulations using parts of published population genetic models. in a streamlined and quality-controlled fashion.
+    simulations using parts of published population genetic models in a streamlined and quality-controlled fashion.
     Having said that, there are a number of interesting results that merit discussion.
@@
     Thus, fitting a mis-specified DFE that assumes $h=0.5$ leads to underestimates of $E[s]$. \revpoint{1}{7}
-    %IG: added this to respond to comment 1.7.
-    %IG: I quote this in the response, so if we edit this, we should copy modified version to the response.
     Our analysis suggests that the distributions of $s$ estimated by these methods
@@
     As shown in Figure \ref{fig:vaquita-dfe}C,
-    while a gamma distribution might fit the conditional distribution of $s$ given $h$, a relationship between $h$ and $s$ might result in a magrinal distribution for $hs$ which is not well approximated by a gamma distribution.
+    while a gamma distribution might fit the conditional distribution of $s$ given $h$, a relationship between $h$ and $s$ might result in a marginal distribution for $hs$ which is not well approximated by a gamma distribution.
 
@@
     Overall, these results are consistent with prior work showing that it is challenging to separately
-    infer the DFE of $s$ along with $h$ \citep{veeramah2014evidence, kyriazis2024constraining, balick2022overcoming} \revpoint{1}{6}
+    infer the DFE of $s$ along with $h$ (\textcite{wright1937distribution, veeramah2014evidence, simons2014deleterious, fuller2019measuring, balick2022overcoming, kyriazis2024constraining}; reviewed in \textcite{fuller2019measuring}) \revpoint{1}{6}
     and that a large proportion ($>5\%$) of recessive deleterious mutations can confound inferences of $s$ \citep{wade2023quantifying}.
@@
 
+    \subsection*{Selective sweeps and rejection sampling}
+
     % NSP: added to address reviewer 1's comments asking for more detail. Not adding
@@
     % these may change.
-    \subsection*{Selective sweeps and rejection sampling}
+    %
+    %% For reference, from the docs on selective_sweep():
+    % Creates a list of extended events corresponding to a selective sweep at a
+    % single site (see :meth:`Contig.add_single_site`) that may subsequently be
+    % passed to the simulation engine. The sequence of events is:
+
+    %     (1) In generation `mutation_generation_ago`, a **neutral** mutation is
+    %         introduced into `population`.
+
+    %     (2) In generation `start_generation_ago`, the mutation becomes
+    %         beneficial (whether globally or locally is controlled by the
+    %         `globally_adaptive` option). If the frequency of the mutation in
+    %         `population` is below `min_freq_at_start` in this generation, the
+    %         simulation will restart from `mutation_generation_ago`.
+
+    %     (3) In the generation after `end_generation_ago`, the mutation reverts to
+    %         neutrality in `population`. If the frequency of the mutation in
+    %         `population` is below `min_freq_at_end` in this generation, the
+    %         simulation will restart from `mutation_generation_ago`.
+
+    % During the sweep, the fitnesses of individuals that are homozygous or
+    % heterozygous for the beneficial mutation are `1 + selection_coeff` and  `1
+    % + selection_coeff * dominance_coeff`, respectively. If `globally_adaptive`
+    % is `True` (the default), then all individuals carrying the mutation are
+    % impacted regardless of the population they belong to. Otherwise, the
+    % mutation only influences fitness within `population`.
+
+    % If the mutation is lost in `population` at any point between
+    % `mutation_generation_ago` and `end_generation_ago`, then the simulation
+    % will restart from `mutation_generation_ago`. The effect of "restarting"
+    % (aka "rejection sampling") is that the resulting simulations are
+    % conditioned on the mutation being extant after time
+    % `mutation_generation_ago` with frequency above `min_freq_at_start` at
+    % `start_generation_ago` and `min_freq_at_end` at `end_generation_ago`.
+
+    % Because the simulation will restart until the allele frequency conditions
+    % are satisfied, care must be taken to ensure that these are reasonable. For
+    % example, if `start_generation_ago` is very close to
+    % `mutation_generation_ago` and `min_freq_at_start` is substantially greater
+    % than zero, the vast majority of proposed allele frequency trajectories will
+    % be rejected and the simulation may take an infeasibly long time to
+    % complete.
 
     Selective sweeps are implemented in the \slim engine by introducing a
-    mutation with a fitness effect of $h s$ where $h$ is the dominance
-    coefficient (as for the DFE model) and $s \geq 0$ is the selection
+    mutation with a fitness effect of $h s$ in heterozygotes and $s$ in homozygotes,
+    where $h$ is the dominance coefficient (as for the DFE model) and $s \geq 0$ is the selection
     coefficient. The allele is introduced in a given population at a specified
@@
     the sweep; this conditioning is accomplished via rejection sampling,
-    whereby the simulation is restarted from the onset of the sweep if the
-    allele drifts out of the population of origin. The conditioning may
+    whereby the simulation is restarted from the time of introduction of the mutation if the
+    allele drifts is lost in the population of origin. The conditioning may
     optionally be extended to reject trajectories where the frequency at the
@@
     \citet[\stdpopsim label \texttt{Gamma\_K17}]{kim2017inference} to exons annotated by the HAVANA group release 104
-    \citep[\stdpopsim label \texttt{ensembl\_havana\_104\_exons}]{ensembl2018}.
+    \citep[\stdpopsim label \texttt{ensembl\_havana\_104\_exons}]{ensembl2018},
+    which covers 2.1\% of the genome. \revpoint{2}{5}
     Each simulation generated 100 diploid genomes from each of three extant human populations (labeled YRI, CEU, and CHB),
@@
     \citet[\stdpopsim label \texttt{Gamma\_R22}]{robinson2022critically} to exons annotated from the vaquita porpoise genome assembly
-    \citet[\stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}]{morin2021}.
+    \citet[\stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}]{morin2021},
+    which covers 2.2\% of the genome. \llabel{vaquita_percent}
     This DFE model implements a relationship between selection and dominance as follows:
@@
     and $p$ is the proportion of sites under selection.
-    Thus, the mean value of $s$ in \slim terms is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$.
+    Thus, the mean value of $s$ in \slim terms is obtained by  $4\mu \alpha \nu L p / \theta$.
     Complete details for the DFE inference analysis can be found in the

### 2025-07-28 review-responses.tex
adding runtime fig and response to rev; fixing revref command to handle spaces after (#92)

@@
   \reply{
-      TODO. 
+    We have now included two additional figures showing the runtimes for both the whole genome simulations (used in demographic and DFE inference) and the 5Mb segment simulations of human chromosome 1 (used in the analyses of sweep detection) at \revref and \llname{sim-sweeps-runtime}.
+    With respect to the sweep simulations, we show that the simulation of sweeps in each of the segments is takes approximately 3 minutes \llname{sim-sweeps-runtime}.
+    Given we simulated 1000 replicates of each one of the 100 segments, this means that the total runtime for the sweep simulations is approximately 300,000 minutes, or 5,000 hours.
+    However, whole chromosome 1 simulations with BGS and sweeps would have been much more time consuming, each taking approximately 200 minutes \revref (or in total, 20,000,000 minutes, or 333,333 hours).
+
   }

### 2025-07-28 analysis2_manuscript.tex
adding runtime fig and response to rev; fixing revref command to handle spaces after (#92)

@@
     at the genome-wide level, but more substantially within exonic regions (Figure \ref{fig:homsap-summary-stats}). \revpoint{1}{9}
+
     Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods:
     \stairway \citep{liu2020stairway}, \msmc \citep{Schiffels2020}, \smcpp \citep{terhorst2017robust} and \gone \citep{santiago2020recent}.
-
     Although the estimates produced by the four methods differ from one another,
@@
     the performance of different methods for detecting selective sweeps from genetic data.
-    Because simulation of selective sweeps is very time consuming and evaluation of statistical power requires
+    Because simulation of selective sweeps is time consuming and evaluation of statistical power requires
     many replicate simulations, we focused on human chromosome 1 and generated replicate simulations on 5cM segments
@@
     where the code utilized is similar to that shown in Figure~\ref{fig:schematic}B.
+    Simulation runtimes are shown in Figure \ref{fig:sim-runtime}. \revpoint{1}{4}
     The complete simulation workflow is available at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/simulation.snake}
@@
 
-    To test methods for detecting selective sweeps, we considered 100 evenly distributed 5cM
+    To test methods for detecting selective sweeps, we considered 100 evenly distributed 5Mb
     segments of human chromosome 1.
-    As recombination rate varies across the genome, the size of the region in base pairs simulated varied, however the recombination distance,
-    which is critical for the linked effects of a sweep, was held constant.
+    We included a 2.5cM buffer on either side of each segment to minimize the effect of chromosome ends on patterns of genetic variation.
     Seven segments in which the central 1Mb had an average recombination rate below $10^{-11}$ were excluded from the analysis.
@@
     used by the three sweep-detection methods (see below).
+    See Figure \ref{fig:sim-sweeps-runtime} for the distribution of runtimes (in minutes) for the simulations of the 100 segments of human chromosome 1. \llabel{sim-sweeps-runtime}
 
@@
 
-%sweep power vs exon density
+%demography sims runtime
 \begin{figure}[h]
@@
 
+%sweep sims runtime
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8 \textwidth]{figures/sweeps/runtime.png}
+    \caption{
+    Distribution of runtimes for the simulated segments of human chromosome 1 under different regimes used in the analyses of sweep detection.
+    }
+    \label{fig:sim-sweeps-runtime}
+\end{figure}
+
 \stopsupplement

### 2025-07-28 review-responses.tex
typo fix

@@
     We have now included two additional figures showing the runtimes for both the whole genome simulations (used in demographic and DFE inference) and the 5Mb segment simulations of human chromosome 1 (used in the analyses of sweep detection) at \revref and \llname{sim-sweeps-runtime}.
-    With respect to the sweep simulations, we show that the simulation of sweeps in each of the segments is takes approximately 3 minutes \llname{sim-sweeps-runtime}.
+    With respect to the sweep simulations, we show that the simulation of sweeps in each of the segments takes approximately 3 minutes \llname{sim-sweeps-runtime}.
     Given we simulated 1000 replicates of each one of the 100 segments, this means that the total runtime for the sweep simulations is approximately 300,000 minutes, or 5,000 hours.

### 2025-07-29 review-responses.tex
feat: added details about overlapping annotations and response

@@
   \reply{
-    TODO. \revref
+    We have added a note to the implementation section about how we handle overlapping annotations in a GFF file.
+    In short, we do not allow for overlapping annotations in a GFF file, and we prioritize the annotation being modeled (e.g., CDS) 
+    over potentially overlapping annotations of other types (e.g., exons). 
+    Further, we do not currently support multiple overlapping annotations of different types.
+    \revref.
   }

### 2025-07-29 analysis2_manuscript.tex
feat: added details about overlapping annotations and response

@@
     genomic elements (e.g., exons, introns, intergenic regions, etc.)
-    whose coordinates are specified using a GFF3 annotation file. \revpoint{2}{1}
+    whose coordinates are specified using a GFF3 annotation file. 
+    For the purposes of \stdpopsim, we do not allow for overlapping annotations -- 
+    genomic elements of the given annotation being modeled (e.g., CDS) are prioritized over overlapping annotations of other types (e.g., exons)
+    and we do not currently support multiple overlapping annotations of different types. \revpoint{2}{1}
     More concretely, consider implementing a model of selection consistent with the

### 2025-07-29 analysis2_manuscript.tex
Apply suggestions from code review
Co-authored-by: Peter Ralph <petrel.harp@gmail.com>

--- 2025-07-25 analysis2_manuscript.tex
@@
 \newcommand{\dadicli}{\texttt{dadi-cli}\xspace}
-\newcommand{\dadi}{\texttt{dadi}\xspace}
 \newcommand{\grapes}{\texttt{GRAPES}\xspace}
@@
     from other evolutionary processes such as mutation, recombination,
-    and genetic drift \citep[e.g.,][]{ohta1973slightly,lewontin1973distribution, kimura1985neutral, Tajima1989, gillespie1991causes, kern2018neutral}. \revpoint{2}{4}
+    and genetic drift \citep[e.g.,][]{ohta1973slightly,lewontin1973distribution, kimura1985neutral, tajima1989statistical,gillespie1991causes, kern2018neutral}. \revpoint{2}{4}
     For instance, changes in population size can lead to fluctuations
@@
     and the targets of recent positive selection.
-    The examples use genomes of both humans (\textit{Homo sapiens})
-    and the critically endangered Vaquita porpoise (\textit{Phocoena sinus}),
-    which provides an interesting comparison due to its small recent population size. \revpoint{1}{3}
 
@@
     if and when it is suitable,
-    namely, when a species of interest lacks a published DFE in the catalog \revpoint{1}{8}.
+    namely, when a species of interest lacks a pubished DFE in the catalog \revpoint{1}{8}.
     For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
@@
     % sweep interface
-    While the \texttt{DFE} class facilitates the simulation of non-neutral
-    variation across a sequence, it does not provide a means of specifying the
-    origin and fate of particular selected alleles.  Thus, we have also
-    implemented an API in \stdpopsim (\texttt{stdpopsim.selective\_sweep}) that
-    enables selective sweeps to be simulated with a high degree of control.
-    This function augments a demographic model by introducing a mutation at a
-    particular time and position along the sequence; sets the selection
-    coefficient of the mutation to a positive value over a given time interval;
-    and conditions the simulation such that the mutation is above a particular
-    frequency at the beginning and end of the sweep period.  The conditioning \llabel{sweep_conditioning}
-    is implemented via rejection sampling; if the advantageous allele is lost
-    or does not have a frequency exceeding a user-specified threshold at the
-    beginning or end of the sweep, the simulation is restarted from the
-    introduction of the advantageous allele.  If there are competing
-    advantageous alleles (e.g.  simultaneous sweeps at distinct loci) then the
-    simulation is restarted from the introduction of the first of these.  As
-    this functionality is implemented on top of the existing \stdpopsim API, it
-    may be combined with other models of selection and demography to simulate
-    the combined effects of multiple, disparate evolutionary processes on
-    genetic variation in a population.
+    While the \texttt{DFE} class facilitates the simulation of non-neutral variation across a sequence,
+    it does not provide a means of specifying the origin and fate of particular selected alleles.
+    Thus, we have also implemented an API in \stdpopsim (\texttt{stdpopsim.selective\_sweep}) that enables 
+    selective sweeps to be simulated with a high degree of control.
+    This function augments a demographic model by introducing an advantageous mutation over a given time interval,
+    at a given position along the sequence; and conditions the simulation such that the mutation is above a particular 
+    frequency at the beginning and end of the sweep period.
+    The conditioning is implemented via rejection sampling; if the frequency of the advantageous allele 
+    drifts out of the prescribed bounds, the simulation is restarted from the introduction of 
+    the advantageous allele. If there are competing advantageous alleles (e.g. simultaneous sweeps at distinct loci) 
+    then the simulation is restarted from the introduction of the first of these.
+    As this functionality is implemented on top of the existing \stdpopsim API, it may be combined with other 
+    models of selection and demography to simulate the combined effects of multiple, disparate evolutionary processes
+    on genetic variation in a population.
 
@@
     To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
-    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
+    we performed a similar analysis using the vaquita porpoise \textit{Phocoena sinus}.
+The vaquita offers an intriguing contast to the demographic history of humans.
+While human poulations continue to expand, the vaquita is critically endangered, with only approxmiately 10 individauls remaining.
+Previous demographic inference from genetic data \textcite{robinson2022critically} found a long-term effective population size of 4,485 indiviuals which then decreased to 2,807 individuals 2,162 generations ago.
+The population then severely decreased even more to its current size in the last few generations.
 We simulated 100 genomes under a two-epoch demographic model inferred for vaquita by \textcite{robinson2022critically},
     and a constant recombination rate across the genome (see \textbf{Methods}).
-    This is particularly interesting because the model was inferred assuming neutrality (with \dadi)
-    and used to inform conservation efforts for this critically endangered species. \llabel{vaquita_note}
     In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
@@
         (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
-        shown for each DFE inferred from all three simulation replicates,
+        shown for each DFE inferred from all three simulated datasets,
         with median values marked by horizontal bars
@@
         (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
-        white bars represent the distribution used in the simulation. Standard error of the mean for each bin is shown in black. 
+        white bars represent the distribution used in the simulation. 
         }
@@
         (C) Binned distribution of $2hs$ implied by the simulated DFE compared with
-        the distribution of $s$ implied by the average DFE inferred for each method from the three simulated replicates (averaging the inferred gamma parameters).
+        the distribution of $s$ implied by the average DFE inferred for each method from the three simulated datasets (averaging the inferred gamma parameters).
         The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
@@
     that enables the simulation of genetic variation in the presence of selection.
-    We have demonstrated the utility of this new API by showing how it can be used to benchmark
+    We have demonstrated the utility of this new API by showing several illustrative examples of how it can be used to benchmark
     the performance of different methods for demographic history inference, the inference of the distribution
@@
     The exploration of performance among methods done here is not meant to be thorough
-    benchmarking.
-    Rather, we include these analyses as examples to highlight the utility of \stdpopsim \llabel{not_the_point}
-    for both benchmarking and other tasks where researchers would benefit from the ability to perform
-    simulations using parts of published population genetic models in a streamlined and quality-controlled fashion.
+    benchmarking, but rather to highlight the utility of the \stdpopsim resource.
     Having said that, there are a number of interesting results that merit discussion.
@@
     Thus, fitting a mis-specified DFE that assumes $h=0.5$ leads to underestimates of $E[s]$. \revpoint{1}{7}
+    %IG: added this to respond to comment 1.7.
+    %IG: I quote this in the response, so if we edit this, we should copy modified version to the response.
     Our analysis suggests that the distributions of $s$ estimated by these methods
@@
     which assume that selection coefficients are gamma-distributed.
-    %DRS: Is the limitatio here that selection coefficients are gamma-distributed, or rather that h is fixed?
     As shown in Figure \ref{fig:vaquita-dfe}C,
-    while a gamma distribution might fit the conditional distribution of $s$ given $h$, a relationship between $h$ and $s$ might result in a marginal distribution for $hs$ which is not well approximated by a gamma distribution.
+    while a gamma distribution might fit the distribution of $s$, a relationship between $h$ and $s$ might result in
+    $hs$ which is not well approximated by a gamma distribution.
 
@@
     Overall, these results are consistent with prior work showing that it is challenging to separately
-    infer the DFE of $s$ along with $h$ (\textcite{wright1937distribution, veeramah2014evidence, simons2014deleterious, fuller2019measuring, balick2022overcoming, kyriazis2024constraining}; reviewed in \textcite{fuller2019measuring}) \revpoint{1}{6}
+    infer the DFE of $s$ along with $h$ \citep{veeramah2014evidence, kyriazis2024constraining, balick2022overcoming} \revpoint{1}{6}
     and that a large proportion ($>5\%$) of recessive deleterious mutations can confound inferences of $s$ \citep{wade2023quantifying}.
@@
     %IG: added this section to address rev's comment about lacking details in implementation
-    %DRS: I think this section looks great.
     \subsection*{Implementing a DFE model using \stdpopsim's \texttt{DFE} class}
@@
 
-    \subsection*{Selective sweeps and rejection sampling}
-
     % NSP: added to address reviewer 1's comments asking for more detail. Not adding
@@
     % these may change.
-    %
-    %% For reference, from the docs on selective_sweep():
-    % Creates a list of extended events corresponding to a selective sweep at a
-    % single site (see :meth:`Contig.add_single_site`) that may subsequently be
-    % passed to the simulation engine. The sequence of events is:
-
-    %     (1) In generation `mutation_generation_ago`, a **neutral** mutation is
-    %         introduced into `population`.
-
-    %     (2) In generation `start_generation_ago`, the mutation becomes
-    %         beneficial (whether globally or locally is controlled by the
-    %         `globally_adaptive` option). If the frequency of the mutation in
-    %         `population` is below `min_freq_at_start` in this generation, the
-    %         simulation will restart from `mutation_generation_ago`.
-
-    %     (3) In the generation after `end_generation_ago`, the mutation reverts to
-    %         neutrality in `population`. If the frequency of the mutation in
-    %         `population` is below `min_freq_at_end` in this generation, the
-    %         simulation will restart from `mutation_generation_ago`.
-
-    % During the sweep, the fitnesses of individuals that are homozygous or
-    % heterozygous for the beneficial mutation are `1 + selection_coeff` and  `1
-    % + selection_coeff * dominance_coeff`, respectively. If `globally_adaptive`
-    % is `True` (the default), then all individuals carrying the mutation are
-    % impacted regardless of the population they belong to. Otherwise, the
-    % mutation only influences fitness within `population`.
-
-    % If the mutation is lost in `population` at any point between
-    % `mutation_generation_ago` and `end_generation_ago`, then the simulation
-    % will restart from `mutation_generation_ago`. The effect of "restarting"
-    % (aka "rejection sampling") is that the resulting simulations are
-    % conditioned on the mutation being extant after time
-    % `mutation_generation_ago` with frequency above `min_freq_at_start` at
-    % `start_generation_ago` and `min_freq_at_end` at `end_generation_ago`.
-
-    % Because the simulation will restart until the allele frequency conditions
-    % are satisfied, care must be taken to ensure that these are reasonable. For
-    % example, if `start_generation_ago` is very close to
-    % `mutation_generation_ago` and `min_freq_at_start` is substantially greater
-    % than zero, the vast majority of proposed allele frequency trajectories will
-    % be rejected and the simulation may take an infeasibly long time to
-    % complete.
-
-    Selective sweeps are implemented in the \slim engine by introducing a
-    mutation with a fitness effect of $h s$ in heterozygotes and $s$ in homozygotes,
-    where $h$ is the dominance coefficient (as for the DFE model) and $s \geq 0$ is the selection
-    coefficient. The allele is introduced in a given population at a specified
-    time and genomic position, and may be either globally or locally adaptive
-    (so that the fitness effect is confined to the population of origin) over a
-    fixed duration. The start of the sweep may be offset from the introduction
-    of the allele, to model selection on standing genetic variation (so-called
-    ``soft sweeps'').
+    \subsection*{Selective sweeps and rejection sampling}
+
+    Selective sweeps are implemented in the \slim engine by
+    introducing a mutation with a fitness effect of $h s$ where $h$ is the
+    dominance coefficient (as for the DFE model) and $s \geq 0$ is the
+    selection coefficient.  The allele is introduced in a given population at
+    a specified time and genomic position, and may be either globally or locally
+    adaptive (so that the fitness effect is confined to the population of
+    origin) over a fixed duration. To model sweeps from standing genetic
+    variation (so-called ``soft sweeps''), the allele may be introduced at a
+    given frequency, in which case carriers are selected at random at the
+    initiation of the sweep.
 
@@
     the sweep; this conditioning is accomplished via rejection sampling,
-    whereby the simulation is restarted from the time of introduction of the mutation if the
-    allele drifts is lost in the population of origin. The conditioning may
-    optionally be extended to reject trajectories where the frequency at the
-    beginning and end of the sweep period is below a specified threshold. If
-    there are competing advantageous alleles (simultaneous sweeps at distinct
-    loci) and one of these does not satisfy the acceptance conditions, then the
-    simulation is restarted from the onset of the first sweep; this avoids
-    conditioning on particular (partial) allele frequency trajectories for the
-    remaining alleles.  A downside of this strategy is that rejection becomes
-    more likely with increasing numbers of simultaneous sweeps or with
-    hard-to-satisfy constraints; but it allows for greater flexibility in sweep
-    specification and for interoperability with the \stdpopsim DFE model.
+    whereby the simulation is restarted from the onset of the sweep if the
+    allele drifts out of the population of origin (or alternatively, below a
+    user-specified frequency). If there are competing advantageous alleles
+    (simultaneous sweeps at distinct loci) and one of these drifts out of
+    bounds, then the simulation is restarted from the onset of the first sweep;
+    this avoids conditioning on particular (partial) allele frequency
+    trajectories for the remaining alleles.  A downside of this strategy is
+    that rejection becomes more likely with increasing numbers of simultaneous
+    sweeps or with hard-to-satisfy constraints; but it allows for greater
+    flexibility in sweep specification and for interoperability with the
+    \stdpopsim DFE model.
 
@@
     \citet[\stdpopsim label \texttt{Gamma\_K17}]{kim2017inference} to exons annotated by the HAVANA group release 104
-    \citep[\stdpopsim label \texttt{ensembl\_havana\_104\_exons}]{ensembl2018},
-    which covers 2.1\% of the genome. \revpoint{2}{5}
+    \citep[\stdpopsim label \texttt{ensembl\_havana\_104\_exons}]{ensembl2018}.
     Each simulation generated 100 diploid genomes from each of three extant human populations (labeled YRI, CEU, and CHB),
@@
     \citet[\stdpopsim label \texttt{Gamma\_R22}]{robinson2022critically} to exons annotated from the vaquita porpoise genome assembly
-    \citet[\stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}]{morin2021},
-    which covers 2.2\% of the genome. \llabel{vaquita_percent}
+    \citet[\stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}]{morin2021}.
     This DFE model implements a relationship between selection and dominance as follows:
@@
     and $p$ is the proportion of sites under selection.
-    Thus, the mean value of $s$ in \slim terms is obtained by  $4\mu \alpha \nu L p / \theta$.
+    Thus, the mean value of $s$ in \slim terms is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$.
     Complete details for the DFE inference analysis can be found in the
@@
         Mean absolute value of selection coefficient ($\lvert E(s) \rvert$, panel A) and shape parameter (panel B)
-        are shown for each DFE inferred from all three simulation replicates
+        are shown for each DFE inferred from all three simulated datasets
         with median values marked by horizontal bars

### 2025-07-29 analysis2_manuscript.tex
edits from Kirk (#93)
* Update analysis2_manuscript.tex

Here are my edits on the revision. I added some context on the vaquita in a couple places. I also switched out the word "benchmark" in a few places to make it seem less like we're claiming we did a proper benchmarking. This is in line with the spirit of the response to the last comment from the first reviewer.

* Apply suggestions from code review

Co-authored-by: Peter Ralph <petrel.harp@gmail.com>

---------

Co-authored-by: Andrew Kern <adkern@uoregon.edu>
Co-authored-by: Peter Ralph <petrel.harp@gmail.com>

--- 2025-07-28 analysis2_manuscript.tex
@@
     To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
-    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
+    we performed a similar analysis using the vaquita porpoise \textit{Phocoena sinus}.
+The vaquita offers an intriguing contast to the demographic history of humans.
+While human poulations continue to expand, the vaquita is critically endangered, with only approxmiately 10 individauls remaining.
+Previous demographic inference from genetic data \textcite{robinson2022critically} found a long-term effective population size of 4,485 indiviuals which then decreased to 2,807 individuals 2,162 generations ago.
+The population then severely decreased even more to its current size in the last few generations.
 We simulated 100 genomes under a two-epoch demographic model inferred for vaquita by \textcite{robinson2022critically},

### 2025-07-29 review-responses.tex
Apply suggestions from code review
Co-authored-by: Peter Ralph <petrel.harp@gmail.com>

--- 2025-07-28 review-responses.tex
@@
   \reply{
-    TODO. \revref
+    We have added a note to the implementation section about how we handle overlapping annotations in a GFF file to \revref.
+    In short, annotations can't overlap (we remove overlapping bits);
+    see the documentation at \url{https://popsim-consortium.github.io/stdpopsim-docs/stable/api.html#stdpopsim.Contig.add_dfe}.
   }

### 2025-07-29 analysis2_manuscript.tex
Apply suggestions from code review
Co-authored-by: Peter Ralph <petrel.harp@gmail.com>

--- 2025-07-29 analysis2_manuscript.tex
@@
     For the purposes of \stdpopsim, we do not allow for overlapping annotations -- 
-    genomic elements of the given annotation being modeled (e.g., CDS) are prioritized over overlapping annotations of other types (e.g., exons)
-    and we do not currently support multiple overlapping annotations of different types. \revpoint{2}{1}
+    any interval covered by an annotation that is added to a given contig
+    is removed from any previously-added annotations. \revpoint{2}{1}
     More concretely, consider implementing a model of selection consistent with the

### 2025-07-30 review-responses.tex
Ilan's final edits - minor tweaks here and there

@@
      %    I think that this addresses the first part of the rev's comment, but details need to also be added for selective sweeps
-     We added a new section to the methods that provides detials on how DFE models are implemented using DFE objects \revref.
-     The end of this section contains a brief description of how the simulation engine uses the information specified in a given DFE object when sampling a new mutation,
-     and how the conditioning for selective sweeps is implemented
-     (indeed, by rejection sampling).
-     We've also added a small note earlier in the paper, at \llname{sweep_conditioning}.
+     We added two sections to the methods: one that provides detials on how DFE models are implemented using DFE objects and one describing how selective sweeps are modeled \revref.
+     The first section contains a brief description of how the simulation engine uses the information specified in a given DFE object when sampling a new mutation.
+     The second section describes how the conditioning for selective sweeps is implemented (indeed, by rejection sampling).
+     We've also added brief clarification regarding the implementation of sweeps earlier in the paper, at \llname{sweep_conditioning}.
   }
@@
   \reply{
-    We have now included two additional figures showing the runtimes for both the whole genome simulations (used in demographic and DFE inference) and the 5Mb segment simulations of human chromosome 1 (used in the analyses of sweep detection) at \revref and \llname{sim-sweeps-runtime}.
-    With respect to the sweep simulations, we show that the simulation of sweeps in each of the segments takes approximately 3 minutes \llname{sim-sweeps-runtime}.
+    We have now included two additional figures showing the runtimes for both the whole genome simulations (used in demographic and DFE inference) and the 5Mb segment simulations of human chromosome 1   (used in the analyses of sweep detection) in Figures \ref{fig:sim-runtime} and \ref{fig:sim-sweeps-runtime}.
+    With respect to the sweep simulations, we show that the simulation of sweeps in each of the segments takes approximately 3 minutes (Figure \ref{fig:sim-sweeps-runtime}).
     Given we simulated 1000 replicates of each one of the 100 segments, this means that the total runtime for the sweep simulations is approximately 300,000 minutes, or 5,000 hours.
-    However, whole chromosome 1 simulations with BGS and sweeps would have been much more time consuming, each taking approximately 200 minutes \revref (or in total, 20,000,000 minutes, or 333,333 hours).
-
+    However, whole chromosome 1 simulations with BGS and sweeps would have been much more time consuming, each taking approximately 200 minutes (Figure \ref{fig:sim-runtime}),
+    or in total, 20,000,000 minutes, or 333,333 hours).
   }
@@
     Thank you for the suggestion. 
-    While it is possible to mix demographies and DFEs in stdpopsim catalog and study the effects, 
+    While it is possible to mix demographies and DFEs in \stdpopsim's catalog and study the effects, 
     this is outside of the motivation for this paper.
@@
   \reply{
-    Thanks for the suggestion. In Fig. S1, $\pi$ is computed across all simulated genomes from all three populations,
+    In Fig. S1, $\pi$ is computed across all simulated genomes from all three populations,
     as we wish to focus on the comparison between diversity estimates in different genomic segments in different simulation scenarios.
@@
   \reply{
-  %IG: I don't think that error bars are appropriate here and I tried to explain why
   The main purpose of the plots in Figure 3C is to show the actual differences between simulated
@@
   easy to judge assess what practical impact the parameter differences have on the realized distributions. 
-  We have added standard error bars to the mean values in Figure 3C, but the variation among the two parameters themselves 
-  is better for directly assessing the differences among methods. 
+  Following the reviewer's suggestion, 
+  we added error bars depicting the standard error to the mean values in Figure 3C (reflecting variation among the three replicate simulations).
+  % 
   Regarding the estimates of $\text{E}[s]$ shown in Figure 3A (and also Figure 4A and associated supplementary figures),
@@
  For simulations with selection we assumed that selected mutations could only fall within exonic regions,
- thus approximately 2.1\% of the genome is under selection in human simulations and 2.2\% in vaquita simulations.
+ which cover 2.1\% of the human genome and 2.2\% of the vaquita genome.
+ In each of these simulations, 70\% of exonic sites are assumed to be under selection (with gamma-distributed selection coefficients)
+ and the remaining 30\% are assumed to be neutral.
  We've added notes to this effect, at \revref{} and \llname{vaquita_percent}.

### 2025-07-30 analysis2_manuscript.tex
feat: added details about overlapping annotations and response (#94)
* feat: added details about overlapping annotations and response

* Apply suggestions from code review

Co-authored-by: Peter Ralph <petrel.harp@gmail.com>

---------

Co-authored-by: Peter Ralph <petrel.harp@gmail.com>
Ilan's final edits - minor tweaks here and there

--- 2025-07-29 analysis2_manuscript.tex
@@
     genomic elements (e.g., exons, introns, intergenic regions, etc.)
-    whose coordinates are specified using a GFF3 annotation file. \revpoint{2}{1}
+    whose coordinates are specified using a GFF3 annotation file. 
     More concretely, consider implementing a model of selection consistent with the
@@
     to provide a rich model of how selection may vary along a chromosome.
-    For example, coding sequences may be associated with one DFE and non-coding sequences in exons with another DFE.
+    For example, coding sequences may be associated with one DFE and non-coding sequences in exons with another DFE. \revpoint{2}{1}
+    %%IG: I moved the comment about GFF overlaps here and edited because it fits here better than in the prev location.
+    We note that if a genomic interval belongs to more than one \texttt{Annotation} object,
+    then it will be eventually associated with the last \texttt{Annotation} object added
+    to the simulated contig and removed from any previously-added annotations. 
     Simulations with selection are then implemented by specifying a species,
@@
     advantageous alleles (e.g.  simultaneous sweeps at distinct loci) then the
-    simulation is restarted from the introduction of the first of these.  As
+    simulation is restarted from the introduction of the first of these (see \textbf{Methods}).  As
     this functionality is implemented on top of the existing \stdpopsim API, it
@@
         alongside the true values of $N_e$ used in simulation (black).
+        The X axis depicts time back from present (in years) and the Y axis depicts $N_e$ (number of individuals).
         }
@@
     To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
-    we performed a similar analysis using the vaquita porpoise \textit{Phocoena sinus}.
+    we performed a similar analysis using the vaquita porpoise \textit{Phocoena sinus}. \llabel{vaquita_note}
     The vaquita offers an intriguing contast to the demographic history of humans.
@@
     This is particularly interesting because the model was inferred assuming neutrality (with \dadi)
-    and used to inform conservation efforts for this critically endangered species. \llabel{vaquita_note}
+    and used to inform conservation efforts for this critically endangered species.
     In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
@@
         and simulated values represented by dashed horizontal lines.
-        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
-        white bars represent the distribution used in the simulation. Standard error of the mean for each bin is shown in black. 
+        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters),
+        with error bars depicting the standard error of the mean for each bin.
+        White bars represent the distribution used in the simulation. 
         }
@@
     %
-    We then applied each method to data simulated with sweeps for two samples (YRI and CEU)
+    We then applied each method to data simulated with sweeps in one of two populations (YRI and CEU)
     in each of the simulated locations along human chromosome 1 (Figure \ref{fig:chr1_power}).
@@
     Interestingly, we do not see a clear correlation between power to detect sweeps and
-    local exon density (Figure \ref{fig:power-exon}).
-    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature (reviewed in \citet{stephan2010genetic}) \revpoint{2}{6}
+    local exon density (Figure \ref{fig:power-exon}).  \revpoint{2}{6}
+    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature (reviewed by \citet{stephan2010genetic}).
     This might suggest that the effects of realistic background selection on genome variation
@@
     (e.g., Figures \ref{fig:human-demography} and \ref{fig:vaquita-demography}).
-    Indeed our results echo those of \citet{marsh2024biases} who found \revpoint{AE}{1}
+    Indeed, our results echo those of \citet{marsh2024biases} who found \revpoint{AE}{1}
     similar results on realistic human simulations that included selection.
@@
     This points to a notable limitation of most existing DFE inference methods,
-    which assume that selection coefficients are gamma-distributed.
-    %DRS: Is the limitatio here that selection coefficients are gamma-distributed, or rather that h is fixed?
+    which assume that fitness effects ($hs$) are gamma-distributed.
     As shown in Figure \ref{fig:vaquita-dfe}C,
-    while a gamma distribution might fit the conditional distribution of $s$ given $h$, a relationship between $h$ and $s$ might result in a marginal distribution for $hs$ which is not well approximated by a gamma distribution.
+    while a gamma distribution might fit the marginal distribution of $s$,
+    a relationship between $h$ and $s$ might result in fitness effects ($hs$),
+    which are not gamma-distributed.
 
@@
     these results suggest the need for care in DFE inference.
-    Kyriazis and Lohmueller \citeyear{kyriazis2024constraining} inferred the DFE of $s$ for nonsynonymous mutations
+    Kyriazis and Lohmueller (\citeyear{kyriazis2024constraining}) inferred the DFE of $s$ for nonsynonymous mutations
     in humans when considering different values of $h$.
@@
     has the following attributes for modeling natural selection in a given set of annotated genomic segments.\revpoint{1}{1}\revpoint{1}{5}
-    As with demographic models, each \texttt{DFE} object is associated with a unique identifier, description strings for documentation,
+    Similarly to other \stdpopsim classes, the \texttt{DFE} class has fields for specifying the model identifier, description strings for documentation,
     and a citation object for the publication containing information on the inference on which the DFE model is based.
-    Each \texttt{DFE} object also specifies a mixture of mutation types,
+    A \texttt{DFE} model is specified using a mixture of mutation types,
     each associated with a frequency and a distribution of the seletion coefficient ($s$).
@@
     whereby the simulation is restarted from the time of introduction of the mutation if the
-    allele drifts is lost in the population of origin. The conditioning may
+    allele is lost in the population of origin. The conditioning may
     optionally be extended to reject trajectories where the frequency at the
@@
     We note that aggressive scaling of forward-in-time simulations
-    can lead to quite different outcomes when selection in involved
-    \cite{uricchio2014robust,ferrari2024towards, dabi2025population},
+    can lead to quite different outcomes when selection is involved
+    \citep{uricchio2014robust,ferrari2024towards, dabi2025population},
     and we caution users against using high degrees of scaling for drawing
@@
     where the code utilized is similar to that shown in Figure~\ref{fig:schematic}B.
-    Simulation runtimes are shown in Figure \ref{fig:sim-runtime}. \revpoint{1}{4}
+    % Simulation runtimes are shown in Figure \ref{fig:sim-runtime}. \revpoint{1}{4}
     The complete simulation workflow is available at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/simulation.snake}
@@
     \citep[\stdpopsim label \texttt{ensembl\_havana\_104\_exons}]{ensembl2018},
-    which covers 2.1\% of the genome. \revpoint{2}{5}
+    which cover 2.1\% of the genome. \revpoint{2}{5}
     Each simulation generated 100 diploid genomes from each of three extant human populations (labeled YRI, CEU, and CHB),
@@
     for these simulations can be found at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_HomSap.yml}.
+    The running times of these simulations depend on chromosome length, the demographic model, and incorporation of selection (Figure \ref{fig:sim-runtime}).
 
@@
     \citet[\stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}]{morin2021},
-    which covers 2.2\% of the genome. \llabel{vaquita_percent}
+    which cover 2.2\% of the genome. \llabel{vaquita_percent}
     This DFE model implements a relationship between selection and dominance as follows:
@@
     This discretized relationship matches that used by \textcite{robinson2022critically},
-    and is easily implemented in \stdpopsim's \texttt{DFE} class (see above). %IG: added ref to newly added methods section
+    and is specified using the following attributes in the \texttt{DFE} class:
+    \texttt{dominance\_coeff\_list}$=(0,0.01,0.1,0.4)$, and
+    \texttt{dominance\_coeff\_breaks}$=(-0.1,-0.01,-0.001)$.
     Each simulation generates 100 diploid genomes,
@@
     used by the three sweep-detection methods (see below).
-    See Figure \ref{fig:sim-sweeps-runtime} for the distribution of runtimes (in minutes) for the simulations of the 100 segments of human chromosome 1. \llabel{sim-sweeps-runtime}
+    See Figure \ref{fig:sim-sweeps-runtime} for the distribution of runtimes for these simulations. \llabel{sim-sweeps-runtime}
 
@@
 HomSap & OOA 5R19  & 0.014 & 0.0111 & \bf{0.0072} & 0.031 & 0.051 & \bf{0.024} \\
-PhoSin & Constant  & 0.024 & 0.025 & 0.024 & \bf{0.23} & 0.25 & 0.23 \\
-PhoSin & Vaquita2Epoch 1R22  & 0.024 & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
+PhoSin & Constant  & \bf{0.024} & 0.025 & \bf{0.024} & \bf{0.23} & 0.25 & \bf{0.23} \\
+PhoSin & Vaquita2Epoch 1R22  & \bf{0.024} & 0.025 & \bf{0.024} & \bf{0.18} & 0.23 & 0.21 \\
 %HomSap & Constant & all sites & 0.0027 & 0.0111 & \bf{0.0018} & \bf{0.00071} & 0.029 & 0.014 \\
@@
     alongside the true values of $N_e$ used in simulation (black).
+    The X axis depicts time back from present (in years) and the Y axis depicts $N_e$ (number of individuals).
 }
@@
     \caption{
-    Runtime of simulated human chromosomes with and without background selection and demography.
+    Simulation runtimes for human chromosomes with and without selection applied to exons and under two demographic models:
+    a single population with constant size (left) and a three-population out-of-Africa model (right).
     Each point represents a single simulation of a single chromosome under conditions labeled by facets.
+    %IG: TODO: add spec of machine
     }
@@
     \caption{
-    Distribution of runtimes for the simulated segments of human chromosome 1 under different regimes used in the analyses of sweep detection.
+    Distribution of simulation runtimes for segments of human chromosome 1 simulated for the sweep detection analysis.
+    Runtimes are summarized separately for each of the six simulation conditions used in the analysis.
+    %IG: TODO: add spec of machine
     }

### 2025-07-30 analysis2_manuscript.tex
Update analysis2_manuscript.tex
Co-authored-by: Andrew Kern <adkern@uoregon.edu>

@@
     local exon density (Figure \ref{fig:power-exon}).  \revpoint{2}{6}
-    This is somewhat surprising given that background selection is expected to somewhat interfere with the sweep signature (reviewed by \citet{stephan2010genetic}).
+    This is somewhat surprising given that background selection may be expected to interfere with the sweep signature (reviewed by \citet{stephan2010genetic}).
     This might suggest that the effects of realistic background selection on genome variation
